---
title: "SGG Raw Plots"
author: "Traci Treleaven and Dave Bosworth"
date: "08/07/2020"
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Code to plot continuous water quality data and check for missing data during the study period.

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

# Import Data

```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/Processed_Data/Continuous"
  )
)

# Import SGG data
sgg_orig <- 
  read_csv(
    file = paste0(sharepoint_path, "/RTM_OUTPUT_SGG_formatted.csv"),
    col_types = paste0("ccc", str_c(rep("dc", 7), collapse = ""))
  )
```

```{r}
#change dateTime from char to special DateTime
sgg_clean <- sgg_orig %>% 
  mutate(DateTime = ymd_hms(DateTime)) 

```

The data in the correct format:
```{r glimpse data}
glimpse(sgg_clean)
```



# Create Plots

## Water Temperature

Simple line plots starting with a high level overview then going by year. We used the `ggplot` function from the ggplot2 package.

Water Temperature
```{r plot one param, warning = FALSE}
p <- sgg_clean %>% 
  ggplot(aes(x = DateTime, y = WaterTemp)) +
  geom_line()

print(p)
```


Water Temperature 2018 (About 4 months of data)
```{r plot WaterTemp 2018 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2018-08-16 00:00:00"), as_datetime("2018-12-31 00:00:00")))
```

Water Temperature 2019
```{r plot WaterTemp 2019 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2019-01-01 00:00:00"), as_datetime("2019-12-31 00:00:00"))) 
```

## Turbidity

Simple line plots starting with a high level overview then going by year. We used the `ggplot` function from the ggplot2 package.

Turbidity
```{r plot Turbidity param, warning = FALSE}
pt <- sgg_clean %>% 
  ggplot(aes(x = DateTime, y = Turbidity)) +
  geom_line()

print(pt)
```

Turbidity 2018 (About 4 months of data)
```{r plot Turbidity 2018 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2018-08-16 00:00:00"), as_datetime("2018-12-31 00:00:00")))
```

Turbidity 2019
```{r plot Turbidity 2019 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2019-01-01 00:00:00"), as_datetime("2019-12-31 00:00:00"))) 
```


## Specific Conductance

Simple line plots starting with a high level overview then going by year. We used the `ggplot` function from the ggplot2 package.

Specific Conductance (SpCnd)
```{r plot SC param, warning = FALSE}
ps <- sgg_clean %>% 
  ggplot(aes(x = DateTime, y = SpCnd)) +
  geom_line()

print(ps)
```

SpCnd 2018 (About 4 months of data)
```{r plot SC 2018 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2018-08-16 00:00:00"), as_datetime("2018-12-31 00:00:00"))) 
```

SpCnd 2019
```{r plot SC 2019 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2019-01-01 00:00:00"), as_datetime("2019-12-31 00:00:00"))) 
```


## Dissolved Oxygen

Simple line plots starting with a high level overview then going by year. We used the `ggplot` function from the ggplot2 package.

Dissolved Oxygen
```{r plot DO param, warning = FALSE}
pd <- sgg_clean %>% 
  ggplot(aes(x = DateTime, y = DO)) +
  geom_line()

print(pd)
```

Zoomed in to remove space with no data (8/18/2018 to 02/11/2019)
```{r plot DO zoomed param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2018-08-18 00:00:00"), as_datetime("2019-02-12 00:00:00"))) 
```

## pH

Simple line plots starting with a high level overview then going by year. We used the `ggplot` function from the ggplot2 package.

pH Total
```{r plot pH param, warning = FALSE}
pH <- sgg_clean %>% 
  ggplot(aes(x = DateTime, y = pH)) +
  geom_line()

print(pH)
```

pH Zoomed in to remove space with no data (8/18/2018 to 12/12/2018)
```{r plot pH zoomed param, warning = FALSE}
pH + scale_x_datetime(limits = c(as_datetime("2018-08-18 00:00:00"), as_datetime("2018-12-13 00:00:00"))) 
```


## Flow

Simple line plots starting with a high level overview then going by year. We used the `ggplot` function from the ggplot2 package.

Flow Total
```{r plot flow param, warning = FALSE}
pf <- sgg_clean %>% 
  ggplot(aes(x = DateTime, y = Flow)) +
  geom_line()

print(pf)
```

Flow 2018 (less than 4 month of data)

A large gap in the data between 10/10/2018-11/7/2018
```{r plot flow 2018 param, warning = FALSE}
pf + scale_x_datetime(limits = c(as_datetime("2018-09-19 00:00:00"), as_datetime("2018-12-31 00:00:00"))) 
```

Flow 2018
```{r plot flow 2019 param, warning = FALSE}
pf + scale_x_datetime(limits = c(as_datetime("2019-01-01 00:00:00"), as_datetime("2019-12-31 00:00:00"))) 
```




## Multiple parameters

Multiple parameters at once on a single plot with `facet_grid`. We removed the Quality Code variables and restructure the data to a long format using `pivot_longer`.
```{r pivot data longer}
sgg_long <- sgg_clean %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(
    cols = Flow:pH,
    names_to = "parameter",
    values_to = "value"
  )
```

Now the data is structured so that the parameter names are in the `parameter` variable and the data values are in the `value` variable. 
```{r display data in long format}
head(sgg_long)
```

A plot with facets. Use the `scales` argument set to `"free_y"` in `facet_grid` to adjust y-axis limits to the data of each parameter.
```{r plot with facets, warning = FALSE}
p2 <- sgg_long %>%
  ggplot(aes(x = DateTime, y = value)) +
  geom_line() +
  facet_grid(
    rows = vars(parameter),
    scales = "free_y"
  )

p2
```

The plot larger so we can see it better. This is done with the `fig.width` and `fig.height` options in the top of the code chunk.
```{r plot with facets larger, warning = FALSE, fig.width = 8.5, fig.height = 11}
p2
```



