---
title: "Code_cont_WQ_LIS_plots"
author: "Amanda Maguire"
date: "July 29, 2020"
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document provides an example of how to use Rmarkdown to bring together text, code, and its output in way to communicate code and results to others. This document also introduces some useful code to use to create simple plots of continuous water quality data.

Code is placed in code chunks like below:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

# Import Data

To start plotting, first we need to import some data of course:
```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/Processed_Data/Continuous"
  )
)

# Import Lisbon data
lis_orig <- 
  read_csv(
    file = paste0(sharepoint_path, "/RTM_OUTPUT_LIS_formatted.csv"),
    col_types = paste0("T", str_c(rep("d", 14), collapse = ""), "c")
  )
```

The data looks as we expect it to:
```{r glimpse data}
glimpse(lis_orig)
```

# Create Plots

## One parameter

Simple line plot of one parameter collected at the Lisbon station. We'll use the `ggplot` function from the ggplot2 package.

WaterTemp plots
```{r plot one param, warning = FALSE}
p <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = WaterTemp)) +
  geom_line()

print(p)
```


2013
```{r plot WaterTemp 2013 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00"))) 
```

2014
```{r plot WaterTemp 2014 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot WaterTemp 2015 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot WaterTemp 2016 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot WaterTemp 2017 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot WaterTemp 2018 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot WaterTemp 2019 param, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```


Turbidity plot
```{r plot Turbidity param, warning = FALSE}
pt <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = Turbidity)) +
  geom_line()

print(pt)
```

2013
```{r plot Turbidity 2013 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00"))) 
```

2014
```{r plot Turbidity 2014 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot Turbidity 2015 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot Turbidity 2016 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot Turbidity 2017 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot Turbidity 2018 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot Turbidity 2019 param, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```


SpCnd plot
```{r plot SpCnd param, warning = FALSE}
ps <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = SpCnd)) +
  geom_line()

print(pt)
```

2013
```{r plot SpCnd 2013 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00"))) 
```

2014
```{r plot SpCnd 2014 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot SpCnd 2015 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot SpCnd 2016 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot SpCnd 2017 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot SpCnd 2018 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot SpCnd 2019 param, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```

pH plots
```{r plot pH param, warning = FALSE}
ph <- lis_orig %>%
  ggplot(aes(x = DateTime, y = pH)) +
  geom_line()

print(ph)
```

2013
```{r plot pH 2013 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00")))
```

2014
```{r plot pH 2014 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot pH 2015 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot pH 2016 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot pH 2017 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot pH 2018 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot pH 2019 param, warning = FALSE}
ph + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```



DO plot
```{r plot DO param, warning = FALSE}
pd <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = DO)) +
  geom_line()

print(po)
```

2013
```{r plot DO 2013 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00")))
```

2014
```{r plot DO 2014 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot DO 2015 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot DO 2016 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot DO 2017 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot DO 2018 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot DO 2019 param, warning = FALSE}
pd + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```


Chla plot
```{r plot Chla param, warning = FALSE}
pc <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = Chla)) +
  geom_line()

print(pc)
```

2013
```{r plot Chla 2013 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00"))) 
```

2014
```{r plot Chla 2014 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot Chla 2015 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot Chla 2016 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot Chla 2017 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot Chla 2018 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot Chla 2019 param, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```

Flow plot
```{r plot Flow param, warning = FALSE}
pflo <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = Flow)) +
  geom_line()

print(pflo)
```

2013
```{r plot Flow 2013 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2013-07-15 00:00:00"), as_datetime("2013-11-05 00:00:00"))) 
```

2014
```{r plot Flow 2014 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2014-07-15 00:00:00"), as_datetime("2014-11-05 00:00:00")))
```

2015
```{r plot Flow 2015 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2015-07-15 00:00:00"), as_datetime("2015-11-05 00:00:00")))
```

2016
```{r plot Flow 2016 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2016-06-15 00:00:00"), as_datetime("2016-11-05 00:00:00")))
```

2017
```{r plot Flow 2017 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2017-07-15 00:00:00"), as_datetime("2017-11-05 00:00:00")))
```

2018
```{r plot Flow 2018 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2018-07-15 00:00:00"), as_datetime("2018-11-05 00:00:00")))
```

2019
```{r plot Flow 2019 param, warning = FALSE}
pflo + scale_x_datetime(limits = c(as_datetime("2019-07-15 00:00:00"), as_datetime("2019-11-05 00:00:00")))
```



## Multiple parameters

This is nice, but we can also plot multiple parameters at once on a single plot with `facet_grid`. But, first we have to remove the Quality Code variables and restructure the data to a long format using `pivot_longer`.
```{r pivot data longer}
lis_long <- lis_orig %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(
    cols = Flow:Chla,
    names_to = "parameter",
    values_to = "value"
  )
```

Now the data is structured so that the parameter names are in the `parameter` variable and the data values are in the `value` variable. 
```{r display data in long format}
head(lis_long)
```

Let's create a plot with facets. Use the `scales` argument set to `"free_y"` in `facet_grid` to adjust y-axis limits to the data of each parameter.
```{r plot with facets, warning = FALSE}
p2 <- lis_long %>%
  ggplot(aes(x = DateTime, y = value)) +
  geom_line() +
  facet_grid(
    rows = vars(parameter),
    scales = "free_y"
  )

p2
```

This is nice, but it would be even nicer to make the plot larger so we can see it better. This is done with the `fig.width` and `fig.height` options in the top of the code chunk.
```{r plot with facets larger, warning = FALSE, fig.width = 8.5, fig.height = 11}
p2
```

If you would like to zoom in on one particular time period, we can use the `scale_x_datetime` function as we did above.
```{r plot with facets change x axis lim, warning = FALSE, fig.height = 8}
p2 + scale_x_datetime(
  limits = c(
    as_datetime("2017-01-01 00:00:00"), 
    as_datetime("2018-01-01 00:00:00")
  )
)
```

