---
title: "Code_cont_WQ_RMB_plots"
author: "Amanda Maguire"
date: "July 27, 2020"
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Code to plot continuous water quality data and check for missing data during the study period.

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

# Import Data

```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/Processed_Data/Continuous"
  )
)

# Import RMB data
rmb_orig <- 
  read_csv(
    file = paste0(sharepoint_path, "/RTM_OUTPUT_RMB_formatted.csv"),
    col_types = paste0("T", str_c(rep("d", 12), collapse = ""), "c")
  )
```

The data looks as we expect it to:
```{r glimpse data}
glimpse(rmb_orig)
```

# Create Plots

## One parameter

Simple line plot of the parameters collected at the RMB station. We'll use the `ggplot` function from the ggplot2 package.

WaterTemp plot
```{r plot WaterTemp param, warning = FALSE}
p <- rmb_orig %>% 
  ggplot(aes(x = DateTime, y = WaterTemp)) +
  geom_line()

print(p)
```

SpCnd plot
```{r plot SpCnd param, warning = FALSE}
ps <- rmb_orig %>%
  ggplot(aes(x = DateTime, y = SpCnd)) +
  geom_line()

print(ps)
```

Turbidity plot
```{r plot Turbidity param, warning = FALSE}
pt <- rmb_orig %>%
  ggplot(aes(x = DateTime, y = Turbidity)) +
  geom_line()

print(pt)
```

Let's take a closer look at the December 2019 sharp increase in Turbidity. The result of a winter storm.
```{r plot Turb param change x axis lim, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2019-11-25 00:00:00"), NA))
```

DO plot
```{r plot DO param, warning = FALSE}
pd <- rmb_orig %>%
  ggplot(aes(x = DateTime, y = DO)) +
  geom_line()

print(pd)
```

pH plot
```{r plot pH param, warning = FALSE}
ph <- rmb_orig %>%
  ggplot(aes(x = DateTime, y = pH)) +
  geom_line()

print(ph)
```

Chla plot
```{r plot CHla param, warning = FALSE}
pc <- rmb_orig %>%
  ggplot(aes(x = DateTime, y = Chla)) +
  geom_line()

print(pc)
```

## Multiple parameters

This is nice, but we can also plot multiple parameters at once on a single plot with `facet_grid`. But, first we have to remove the Quality Code variables and restructure the data to a long format using `pivot_longer`.
```{r pivot data longer}
rmb_long <- rmb_orig %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(
    cols = WaterTemp:Chla,
    names_to = "parameter",
    values_to = "value"
  ) 
```

Now the data is structured so that the parameter names are in the `parameter` variable and the data values are in the `value` variable. 
```{r display data in long format}
head(rmb_long)
```

Let's create a plot with facets. Use the `scales` argument set to `"free_y"` in `facet_grid` to adjust y-axis limits to the data of each parameter.
```{r plot with facets, warning = FALSE}
p2 <- rmb_long %>%
  ggplot(aes(x = DateTime, y = value)) +
  geom_line() +
  facet_grid(
    rows = vars(parameter),
    scales = "free_y"
  )

p2
```

This is nice, but it would be even nicer to make the plot larger so we can see it better. This is done with the `fig.width` and `fig.height` options in the top of the code chunk.
```{r plot with facets larger, fig.width = 8.5, fig.height = 11}
p2
```
