---
title: "Example RTM Plots"
author: "Dave Bosworth"
date: "7/14/2020"
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document provides an example of how to use Rmarkdown to bring together text, code, and its output in way to communicate code and results to others. This document also introduces some useful code to use to create simple plots of continuous water quality data.

Code is placed in code chunks like below:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

# Import Data

To start plotting, first we need to import some data of course:
```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/Processed_Data/Continuous"
  )
)

# Import Lisbon data
lis_orig <- read_csv(paste0(sharepoint_path, "/RTM_OUTPUT_LIS_formatted.csv"))
```

Shoot, we got some errors while importing the data. We will need to define the column types within the `read_csv` function.
```{r import data with defined col types}
# Import Lisbon data
lis_orig <- 
  read_csv(
    file = paste0(sharepoint_path, "/RTM_OUTPUT_LIS_formatted.csv"),
    col_types = paste0("T", str_c(rep("d", 14), collapse = ""), "c")
  )
```

Now, the data looks as we expect it to:
```{r glimpse data}
glimpse(lis_orig)
```

# Create Plots

## One parameter

To start, let's make a simple line plot of one of the parameters collected at the I-80 station. We'll use the `ggplot` function from the ggplot2 package.
```{r plot one param, warning = FALSE}
p <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = WaterTemp)) +
  geom_line()

print(p)
```

Hmm, it looks like there is some empty space on the left side of the plot without data. We can change the x-axis limits for the plot.
```{r plot one param change x axis lim, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2013-01-01 00:00:00"), NA))
```

We can use this same concept to zoom in on particular time periods.
```{r plot one param change x axis lim 2, warning = FALSE}
p + scale_x_datetime(
  limits = c(
    as_datetime("2017-01-01 00:00:00"), 
    as_datetime("2018-01-01 00:00:00")
  )
)
```

## Multiple parameters

This is nice, but we can also plot multiple parameters at once on a single plot with `facet_grid`. But, first we have to remove the Quality Code variables and restructure the data to a long format using `pivot_longer`.
```{r pivot data longer}
lis_long <- lis_orig %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(
    cols = Flow:Chla,
    names_to = "parameter",
    values_to = "value"
  ) %>% 
  filter(!is.na(value))
```

Now the data is structured so that the parameter names are in the `parameter` variable and the data values are in the `value` variable. 
```{r display data in long format}
head(lis_long)
```

Let's create a plot with facets. Use the `scales` argument set to `"free_y"` in `facet_grid` to adjust y-axis limits to the data of each parameter.
```{r plot with facets}
p2 <- lis_long %>%
  ggplot(aes(x = DateTime, y = value)) +
  geom_line() +
  facet_grid(
    rows = vars(parameter),
    scales = "free_y"
  )

p2
```

This is nice, but it would be even nicer to make the plot larger so we can see it better. This is done with the `fig.width` and `fig.height` options in the top of the code chunk.
```{r plot with facets larger, fig.width = 8.5, fig.height = 11}
p2
```

If you would like to zoom in on one particular time period, we can use the `scale_x_datetime` function as we did above.
```{r plot with facets change x axis lim, warning = FALSE, fig.height = 8}
p2 + scale_x_datetime(
  limits = c(
    as_datetime("2017-01-01 00:00:00"), 
    as_datetime("2018-01-01 00:00:00")
  )
)
```

