---
title: "Raw Data Plots - Example"
author: "Dave Bosworth"
date: "8/31/2020"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This paragraph isn't necessary in the actual code:**

Here is some example code to use for updating the Rmd files containing the plots of raw continuous water quality data. This code adds some functions to increase efficiency and consistency and tab navigation for the plots. The example uses the RD22 station.

# Purpose

Create plots of the raw continuous water quality data for the RD22 station. The data in these plots has been cleaned and formatted, but not inspected for suspect values. The purpose of these plots is to allow for us to visually inspect the data for suspect or unreliable values during the study period (2011-2019). The final QA/QC steps for this data can be found in the xxx.Rmd file.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

```{r load functions, message = FALSE, warning = FALSE}
# Source global WQ functions
source("Water_Quality/global_wq_funcs.R")

# Source continuous WQ raw data plot functions
source("Water_Quality/Continuous_WQ/Plots_Raw_Data/rtm_raw_plots_funcs.R")
```

```{r define file paths}
# Define main NDFA file path for WQ subteam (assumes synced with SharePoint)
fp_fastr <- "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/"

# Define relative file path for processed continuous WQ data file for RD22
fp_rel_rtm_data <- paste0(fp_fastr, "Processed_Data/Continuous/RTM_OUTPUT_RD22_formatted.csv")

# Define absolute file path
fp_abs_rtm_data <- get_abs_path(fp_rel_rtm_data)
```

```{r set system tz as PST}
# Set System Timezone as "Etc/GMT+8" (PST) to make it consistent with the data 
Sys.setenv(TZ = "Etc/GMT+8")
```

# Import and Prepare Data

```{r import data}
# Import data
data_orig <- import_rtm_data(fp_abs_rtm_data, 7)
```

```{r prepare data, message = FALSE}
# Clean original data
data_clean <- data_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # pivot data longer for easier plotting
  select(-c(StationCode, ends_with("_Qual"))) %>% 
  pivot_longer(
    cols = -DateTime,
    names_to = "parameter",
    values_to = "value"
  ) %>% 
  # add yr and y-axis label variables
  mutate(yr = year(DateTime)) %>% 
  add_yaxis_var(parameter)
```

# Raw Data Plots {.tabset .tabset-pills}

## Chla {.tabset .tabset-pills}

```{r filter chla data, include = FALSE}
# Use this code to filter to your parameter of interest and remove any excess data at the beginning of the period of record
data_clean_chla <- data_clean %>% 
  filter(
    parameter == "Chla",
    yr >= 2013
  )
```

### All Years

```{r plot chla all years, echo = FALSE, warning = FALSE}
# Use the create_ts_plot function to create your plot
data_clean_chla %>% create_ts_plot(.$yaxis_lab[1])
```

### 2013

```{r plot chla 2013, echo = FALSE}
# Filter by yr and remove NA values before creating your plot
data_clean_chla %>% 
  filter(
    yr == 2013,
    !is.na(value)
  ) %>% 
  create_ts_plot(.$yaxis_lab[1])
```

### 2014

```{r plot chla 2014, echo = FALSE}
# Follow the same pattern for creating the plot for 2014
data_clean_chla %>% 
  filter(
    yr == 2014,
    !is.na(value)
  ) %>% 
  create_ts_plot(.$yaxis_lab[1])
```

## DO {.tabset .tabset-pills}

```{r filter do data, include = FALSE}
# Follow the same pattern for your next parameter
data_clean_do <- data_clean %>% 
  filter(
    parameter == "DO",
    yr >= 2014
  )
```

### All Years

```{r plot do all years, echo = FALSE, warning = FALSE}
# Use the create_ts_plot function to create your plot
data_clean_do %>% create_ts_plot(.$yaxis_lab[1])
```

### 2014

```{r plot do 2014, echo = FALSE}
# Filter by yr and remove NA values before creating your plot
data_clean_do %>% 
  filter(
    yr == 2014,
    !is.na(value)
  ) %>% 
  create_ts_plot(.$yaxis_lab[1])
```

# Extra Instructions

**This section isn't necessary in the actual code**

Follow the pattern above for the Chlorophyll and DO data for the remaining parameters and years.

