---
title: "zoop_LIS_flow_models"
author: "Laura"
date: "2026-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(mgcv)
library(ggeffects)
library(emmeans)
library(multcomp)
library(gratia)
library(patchwork)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```


# Import Data

```{r import data}

# Import zooplankton data
df_zoop <- read_csv(here("Zoop_code/Jesse-zoop-analyses/zoopNDFA7.csv"))

# Import weekly average flow data
df_flow <- readRDS(here("manuscript_synthesis/data/processed/flow_week_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare zoop data}
#add week of year to zoop data
df_zoop_weekly<-df_zoop%>%mutate(Week=week(Date))

# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

#create list of flow stations that we will use in analysis
flow_stations<-c("LIS", "RD22")

# We will use LIS flow data as a proxy for STTD, LIB, and RVB
df_flow_c <- df_flow %>% filter(StationCode%in%flow_stations)
#define stations for which we want to use LIS flow data
df_zoop_stations<-c("STTD", "LIB", "RVB") 


# Prepare zooplankton and flow data for analysis
df_zoop2 <- df_zoop_weekly %>%
  select(StationCode, Year, Week, cpue) %>%
  drop_na(cpue) %>%
  filter(
    StationCode %in% sta_order,
    Year %in% 2014:2019
  ) %>%
  # Assign the appropriate flow station BEFORE joining
  mutate(
    FlowStation = case_when(
      StationCode == "RD22" ~ "RD22",
      StationCode %in% c("STTD", "LIB", "RVB") ~ "LIS"
    )
  ) %>%
  # Now join on Year, Week, AND the specific flow station
  left_join(
    df_flow_c,
    by = join_by(Year, Week, FlowStation == StationCode)
  ) %>%
  drop_na(Flow) %>%
  mutate(
    zoop_log = log(cpue),
    Year_fct = factor(Year),
    Station_fact = factor(StationCode, levels = sta_order)
  ) %>%
  arrange(Station_fact, Year, Week)


df_zoop2 <- df_zoop_weekly %>% 
  select(StationCode, Year, Week, cpue) %>% 
  drop_na(cpue) %>% 
  # Filter to only include representative stations for 4 habitat types (RD22,
    # STTD, LIB, RVB) and only include years 2014-2019
  filter(
    StationCode %in% sta_order,
    Year %in% 2014:2019
  ) %>% 
  # Join flow data to zoop data
  left_join(df_flow_c, by = join_by(Year, Week))%>%
  #remove rows with RD22 as flow and STTD, LIB, or RVB as zoop
  filter(!(StationCode.x %in% df_zoop_stations & StationCode.y == "RD22")) %>%
  #remove rows with LIS as flow and RD22 as zoop
  filter(!(StationCode.x == "RD22" & StationCode.y == "LIS")) %>% 
  # Remove all NA flow values
  drop_na(Flow) %>% 
  mutate(
    # Scale and log transform chlorophyll values
    zoop_log = log(cpue),
    # Add a column for Year as a factor for the models
    Year_fct = factor(Year),
    #add a column for station as a factor
    Station_fact=factor(StationCode.x, levels = sta_order)
  ) %>% 
  arrange(Station_fact, Year, Week) 

#verify that the join worked as expected
# Full diagnostic summary
df_zoop2 %>%
  group_by(StationCode, FlowStation, Year) %>%
  summarise(
    n_weeks = n_distinct(Week),
    n_rows = n(),
    has_duplicates = n() > n_distinct(Week),
    flow_range = paste(round(min(Flow)), "-", round(max(Flow))),
    .groups = "drop"
  ) %>%
  print(n = Inf)

```

# Initial model with two-way interactions and a smooth term for week of year

```{r create gams, warning = FALSE}
# Run GAM for all stations

all_stations_gam<-gam(zoop_log ~ (Year_fct + Flow + Station_fact)^2 + s(Week, bs = 'cc', k = 5), data = df_zoop2, method = "REML", knots =list(week = c(0, 52)))
summary(all_stations_gam)
```

# Model Results

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova <- anova(all_stations_gam)
ndf_gam_anova

```
#Model diagnostics
```{r gam diagnostics, warning = FALSE}
cat("\n=== Model Diagnostics ===\n")
cat("AIC:", AIC(all_stations_gam), "\n")
cat("BIC:", BIC(all_stations_gam), "\n")
cat("Deviance explained:", 
    round(summary(all_stations_gam)$dev.expl * 100, 2), "%\n")

```
# Linear model with two-way interactions

```{r create lms, warning = FALSE}
# Run LM for all stations

all_stations_lm<-lm(zoop_log ~ (Year_fct + Flow + Station_fact)^2, data = df_zoop2)
summary(all_stations_lm)
```
## ANOVA Tables

```{r lm anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_lm_anova <- anova(all_stations_lm)
ndf_lm_anova

```
#Model diagnostics
```{r lm diagnostics, warning = FALSE}
cat("\n=== Model Diagnostics ===\n")
cat("AIC:", AIC(all_stations_lm), "\n")
cat("BIC:", BIC(all_stations_lm), "\n")

```

## Effect of Flow by Year: below code is in development

