---
title: "zoop_LIS_flow_models"
author: "Laura"
date: "2026-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(mgcv)
library(ggeffects)
library(emmeans)
library(multcomp)
library(gratia)
library(patchwork)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```


```{r functions and global code}
# Calculate effects of flow on zooplankton for each year holding the
  # non-focal variables constant - marginal effects/adjusted predictions
calc_flow_yr_marg_eff <- function(df_data, m_gam) {
  # Calculate min and max flows for each year to narrow down x-axis in the
    # effects plots
  df_flow_yr_summ <- df_data %>% 
    summarize(
      Flow_min = min(Flow),
      Flow_max = max(Flow),
      .by = c(Year_fct)
    ) %>% 
    mutate(
      Flow_buffer = (Flow_max - Flow_min) * 0.05,
      Flow_min = Flow_min - Flow_buffer,
      Flow_max = Flow_max + Flow_buffer
    ) %>% 
    select(-Flow_buffer)
  
  # Calculate marginal effects
  as.data.frame(
    predict_response(m_gam, terms = c("Flow", "Year_fct"), margin = "marginalmeans"),
    terms_to_colnames = TRUE
  ) %>% 
  as_tibble() %>% 
  # Narrow down range of flow values for each station
  left_join(df_flow_yr_summ, by = join_by(Year_fct)) %>% 
  filter(Flow >= Flow_min & Flow <= Flow_max)
}

# Create effects plot of flow on zooplankton by year
plot_eff_flow_by_yr <- function(df_eff, df_data) {
  df_eff %>% 
    ggplot(aes(x = Flow, y = predicted)) +
    geom_point(data = df_data, aes(y = zoop_log), alpha = 0.6) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
    facet_wrap(vars(Year_fct), scales = "free_x", nrow = 1) +
    theme_bw() +
    labs(
      title = df_data$StationCode[1],
      x = "Flow (cfs)",
      y = "log(Zooplankton abundance)"
    ) +
    scale_x_continuous(labels = label_comma(), breaks = breaks_extended()) +
    theme(
      axis.text = element_text(size = 8),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
    )
}

# Create boxplots of zooplankton by year showing Tukey post-hoc results
boxplot_cont_yr <- function(em_obj, df_data, xlab = TRUE) {
  # Create table of contrasts and convert it to a tibble for plot
  df_yr_cont <- em_obj %>% 
    cld(sort = FALSE, Letters = letters) %>% 
    as_tibble() %>% 
    mutate(
      group = str_remove_all(.group, fixed(" ")),
    ) %>% 
    # Add min and max values of observed data to the Tukey post-hoc results and
      # calculate vertical positioning of letters
    left_join(
      df_data %>% 
        summarize(
          max_val = max(zoop_log),
          min_val = min(zoop_log),
          .by = Year_fct
      ), 
      by = join_by(Year_fct)
    ) %>% 
    mutate(
      max_val = if_else(upper.CL > max_val, upper.CL, max_val),
      y_pos = max_val + (max_val - min_val) / 10,
      # Make all post-hoc contrast letters at same height equal to max of all
      y_pos = max(y_pos)
    ) %>% 
    select(
      Year_fct,
      emmean,
      lower.CL,
      upper.CL,
      group,
      y_pos
    )
  
  # Create boxplot showing Tukey post-hoc results
  df_yr_cont %>% 
    ggplot(
      aes(
        x = Year_fct,
        y = emmean,
        ymin = lower.CL,
        ymax = upper.CL
      )
    ) +
    geom_boxplot(
      data = df_data,
      aes(x = Year_fct, y = zoop_log),
      inherit.aes = FALSE
    ) +
    geom_crossbar(
      color = "grey82", 
      fill = "grey", 
      alpha = 0.7, 
      linewidth = 0.1
    ) +
    geom_point(color = "red") +
    geom_text(aes(y = y_pos, label = group), size = 3) +
    labs(
      title = df_data$StationCode[1],
      x = "Year",
      y = "log(Zooplankton abundance)"
    ) +
    theme_bw() +
    theme(axis.text = element_text(size = 8))
}

# Create smooth of Week plot
plot_smooth_week <- function(m_gam, station) {
  ls_plt <- draw(m_gam, residuals = TRUE, rug = FALSE, wrap = FALSE)
  
  # Extract plot from list and format
  chuck(ls_plt, 1) + 
    theme_bw() + 
    labs(title = station, caption = NULL)
}
```

# Import Data

```{r import data}

# Import zooplankton data
df_zoop <- read_csv(here("Zoop_code/Jesse-zoop-analyses/zoopNDFA7.csv"))

# Import weekly average flow data
df_flow <- readRDS(here("manuscript_synthesis/data/processed/flow_week_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare zoop data}

#add week of year to zoop data
df_zoop_weekly<-df_zoop%>%mutate(Week=week(Date))

# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

#create list of flow stations that we will use in analysis
flow_stations<-c("LIS", "RD22")

# We will use LIS flow data as a proxy for STTD, LIB, and RVB
df_flow_c <- df_flow %>% filter(StationCode%in%flow_stations)
df_zoop_stations<-c("STTD", "LIB", "RVB") #define stations for which we want to use LIS flow data


# Prepare zooplankton and flow data for analysis
df_zoop2 <- df_zoop_weekly %>%
  select(StationCode, Year, Week, cpue) %>%
  drop_na(cpue) %>%
  filter(
    StationCode %in% sta_order,
    Year %in% 2014:2019
  ) %>%
  # Map each zoop station to its corresponding flow station BEFORE joining
  mutate(
    FlowStation = case_when(
      StationCode == "RD22"                        ~ "RD22",
      StationCode %in% df_zoop_stations            ~ "LIS"  # STTD, LIB, RVB → LIS
    )
  ) %>%
  # Now join on Year, Week, AND the mapped flow station — no many-to-many!
  left_join(
    df_flow_c,
    by = join_by(Year, Week, FlowStation == StationCode)
  ) %>%
  drop_na(Flow) %>%
  mutate(
    zoop_log  = log(cpue),
    Year_fct  = factor(Year))%>%
  arrange(StationCode, Year, Week) %>% 
  group_by(StationCode, Year) %>%
  mutate(flow_lag1 = lag(Flow, n=1),
    flow_lag2 = lag(Flow, n=2),
    flow_lag3 = lag(Flow, n=3)) %>%
  split(~ StationCode)

# Combine all data sets into a nested data frame to be used for analysis
ndfa_zoop <- 
  tibble(
    StationCode = sta_order,
    df_data = list(df_zoop2$RD22, df_zoop2$STTD, df_zoop2$LIB, df_zoop2$RVB)
  ) %>% 
  # Apply factor order to StationCode
  mutate(StationCode = factor(StationCode, levels = sta_order)) %>% 
  arrange(StationCode)
```

## Create GAMs with interaction terms

```{r create gams, warning = FALSE}
# Define model formulae
f_base <- " zoop_log ~ Year_fct * Flow + Year_fct + s(Week, bs = 'cc', k = 5)"
f_lag1 <- " zoop_log ~ Year_fct * flow_lag1 + Year_fct + s(Week, bs = 'cc', k = 5)"
f_lag2 <- " zoop_log ~ Year_fct * flow_lag2 + Year_fct + s(Week, bs = 'cc', k = 5)"
f_lag3 <- " zoop_log ~ Year_fct * flow_lag3 + Year_fct + s(Week, bs = 'cc', k = 5)"

# Create gam models for each station
ndf_gam <- ndfa_zoop %>% mutate(
    str_formula = case_when(
      StationCode == "RD22" ~ paste(f_base),
      StationCode == "STTD" ~ paste(f_lag1),
      StationCode == "LIB" ~ paste(f_lag2),
      StationCode == "RVB" ~ paste(f_lag3)
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "ML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

## Model summaries for GAMs with interaction terms

```{r model summaries}
library(broom)

#append model summaries
ndf_gam <- ndf_gam %>%
  mutate(
    # Tidy parametric coefficients
    tidy_params = map(model, ~ tidy(.x, parametric = TRUE)),

    # Tidy smooth terms
    tidy_smooths = map(model, ~ tidy(.x, parametric = FALSE)),

    # Model-level fit statistics (AIC, R-sq, deviance explained, etc.)
    glance_stats = map(model, glance)
  )

#print summaries by station
walk2(ndf_gam$StationCode, ndf_gam$model, ~ {
  cat("\n", strrep("=", 50), "\n")
  cat("Station:", .x, "\n")
  cat(strrep("=", 50), "\n")
  print(summary(.y))
})

#unnest to create a single comparison table
fit_table <- ndf_gam %>%
  select(StationCode, glance_stats) %>%
  unnest(glance_stats) %>%
  select(StationCode, AIC, BIC, adj.r.squared, deviance, df.residual)

print(fit_table)

```


## Diagnostic plots for GAMs with interaction terms

```{r diagnostic plots}

library(gratia)

diag_plots <- ndf_gam %>%
  mutate(
    diag_plot = map2(model, StationCode, ~ 
      appraise(.x) + 
        patchwork::plot_annotation(title = paste("Station:", .y))
    )
  )

# Save all to PDF
pdf("gam_diagnostics_gratia.pdf", width = 10, height = 8)
walk(diag_plots$diag_plot, print)
dev.off()


# Using gratia::draw() for partial effects of smooth terms
smooth_plots <- ndf_gam %>%
  mutate(
    smooth_plot = map2(model, StationCode, ~
      draw(.x) +
        patchwork::plot_annotation(title = paste("Smooths — Station:", .y))
    )
  )

pdf("gam_smooths.pdf", width = 10, height = 6)
walk(smooth_plots$smooth_plot, print)
dev.off()
```

## GAMs without interaction terms

```{r create gams without interaction terms, warning = FALSE}
# Define model formulae
f2_base <- " zoop_log ~ Flow + Year_fct + s(Week, bs = 'cc', k = 5)"
f2_lag1 <- " zoop_log ~ flow_lag1 + Year_fct + s(Week, bs = 'cc', k = 5)"
f2_lag2 <- " zoop_log ~ flow_lag2 + Year_fct + s(Week, bs = 'cc', k = 5)"
f2_lag3 <- " zoop_log ~ flow_lag3 + Year_fct + s(Week, bs = 'cc', k = 5)"

# Create gam models for each station
ndf_gam2 <- ndfa_zoop %>% mutate(
    str_formula = case_when(
      StationCode == "RD22" ~ paste(f2_base),
      StationCode == "STTD" ~ paste(f2_lag1),
      StationCode == "LIB" ~ paste(f2_lag2),
      StationCode == "RVB" ~ paste(f2_lag3)
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "ML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

## Model summaries for GAMs without interaction terms

```{r model summaries}
library(broom)

#append model summaries
ndf_gam2 <- ndf_gam2 %>%
  mutate(
    # Tidy parametric coefficients
    tidy_params = map(model, ~ tidy(.x, parametric = TRUE)),

    # Tidy smooth terms
    tidy_smooths = map(model, ~ tidy(.x, parametric = FALSE)),

    # Model-level fit statistics (AIC, R-sq, deviance explained, etc.)
    glance_stats = map(model, glance),
    k.check= map(model, k.check))

#print summaries by station
walk2(ndf_gam2$StationCode, ndf_gam2$model, ~ {
  cat("\n", strrep("=", 50), "\n")
  cat("Station:", .x, "\n")
  cat(strrep("=", 50), "\n")
  print(summary(.y))
})

#unnest to create a single comparison table
fit_table2 <- ndf_gam2 %>%
  select(StationCode, glance_stats) %>%
  unnest(glance_stats) %>%
  select(StationCode, AIC, BIC, adj.r.squared, deviance, df.residual)

print(fit_table2)
print(ndf_gam2$k.check)

```

## Diagnostic plots for GAMs without interaction terms

```{r diagnostic plots}

library(gratia)

diag_plots2 <- ndf_gam2 %>%
  mutate(
    diag_plot = map2(model, StationCode, ~ 
      appraise(.x) + 
        patchwork::plot_annotation(title = paste("Station:", .y))
    )
  )

# Save all to PDF
pdf("gam_diagnostics_gratia2.pdf", width = 10, height = 8)
walk(diag_plots2$diag_plot, print)
dev.off()


# Using gratia::draw() for partial effects of smooth terms
smooth_plots2 <- ndf_gam2 %>%
  mutate(
    smooth_plot = map2(model, StationCode, ~
      draw(.x) +
        patchwork::plot_annotation(title = paste("Smooths — Station:", .y))
    )
  )

pdf("gam_smooths2.pdf", width = 10, height = 6)
walk(smooth_plots2$smooth_plot, print)
dev.off()


```

## lower AIC for GAMs without interaction terms, so moving forward with those models

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova <- ndf_gam2 %>% 
  mutate(
    an_gam = map(model, anova),
    # Parametric terms
    df_parametric = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "pTerms.table"), rownames = "Term")
    ),
    # Smooth terms
    df_smooth = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "s.table"), rownames = "Term") %>% 
        rename(df = edf) %>% 
        select(-Ref.df)
    ),
    df_anova_tbl = map2(df_parametric, df_smooth, bind_rows)
  )

# Combine ANOVA tables and format for export
df_gam_anova_tbl_c <- ndf_gam_anova %>% 
  select(StationCode, df_anova_tbl) %>% 
  unnest(df_anova_tbl) %>% 
  transmute(
    Station = StationCode,
    Term,
    across(
      c("df", "F"),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    )) %>%
  bind_rows(
    ndf_gam_anova %>%
      transmute(
        Station = as.character(StationCode),
        df      = NA_character_,
        F       = NA_character_,
        p_value = NA_character_
      )
  ) 

df_gam_anova_tbl_c

df_gam_anova_tbl_c %>% 
  mutate(across(c("df", "F", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv("zoop_LIS_flow_GAM_anova.csv")


```

## Flow Effect plots
```{r flow effect plots}

flow_var_lookup <- c(
  RD22 = "Flow",
  STTD = "flow_lag1",
  LIB  = "flow_lag2",
  RVB  = "flow_lag3"
)

ndf_gam2 <- ndf_gam2 %>%
  mutate(
    flow_var = flow_var_lookup[StationCode],

    pred_data = pmap(
      list(model, df_data, flow_var),
      \(mod, dat, fv) {
        flow_seq <- seq(min(dat[[fv]], na.rm = TRUE),
                        max(dat[[fv]], na.rm = TRUE),
                        length.out = 200)

        ref_year <- levels(dat$Year_fct)[1]

        # Use setNames() to name the flow column correctly from the start
        nd <- tibble(
          flow_val = flow_seq,
          Year_fct = factor(ref_year, levels = levels(dat$Year_fct)),
          Week     = median(dat$Week, na.rm = TRUE)
        ) %>%
          setNames(c(fv, "Year_fct", "Week"))  # rename flow_val -> fv here

        preds <- predict(mod, newdata = nd, type = "link", se.fit = TRUE)

        # Pull flow column back out by name for plotting
        nd %>%
          mutate(
            flow_val = .[[fv]],  # extract by name for ggplot
            fit      = preds$fit,
            lower    = fit - 1.96 * preds$se.fit,
            upper    = fit + 1.96 * preds$se.fit
          )
      }
    )
  )

station_plots <- ndf_gam2 %>%
  mutate(
    plot = pmap(
      list(pred_data, StationCode, flow_var),
      \(pd, stn, fv) {
        ggplot(pd, aes(x = flow_val, y = fit)) +
          geom_ribbon(aes(ymin = lower, ymax = upper),
                      fill = "steelblue", alpha = 0.25) +
          geom_line(color = "steelblue", linewidth = 1) +
          labs(
            title = stn,
            x     = paste0("Flow (cfs)"),
            y     = "Zooplankton density"
          ) +
          theme_bw(base_size = 11) +
          theme(plot.title = element_text(face = "plain"))
      }
    )
  )

# Combine into 2x2 grid
library(cowplot)

# Step 1 — Strip only axis titles from individual plots, leave tick labels alone
cleaned_plots <- map(station_plots$plot, ~ .x &
  theme(axis.title = element_blank())
)

# Step 2 — Combine with patchwork, no axis collecting
inner_plot <- wrap_plots(cleaned_plots, ncol = 2) &
  theme(plot.margin = margin(4, 4, 4, 4))

# Step 3 — Add true margin y-axis label
combined_with_y <- plot_grid(
  ggdraw() +
    draw_label("log10(Zooplankton density)", angle = 90,
               fontface = "plain", size = 15),
  inner_plot,
  ncol       = 2,
  rel_widths = c(0.05, 1)
)

# Step 4 — Add true margin x-axis label
combined_plot <- plot_grid(
  combined_with_y,
  ggdraw(clip="off") +
    draw_label("Flow (cfs)", fontface = "plain", size = 15)+
  theme(plot.margin = margin(t = -3, 0, 0, 0)), 
  ncol        = 1,
  rel_heights = c(1, 0.04)
)

combined_plot
```

## s(Week) Plots

```{r gam sWeek plots, warning = FALSE}
# Create smooth of Week plots for each station
ndf_gam_s_week_plt <- ndf_gam2 %>% 
  mutate(plt_s_week = map2(model, StationCode, plot_smooth_week))

# Combine smooth of Week plots together using patchwork
plt_gam_s_week <- 
  wrap_plots(ndf_gam_s_week_plt$plt_s_week, ncol = 2) + 
  plot_layout(axis_titles = "collect") +
  plot_annotation(caption = "Basis: Cyclic CRS")

plt_gam_s_week
```
## Create boxplots of density by year

```{r boxplots}
ndf_gam2 <- ndf_gam2 %>%
  mutate(
    boxplot = map2(
      df_data, StationCode,
      \(dat, stn) {
        ggplot(dat, aes(x = Year_fct, y = zoop_log)) +
          geom_boxplot(fill = "purple", alpha = 0.6, outlier.shape = 19) +
          geom_jitter(width = 0.15, alpha = 0.4, size = 1.5, color = "gray30") +
          labs(
            title = stn,
            x     = "Year",
            y     = "log(Zooplankton Density)"
          ) +
          theme_bw(base_size = 11) +
          theme(
            plot.title     = element_text(face = "plain"),
            axis.text.x    = element_text(angle = 45, hjust = 1)
          )
      }
    )
  )

walk2(
  ndf_gam2$StationCode, ndf_gam2$boxplot,
  \(stn, plt) {
    ggsave(
      filename = paste0("boxplot_", stn, ".png"),
      plot     = plt,
      width    = 8,
      height   = 5,
      dpi      = 300
    )
  }
)

# Combine into 2x2 grid

# Step 1 — Build individual boxplots with no axis titles
cleaned_boxplots <- map(ndf_gam2$boxplot, ~ .x +
  theme(axis.title = element_blank())
)

# Step 2 — Combine with patchwork, no axis collecting
inner_boxplot <- wrap_plots(cleaned_boxplots, ncol = 2) &
  theme(plot.margin = margin(4, 4, 4, 4))

# Step 3 — Add true margin y-axis label
boxplot_with_y <- plot_grid(
  ggdraw() +
    draw_label("log10(Zooplankton density)", angle = 90,
               fontface = "plain", size = 15),
  inner_boxplot,
  ncol       = 2,
  rel_widths = c(0.05, 1)
)

# Step 4 — Add true margin x-axis label
combined_boxplot <- plot_grid(
  boxplot_with_y,
  ggdraw() +
    draw_label("Year", fontface = "plain", size = 15, vjust = 0),
  ncol        = 1,
  rel_heights = c(1, 0.04)
)

combined_boxplot


```

# Export Results

```{r export results}

# Export effects plots of smooth of Week
ggsave(
  file.path(fp_figures, "zoop_LIS_flow_smooth_week.jpg"),
  plot = plt_gam_s_week,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export combined ANOVA tables for GAMs
df_gam_anova_tbl_c %>% 
  mutate(across(c("df", "F", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "zoop_LIS_flow_GAM_anova.csv"))

# Export slopes and p-values of the modeled effects of flow by year for each station
df_gam_eff_flow_stat %>% 
  mutate(across(c("Flow_slope", "SE", "df", "t_ratio", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "zoop_LIS_flow_GAM_eff_flow_stat.csv"))

# Export pairwise Tukey post-hoc year contrasts
df_gam_cont_yr_stat %>% 
  mutate(across(c("Estimate", "SE", "df", "t_ratio", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "zoop_LIS_flow_GAM_year_contrasts_stat.csv"))
```