---
title: "zoop_LIS_flow_models"
author: "Laura"
date: "2026-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(mgcv)
library(ggeffects)
library(emmeans)
library(multcomp)
library(gratia)
library(patchwork)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```


```{r functions and global code}
# Calculate effects of flow on zooplankton for each year holding the
  # non-focal variables constant - marginal effects/adjusted predictions
calc_flow_yr_marg_eff <- function(df_data, m_gam) {
  # Calculate min and max flows for each year to narrow down x-axis in the
    # effects plots
  df_flow_yr_summ <- df_data %>% 
    summarize(
      Flow_min = min(Flow),
      Flow_max = max(Flow),
      .by = c(Year_fct)
    ) %>% 
    mutate(
      Flow_buffer = (Flow_max - Flow_min) * 0.05,
      Flow_min = Flow_min - Flow_buffer,
      Flow_max = Flow_max + Flow_buffer
    ) %>% 
    select(-Flow_buffer)
  
  # Calculate marginal effects
  as.data.frame(
    predict_response(m_gam, terms = c("Flow", "Year_fct"), margin = "marginalmeans"),
    terms_to_colnames = TRUE
  ) %>% 
  as_tibble() %>% 
  # Narrow down range of flow values for each station
  left_join(df_flow_yr_summ, by = join_by(Year_fct)) %>% 
  filter(Flow >= Flow_min & Flow <= Flow_max)
}

# Create effects plot of flow on zooplankton by year
plot_eff_flow_by_yr <- function(df_eff, df_data) {
  df_eff %>% 
    ggplot(aes(x = Flow, y = predicted)) +
    geom_point(data = df_data, aes(y = zoop_log), alpha = 0.6) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
    facet_wrap(vars(Year_fct), scales = "free_x", nrow = 1) +
    theme_bw() +
    labs(
      title = df_data$StationCode[1],
      x = "Flow (cfs)",
      y = "log(Zooplankton abundance)"
    ) +
    scale_x_continuous(labels = label_comma(), breaks = breaks_extended()) +
    theme(
      axis.text = element_text(size = 8),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
    )
}

# Create boxplots of zooplankton by year showing Tukey post-hoc results
boxplot_cont_yr <- function(em_obj, df_data, xlab = TRUE) {
  # Create table of contrasts and convert it to a tibble for plot
  df_yr_cont <- em_obj %>% 
    cld(sort = FALSE, Letters = letters) %>% 
    as_tibble() %>% 
    mutate(
      group = str_remove_all(.group, fixed(" ")),
    ) %>% 
    # Add min and max values of observed data to the Tukey post-hoc results and
      # calculate vertical positioning of letters
    left_join(
      df_data %>% 
        summarize(
          max_val = max(zoop_log),
          min_val = min(zoop_log),
          .by = Year_fct
      ), 
      by = join_by(Year_fct)
    ) %>% 
    mutate(
      max_val = if_else(upper.CL > max_val, upper.CL, max_val),
      y_pos = max_val + (max_val - min_val) / 10,
      # Make all post-hoc contrast letters at same height equal to max of all
      y_pos = max(y_pos)
    ) %>% 
    select(
      Year_fct,
      emmean,
      lower.CL,
      upper.CL,
      group,
      y_pos
    )
  
  # Create boxplot showing Tukey post-hoc results
  df_yr_cont %>% 
    ggplot(
      aes(
        x = Year_fct,
        y = emmean,
        ymin = lower.CL,
        ymax = upper.CL
      )
    ) +
    geom_boxplot(
      data = df_data,
      aes(x = Year_fct, y = zoop_log),
      inherit.aes = FALSE
    ) +
    geom_crossbar(
      color = "grey82", 
      fill = "grey", 
      alpha = 0.7, 
      linewidth = 0.1
    ) +
    geom_point(color = "red") +
    geom_text(aes(y = y_pos, label = group), size = 3) +
    labs(
      title = df_data$StationCode[1],
      x = "Year",
      y = "log(Zooplankton abundance)"
    ) +
    theme_bw() +
    theme(axis.text = element_text(size = 8))
}

# Create smooth of Week plot
plot_smooth_week <- function(m_gam, station) {
  ls_plt <- draw(m_gam, residuals = TRUE, rug = FALSE, wrap = FALSE)
  
  # Extract plot from list and format
  chuck(ls_plt, 1) + 
    theme_bw() + 
    labs(title = station, caption = NULL)
}
```

# Import Data

```{r import data}

# Import zooplankton data
df_zoop <- read_csv(here("Zoop_code/Jesse-zoop-analyses/zoopNDFA7.csv"))

# Import weekly average flow data
df_flow <- readRDS(here("manuscript_synthesis/data/processed/flow_week_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare zoop data}

#add week of year to zoop data
df_zoop_weekly<-df_zoop%>%mutate(Week=week(Date))

# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

#create list of flow stations that we will use in analysis
flow_stations<-c("LIS", "RD22")

# We will use LIS flow data as a proxy for STTD, LIB, and RVB
df_flow_c <- df_flow %>% filter(StationCode%in%flow_stations)
df_zoop_stations<-c("STTD", "LIB", "RVB") #define stations for which we want to use LIS flow data


# Prepare zooplankton and flow data for analysis
df_zoop2 <- df_zoop_weekly %>%
  select(StationCode, Year, Week, cpue) %>%
  drop_na(cpue) %>%
  filter(
    StationCode %in% sta_order,
    Year %in% 2014:2019
  ) %>%
  # Map each zoop station to its corresponding flow station BEFORE joining
  mutate(
    FlowStation = case_when(
      StationCode == "RD22"                        ~ "RD22",
      StationCode %in% df_zoop_stations            ~ "LIS"  # STTD, LIB, RVB → LIS
    )
  ) %>%
  # Now join on Year, Week, AND the mapped flow station — no many-to-many!
  left_join(
    df_flow_c,
    by = join_by(Year, Week, FlowStation == StationCode)
  ) %>%
  drop_na(Flow) %>%
  mutate(
    zoop_log  = log(cpue),
    Year_fct  = factor(Year), 
    flow_lag1 = lag(Flow),
    flow_lag2 = lag(flow_lag1),
    flow_lag3 = lag(flow_lag2)
  ) %>%
  arrange(StationCode, Year, Week) %>%
  split(~ StationCode)

# Combine all data sets into a nested data frame to be used for analysis
ndfa_zoop <- 
  tibble(
    StationCode = sta_order,
    df_data = list(df_zoop2$RD22, df_zoop2$STTD, df_zoop2$LIB, df_zoop2$RVB)
  ) %>% 
  # Apply factor order to StationCode
  mutate(StationCode = factor(StationCode, levels = sta_order)) %>% 
  arrange(StationCode)
```

# Evaluate collinearity between lagged flow terms
```{r, evaluate collinearity between lagged flow terms}
cor(df_zoop2$STTD$Flow, df_zoop2$STTD$flow_lag1)

cor(df_zoop2$STTD$Flow, df_zoop2$STTD$flow_lag2)

```
# Create GAMs- no lag on flow variable

```{r create gams, warning = FALSE}
# Define base model formula
f_base <- " zoop_log ~ Year_fct * Flow + Year_fct + s(Week, bs = 'cc', k = 5)"



# Create gam models for each station
ndf_gam <- ndfa_zoop %>% mutate(
    str_formula = case_when(
      StationCode == "LIB" ~ paste(f_base),
      StationCode == "RVB" ~ paste(f_base),
      TRUE ~ f_base
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "ML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

# Model Results- no lag on flow variable

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova <- ndf_gam %>% 
  mutate(
    an_gam = map(model, anova),
    # Extract AIC from each model
    aic = map_dbl(model, AIC),
    # Parametric terms
    df_parametric = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "pTerms.table"), rownames = "Term")
    ),
    # Smooth terms
    df_smooth = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "s.table"), rownames = "Term") %>% 
        rename(df = edf) %>% 
        select(-Ref.df)
    ),
    df_anova_tbl = map2(df_parametric, df_smooth, bind_rows)
  )

# Combine ANOVA tables and format for export
df_gam_anova_tbl_c <- ndf_gam_anova %>% 
  select(StationCode, df_anova_tbl, aic) %>% 
  unnest(df_anova_tbl) %>% 
  transmute(
    Station = StationCode,
    Term,
    across(
      c("df", "F"),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    ),
    aic) %>%
  # Add AIC as a summary row per station
  bind_rows(
    ndf_gam_anova %>%
      transmute(
        Station = as.character(StationCode),
        Term    = "AIC",
        df      = NA_character_,
        F       = NA_character_,
        p_value = NA_character_,
        aic
      )
  ) %>%
  arrange(Station, Term == "AIC")  # Keep AIC rows at the bottom per station

df_gam_anova_tbl_c
```
# Create GAMs- one week lag on flow variable

```{r create gams, warning = FALSE}
# Define base model formula
f_lag1 <- " zoop_log ~ Year_fct * flow_lag1 + Year_fct + s(Week, bs = 'cc', k = 5)"



# Create gam models for each station
ndf_gam_lag1 <- ndfa_zoop %>% mutate(
    str_formula = case_when(
      StationCode == "LIB" ~ paste(f_lag1),
      StationCode == "RVB" ~ paste(f_lag1),
      TRUE ~ f_lag1
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "ML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

# Model Results- one week lag on flow variable

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova_lag1 <- ndf_gam_lag1 %>% 
  mutate(
    an_gam = map(model, anova),
    # Extract AIC from each model
    aic = map_dbl(model, AIC),
    # Parametric terms
    df_parametric = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "pTerms.table"), rownames = "Term")
    ),
    # Smooth terms
    df_smooth = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "s.table"), rownames = "Term") %>% 
        rename(df = edf) %>% 
        select(-Ref.df)
    ),
    df_anova_tbl = map2(df_parametric, df_smooth, bind_rows)
  )

# Combine ANOVA tables and format for export
df_gam_anova_lag1_tbl_c <- ndf_gam_anova_lag1 %>% 
  select(StationCode, df_anova_tbl, aic) %>% 
  unnest(df_anova_tbl) %>% 
  transmute(
    Station = StationCode,
    Term,
    across(
      c("df", "F"),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    ),
    aic
  )%>%
   # Add AIC as a summary row per station
  bind_rows(
    ndf_gam_anova_lag1 %>%
      transmute(
        Station = as.character(StationCode),
        Term    = "AIC",
        df      = NA_character_,
        F       = NA_character_,
        p_value = NA_character_,
        aic
      )
  ) %>%
  arrange(Station, Term == "AIC")  # Keep AIC rows at the bottom per station


df_gam_anova_lag1_tbl_c
```
# Create GAMs- two week lag on flow variable

```{r create gams, warning = FALSE}
# Define base model formula
f_lag2 <- " zoop_log ~ Year_fct * flow_lag2 + Year_fct + s(Week, bs = 'cc', k = 5)"



# Create gam models for each station
ndf_gam_lag2 <- ndfa_zoop %>% mutate(
    str_formula = case_when(
      StationCode == "LIB" ~ paste(f_lag2),
      StationCode == "RVB" ~ paste(f_lag2),
      TRUE ~ f_lag2
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "ML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

# Model Results- two week lag on flow variable

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova_lag2 <- ndf_gam_lag2 %>% 
  mutate(
    an_gam = map(model, anova),
    # Extract AIC from each model
    aic = map_dbl(model, AIC),
    # Parametric terms
    df_parametric = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "pTerms.table"), rownames = "Term")
    ),
    # Smooth terms
    df_smooth = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "s.table"), rownames = "Term") %>% 
        rename(df = edf) %>% 
        select(-Ref.df)
    ),
    df_anova_tbl = map2(df_parametric, df_smooth, bind_rows)
  )

# Combine ANOVA tables and format for export
df_gam_anova_lag2_tbl_c <- ndf_gam_anova_lag2 %>% 
  select(StationCode, df_anova_tbl, aic) %>% 
  unnest(df_anova_tbl) %>% 
  transmute(
    Station = StationCode,
    Term,
    across(
      c("df", "F"),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    ),
    aic
  ) %>%
   # Add AIC as a summary row per station
  bind_rows(
    ndf_gam_anova_lag2 %>%
      transmute(
        Station = as.character(StationCode),
        Term    = "AIC",
        df      = NA_character_,
        F       = NA_character_,
        p_value = NA_character_,
        aic
      )
  ) %>%
  arrange(Station, Term == "AIC")  # Keep AIC rows at the bottom per station

df_gam_anova_lag2_tbl_c
```
# Create GAMs- three week lag on flow variable

```{r create gams, warning = FALSE}
# Define base model formula
f_lag3 <- " zoop_log ~ Year_fct * flow_lag3 + Year_fct + s(Week, bs = 'cc', k = 5)"



# Create gam models for each station
ndf_gam_lag3 <- ndfa_zoop %>% mutate(
    str_formula = case_when(
      StationCode == "LIB" ~ paste(f_lag3),
      StationCode == "RVB" ~ paste(f_lag3),
      TRUE ~ f_lag3
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "ML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

# Model Results- three week lag on flow variable

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova_lag3 <- ndf_gam_lag3 %>% 
  mutate(
    an_gam = map(model, anova),
    # Extract AIC from each model
    aic = map_dbl(model, AIC),
    # Parametric terms
    df_parametric = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "pTerms.table"), rownames = "Term")
    ),
    # Smooth terms
    df_smooth = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "s.table"), rownames = "Term") %>% 
        rename(df = edf) %>% 
        select(-Ref.df)
    ),
    df_anova_tbl = map2(df_parametric, df_smooth, bind_rows)
  )

# Combine ANOVA tables and format for export
df_gam_anova_lag3_tbl_c <- ndf_gam_anova_lag3 %>% 
  select(StationCode, df_anova_tbl, aic) %>% 
  unnest(df_anova_tbl) %>% 
  transmute(
    Station = StationCode,
    Term,
    across(
      c("df", "F"),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    ),
    aic
  ) %>%
   # Add AIC as a summary row per station
  bind_rows(
    ndf_gam_anova_lag3 %>%
      transmute(
        Station = as.character(StationCode),
        Term    = "AIC",
        df      = NA_character_,
        F       = NA_character_,
        p_value = NA_character_,
        aic
      )
  ) %>%
  arrange(Station, Term == "AIC")  # Keep AIC rows at the bottom per station

df_gam_anova_lag3_tbl_c
```

## Extract AIC values into their own table

## Run model diagnostics on top fitting models

## Start a new script or modify the below to create the final tables and figures for the MS.
Effect of Flow by Year. Will show RD22 lag1, STTD no lag, RVB no lag, and LIB 2 week lag of flow

### Effect Plots for top models

```{r gam flow year effect plots, warning = FALSE, fig.height = 9}
# Create effects plot of flow on zooplankton by year for each station
ndf_gam_eff_flow_plt <- ndf_gam %>% 
  mutate(
    df_marg_eff = map2(df_data, model, calc_flow_yr_marg_eff),
    plt_marg_eff = map2(df_marg_eff, df_data, plot_eff_flow_by_yr)
  )

# Combine effects plots together using patchwork
plt_gam_eff_flow <- 
  wrap_plots(ndf_gam_eff_flow_plt$plt_marg_eff, ncol = 1) + 
  plot_layout(axes = "collect")

plt_gam_eff_flow
```
### Slopes and p-values

```{r gam flow year effect stats}
# Calculate slopes and p-values of the modeled effects of flow on zoop by
  # year for each station
ndf_gam_eff_flow_stat <- ndf_gam %>% 
  mutate(
    em_gam_flow = map(model, \(x) emtrends(x, ~ Year_fct, var = "Flow")),
    df_stat = map(em_gam_flow, \(x) as_tibble(test(x)))
  )

# Combine slopes and p-values into one data frame for export
df_gam_eff_flow_stat <- ndf_gam_eff_flow_stat %>% 
  select(StationCode, df_stat) %>% 
  unnest(df_stat) %>% 
  # Format for export
  rename(
    Station = StationCode,
    Year = Year_fct,
    Flow_slope = Flow.trend,
    t_ratio = t.ratio,
    p_value = p.value
  ) %>% 
  mutate(
    p_value = if_else(
      p_value < 0.001, 
      "< 0.001", 
      formatC(p_value, digits = 3, format = "f")
    ),
    across(
      where(is.numeric),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    )
  )

df_gam_eff_flow_stat
```
## Year Contrasts

### Boxplots

```{r gam year contrast boxplots, warning = FALSE}
# Calculate estimated marginal means of year for each station
ndf_gam_cont_yr <- ndf_gam %>% 
  mutate(em_gam_yr = map(model, \(x) emmeans(x, ~ Year_fct)))

# Create boxplots showing Tukey post-hoc results of year contrasts for each station
ndf_gam_cont_yr_plt <- ndf_gam_cont_yr %>% 
  mutate(plt_cont_yr = map2(em_gam_yr, df_data, boxplot_cont_yr))

# Combine boxplots together using patchwork
plt_gam_cont_yr <- 
  wrap_plots(ndf_gam_cont_yr_plt$plt_cont_yr, ncol = 2) + 
  plot_layout(axes = "collect")

plt_gam_cont_yr
```
### Statistics

```{r gam year contrast stats}
# Generate pairwise Tukey post-hoc year contrasts for each station
ndf_gam_cont_yr_stat <- ndf_gam_cont_yr %>% 
  mutate(df_stat = map(em_gam_yr, \(x) as_tibble(pairs(x))))

# Combine pairwise Tukey post-hoc year contrasts into one data frame for export
df_gam_cont_yr_stat <- ndf_gam_cont_yr_stat %>% 
  select(StationCode, df_stat) %>% 
  unnest(df_stat) %>% 
  # Format for export
  rename(
    Station = StationCode,
    Contrast = contrast,
    Estimate = estimate,
    t_ratio = t.ratio,
    p_value = p.value
  ) %>% 
  mutate(
    p_value = if_else(
      p_value < 0.001, 
      "< 0.001", 
      formatC(p_value, digits = 3, format = "f")
    ),
    across(
      where(is.numeric),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    )
  )

print(df_gam_cont_yr_stat, n = Inf)
```
## s(Week) Plots

```{r gam sWeek plots, warning = FALSE}
# Create smooth of Week plots for each station
ndf_gam_s_week_plt <- ndf_gam %>% 
  mutate(plt_s_week = map2(model, StationCode, plot_smooth_week))

# Combine smooth of Week plots together using patchwork
plt_gam_s_week <- 
  wrap_plots(ndf_gam_s_week_plt$plt_s_week, ncol = 2) + 
  plot_layout(axis_titles = "collect") +
  plot_annotation(caption = "Basis: Cyclic CRS")

plt_gam_s_week
```
# Export Results

```{r export results}
# Define file path for figures and tables for the manuscript
fp_figures <- here("manuscript_synthesis/results/figures")
fp_tables <- here("manuscript_synthesis/results/tables")

# Export effects plot of flow by year for each station
ggsave(
  file.path(fp_figures, "zoop_LIS_flow_eff_plot_by_year_each_sta.jpg"),
  plot = plt_gam_eff_flow,
  width = 7,
  height = 9,
  units = "in"
)

ggsave(
  file.path(fp_figures, "zoop_LIS_flow_eff_plot_by_year_each_sta.pdf"),
  plot = plt_gam_eff_flow,
  width = 7,
  height = 9,
  units = "in"
)

# Export boxplots of year contrasts for each station
ggsave(
  file.path(fp_figures, "zoop_LIS_flow_year_contrasts_each_sta.jpg"),
  plot = plt_gam_cont_yr,
  width = 6.5,
  height = 5,
  units = "in"
)

ggsave(
  file.path(fp_figures, "zoop_LIS_flow_year_contrasts_each_sta.pdf"),
  plot = plt_gam_cont_yr,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export effects plots of smooth of Week
ggsave(
  file.path(fp_figures, "zoop_LIS_flow_smooth_week.jpg"),
  plot = plt_gam_s_week,
  width = 6.5,
  height = 5,
  units = "in"
)

ggsave(
  file.path(fp_figures, "zoop_LIS_flow_smooth_week.pdf"),
  plot = plt_gam_s_week,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export combined ANOVA tables for GAMs
df_gam_anova_tbl_c %>% 
  mutate(across(c("df", "F", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "zoop_LIS_flow_GAM_anova.csv"))

# Export slopes and p-values of the modeled effects of flow by year for each station
df_gam_eff_flow_stat %>% 
  mutate(across(c("Flow_slope", "SE", "df", "t_ratio", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "zoop_LIS_flow_GAM_eff_flow_stat.csv"))

# Export pairwise Tukey post-hoc year contrasts
df_gam_cont_yr_stat %>% 
  mutate(across(c("Estimate", "SE", "df", "t_ratio", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "zoop_LIS_flow_GAM_year_contrasts_stat.csv"))
```