```{r}
# C-WQ Regressions for Hypothesis Table
# Authors: Sarah Perry, Dave Bosworth
# Contacts: seperry83@gmail.com; David.Bosworth@water.ca.gov
```

```{r, message=FALSE}
# import packages
library(tidyverse)
library(lubridate)
library(lmerTest)
library(emmeans)
library(car)
library(magrittr)
library(timetk)
source('global_ndfa_funcs.R')
source('Water_Quality/global_wq_funcs.R')

# import data
fp_wq <- ndfa_abs_sp_path('/WQ_Subteam/Processed_Data/Continuous/RTM_INPUT_all_2021-04-20.csv')
fp_region <- ndfa_abs_sp_path('/WQ_Subteam/Processed_Data/Continuous/NDFA_Cont_WQ_Stations.csv')

df_wq <- import_rtm_data(fp_wq, num_params = 10)
df_region <- read_csv(fp_region)

# remove excess cols in df_region
df_region %<>% select(StationCode, BroadRegion)

# Create vector of sites we aren't including (not available for all years)
remove <- c('SRH', 'SHR', 'SRV', 'WWT', 'DWT', 'SDI', 'SGG') # keep TOE, LIBCUT for now

# Clean and Prepare df_wq for analysis
df_wq_clean <- df_wq %>% 
  # Parse DateTime variable and create Date and Year variables
  mutate(
    DateTime = ymd_hms(DateTime),
    Date = date(DateTime),
    Year = year(DateTime)
  ) %>% 
  # Remove sites, parameters, and years we aren't including in the analysis
  filter(
    !StationCode %in% remove,
    Year > 2012
  ) %>% 
  select(
    !c(
      starts_with("Flow"), 
      starts_with("fDOM"), 
      starts_with("Nitrate"), 
      starts_with("WaterTemp")
    )
  ) %>% 
  # Pivot data longer
  select(!ends_with("_Qual")) %>% 
  pivot_longer(
    cols = -c(StationCode, DateTime, Date, Year),
    names_to = "Analyte",
    values_to = "Result"
  ) %>% 
  # Remove NA values
  filter(!is.na(Result)) %>% 
  # Calculate daily averages
  group_by(StationCode, Analyte, Year, Date) %>% 
  summarize(Result = mean(Result)) %>% 
  # Fill in missing days for each StationCode, Analyte, Year combination
  complete(Date = seq.Date(min(Date), max(Date), by = "1 day")) %>% 
  ungroup() %>% 
  # Log transform data
  mutate(Result = log(Result)) %>% 
  # Create lags
  arrange(StationCode, Analyte, Year, Date) %>% 
  group_by(StationCode, Analyte, Year) %>% 
  mutate(
    lag1 = lag(Result),
    lag2 = lag(Result, 2),
    lag3 = lag(Result, 3)
  ) %>% 
  ungroup() %>% 
  # Add Flow Action Periods, Regions, Month, and Week variables
  ndfa_action_periods() %>% 
  left_join(df_region) %>% 
  mutate(
    Month = as.character(month(Date)),
    Week = as.character(week(Date)),
    # Convert Year to character so it runs correctly in the model
    Year = as.character(Year)
  )
```

```{r create diag plots}
analytes <- unique(df_wq_clean$Analyte)

# Define file path for ANOVA Diagnostics Plots
fp_anova_diag <- ndfa_abs_sp_path("WQ_Subteam/Plots/Continuous/ANOVA_Diagnostics")

# Create seasonal Diagnostics plots first
pdf(file.path(fp_anova_diag, "Seasonal_Diagnostics.pdf"))
  for (ana in analytes) {
    df_temp <- df_wq_clean %>% filter(Analyte == ana)
    
    plt <- df_temp %>% 
      plot_seasonal_diagnostics(
        Date, 
        Result, 
        .feature_set = c("week", "month.lbl"), 
        .interactive = FALSE
      ) +
      ggtitle(ana)
    
    print(plt)
  }
dev.off()  

# Create ANOVA Diagnostics plots for no, 1, 2, and 3 lag terms without a seasonal term
for (ana in analytes) {
  df_temp <- df_wq_clean %>% filter(Analyte == ana)
  
  # Create models
  lmer_fit0 <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + (1|StationCode), 
      data = df_temp
    )
  lmer_fit1 <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + (1|StationCode), 
      data = df_temp
    )
  lmer_fit2 <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + (1|StationCode), 
      data = df_temp
    )
  lmer_fit3 <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + (1|StationCode), 
      data = df_temp
    )
  
  # Create and save ANOVA Diagnostics plots
  pdf(paste0(fp_anova_diag, "/", ana, "_daily_avg_sea-none.pdf"))
    acf(residuals(lmer_fit0), na.action = na.pass, main = paste0(ana, "- no lag terms"))
    acf(residuals(lmer_fit1), na.action = na.pass, main = paste0(ana, "- 1 lag term"))
    acf(residuals(lmer_fit2), na.action = na.pass, main = paste0(ana, "- 2 lag terms"))
    acf(residuals(lmer_fit3), na.action = na.pass, main = paste0(ana, "- 3 lag terms"))
    plot(residuals(lmer_fit0), main = paste0(ana, "- no lag terms"))
    plot(residuals(lmer_fit1), main = paste0(ana, "- 1 lag term"))
    plot(residuals(lmer_fit2), main = paste0(ana, "- 2 lag terms"))
    plot(residuals(lmer_fit3), main = paste0(ana, "- 3 lag terms"))
    qqnorm(residuals(lmer_fit0), main = paste0("Normal Q−Q Plot- ", ana, ", no lag terms"))
    qqnorm(residuals(lmer_fit1), main = paste0("Normal Q−Q Plot- ", ana, ", 1 lag term"))
    qqnorm(residuals(lmer_fit2), main = paste0("Normal Q−Q Plot- ", ana, ", 2 lag terms"))
    qqnorm(residuals(lmer_fit3), main = paste0("Normal Q−Q Plot- ", ana, ", 3 lag terms"))
    hist(residuals(lmer_fit0), breaks = 15)
    hist(residuals(lmer_fit1), breaks = 15)
    hist(residuals(lmer_fit2), breaks = 15)
    hist(residuals(lmer_fit3), breaks = 15)
  dev.off()
  
  # Print off results of Portmanteau tests to look for autocorrelation among the residuals
  print(ana)
  print(stats::Box.test(residuals(lmer_fit0), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit1), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit2), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit3), lag = 20, type = 'Ljung-Box'))
}

# Chla: 2 lag terms is best
# DO: 2 lag terms is best (lag 3 is significant)
# pH: 3 lag terms is best
# SpCnd: 3 lag terms is best
# Turbidity: 3 lag terms is best

# Next, we'll add seasonal terms to these best-fitting models to see if they improve anything
# Chla and DO with 2 lag terms
analytes2 <- c("Chla", "DO")
for (ana in analytes2) {
  df_temp <- df_wq_clean %>% filter(Analyte == ana)
  
  # Create models
  lmer_fit2 <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + (1|StationCode), 
      data = df_temp
    )
  lmer_fit2_month <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + Month + (1|StationCode), 
      data = df_temp
    )
  lmer_fit2_week <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + Week + (1|StationCode), 
      data = df_temp
    )
  # Create and save ANOVA Diagnostics plots
  pdf(paste0(fp_anova_diag, "/", ana, "_daily_avg_2lag_seasonal.pdf"))
    acf(
      residuals(lmer_fit2), 
      na.action = na.pass, 
      main = paste0(ana, "- 2 lags, no seasonal terms")
    )
    acf(
      residuals(lmer_fit2_month), 
      na.action = na.pass, 
      main = paste0(ana, "- 2 lags, Month seasonal term")
    )
    acf(
      residuals(lmer_fit2_week), 
      na.action = na.pass, 
      main = paste0(ana, "- 2 lags, Week seasonal term")
    )
    plot(residuals(lmer_fit2), main = paste0(ana, "- 2 lags, no seasonal terms"))
    plot(residuals(lmer_fit2_month), main = paste0(ana, "- 2 lags, Month seasonal term"))
    plot(residuals(lmer_fit2_week), main = paste0(ana, "- 2 lags, Week seasonal term"))
    qqnorm(
      residuals(lmer_fit2), 
      main = paste0("Normal Q−Q Plot- ", ana, ", 2 lags, no seasonal terms")
    )
    qqnorm(
      residuals(lmer_fit2_month), 
      main = paste0("Normal Q−Q Plot- ", ana, ", 2 lags, Month seasonal term")
    )
    qqnorm(
      residuals(lmer_fit2_week), 
      main = paste0("Normal Q−Q Plot- ", ana, ", 2 lags, Week seasonal term")
    )
    hist(residuals(lmer_fit2), breaks = 15)
    hist(residuals(lmer_fit2_month), breaks = 15)
    hist(residuals(lmer_fit2_week), breaks = 15)
  dev.off()
  
  # Print off results of Portmanteau tests to look for autocorrelation among the residuals
  print(ana)
  print(stats::Box.test(residuals(lmer_fit2), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit2_month), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit2_week), lag = 20, type = 'Ljung-Box'))
}
# 2 lags, no seasonal terms is best for Chla and DO

# DO, pH, SpCnd, and Turbidity with 3 lag terms
analytes3 <- c("DO", "pH", "SpCnd", "Turbidity")
for (ana in analytes3) {
  df_temp <- df_wq_clean %>% filter(Analyte == ana)
  
  # Create models
  lmer_fit3 <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + (1|StationCode), 
      data = df_temp
    )
  lmer_fit3_month <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + Month + (1|StationCode), 
      data = df_temp
    )
  lmer_fit3_week <- 
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + Week + (1|StationCode), 
      data = df_temp
    )
  # Create and save ANOVA Diagnostics plots
  pdf(paste0(fp_anova_diag, "/", ana, "_daily_avg_3lag_seasonal.pdf"))
    acf(
      residuals(lmer_fit3), 
      na.action = na.pass, 
      main = paste0(ana, "- 3 lags, no seasonal terms")
    )
    acf(
      residuals(lmer_fit3_month), 
      na.action = na.pass, 
      main = paste0(ana, "- 3 lags, Month seasonal term")
    )
    acf(
      residuals(lmer_fit3_week), 
      na.action = na.pass, 
      main = paste0(ana, "- 3 lags, Week seasonal term")
    )
    plot(residuals(lmer_fit3), main = paste0(ana, "- 3 lags, no seasonal terms"))
    plot(residuals(lmer_fit3_month), main = paste0(ana, "- 3 lags, Month seasonal term"))
    plot(residuals(lmer_fit3_week), main = paste0(ana, "- 3 lags, Week seasonal term"))
    qqnorm(
      residuals(lmer_fit3), 
      main = paste0("Normal Q−Q Plot- ", ana, ", 3 lags, no seasonal terms")
    )
    qqnorm(
      residuals(lmer_fit3_month), 
      main = paste0("Normal Q−Q Plot- ", ana, ", 3 lags, Month seasonal term")
    )
    qqnorm(
      residuals(lmer_fit3_week), 
      main = paste0("Normal Q−Q Plot- ", ana, ", 3 lags, Week seasonal term")
    )
    hist(residuals(lmer_fit3), breaks = 15)
    hist(residuals(lmer_fit3_month), breaks = 15)
    hist(residuals(lmer_fit3_week), breaks = 15)
  dev.off()
  
  # Print off results of Portmanteau tests to look for autocorrelation among the residuals
  print(ana)
  print(stats::Box.test(residuals(lmer_fit3), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit3_month), lag = 20, type = 'Ljung-Box'))
  print(stats::Box.test(residuals(lmer_fit3_week), lag = 20, type = 'Ljung-Box'))
}
# 3 lags, no seasonal terms is best for pH, SpCnd, and Turbidity
```

We are going to use the following model structures:

* Chla: Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + (1|StationCode)
* DO: Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + (1|StationCode)
* pH: Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + (1|StationCode)
* SpCnd: Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + (1|StationCode)
* Turbidity: Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + (1|StationCode)

```{r all analytes}
# Create function to make lmer model object with either 2 or 3 lag terms
create_lmer_obj <- function(df, num_lags = c(2, 3)) {
  # Define options for num_lags argument
  match.arg(num_lags, c(2, 3))
  
  # Create lmer model object
  if (num_lags == 2) {
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + (1|StationCode),
      data = df
    )
  } else {
    lmer(
      Result ~ Year + FlowActionPeriod + BroadRegion + lag1 + lag2 + lag3 + (1|StationCode),
      data = df
    )
  }
}

# Create function to run Anova F-test with type II sum of squares on a lmer object and format it
run_anova2 <- function(lmer_obj) {
  aov_lmer <- Anova(lmer_obj, type = 'II', test.statistic = 'F')
  
  aov_lmer %>% 
    rownames_to_column(var = "Variable") %>% 
    mutate(
      Significance = case_when(
        `Pr(>F)` >= 0 & `Pr(>F)` < 0.001 ~ '***',
        `Pr(>F)` >= 0.001 & `Pr(>F)` < 0.01 ~ '**',
        `Pr(>F)` >= 0.01 & `Pr(>F)` < 0.05 ~ '*',
        `Pr(>F)` >= 0.05 & `Pr(>F)` < 0.1 ~ '.',
        `Pr(>F)` >= 0.1 & `Pr(>F)` <= 1 ~ ' ',
      )
    )
}

# Create function to format emmeans pairwise results
format_emm_pw <- function(emm_obj) {
  emm_cont <- test(emm_obj)$contrasts
  emm_cont %>% 
    mutate(
      Significance = case_when(
        p.value >= 0 & p.value < 0.001 ~ '***',
        p.value >= 0.001 & p.value < 0.01 ~ '**',
        p.value >= 0.01 & p.value < 0.05 ~ '*',
        p.value >= 0.05 & p.value < 0.1 ~ '.',
        p.value >= 0.1 & p.value <= 1 ~ ' ',
      )
    )
}

# Create function to run emmeans pairwise tests with Sidak adjustment method
run_emm_pw_all <- function(lmer_obj) {
  lmer_rg <- ref_grid(lmer_obj, pbkrtest.limit = 15000, lmerTest.limit = 15000)
  
  # Run pairwise test on Year variable
  lmer_emm_yr <- emmeans(
    lmer_rg, 
    specs = pairwise ~ Year:Year, 
    adjust = 'sidak'
  ) 
  cont_yr <- format_emm_pw(lmer_emm_yr)
  
  # Run pairwise test on FlowActionPeriod variable
  lmer_emm_fa <- emmeans(
    lmer_rg, 
    specs = pairwise ~ FlowActionPeriod:FlowActionPeriod, 
    adjust = 'sidak'
  ) 
  cont_fa <- format_emm_pw(lmer_emm_fa)
  
  # Run pairwise test on BroadRegion variable
  lmer_emm_region <- emmeans(
    lmer_rg, 
    specs = pairwise ~ BroadRegion:BroadRegion, 
    adjust = 'sidak'
  ) 
  cont_region <- format_emm_pw(lmer_emm_region)
  
  # Return a named list of the results
  lmer_emm_list <- list(
    "Year" = cont_yr, 
    "FlowActionPeriod" = cont_fa,
    "BroadRegion" = cont_region
  )
  
  return(lmer_emm_list)
}

# Create models for each parameter
df_models <- df_wq_clean %>% 
  nest(df = -Analyte) %>% 
  mutate(
    n_lags = as.character(if_else(Analyte %in% c("Chla", "DO"), 2, 3)),
    Model = map2(df, n_lags, create_lmer_obj)
  )

# Run Anova F-tests with type II sum of squares on each model
df_models_aov <- df_models %>% mutate(aov_Model = map(Model, run_anova2))

# Run emmeans pairwise tests on each model and variable, then separate results by variable
df_emm_test <- df_models %>% 
  mutate(emm_pw_all = map(Model, run_emm_pw_all)) %>% 
  mutate(
    emm_pw_yr = map(emm_pw_all, ~chuck(.x, "Year")),
    emm_pw_fa = map(emm_pw_all, ~chuck(.x, "FlowActionPeriod")),
    emm_pw_region = map(emm_pw_all, ~chuck(.x, "BroadRegion"))
  )

# Export results as .csv files
fp_anova_results <- ndfa_abs_sp_path("WQ_Subteam/Analysis_Results/CWQ")

df_models_aov %>% 
  select(Analyte, aov_Model) %>% 
  unnest(aov_Model) %>% 
  write_excel_csv(file.path(fp_anova_results, "cwq_aov_daily_avg_sea-none.csv"))

df_emm_test %>% 
  select(Analyte, emm_pw_yr) %>% 
  unnest(emm_pw_yr) %>% 
  write_excel_csv(file.path(fp_anova_results, "cwq_emmeans_year_daily_avg_sea-none.csv"))

df_emm_test %>% 
  select(Analyte, emm_pw_fa) %>% 
  unnest(emm_pw_fa) %>%  
  write_excel_csv(file.path(fp_anova_results, "cwq_emmeans_flowaction_daily_avg_sea-none.csv"))

df_emm_test %>% 
  select(Analyte, emm_pw_region) %>% 
  unnest(emm_pw_region) %>%  
  write_excel_csv(file.path(fp_anova_results, "cwq_emmeans_region_daily_avg_sea-none.csv"))
```

