---
title: "FASTR - Phytoplankton Data Processing"
author: "Ted Flynn"
date: "2024-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(knitr)))
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(vegan)))
suppressWarnings(suppressMessages(library(janitor)))
suppressWarnings(suppressMessages(library(here)))

```

## Import phyto data collected for FASTR project

These steps import raw phytoplankton data collected by DWR's Aquatic Ecology group as part of the FASTR project and combines them into a single data frame. 

```{r import FASTR, echo=FALSE, message=FALSE}
# Import AEU data files (comment out when finished)

files_AEU <- list.files(path = here("phyto-final","data","csv"), 
                   pattern = "\\.csv", 
                   full.names = T)

df_phyto_AEU <- map(files_AEU, ~read_csv(.x, show_col_types = FALSE)) %>% 
  list_rbind()

# Remove empty rows
df_phyto_AEU <- df_phyto_AEU %>% filter_all(any_vars(!is.na(.)))

# Remove weird row with only a single zero in biovolume
df_phyto_AEU <- df_phyto_AEU %>% drop_na(MethodCode)

# Clean up column names
df_phyto_AEU <- df_phyto_AEU %>% clean_names(case = "big_camel")

# Filter out samples from other projects. Read in station names with flags and 
# merge together.
df_keepers <- read_csv(here("phyto-final","CSVs", "station_names_flagged.csv"),
                       show_col_types = FALSE)

df_phyto_AEU <- left_join(df_phyto_AEU, df_keepers)

rm(df_keepers)

# Remove data unrelated to project
df_phyto_AEU <- df_phyto_AEU %>% filter(Flag == "Keep")

# Remove flag column
df_phyto_AEU <- df_phyto_AEU %>% select(!(Flag))

# Confirm station IDs
unique(df_phyto_AEU$StationCode) # Shows only 10 FASTR stations
table(df_phyto_AEU$StationCode)

# Remove blank columns
df_phyto_AEU <- df_phyto_AEU %>% select_if(~ !all(is.na(.)))

# Remove individual measurement columns as they are only present in a small 
# number of samples.
df_phyto_AEU <- df_phyto_AEU %>% select(!(Dimension:DepthM))

# Remove secondary and tertiary GALD measurements as they don't vary much and 
# aren't present in EMP data
df_phyto_AEU <- df_phyto_AEU %>% select(!(Gald2:Gald3))

# Correct GALD, which is imported into two separate columns. Test to see if NAs 
# are 'either/or' and that there aren't some rows with a value in both GALD and 
# GALD 1

sum(is.na(df_phyto_AEU$Gald)) # Total is 1791
sum(is.na(df_phyto_AEU$Gald1)) # Total is 4535

# Sum of NAs is 6326 which is the same as the number of rows in the df. This
# shows that there aren't any rows with two values so that we can Combine them 
# without any issues.

# Move Gald1 column
df_phyto_AEU <- df_phyto_AEU %>% relocate(Gald1, .after = Gald)

# Combine both GALD columns
df_phyto_AEU <- df_phyto_AEU %>%
  rowwise() %>%
  mutate(GALD.Tot = sum(c_across(Gald:Gald1), na.rm = TRUE))

# Check if rows have two NAs or no NAs in the GALD columns
# test <- df_phyto_AEU %>%
#   mutate(GALD.Value = case_when(is.na(Gald) & is.na(Gald1) ~ "Fix",
#                                 TRUE ~ "Okay"))

# Remove old GALD columns and rename GALD.Tot
df_phyto_AEU <- df_phyto_AEU %>% 
  select(!(Gald:Gald1)) %>%
  rename("GALD" = "GALD.Tot")

# Remove MethodCode column that just says "Phyto" 
df_phyto_AEU <- df_phyto_AEU %>% select(!(MethodCode))

# Add column to indicate what study it came from
df_phyto_AEU <- df_phyto_AEU %>% mutate(Study = "FASTR", .after = StationCode)

```

## Import EMP Data

Import phytoplankton data collected by the Environmental Monitoring Program. 

```{r import EMP, echo=FALSE, message=FALSE}

# Import EMP data files
phyto_files_EMP <- list.files(path = here("phyto-final","data","EMP","csv"), 
                   pattern = "\\.csv", 
                   full.names = T)

df_phyto_EMP <- map(phyto_files_EMP, ~read_csv(.x, show_col_types = FALSE)) %>% 
  list_rbind()

# Read in files with non-standard headers individually
df_Dec2021 <- read_csv(here("phyto-final","data","EMP","oddballs","December 2021.csv"), 
                       show_col_types = FALSE)
df_Nov2021 <- read_csv(here("phyto-final","data","EMP","oddballs","November 2021.csv"),
                       show_col_types = FALSE)
df_Sep2013 <- read_csv(here("phyto-final","data","EMP","oddballs","September 2013.csv"), 
                       show_col_types = FALSE)
df_Nov2013 <- read_csv(here("phyto-final","data","EMP","oddballs","November 2013.csv"), 
                       show_col_types = FALSE)

# Combine like oddball dfs
df_phyto2013 <- bind_rows(df_Sep2013, df_Nov2013)
df_phyto2021 <- bind_rows(df_Dec2021, df_Nov2021)

# Remove individual dfs
rm(df_Dec2021)
rm(df_Nov2021)
rm(df_Nov2013)
rm(df_Sep2013)

# Rename headers to match standard BSA headers Oddballs actually have the 
# "correct" name of Total Cells rather than the incorrect "Number of cells per 
# unit"

df_phyto2013 <- df_phyto2013 %>%
  rename("Number of cells per unit" = "Total Cells Counted")

df_phyto2021 <- df_phyto2021 %>%
  rename("Number of cells per unit" = "Total Number of Cells") %>%
  rename("Unit Abundance" = "Unit Abundance (# of Natural Units)")

# Combine oddball files with others
df_phyto_EMP <- bind_rows(df_phyto_EMP, df_phyto2013)
df_phyto_EMP <- bind_rows(df_phyto_EMP, df_phyto2021)

# Remove unneeded dfs
rm(df_phyto2013)
rm(df_phyto2021)

# Remove empty rows
df_phyto_EMP <- df_phyto_EMP %>% filter_all(any_vars(!is.na(.)))

# Correct GALD, which is imported into two separate columns
# Test to see if NAs are 'either/or' and that there aren't some rows with a 
# value in both GALD and GALD 1

sum(is.na(df_phyto_EMP$GALD)) # Total is 5880
sum(is.na(df_phyto_EMP$`GALD 1`)) # Total is 8072

# Sum of NAs is 13952 which is the same as the number of rows in the df.
# This shows that there aren't any rows with two values so that we can 
# Combine them without any issues.

# Move GALD header
df_phyto_EMP <- df_phyto_EMP %>% relocate(`GALD 1`, .after = GALD)

# Combine both GALD columns
df_phyto_EMP <- df_phyto_EMP %>%
  rowwise() %>%
  mutate(GALD.Tot = sum(c_across(GALD:`GALD 1`), na.rm = TRUE))

# Remove old GALD columns and rename GALD.Tot
df_phyto_EMP <- df_phyto_EMP %>% 
  select(!(GALD:`GALD 1`)) %>%
  rename("GALD" = "GALD.Tot")

# Clean up column names
df_phyto_EMP <- df_phyto_EMP %>% clean_names(case = "big_camel")

df_phyto_EMP <- df_phyto_EMP %>% rename("GALD" = "Gald")

# Remove blank columns
df_phyto_EMP <- df_phyto_EMP %>% select_if(~ !all(is.na(.)))

# Remove columns that just have the method code "Phyto" as well as 
# pre-calculated organisms per mL
df_phyto_EMP <- df_phyto_EMP %>% select(SampleDate:Biovolume10,GALD)

# Add column to indicate what study it came from
df_phyto_EMP <- df_phyto_EMP %>% mutate(Study = "EMP", .after = StationCode)

```

## Combine EMP and AEU data

```{r combine data, echo=FALSE, message=FALSE}

df_phyto <- bind_rows(df_phyto_EMP, df_phyto_AEU)

# Average all 10 biovolume measurements for each taxon
df_phyto <- df_phyto %>% 
  rowwise() %>% 
  mutate(BV.Avg = mean(c_across(Biovolume1:Biovolume10), na.rm = T)) %>% 
  select(!(Biovolume1:Biovolume10)) # Remove Individual Biovolume Columns

# Remove unneeded columns
df_phyto <- df_phyto %>% select(!c("BsaTin","DiatomSoftBody"))
df_phyto <- df_phyto %>% select(!(ColonyFilamentIndividualGroupCode:Shape))
df_phyto <- df_phyto %>% select(!(VolumeReceivedML:NumberOfFieldsCounted))

# Fix samples with missing times
# Just replace with 12:00:00 b/c aren't doing any time-based analyses
df_phyto <- df_phyto %>% replace_na(list(SampleTime = "12:00:00"))

# Get dates in the right format. Some are 1/1/14 and others 1/1/2014.
df_phyto$SampleDate <- mdy(df_phyto$SampleDate)

# Combine date and time column
df_phyto <- df_phyto %>% unite(DateTime, c("SampleDate","SampleTime"), sep = " ") #, remove = FALSE, na.rm = FALSE)

df_phyto$DateTime <- as_datetime(df_phyto$DateTime, 
                              tz = "US/Pacific",
                              format = c("%Y-%m-%d %H:%M:%OS"))

# Check for missing dates
df_phyto %>% filter(is.na(DateTime)) # No missing dates

# Correct BSA header
df_phyto <- df_phyto %>% rename("TotalCells" = "NumberOfCellsPerUnit")

# Calculate Unit Density & Biovolume Density
df_phyto <- df_phyto %>%
  mutate(Units.per.mL = UnitAbundance * Factor) %>%
  mutate(BV.um3.per.mL= TotalCells * BV.Avg * Factor) %>% 
  mutate(Cells.per.mL = TotalCells * Factor) 

# Remove columns used for density calcs
# Remove Biovolume Columns
df_phyto <- df_phyto %>% select(!(c(Factor,UnitAbundance:TotalCells)))

# Add column for year and month for highlighting data
df_phyto <- df_phyto %>% 
  mutate(Year = year(df_phyto$DateTime)) %>%
  mutate(Month = month(df_phyto$DateTime, label = T))

# Order month in calendar order rather than (default) alphabetical
df_phyto$Month = factor(df_phyto$Month, levels = month.abb)

# Reorder date/time columns
df_phyto <- df_phyto %>% 
  relocate(Year, .after = DateTime) %>% 
  relocate(Month, .after = DateTime)

# Remove 2013 data (FASTR data is very limited)
df_phyto <- df_phyto %>% filter(Year != 2013)

```

## Fix Station Names

```{r data cleaning, echo=FALSE, message=FALSE}

# Fix EMP site names
df_phyto$StationCode <- gsub("EZ6 SAC","EZ6",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ6SAC","EZ6",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ6SJR","EZ6-SJR",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ2SAC","EZ2",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ2 SAC","EZ2",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ2SJR","EZ2-SJR",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ2 SJR","EZ2-SJR",df_phyto$StationCode)
df_phyto$StationCode <- gsub("EZ6 SJR","EZ6-SJR",df_phyto$StationCode)
df_phyto$StationCode <- gsub("D16-Twitchell","D16",df_phyto$StationCode)
df_phyto$StationCode <- gsub("D16-Twitchel","D16",df_phyto$StationCode)
df_phyto$StationCode <- gsub("D16 - Twitchell","D16",df_phyto$StationCode)
df_phyto$StationCode <- gsub("D16 Twitchell","D16",df_phyto$StationCode)
df_phyto$StationCode <- gsub("NZ328","NZ325",df_phyto$StationCode) # Typo in August 2019
df_phyto$StationCode <- gsub("C3A-HOOD","C3A",df_phyto$StationCode)
df_phyto$StationCode <- gsub("C3A Hood","C3A",df_phyto$StationCode)
df_phyto$StationCode <- gsub("C3A- Hood","C3A",df_phyto$StationCode)
df_phyto$StationCode <- gsub("C3A-Hood","C3A",df_phyto$StationCode)
df_phyto$StationCode <- gsub("NZ542","NZS42",df_phyto$StationCode)
df_phyto$StationCode <- gsub("E26","EZ6",df_phyto$StationCode)
df_phyto$StationCode <- gsub("E22","EZ2",df_phyto$StationCode) # Typo in May 2018

# Fix AEU site names
df_phyto$StationCode <- gsub("SHER","SHR",df_phyto$StationCode)
df_phyto$StationCode <- gsub("BL-5","BL5",df_phyto$StationCode)

# Remove extra Microcystis tows at D19
df_phyto <- df_phyto %>% filter(StationCode != "D19 MC Tow")

```

## Add Taxonomy Data and Fix Errors

In Fall 2016, taxonomists began classifying the species Chroococcus microscopicus as Eucapsis microscopica. This is one of the most dominant species in this samples, so all taxa previously classified asC. microscopicus will be re-named E. microscopica. Also, the taxon Plagioselmis lacustris is inconsistently named, appearing sometimes as Rhodomonas lacustris. Change to Rhodomonas lacustris to avoid having these show up as separate genera. 

We also use a CSV of group classification names matched up with the genera occurring here to add higher-level taxonomic resolution.

```{r taxonomy, echo=FALSE, message=FALSE}

# Rename Chroococcus microscopicus as Eucapsis microscopica
df_phyto <- df_phyto %>% 
  mutate(Taxon = case_when(Taxon == 'Chroococcus microscopicus' ~ 'Eucapsis microscopica',
                           TRUE ~ Taxon)) %>%
  mutate(Genus = case_when(Taxon == 'Eucapsis microscopica' ~ 'Eucapsis',
                           TRUE ~ Genus))

# Rename Plagioselmis lacustris as Rhodomonas lacustris
df_phyto <- df_phyto %>% 
  mutate(Taxon = case_when(Taxon == 'Plagioselmis lacustris' ~ 'Rhodomonas lacustris',
                           TRUE ~ Taxon)) %>%
  mutate(Genus = case_when(Taxon == 'Rhodomonas lacustris' ~ 'Rhodomonas',
                           TRUE ~ Genus))

# Correct the genus label for a Chlorella entry
df_phyto$Genus <- gsub("cf Chlorella","Chlorella",df_phyto$Genus)

# Read in CSV with manually-added WoRMS classification
df_taxa <- read_csv(here("phyto-final","CSVs","phyto_group_classification.csv"),
                    show_col_types = FALSE)

df_phyto <- left_join(df_phyto, df_taxa)

# Check if any groups are missing
sum(is.na(df_phyto$Group)) # none missing 

# Reorder Group classification column
df_phyto <- df_phyto %>% 
  relocate(Group, .before = Taxon)

# Read in Region data for EMP stations
df_region <- read_csv(here("phyto-final","CSVs","station_regions.csv"),
                      show_col_types = FALSE)

# Combine region data and phyto data
df_phyto <- left_join(df_phyto, df_region)

# Reorder column
df_phyto <- df_phyto %>% relocate(Region, .after = StationCode)

# check if there are any NAs in Region after the join
table(is.na(df_phyto$Region)) # no NAs

# Remove data frames with old data
rm(df_region)
rm(df_phyto_EMP)
rm(df_phyto_AEU)

```
## Estimate Phytoplankton Biomass and LCEFA Abundance

Convert biovolume to biomass using relationships from Menden-Deuer and Lussard (2000, doi: 10.4319/lo.2000.45.3.0569). This is only relevant to group-level data. Use biovolume measurements to calculate the abundance of long-chain essential fatty acids for each group based on the method in Galloway & Winder (2015, doi: 10.1371/journal.pone.0130053). 


```{r biomass calcs, echo=FALSE, message=FALSE}

# Calculate Biomass (pg-C per cell) from Biovolume (um^3 per cell)
df_phyto <- df_phyto %>% 
  mutate(Biomass.pg.C = case_when(Group == "Diatoms" ~ 0.288 * (BV.Avg^0.811), 
                                  Group != "Diatoms" ~ 0.216 * (BV.Avg^0.939)))

# Convert pg to ug (ug-C per mL)
df_phyto <- df_phyto %>% 
  mutate(Biomass.ug.C = Biomass.pg.C / 10^6, .keep = "unused")

# Calculate Biomass Density (pg-C per mL)
df_phyto <- df_phyto %>% 
  mutate(BM.ug.per.mL = Biomass.ug.C * Cells.per.mL)

df_phyto <- df_phyto %>% 
  mutate(LCEFA.per.mL = case_when(Group == 'Diatoms' ~ BM.ug.per.mL*2.77/100,
                                  Group == 'Cyanobacteria' ~ BM.ug.per.mL*0.02/100,
                                  Group == 'Green Algae' ~ BM.ug.per.mL*0.52/100,
                                  Group == 'Cryptophytes' ~ BM.ug.per.mL*2.13/100,
                                  Group == 'Dinoflagellates' ~ BM.ug.per.mL*3.82/100,
                                  TRUE ~ 0))

```

## Add Regional Metadata
```{r regions, echo=FALSE, message=FALSE}

# Select only stations in FASTR report
df_phyto <- df_phyto %>%
  filter(!Region %in% c("San Pablo Bay",
                        "Grizzly and Suisun Bay",
                        "Entrapment Zone"))

# Set display levels for different regions
unique(df_phyto$Region)

region.levels <- c("Upstream","Downstream","Northern Interior Delta","Confluence",
                   "Central Delta","Southern Interior Delta")

# Set levels so that Upstream stations appear before Downstream in graphs
df_phyto$Region <- factor(df_phyto$Region, levels = region.levels)

# Import flow pulse dates and categories
df_flow_designation <- read_csv(here("phyto-final",
                                     "CSVs",
                                     "FlowDatesDesignations_45days.csv"),
                                show_col_types = FALSE)

# Remove column with net flow days. Won't be used in any analyses.
df_flow_designation <- df_flow_designation %>% select(!(NetFlowDays))

# Format dates as dates in each of the 4 date columns
df_flow_designation <- df_flow_designation %>% 
  mutate(across(PreFlowStart:PostFlowEnd, mdy))

# Merge data from FlowDesignation Table (Water Year Type, Flow days and type)
df_phyto <- inner_join(df_phyto, df_flow_designation, by = "Year")   

# Label samples with the appropriate phase (Pre-During-Post) and filter
# FASTR phyto data to include only samples collected within the flow pulse window
df_phyto <- df_phyto %>%
  mutate(ActionPhase = ifelse(DateTime > PreFlowStart & DateTime < PreFlowEnd, "Before", NA)) %>%
  mutate(ActionPhase = replace(ActionPhase, DateTime >= PreFlowEnd & DateTime < PostFlowStart, "During")) %>% 
  mutate(ActionPhase = replace(ActionPhase, DateTime >= PostFlowStart & DateTime < PostFlowEnd, "After")) %>% 
  filter(!is.na(ActionPhase)) %>%
  select(-c(PreFlowStart:PostFlowEnd))

# Order the new ActionPhase so that it plots in the order Pre < During < Post
phase.order <- c("Before","During","After")
df_phyto$ActionPhase <- factor(as.character(df_phyto$ActionPhase), 
                                     levels = phase.order)

# Add in Flow Pulse Category Data
df_pulse_category <- read_csv(here("phyto-final","CSVs","FlowPulseType.csv"),
                              show_col_types = FALSE)

df_pulse_category <- df_pulse_category %>% 
  rename("FlowPulseCategory" = "FlowPulseType")

# Add Flow Pulse Category to phytoplankton data frame
df_phyto <- left_join(df_phyto, df_pulse_category, by = "Year")

## Relocate metadata columns
df_phyto <- df_phyto %>%
  relocate(FlowPulseCategory, .after = Study) %>%
  relocate(WYType:ActionPhase, .after = Study) %>% 
  relocate(Biomass.ug.C, .after = BV.Avg)

```


## Analyze FASTR Data for Outliers

```{r outlier analysis, echo=FALSE, message=FALSE}

# Subset combined dataset to FASTR-only
df_phyto_FASTR <- df_phyto %>% filter(Study == "FASTR")

# Identify if outliers are present
boxplot(df_phyto_FASTR$BV.um3.per.mL)

# Identify 0.1% and 99.9% percentiles of the data
quartiles <- quantile(df_phyto_FASTR$BV.um3.per.mL, 
                      probs=c(0.001, 0.999), 
                      na.rm = TRUE)

# List upper cutoff (99.9%)
cutoff <- quartiles[2]

# Filter FASTR dataset to show all taxa above this 99.9% cutoff
df_outliers <- df_phyto_FASTR %>% 
  filter(BV.um3.per.mL > cutoff)

df_outliers %>% 
  select(DateTime,Taxon,BV.um3.per.mL) %>% 
  kable(caption = "Al")## 4 of top 6 are Spirogyra. 

df_phyto_FASTR %>% filter(Genus == "Spirogyra")# Only 5 total samples w/ Spirogyra

# Decision: Remove Spirogyra from datasets
# 4/5 are above 99.9% cutoff, 5th is also highly abundant
# These are benthic algae that clump, more likely to be "bullseye" samples
# that got a big blob or clump

# Remove Spirogyra from FASTR and combined dataset
df_phyto_FASTR <- df_phyto_FASTR %>% 
  filter(Genus != "Spirogyra")

df_phyto <- df_phyto %>% 
  filter(Genus != "Spirogyra")

```

## Add FASTR-Specific Metadata

This section adds metadata specific to the FASTR project to a subset of the data. EMP data is not included in this data frame as the categories do not apply to it. 

Code for Merging Flow Designations was provided by Cat Pien and includes 45 days limit on either end of the flow pulse. Code was tweaked by TF in May 2022 to include greater-than or equal-to signs avoid inadvertently removing samples falling on the exact dates of pre- and post-flow phases. 

```{r FASTR metadata, echo=FALSE, message=FALSE}

# Create a list of stations used for FASTR
stations <- c("RMB","RCS","RD22","I80","LIS","STTD","BL5","LIB","RYI","RVB")

# Re-order stations from North to South
df_phyto_FASTR$StationCode <- factor(as.character(df_phyto_FASTR$StationCode),
                                     levels = stations)

```

## Create New Data Frames summarizing by Genus and Algal Group

```{r summarize phyto data, echo=FALSE, message=FALSE}

# Summarize by algal group
df_phyto_FASTR_grp <- df_phyto_FASTR %>%
  group_by(Year, Month, Region, ActionPhase, DateTime, StationCode, Group) %>%
  summarize(across(Units.per.mL:LCEFA.per.mL, ~sum(.x, na.rm = TRUE))) %>%
  ungroup

# Summarize by genus
df_phyto_FASTR_gen <- df_phyto_FASTR %>%
  group_by(Year, Month, Region, ActionPhase, DateTime, StationCode, Genus) %>%
  summarize(across(Units.per.mL:LCEFA.per.mL, ~sum(.x, na.rm = TRUE))) %>%
  ungroup

# Set group display order
group.order <- c("Diatoms",
                 "Cyanobacteria",
                 "Green Algae",
                 "Cryptophytes",
                 "Ciliates",
                 "Dinoflagellates",
                 "Golden Algae",
                 "Other")

df_phyto_FASTR_grp$Group <- factor(as.character(df_phyto_FASTR_grp$Group),
                                   levels =  group.order)

# Summarize totals
df_phyto_FASTR_sum <- df_phyto_FASTR %>%
  group_by(Year, Month, DateTime, StationCode, Region, WYType, FlowPulseType, FlowPulseCategory, ActionPhase) %>%
  summarize(across(Units.per.mL:LCEFA.per.mL, ~sum(.x, na.rm = TRUE))) %>%
  rename("Total.Units.per.mL" = "Units.per.mL") %>%
  rename("Total.BV.per.mL" = "BV.um3.per.mL") %>%
  rename("Total.Cells.per.mL" = "Cells.per.mL") %>%
  rename("Total.BM.per.mL" = "BM.ug.per.mL") %>%
  rename("Total.LCEFA.per.mL" = "LCEFA.per.mL") %>%
  ungroup

# Separate out data frames by density type and add zeros
# Biovolume Density + Group
df_phyto_FASTR_grp_BV <- df_phyto_FASTR_grp %>% 
  select(Year:Group,BV.um3.per.mL)

# Add zeros for taxa that weren't detected
temp <- pivot_wider(df_phyto_FASTR_grp_BV, 
                    names_from = Group,
                    values_from = BV.um3.per.mL,
                    values_fill = 0)

df_phyto_FASTR_grp_BV <- pivot_longer(temp,
                             cols = Cyanobacteria:last_col(),
                             names_to = "Group",
                             values_to = "BV.um3.per.mL")

# Create Biovolume-only data frame at genus level
df_phyto_FASTR_gen_BV <- df_phyto_FASTR_gen %>% 
  select(Year:Genus,BV.um3.per.mL)

temp <- pivot_wider(df_phyto_FASTR_gen_BV,
                    names_from = Genus,
                    values_from = BV.um3.per.mL,
                    values_fill = 0)

df_phyto_FASTR_gen_BV <- pivot_longer(temp,
                                      cols = Chlorella:last_col(),
                                      names_to = "Genus",
                                      values_to = "BV.um3.per.mL")

rm(temp)

```

## Create Condensed Data Frame for Most Abundant Genera
```{r condensed genera, echo=FALSE, message=FALSE}

## Calculate relative abundance of each genus within a group
df_phyto_FASTR_gen_RA <- df_phyto_FASTR_gen_BV %>%
  group_by(Genus) %>%
  summarize(Mean.BV.per.mL = mean(BV.um3.per.mL)) %>%
  mutate(MeanRelAbund = Mean.BV.per.mL/sum(Mean.BV.per.mL)) %>%
  ungroup()

# Highlight most abundant genera
df_phyto_FASTR_gen_RA <- df_phyto_FASTR_gen_RA %>%
  mutate(Type = case_when(MeanRelAbund > 0.05 ~ Genus,
                          TRUE ~ 'Other'))

# lump together all "other" taxa
df_phyto_FASTR_gen_RA_tot <- df_phyto_FASTR_gen_RA %>%
  group_by(Type) %>%
  summarize(MeanRelAbund = sum(MeanRelAbund)) %>%
  ungroup()

# Create cross-walk table for what taxa are lumped into the Other category
df_gen_types <- df_phyto_FASTR_gen_RA %>% select(Genus, Type)

# Add classification of "Named Genus" or "Other to main df
df_phyto_FASTR_gen_BV <- left_join(df_phyto_FASTR_gen_BV,
                                   df_gen_types,
                                   by = "Genus")

# Add higher-level taxonomy info
df_phyto_FASTR_gen_BV <- left_join(df_phyto_FASTR_gen_BV,
                                   df_taxa,
                                   by = "Genus")

# Reorganize columns
df_phyto_FASTR_gen_BV <- df_phyto_FASTR_gen_BV %>% 
  relocate(Group, .before = Genus) %>% 
  relocate(Type, .before = Genus)

# lump together all "other" taxa
df_phyto_FASTR_top_taxa_BV <- df_phyto_FASTR_gen_BV %>%
  group_by(Year, Month, Region, ActionPhase, DateTime, StationCode, Group, Type) %>%
  summarize(BV.um3.per.mL = sum(BV.um3.per.mL)) %>%
  ungroup()

```


## Multivariate Calculations 
Use the vegan package in R to calculate the Bray-Curtis similarity coefficients for phytoplankton community comparisons in each year (2014 - 2019). A good rule of thumb: stress < 0.05 provides an excellent representation in reduced dimensions, < 0.1 is great, < 0.2 is good/ok, and stress < 0.3 provides a poor representation.

```{r multivariate, echo=FALSE, message=FALSE}

set.seed(2390)

# Generate NMDS data with metaMDS by each year separately
years <- unique(df_phyto_FASTR_gen_BV$Year) 
years <- sort(years, decreasing = F, na.last = T)

# Create blank data frame to fill stresses in
stresses <- data.frame(Year = years, stress = NA)
ls_dfs <- list()

for (i in 1:length(years)) {
  genw <- pivot_wider(df_phyto_FASTR_gen_BV, 
                      names_from = "Genus", 
                      values_from = "BV.um3.per.mL",
                      values_fill = 0)
  
  genw <- genw %>%
    filter(Year == years[i])
  
  # Calculate the nMDS using vegan 

  phyto.NMDS <- metaMDS(
    comm = genw[c(7:139)],
    distance = "bray",
    k = 3,
    trymax = 500
    #trace = F,
    #autotransform = F
  )
  
  stresses$stress[which(stresses$Year == years[i])] <- phyto.NMDS$stress
  
  #look at Shepard plot which shows scatter around the regression between the interpoint distances 
  #in the final configuration (i.e., the distances between each pair of communities) against their 
  #original dissimilarities.
  stressplot(phyto.NMDS)
  
  # Using the scores function from vegan to extract the site scores and convert to a data.frame
  data.scores <- as_tibble(scores(phyto.NMDS, display = "sites"))
  
  # Combine metadata with NMDS data scores to plot in ggplot
  meta <- genw %>% select(1:6)
  meta <- cbind(meta, data.scores)
  
  # Read in years as a character otherwise it shows up as a number and gets displayed as a gradient
  meta$Year <- as.character(meta$Year)
  
  ls_dfs[[i]] <- meta
  
}


df_phyto_NMDS <- do.call(rbind, ls_dfs)

# Output stresses for us in plotting NMDS
write_csv(stresses, file = here("phyto-final","analyses","phyto_NMDS_stress.csv"))

```

## Save RData files for use in analysis 

```{r save RData, echo=FALSE, message=FALSE}

## Save data files
save(df_phyto, file = here("phyto-final","RData","df_phyto.RData"))
save(df_phyto_FASTR_sum, file = here("phyto-final","RData","df_phyto_FASTR_sum.RData"))
save(df_phyto_FASTR_gen, file = here("phyto-final","RData","df_phyto_FASTR_gen.RData"))
save(df_phyto_FASTR_gen_BV, file = here("phyto-final","RData","df_phyto_FASTR_gen_BV.RData"))
save(df_phyto_FASTR_top_taxa_BV, file = here("phyto-final","RData","df_phyto_FASTR_top_taxa_BV.RData"))
save(df_phyto_FASTR_grp, file = here("phyto-final","RData","df_phyto_FASTR_grp.RData"))
save(df_phyto_FASTR_grp_BV, file = here("phyto-final","RData","df_phyto_FASTR_grp_BV.RData"))
save(df_phyto_NMDS, file = here("phyto-final","RData","df_phyto_NMDS.RData"))

```

