---
title: "FASTR - Phytoplankton Analyses"
author: "Ted Flynn"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(knitr)))
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(vegan)))
suppressWarnings(suppressMessages(library(janitor)))
suppressWarnings(suppressMessages(library(here)))

# Set location to save plots
plots <- here("phyto-final","plots")

```

## Import data from Processing Scripts

```{r import data}

# Read in RData files processed in the data processing Rmd file
load(file = here("phyto-final","RData","df_phyto.RData"))
load(file = here("phyto-final","RData","df_phyto_FASTR_sum.RData"))
load(file = here("phyto-final","RData","df_phyto_FASTR_gen.RData"))
load(file = here("phyto-final","RData","df_phyto_FASTR_gen_BV.RData"))
load(file = here("phyto-final","RData","df_phyto_FASTR_grp.RData"))
load(file = here("phyto-final","RData","df_phyto_FASTR_grp_BV.RData"))
load(file = here("phyto-final","RData","df_phyto_NMDS.RData"))

```

## Create Plots for Paper
```{r Total Biovolume Plots, echo=FALSE, message=FALSE}

# Plot Upstream and Downstream Total BV by ActionPhase for each year
fig1 <- ggplot(data = df_phyto_FASTR_sum, 
               aes(y = log10(Total.BV.per.mL),
                   x = Region,
                   fill = ActionPhase)) +
  geom_boxplot() 

fig1 +
  labs(title = "Total Phytoplankton Abundance by Year",
       y = bquote(Log[10]~'Biovolume Density'~(um^3~mL^-1)), 
       x = "Sampling Region",
       fill = "Pulse Period") +
  facet_wrap(Year ~ ., ncol = 2) +
  scale_fill_manual(values = c("#4DAF4A",
                               "#984EA3",
                               "#A3D0D4"),
                    labels=c("Before", "During", "After"))

ggsave(path = plots,
       filename = "fig1_log_phyto_biovolume_by_year_and_AP.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=4,
       width=6, 
       dpi="print")

# Plot Total BV by Flow Pulse Type at Each Station

fig2 <- ggplot(data = df_phyto_FASTR_sum, 
               aes(y= log10(Total.BV.per.mL),
                   x = FlowPulseCategory,
                   color = ActionPhase)) +
  geom_boxplot()

fig2 +
  labs(title = "Total Phytoplankton Abundance by Flow Pulse Type",
       y = bquote(Log[10]~'Biovolume Density'~(um^3~mL^-1)), 
       x = "Flow Pulse Category") +
  facet_wrap(StationCode ~ ., ncol = 2)

ggsave(path = plots,
       filename = "fig2_log_phyto_biovolume_by_site_and_pulse_type.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=8,
       width=6.5, 
       dpi="print")

# Stacked Bar Charts for Biovolume ---------------------------------------------

# Upstream/Downstream
fig4 <- ggplot(df_phyto_FASTR_grp_BV, aes(x = ActionPhase,
                                       y = BV.um3.per.mL,
                                       fill = Group)) + 
  geom_bar(position = "stack",  
           width = 0.6, 
           stat = "summary", 
           fun = "mean") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0)) 

fig4 + 
  labs(x = NULL, 
       y = bquote('Average Biovolume Density'~(um^3~mL^-1)), 
       title = paste0("Abundance of Phytoplankton Groups During Flow Pulses")) + 
  theme(panel.background = element_rect(fill = "white", linetype = 0)) + 
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank()) +
  scale_fill_manual(values = c("#E41A1C",
                               "#377EB8",
                               "#4DAF4A",
                               "#984EA3",
                               "#FF7F00",
                               "#FFFF33",
                               "#A65628",
                               "#F781BF")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  facet_grid(Region~Year) # dir = v makes order of station go "north to south"

ggsave(path = plots,
       filename = paste0("fig4_phyto_group_biovolume_by_year.png"), 
       device = "png",
       scale=1.0, 
       units="in",
       height=5,
       width=7.5, 
       dpi="print")

```

## Plot Yearly Abundance of Phytoplankton

```{r plot yearly abundance}

## Create biovolume plots for each year
years <- unique(df_phyto_FASTR_grp_BV$Year) 
years <- sort(years, decreasing = F, na.last = T)

for (year in years) {
  df_temp <- df_phyto_FASTR_grp_BV %>%
    filter(Year == year)
  
  biovolume.plot <- ggplot(df_phyto_FASTR_grp_BV, 
                           aes(x = ActionPhase, 
                               y = BV.um3.per.mL, 
                               fill = Group)) +
    geom_bar(data = subset(df_phyto_FASTR_grp_BV, Year == year), 
             position = "stack",  
             width = 0.6, 
             stat = "summary", 
             fun = "mean") +
    theme(panel.background = element_rect(fill = "white", linetype = 0)) + 
    theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank()) +
    labs(x = "Pulse Period", 
         y = bquote('Average Biovolume Density'~(um^3~mL^-1)), 
         title = paste0("Abundance of Phytoplankton Groups During Flow Pulses - ",year)) 
  
  biovolume.plot +
    scale_fill_manual(values = c("#E41A1C",
                                 "#377EB8",
                                 "#4DAF4A",
                                 "#984EA3",
                                 "#FF7F00",
                                 "#FFFF33",
                                 "#A65628",
                                 "#F781BF")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.7)) +
    facet_wrap(StationCode ~ ., ncol = 5, dir = "h")
  
  ggsave(path = plots,
         filename = paste0("phyto_avg_BV_by_station_",year,".png"), 
         device = "png",
         scale=1.0, 
         units="in",
         height=3.5,
         width=7, 
         dpi="print")
  
  rm(df_temp)
  
}

```

## NMDS Plots 

```{r NMDS, echo=FALSE, message=FALSE}

# Plot NMDS Results

# Create a ggplot2 theme for the NMDS plots
theme_update(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.border = element_rect(fill = NA, colour = "black"),
    strip.background = element_rect(fill = "gray", colour = "black"),
    legend.position = "bottom",
    legend.key = element_rect(fill = "white", colour = NA)
    )

fig.NMDS.Reg <- ggplot(df_phyto_NMDS,
                       aes(x = NMDS1, y = NMDS2, color = Region)) +
  geom_point(size = 3) +
  stat_ellipse() + 
  labs(color = "Region",
       x = NULL,
       y = NULL)

fig.NMDS.Reg +
  facet_wrap(Year ~ ., ncol = 3, dir = "h") +
  scale_color_brewer(palette = "Set1")

ggsave(path = plots,
       filename = paste0("phyto_NMDS_biovolume_by_reg.png"), 
       device = "png",
       scale=1.0, 
       units="in",
       height=4.5,
       width=5.5, 
       dpi="print")

# Create plot colored by ActionPhase
fig.NMDS.AP <- ggplot(df_phyto_NMDS, 
                      aes(x = NMDS1, y = NMDS2, color = ActionPhase)) +
  geom_point(size = 3) +
  stat_ellipse() + 
  labs(color = "Flow Pulse Period",
       x = NULL,
       y = NULL)

fig.NMDS.AP +
  facet_wrap(Year ~ ., ncol = 3, dir = "h") +
  scale_color_brewer(palette = "Set1") 

ggsave(path = plots,
       filename = paste0("phyto_NMDS_biovolume_by_AP.png"), 
       device = "png",
       scale=1.0, 
       units="in",
       height=4.5,
       width=5.5, 
       dpi="print")

```

## Plot Abundance of Cyclotella
SIMPER analysis showed that the genus Cyclotella was a primary differentiator between the upstream and downstream regions. This code creates a boxplot to look at the differential abundance between each region. 

```{r cyclotella, echo=FALSE, message=FALSE}

# Compare Cyclotella Abundance Upstream and Downstream -------------------------

theme_set(theme_bw())

p_cyc <- ggplot(df_phyto_FASTR_gen_BV,
                aes(x = Region, 
                    y = log10(BV.um3.per.mL+1))) +
  geom_jitter(data = subset(df_phyto_FASTR_gen_BV, Genus == "Cyclotella"),
              width = 0.1) +
  geom_boxplot(data = subset(df_phyto_FASTR_gen_BV, Genus == "Cyclotella"),
              width = 0.1,
              outlier.shape = NA)

p_cyc +
  labs(y = bquote(Log[10]~'Biovolume Density'~(um^3~mL^-1)), 
       x = NULL) +
  facet_wrap(Year ~ ., ncol = 3)

ggsave(path = plots,
       filename = "cyclotella_abundance.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=3.5,
       width=6.5, 
       dpi="print")

```

## Plot abundance of LCEFA 

```{r LCEFA, echo=FALSE, message=FALSE}

# Separate out data frames by density type and add zeros
# Biovolume Density + Group
df_phyto_FASTR_grp_LCEFA <- df_phyto_FASTR_grp %>% 
  select(Year:Group,LCEFA.per.mL)

# Add zeros for taxa that weren't detected
temp <- pivot_wider(df_phyto_FASTR_grp_LCEFA, 
                    names_from = Group,
                    values_from = LCEFA.per.mL,
                    values_fill = 0)

#Remove taxa that aren't used for LCEFA calculations
temp <- temp %>% select(!c("Ciliates","Other","Golden Algae"))

df_phyto_FASTR_grp_LCEFA <- pivot_longer(temp,
                             cols = Cyanobacteria:last_col(),
                             names_to = "Group",
                             values_to = "LCEFA.per.mL")

# Create stacked bar chart for yearly LCEFA abundance by Action Phase
# and Upstream/Downstream
fig12 <- ggplot(df_phyto_FASTR_grp_LCEFA, aes(x = ActionPhase,
                                              y = LCEFA.per.mL,
                                              fill = Group)) + 
  geom_bar(position = "stack",  
           width = 0.6, 
           stat = "summary", 
           fun = "mean") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0)) 

fig12 + 
  labs(x = NULL, 
       y = bquote('Average LCEFA'~(mu*g~mL^-1)))
  theme(panel.background = element_rect(fill = "white", linetype = 0)) + 
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank()) +
  scale_fill_manual(values = c("#377EB8",
                               "#4DAF4A",
                               "#984EA3",
                               "#FF7F00",
                               "#A65628")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  facet_grid(Region~Year) # dir = v makes order of station go "north to south"

ggsave(path = plots,
       filename = paste0("fig12_phyto_group_LCEFA_by_year.png"), 
       device = "png",
       scale=1.0, 
       units="in",
       height=5,
       width=7.5, 
       dpi="print")

# Plot LCEFA at each station before, during and after flow pulses in 2016
p_LCEFA_by_station <- ggplot(df_phyto_FASTR_grp_LCEFA, 
                             aes(x = StationCode,
                                 y = LCEFA.per.mL,
                                 fill = Group)) + 
  geom_bar(data = subset(df_phyto_FASTR_grp_LCEFA, Year == "2016"),
           position = "stack",
           width = 0.6,
           stat = "summary",
           fun = "mean") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0)) 

p_LCEFA_by_station + 
  labs(x = NULL, 
       y = bquote('Average LCEFA'~(mu*g~mL^-1))) +
  theme(panel.background = element_rect(fill = "white", linetype = 0)) + 
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank()) +
  scale_fill_manual(values = c("#377EB8",
                               "#4DAF4A",
                               "#984EA3",
                               "#FF7F00",
                               "#A65628")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  facet_wrap(ActionPhase ~ ., ncol = 1, dir = "v") # dir = v makes order of station go "north to south"

ggsave(path = plots,
       filename = paste0("phyto_group_LCEFA_2016.png"), 
       device = "png",
       scale=1.0, 
       units="in",
       height=5,
       width=7.5, 
       dpi="print")

```


## Calculate ANOVAs

```{r anovas}

suppressWarnings(suppressMessages(library(ggpubr)))
suppressWarnings(suppressMessages(library(car)))
suppressWarnings(suppressMessages(library(rstatix)))
suppressWarnings(suppressMessages(library(visreg)))
suppressWarnings(suppressMessages(library(emmeans)))

## Change Year category to factor
df_phyto_FASTR_sum$Year <- as.factor(df_phyto_FASTR_sum$Year)
df_phyto_FASTR_sum$FlowPulseType <- as.factor(df_phyto_FASTR_sum$FlowPulseType)

## Create QQ Plot to check for normality
ggqqplot(log10(df_phyto_FASTR_sum$Total.BV.per.mL))

## View histogram to check for normality
hist(log10(df_phyto_FASTR_sum$Total.BV.per.mL))

## Run Shapiro-Wilks test to check whether data is normally distributed
shapiro.test(log10(df_phyto_FASTR_sum$Total.BV.per.mL))

## Run Levene test to compare variance
df_phyto_FASTR_sum %>% levene_test(log10(Total.BV.per.mL)~FlowPulseType) # p = 0.901
df_phyto_FASTR_sum %>% levene_test(log10(Total.BV.per.mL)~Year) # p = 0.980
df_phyto_FASTR_sum %>% levene_test(log10(Total.BV.per.mL)~Region) # p = 0.0230

## Run 3-way ANOVA
phyto.sum.aov <- df_phyto_FASTR_sum %>% anova_test(log10(Total.BV.per.mL) ~ Region*Year*ActionPhase)
phyto.sum.aov

## Rosie Code
L1 <- lm(log10(Total.BV.per.mL)~ Region*Year + Region*ActionPhase + ActionPhase*Year, 
         data = df_phyto_FASTR_sum)

summary(L1)

Anova(L1, type = 2)

visreg(L1)

visreg.AP.Region <- visreg(L1, xvar = "ActionPhase", by = "Year", gg = T)

visreg.AP.Region +
  facet_wrap(Year ~ ., ncol = 3, dir = "h")

ggsave(path = plots,
       filename = "visreg.AP.Region.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=3.5,
       width=7, 
       dpi="print")

visreg(L1, xvar = "Region", by = "Year")
visreg(L1, xvar = "Region", by = "ActionPhase")
visreg(L1, xvar = "ActionPhase", by = "Region")

emmeans(L1, pairwise ~ Region:ActionPhase)
emmeans(L1, pairwise ~ ActionPhase)
emmeans(L1, pairwise ~ Year:ActionPhase)

detach("package:ggpubr", unload=TRUE)
detach("package:car", unload=TRUE)
detach("package:rstatix", unload=TRUE)
detach("package:visreg", unload=TRUE)
detach("package:emmeans", unload=TRUE)

```

## PERMANOVA Comparisons

```{r PERMANOVA}

# Calculate PERMANOVA comparisons for phytoplankton communities by Region and Year
genw <- pivot_wider(df_phyto_FASTR_gen_BV, 
                    names_from = "Genus", 
                    values_from = "BV.um3.per.mL",
                    values_fill = 0)

adon.results <- adonis2(genw[c(7:139)] ~ genw$Region + genw$Year, 
                        strata = genw$StationCode,
                        method = "bray",
                        perm = 999)

adon.results <- adonis2(genw[c(7:139)] ~ genw$Region*genw$ActionPhase + genw$Year*genw$ActionPhase + genw$Region*genw$Year, 
                        strata = genw$StationCode,
                        method = "bray",
                        perm = 999)

dm_phyto_gen <- vegdist(genw[c(7:139)], method = "bray")

bd <- betadisper(dm_phyto_gen, genw$Region)

anova(bd)
permutest(bd)

```
## Create map of EMP stations for Supplemental Material

```{r, maps, echo=FALSE, message=FALSE}

suppressWarnings(suppressMessages(library(deltamapr)))
suppressWarnings(suppressMessages(library(sf)))
suppressWarnings(suppressMessages(library(ggrepel)))
suppressWarnings(suppressMessages(library(maps)))

# Read EMP Station location data (GPS coordinates)
df_EMP <- read_csv(here("phyto-final","CSVs","EMP_Stations_All.csv"))

df_EMP$Longitude <- df_EMP$Longitude * -1

# Read region data
df_regions <- read_csv(here("phyto-final","CSVs","station_regions.csv"))

# Rename header for StationCode & select phyto stations
df_EMP <- df_EMP %>% 
  rename("StationCode" = "Station ID") %>%
  filter(StationType == "Phytoplankton")

df_EMP <- df_EMP %>% select(StationCode:Longitude)

# Combine region data and phyto data
df_EMP <- left_join(df_EMP, df_regions)

# Select only stations in FASTR report
df_EMP <- df_EMP %>%
  filter(!Region %in% c("San Pablo Bay",
                        "Grizzly and Suisun Bay",
                        "Entrapment Zone"))

#Import data about city locations
CA <- map_data("world") %>% filter(subregion=="California")
cities <- world.cities %>% filter(country.etc=="USA")
cities$long <- cities$long * -1

## Plot EMP station maps
plot <- ggplot(WW_Delta) + 
  geom_sf(fill = "lightblue") + 
  geom_point(data = df_EMP,
             aes(x = Longitude,
                 y = Latitude,
                 fill = Region,
                 size = 3),
             pch = 21,
             color = "black") +
  geom_point(data = cities %>% arrange(pop) %>% tail(250),
             aes(x = long,
                 y = lat)) +
  geom_text_repel(data = cities %>% arrange(pop) %>% tail(250), 
                  aes(x = long,
                      y = lat, 
                      label = name)) +
  ylim(37.65, 38.4) +
  xlim(122.0, 121.2) +
  scale_fill_brewer(palette = "Dark2") +
  theme_bw()

plot + 
  labs(x = NULL,
       y = NULL,
       fill = "Region") +
  guides(size = "none") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.25, 0.16))

ggsave(path=plots,
       filename = "map-EMP-phyto-stations.png", 
       device = "png",
       scale=1.0, 
       units="in",
       height=5.5,
       width=6.5, 
       dpi="print")

```

