---
title: "NDFA Continuous WQ Data: QAQC and Integration of year-round data"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document provides the code and decisions made to perform final QA steps on the additional continuous water quality data collected outside of the period of interest of the North Delta Flow Action synthesis project. We only checked and validated this additional data for the four stations with year-round data available: LIB, LIS, RVB, and STTD (2016-2019). The "rtm_QA_and_combine_data_filt_dates.Rmd" file contains the code used to QA and validate the data within the period of interest. 

The data were collected by multiple groups including USGS, the WQES section within the NCRO office at DWR, and the EMP section within DES at DWR. The final product of this document is a single integrated csv file containing the continuous water quality data for the four stations with year-round data. 

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(magrittr)
```

```{r load functions, message = FALSE, warning = FALSE}
# Source global NDFA functions
source("global_ndfa_funcs.R")
source("Water_Quality/global_wq_funcs.R")

# Source continuous WQ QA processing functions
source("Water_Quality/Continuous_WQ/Data_Cleaning/rtm_QA_data_funcs.R")
```

```{r define file paths}
# Define relative file path for processed continuous WQ data files saved on NDFA SharePoint - year-round data
fp_rel_wq_proc_all <- "WQ_Subteam/Processed_Data/Continuous/All_Dates"

# Define relative file path for file containing all QA'ed and cleaned continuous WQ data for the NDFA period of interest
fp_rel_rtm_data_qa <- "WQ_Subteam/Processed_Data/Continuous/RTM_INPUT_all_2021-04-20.csv"

# Define relative file path for the file containing the final QA'ed and integrated continuous WQ data to be saved on NDFA SharePoint
fp_rel_wq_qa_f <- "WQ_Subteam/Processed_Data/Continuous"

# Define absolute file paths
fp_abs_wq_proc_all <- ndfa_abs_sp_path(fp_rel_wq_proc_all)
fp_abs_rtm_data_qa <- ndfa_abs_sp_path(fp_rel_rtm_data_qa)
fp_abs_wq_qa_f <- ndfa_abs_sp_path(fp_rel_wq_qa_f)
```

```{r set system tz as PST}
# Set System Timezone as "Etc/GMT+8" (PST) to make it consistent with all data 
Sys.setenv(TZ = "Etc/GMT+8")
```

# Import and Prepare Data

```{r import data}
# Import QA'ed and cleaned continuous WQ data for the NDFA period of interest
rtm_wq_qa_orig <- import_rtm_data(fp_abs_rtm_data_qa, 10)

# Create a vector of all relevant file paths for the processed continuous WQ data - year-round data
rtm_wq_proc_fp <- dir(fp_abs_wq_proc_all, pattern = "_LIB_|_LIS_|_RVB_|_STTD_", full.names = TRUE)

# Create a vector of the number of parameters within each station's dataset
n_params <- c(11, 7, 8, 6)

# Import processed continuous WQ data into a dataframe
rtm_wq_proc_orig <- map2_dfr(rtm_wq_proc_fp, n_params, .f = import_rtm_data)
```

We need to prepare both the raw year-round data and the QA'ed data so that they can be combined only adding the raw year-round data outside of the NDFA period of interest.

```{r prepare raw data yr round, message = FALSE}
# Prepare raw year-round data
rtm_wq_proc_clean <- rtm_wq_proc_orig %>% 
  # only keep a subset of the parameters (Chla, DO, pH, SpCnd, Turbidity, WaterTemp)
  select(
    DateTime, 
    StationCode, 
    starts_with(c("Chla", "DO", "pH", "SpCnd", "Turbidity", "WaterTemp")) & !contains("RFU")
  ) %>% 
  # parse date-time variable and define tz as PST; add date and year variables
  mutate(
    DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8"),
    Date = date(DateTime),
    Year = year(DateTime)
  ) %>% 
  # add flow action periods to the data frame but keep all data
  ndfa_action_periods(na_action_remove = FALSE) %>% 
  # only keep data outside of the NDFA period of interest
  filter(is.na(FlowActionPeriod)) %>% 
  # remove unnecessary variables
  select(-c(Date, Year, FlowActionPeriod))
```

```{r prepare qa data}
# Prepare QA'ed and cleaned continuous WQ data to be combined to the raw year-round data
rtm_wq_qa_clean <- rtm_wq_qa_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # only include LIB, LIS, RVB, and STTD
  filter(StationCode %in% c("LIB", "LIS", "RVB", "STTD")) %>% 
  # only keep a subset of the parameters (Chla, DO, pH, SpCnd, Turbidity, WaterTemp)
  select(
    DateTime, 
    StationCode, 
    starts_with(c("Chla", "DO", "pH", "SpCnd", "Turbidity", "WaterTemp"))
  )
```

The data is now ready to be combined.

```{r combine data}
# Combine the raw the QA'ed and cleaned data
rtm_wq_all <- bind_rows(rtm_wq_proc_clean, rtm_wq_qa_clean)
```

To be able to work on the data for each station separately and in an organized way, we will remove the rows where all of the data values are missing and nest the data by `StationCode`.

```{r prepare data for indiv qa}
rtm_wq_all_nest <- rtm_wq_all %>% 
  filter(!if_all(!c(DateTime, StationCode), is.na)) %>% 
  arrange(StationCode, DateTime) %>% 
  nest(df = -StationCode)
```

# QA Processing

Our process to QA the continuous water quality data includes flagging and excluding suspect data for individual stations and parameters through visual inspection of the plots of raw data and through information from the field and sonde verification records.

## LIB

```{r pull out lib data}
# Pull out data for LIB
lib_orig <- pull_df_sta(rtm_wq_all_nest, "LIB")
```

Specific conductance suddenly dropped to values close to zero on the morning of 12/28/2017. We will flag and exclude all water quality data (Chla, DO, pH, SpCnd, Turbidity, and WaterTemp) collected during this time period since the sonde may have been out of the water during this time.

```{r plot lib chla do ph spcnd turb wt dec 2017 before}
# Plot LIB Chla, DO, pH, SpCnd, Turbidity, and WaterTemp data for Dec 2017 - raw data
lib_orig %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(cols = where(is.numeric), names_to = "parameter", values_to = "value") %>% 
  filter(
    year(DateTime) == 2017, 
    month(DateTime) == 12
  ) %>% 
  ggplot(aes(DateTime, value)) +
    geom_line() +
    rtm_xaxis_scale +
    facet_grid(rows = vars(parameter), scales = "free_y")
```

```{r rm lib chla do ph spcnd turb wt 2017, message = FALSE}
# Flag and exclude Chla, DO, pH, SpCnd, Turbidity, and WaterTemp values in Dec 2017
lib_clean_v1 <- lib_orig %>% 
  mutate(
    across(
      where(is.numeric),
      ~replace(
        .x, 
        DateTime >= "2017-12-28 9:45:00" & DateTime <= "2017-12-28 12:30:00", 
        NA_real_
      )
    )
  ) %>% 
  mutate(
    across(
      ends_with("Qual"),
      ~replace(
        .x, 
        DateTime >= "2017-12-28 9:45:00" & DateTime <= "2017-12-28 12:30:00", 
        "unreliable"
      )
    )
  )
```

The Chla, DO, pH, SpCnd, Turbidity, and WaterTemp values are excluded from the data set now:
```{r plot lib chla do ph spcnd turb wt dec 2017 after}
# Plot LIB Chla, DO, pH, SpCnd, Turbidity, and WaterTemp data for Dec 2017 - after exclusion
lib_clean_v1 %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(cols = where(is.numeric), names_to = "parameter", values_to = "value") %>% 
  filter(
    year(DateTime) == 2017, 
    month(DateTime) == 12
  ) %>% 
  ggplot(aes(DateTime, value)) +
    geom_line() +
    rtm_xaxis_scale +
    facet_grid(rows = vars(parameter), scales = "free_y")
```

pH suddenly dropped in the morning of 1/11/2018. We will flag and exclude all available water quality data (Chla, DO, pH, and WaterTemp) collected during this time period since the sonde may have been out of the water for maintenance during this time.

```{r plot lib chla do ph wt jan 2018 before}
# Plot LIB Chla, DO, pH, and WaterTemp data for Jan 2018 - raw data
lib_clean_v1 %>% 
  select(!(c(SpCnd, Turbidity) | ends_with("_Qual"))) %>% 
  pivot_longer(cols = where(is.numeric), names_to = "parameter", values_to = "value") %>% 
  filter(
    year(DateTime) == 2018,
    month(DateTime) == 1
  ) %>% 
  ggplot(aes(DateTime, value)) +
    geom_line() +
    rtm_xaxis_scale +
    facet_grid(rows = vars(parameter), scales = "free_y")
```

```{r rm lib chla do ph wt 2018}
# Flag and exclude Chla, DO, pH, and WaterTemp values in Jan 2018
lib_clean_v2 <- lib_clean_v1 %>% 
  mutate(
    across(
      c("Chla", "DO", "pH", "WaterTemp"),
      ~replace(.x, DateTime == "2018-01-11 10:00:00", NA_real_)
    )
  ) %>% 
  mutate(
    across(
      c("Chla_Qual", "DO_Qual", "pH_Qual", "WaterTemp_Qual"),
      ~replace(.x, DateTime == "2018-01-11 10:00:00", "unreliable")
    )
  )
```

The Chla, DO, pH, and WaterTemp values are excluded from the data set now:
```{r plot lib chla oct 2013 after}
# Plot LIB Chla, DO, pH, and WaterTemp data for Jan 2018 - after exclusion
lib_clean_v2 %>% 
  select(!(c(SpCnd, Turbidity) | ends_with("_Qual"))) %>% 
  pivot_longer(cols = where(is.numeric), names_to = "parameter", values_to = "value") %>% 
  filter(
    year(DateTime) == 2018,
    month(DateTime) == 1
  ) %>% 
  ggplot(aes(DateTime, value)) +
    geom_line() +
    rtm_xaxis_scale +
    facet_grid(rows = vars(parameter), scales = "free_y")
```

The turbidity data is noisy and suspect at LIB on 12/20/2013 directly after a gap in the data. We will flag and exclude this data.

```{r plot lib turb dec 2013 before}
# Plot LIB Turbidity data for Dec 2013 - raw data
lib_clean_v2 %>% 
  filter(
    year(DateTime) == 2013,
    month(DateTime) == 12
  ) %>% 
  plot_indiv_param(Turbidity)
```

```{r rm lib turb 2013}
# Flag and exclude Turbidity values in Dec 2013
lib_clean_v3 <- lib_clean_v2 %>% 
  rm_flag_val_range(Turbidity, "2013-12-20 00:00:00", "2013-12-20 03:30:00")
```

The Turbidity data is excluded from the data set now:
```{r plot lib turb dec 2013 after}
# Plot LIB Turbidity data for Dec 2013 - after exclusion
lib_clean_v3 %>% 
  filter(
    year(DateTime) == 2013,
    month(DateTime) == 12
  ) %>% 
  plot_indiv_param(Turbidity)
```

There are a few issues with the Chlorophyll data for LIB:

* There is some suspicious data directly before a gap in the data in Nov 2013
* There are a few extreme outliers in Dec 2017 and Jan 2018
* The data is very noisy and suspect from 5/8/2018 to 7/12/2018
* There are some values that are equal to zero in Nov 2018 - Jan 2019

This data will be flagged and excluded from the LIB data set.

```{r plot lib chla nov dec 2013 before}
# Plot LIB chlorophyll data for Nov and Dec 2013 - raw data
lib_clean_v3 %>% 
  filter(
    year(DateTime) == 2013,
    month(DateTime) >= 11 & month(DateTime) <= 12
  ) %>% 
  plot_indiv_param(Chla)
```

```{r plot lib chla dec 2017 jan 2018 before}
# Plot LIB chlorophyll data for Dec 2017-Jan 2018 - raw data
lib_clean_v3 %>% 
  filter(date(DateTime) >= "2017-12-01" & date(DateTime) <= "2018-01-31") %>% 
  plot_indiv_param(Chla)
```

```{r plot lib chla may aug 2018 before}
# Plot LIB chlorophyll data for May-Aug 2018 - raw data
lib_clean_v3 %>% 
  filter(
    year(DateTime) == 2018,
    month(DateTime) >= 5 & month(DateTime) <= 8
  ) %>% 
  plot_indiv_param(Chla)
```

```{r plot lib chla nov 2018 jan 2019 before}
# Plot LIB chlorophyll data for Nov 2018-Jan 2019 - raw data
lib_clean_v3 %>% 
  filter(date(DateTime) >= "2018-11-01" & date(DateTime) <= "2019-01-31") %>% 
  plot_indiv_param(Chla)
```

```{r rm lib chla 2013 2017 2018 and 2019}
# Flag and exclude chlorophyll values in 2013, 2017, 2018, and 2019
lib_clean_v4 <- lib_clean_v3 %>% 
  rm_flag_val_range(Chla, "2013-11-21 14:30:00", "2013-11-22 09:00:00") %>% 
  rm_flag_val_range(Chla, "2017-12-07 04:30:00", "2017-12-07 07:15:00") %>% 
  rm_flag_val_one(Chla, "2018-01-05 03:15:00") %>% 
  rm_flag_val_one(Chla, "2018-01-05 20:30:00") %>%
  rm_flag_val_range(Chla, "2018-05-08 11:00:00", "2018-07-12 23:45:00")

# Flag and exclude chlorophyll values equal to zero in Nov 2018-Jan 2019
lib_clean_v5 <- lib_clean_v4 %>%
  mutate(
    Chla_Qual = replace(Chla_Qual, Chla == 0, "unreliable"),
    Chla = replace(Chla, Chla == 0, NA_real_)
  )
```

The chlorophyll data is excluded from the data set now:
```{r plot lib chla nov dec 2013 after}
# Plot LIB chlorophyll data for Nov and Dec 2013 - after exclusion
lib_clean_v5 %>% 
  filter(
    year(DateTime) == 2013,
    month(DateTime) >= 11 & month(DateTime) <= 12
  ) %>% 
  plot_indiv_param(Chla)
```

```{r plot lib chla dec 2017 jan 2018 after}
# Plot LIB chlorophyll data for Dec 2017-Jan 2018 - after exclusion
lib_clean_v5 %>% 
  filter(date(DateTime) >= "2017-12-01" & date(DateTime) <= "2018-01-31") %>% 
  plot_indiv_param(Chla)
```

```{r plot lib chla may aug 2018 after}
# Plot LIB chlorophyll data for May-Aug 2018 - after exclusion
lib_clean_v5 %>% 
  filter(
    year(DateTime) == 2018,
    month(DateTime) >= 5 & month(DateTime) <= 8
  ) %>% 
  plot_indiv_param(Chla)
```

```{r plot lib chla nov 2018 jan 2019 after}
# Plot LIB chlorophyll data for Nov 2018-Jan 2019 - after exclusion
lib_clean_v5 %>% 
  filter(date(DateTime) >= "2018-11-01" & date(DateTime) <= "2019-01-31") %>% 
  plot_indiv_param(Chla)
```


## LIS

```{r pull out lis data}
# Pull out data for LIS
lis_orig <- pull_df_sta(rtm_wq_all_nest, "LIS")
```

The chlorophyll data is very noisy and suspect at LIS from 1/18/2019 to 1/22/2019. We will flag and exclude this data.

```{r plot lis chla jan 2019 before}
# Plot LIS chlorophyll data for Jan 2019 - raw data
lis_orig %>% 
  filter(
    year(DateTime) == 2019,
    month(DateTime) == 1
  ) %>% 
  plot_indiv_param(Chla)
```

```{r rm lis chla 2019}
# Flag and exclude chlorophyll values in 2019
lis_clean_v1 <- lis_orig %>% 
  rm_flag_val_range(Chla, "2019-01-18 07:15:00", "2019-01-22 03:30:00")
```

The chlorophyll data is excluded from the data set now:
```{r plot lis chla jan 2019 after}
# Plot LIS chlorophyll data for Jan 2019 - after exclusion
lis_clean_v1 %>% 
  filter(
    year(DateTime) == 2019,
    month(DateTime) == 1
  ) %>% 
  plot_indiv_param(Chla)
```


## STTD

```{r pull out sttd data}
# Pull out data for STTD and only include 2016-2019
sttd_orig <- 
  pull_df_sta(rtm_wq_all_nest, "STTD") %>% 
  filter(year(DateTime) >= 2016)
```

There is an extremely low DO outlier collected at STTD on 3/3/2019. We will flag and exclude this data.

```{r plot sttd do march 2019 before}
# Plot STTD DO data for March 2019 - raw data
sttd_orig %>% 
  filter(
    year(DateTime) == 2019,
    month(DateTime) == 3
  ) %>% 
  plot_indiv_param(DO)
```

```{r rm sttd do 2019}
# Flag and exclude DO value in 2019
sttd_clean_v1 <- sttd_orig %>% rm_flag_val_one(DO, "2019-03-03 23:45:00")
```

The DO value is excluded from the data set now:
```{r plot sttd do march 2019 after}
# Plot STTD DO data for March 2019 - after exclusion
sttd_clean_v1 %>% 
  filter(
    year(DateTime) == 2019,
    month(DateTime) == 3
  ) %>% 
  plot_indiv_param(DO)
```

There are a few suspicious pH values collected at STTD at the edges of gaps in the data on 12/12/2017 and 3/20/2018. We will flag and exclude this data.

```{r plot sttd ph nov dec 2017 before}
# Plot STTD pH data for Nov-Dec 2017 - raw data
sttd_clean_v1 %>% 
  filter(
    year(DateTime) == 2017,
    month(DateTime) >= 11 & month(DateTime) <= 12
  ) %>% 
  plot_indiv_param(pH)
```

```{r plot sttd ph march 2018 before}
# Plot STTD pH data for March 2018 - raw data
sttd_clean_v1 %>% 
  filter(
    year(DateTime) == 2018,
    month(DateTime) == 3
  ) %>% 
  plot_indiv_param(pH)
```

```{r rm sttd pH 2017 2018}
# Flag and exclude pH values in 2017 and 2018
sttd_clean_v2 <- sttd_clean_v1 %>% 
  rm_flag_val_one(pH, "2017-12-12 14:15:00") %>% 
  rm_flag_val_one(pH, "2018-03-20 08:45:00")
```

The DO value is excluded from the data set now:
```{r plot sttd ph nov dec 2017 after}
# Plot STTD pH data for Nov-Dec 2017 - after exclusion
sttd_clean_v2 %>% 
  filter(
    year(DateTime) == 2017,
    month(DateTime) >= 11 & month(DateTime) <= 12
  ) %>% 
  plot_indiv_param(pH)
```

```{r plot sttd ph march 2018 after}
# Plot STTD pH data for March 2018 - after exclusion
sttd_clean_v2 %>% 
  filter(
    year(DateTime) == 2018,
    month(DateTime) == 3
  ) %>% 
  plot_indiv_param(pH)
```

There are some extremely low and noisy turbidity values collected at STTD in January 2018 most likely due to sensor interference or malfunction. We will flag and exclude this data.

```{r plot sttd turb dec 2017 feb 2018 before}
# Plot STTD Turbidity data for Dec 2017-Feb 2018 - raw data
sttd_clean_v2 %>% 
  filter(date(DateTime) >= "2017-12-01" & date(DateTime) <= "2018-02-28") %>% 
  plot_indiv_param(Turbidity)
```

```{r rm sttd turb 2018}
# Flag and exclude Turbidity values in 2018
sttd_clean_v3 <- sttd_clean_v2 %>% 
  rm_flag_val_range(Turbidity, "2018-01-13 16:45:00", "2018-01-30 10:15:00")
```

The Turbidity values are excluded from the data set now:
```{r plot sttd turb dec 2017 feb 2018 after}
# Plot STTD Turbidity data for Dec 2017-Feb 2018 - after exclusion
sttd_clean_v3 %>% 
  filter(date(DateTime) >= "2017-12-01" & date(DateTime) <= "2018-02-28") %>% 
  plot_indiv_param(Turbidity)
```

There are a few issues with the Chlorophyll data for STTD:

* There are some suspect values in Dec 2016 just before a large gap in the data
* There is one value close to zero on 1/6/2017 after a large data gap
* There are some extremely low and noisy values in Jan 2018 

This data will be flagged and excluded from the STTD data set.

```{r plot sttd chla dec 2016 jan 2017 before}
# Plot STTD chlorophyll data for Dec 2016-Jan 2017 - raw data
sttd_clean_v3 %>% 
  filter(date(DateTime) >= "2016-12-01" & date(DateTime) <= "2017-01-31") %>% 
  plot_indiv_param(Chla)
```

```{r plot sttd chla dec 2017 feb 2018 before}
# Plot STTD chlorophyll data for Dec 2017-Feb 2018 - raw data
sttd_clean_v3 %>% 
  filter(date(DateTime) >= "2017-12-01" & date(DateTime) <= "2018-02-28") %>% 
  plot_indiv_param(Chla)
```

```{r rm sttd chla 2016 2017 2018}
# Flag and exclude chlorophyll values in 2016, 2017, and 2018
sttd_clean_v4 <- sttd_clean_v3 %>% 
  rm_flag_val_range(Chla, "2016-12-18 17:45:00", "2016-12-19 00:45:00") %>% 
  rm_flag_val_one(Chla, "2017-01-06 12:45:00") %>% 
  rm_flag_val_range(Chla, "2018-01-12 13:15:00", "2018-01-30 10:15:00")
```

The chlorophyll values are excluded from the data set now:
```{r plot sttd chla dec 2016 jan 2017 after}
# Plot STTD chlorophyll data for Dec 2016-Jan 2017 - after exclusion
sttd_clean_v4 %>% 
  filter(date(DateTime) >= "2016-12-01" & date(DateTime) <= "2017-01-31") %>% 
  plot_indiv_param(Chla)
```

```{r plot sttd chla dec 2017 feb 2018 after}
# Plot STTD chlorophyll data for Dec 2017-Feb 2018 - after exclusion
sttd_clean_v4 %>% 
  filter(date(DateTime) >= "2017-12-01" & date(DateTime) <= "2018-02-28") %>% 
  plot_indiv_param(Chla)
```


## Combine cleaned data

Combine individually cleaned data for stations back together.

```{r combine indiv clean data}
mod_stations <- c("LIB", "LIS", "STTD")

# Create a nested data frame of all cleaned data sets for the individual stations
indiv_clean_rtm <- 
  tibble(
    StationCode = mod_stations,
    df = list(lib_clean_v5, lis_clean_v1, sttd_clean_v4)
  )

# Combine all cleaned data
rtm_wq_all_clean <- rtm_wq_all_nest %>% 
  filter(!StationCode %in% mod_stations) %>% 
  bind_rows(indiv_clean_rtm) %>% 
  unnest(df) %>% 
  # sort by StationCode, DateTime
  arrange(StationCode, DateTime) %>% 
  # apply a consistent variable order 
  select(
    StationCode,
    DateTime,
    starts_with("Wa"),
    starts_with("Tu"),
    starts_with("Sp"),
    starts_with("DO"),
    starts_with("pH"),
    starts_with("Chla")
  )
```


# Export data

Export QA'ed data to the Processed_Data/Continuous folder as a single integrated .csv file containing the continuous water quality data for the four stations with year-round data.

```{r export all qa data as csv, eval = FALSE}
# Code chunk is set to eval = FALSE, so this code is not executed when this file is knitted
# change eval option to TRUE when you want to export the data when knitting this file
rtm_wq_all_clean %>% 
  write_excel_csv(file.path(fp_abs_wq_qa_f, "RTM_INPUT_yr_round_2021-09-02.csv"), na = "")
```

End of script

