---
title: "FISH_MAN_Analysis_Models_Tow"
author: "Catarina Pien"
date: "12/15/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Code for FASTR fish models for tow data
Fish ~ Region + Action Phase + Year

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r load, echo = FALSE, results = FALSE, warning = FALSE, message = FALSE}

rm(list=ls(all=TRUE))

library(tidyverse)
library(lubridate) # Datetime
library(pscl) # Run zips and zinbs
library(visreg) # Visualize models
library(viridis) # Color palette
library(emmeans) # post hoc
library(MASS) # negbin modelv
library(countreg) # zeroinfl and hurdle functions, qqrplot
library(gridExtra)
library(grid)
library(dplyr)
library(Polychrome)
library(plotly)
library(lme4)
```

## Import data
* Fish_ndfa: integrated fish dataset
* FlowDesignation: Assigns action phases by date
* Regions: Assigns regions to fish stations
* Merge all together
* Define variable structures
* Order action phases
```{r}
#setwd("C:/Users/cpien/OneDrive - California Department of Water Resources/Work/FASTR/FASTR/ND-FASTR")
Fish_ndfa_0 <- read.csv("fish data/FISH_MAN_allIEPsurveys_20201030.csv")
Fish_ndfa <- dplyr::select(Fish_ndfa_0, -c(Latitude, Longitude, DO, Turbidity))
FlowDesignation <- read.csv("fish data/FlowDatesDesignations_45days.csv")
Regions <- read.csv("fish data/Stations_Fish_NDFA_2021-01-13.csv")
# Regions_min <- Regions %>%
#   rename(Survey = Survey.x) %>%
#   dplyr::select(-c(Intersect, geometry, RegionMult))

# Look at data
head(Fish_ndfa)
str(Fish_ndfa)
```


Modify the EDSM Stations - did not end up including this 
```{r}
library(nngeo)
library(sp)

setwd("C:/Users/cpien/OneDrive - California Department of Water Resources/Work/FASTR/FASTR/ND-FASTR")
#Regions2 <- Regions %>% 
#  mutate(StationType = ifelse(Survey.x == "EDSM", "Random", "Fixed"))
#
#EDSM <- filter(Regions2, Survey.x == "EDSM")
#StationRest <- filter(Regions, Survey.x != "EDSM")
#
## Change to shapefile - not sure if this is necessary but couldn't figure out otherwise how to get it to work
# coordinates(EDSM) = ~Longitude+Latitude
# proj4string(EDSM) = CRS("+proj=longlat +datum=WGS84")
# raster::shapefile(EDSM, "fish_code/R_write/EDSMShapefile.shp")
# 
# coordinates(StationRest) = ~Longitude+Latitude
# proj4string(StationRest) = CRS("+proj=longlat +datum=WGS84")
# raster::shapefile(StationRest, "fish_code/R_write/StationRestShapefile.shp")

EDSM_s <- st_read("fish_code/R_write/EDSMShapefile.shp")
StationRest_s <- st_read("fish_code/R_write/StationRestShapefile.shp")

# Run nearest neighbor
NN <- st_nn(EDSM_s, StationRest_s, progress = FALSE)

# Convert list into a data frame
df <- do.call(rbind.data.frame, NN)
colnames(df) <-  "Row"

# Extract the rest of the dataset based on the row number
result <- StationRest_s[df$Row,]

# Merge dataframes
EDSM_nn <- cbind(EDSM_s, result) %>%
  dplyr::select(c(2, 9)) %>%
  rename(Neighbor = StatnCd.1,
         StationCode = StatnCd)
Regions3 <- left_join(Regions, EDSM_nn, by = "StationCode") %>%
  mutate(StationCodeF = ifelse(is.na(Neighbor), StationCode, Neighbor))
Regions_min2 <- Regions3 %>%
  rename(Survey = Survey.x) %>%
  dplyr::select(-c(Intersect, geometry.x, geometry.y, RegionMult, Neighbor))

summary(Regions_min2)
```


```{r import and filter dates, warning = FALSE, message = FALSE, results = FALSE}
# Add variables
Fish_ndfa$Date <- ymd(Fish_ndfa$Date)
Fish_ndfa$Month <- month(Fish_ndfa$Date)
Fish_ndfa$Day <- day(Fish_ndfa$Date)
FlowDesignation$PreFlowStart <- mdy(FlowDesignation$PreFlowStart)
FlowDesignation$PreFlowEnd <- mdy(FlowDesignation$PreFlowEnd)
FlowDesignation$PostFlowStart <- mdy(FlowDesignation$PostFlowStart)
FlowDesignation$PostFlowEnd <- mdy(FlowDesignation$PostFlowEnd)

# Merge data from FlowDesignation Table (Water Year Type, Flow days and type)
# Filter only Pre-During-Post Flow Action Data. 
# We have a during action, and then Pre = 30 days before/Post = 30 days after
Fish_all0 <- inner_join(Fish_ndfa,FlowDesignation, by = "Year")
Fish_allT1 <- left_join(Fish_all0, Regions_min2, by = c("StationCode", "Survey"))

# anti_join(Fish_all1, Fish_all0, by = colnames(Fish_all0))

Fish_allT2 <- Fish_allT1 %>%
   mutate(ActionPhase = ifelse(Date > PreFlowStart & Date<PreFlowEnd, "Pre", NA)) %>%
   mutate(ActionPhase = replace(ActionPhase, Date > PreFlowEnd & Date < PostFlowStart, "During")) %>% 
   mutate(ActionPhase = replace(ActionPhase, Date > PostFlowStart & Date < PostFlowEnd, "Post")) %>%
  filter(!is.na(ActionPhase)) %>%
   dplyr::select(-c(PreFlowStart:PostFlowEnd)) %>%
   arrange(Date, Survey, StationCode, CommonName)

# Define variable structures
Fish_allT2$Date <- ymd(Fish_allT2$Date)
Fish_allT2$fYear <- as.factor(Fish_allT2$Year)
Fish_allT2$WYType <- as.factor(Fish_allT2$WYType)
Fish_allT2$FlowPulseType <- as.factor(Fish_allT2$FlowPulseType)
Fish_allT2$ActionPhase <- as.factor(Fish_allT2$ActionPhase)
Fish_allT2$X1 <- NULL
Fish_allT2$Survey <- as.factor(Fish_allT2$Survey)
Fish_allT2$StationCode <- as.factor(Fish_allT2$StationCode)
Fish_allT2$MethodCode <- as.factor(FFish_allT2ethodCode)

# Renaming
Fish_allT <- Fish_allT2 %>%
  dplyr::rename(PulsePhase = ActionPhase,
                StationCodeInitial = StationCode,
                StationCode = StationCodeF) %>%
  mutate(CommonName = replace(CommonName, CommonName=="Inland Silverside", "Mississippi Silverside")) %>%
  mutate(CommonName = replace(CommonName, CommonName == "Tridentiger spp_", "Tridentiger spp."))


# Order of factors, renaming
Fish_allT$Region <- factor(Fish_allT$Region, levels = c("Upstream", "Downstream", "MiddleSacRiver"))
Fish_all$ActionPhase <-  factor(Fish_allT$PulsePhase, levels = c("Pre", "During", "Post"))
levels(Fish_allT$PulsePhase) = c("Before", "During", "After")
str(Fish_allT)



```


## Data Prep
* Filter to just tow data and year starting 2013
* Fill in zeros for unlisted species using complete cases
* Calculate CPUE 
* Organize data

```{r tow, warning = FALSE, message = FALSE}
# Filter Dataset
Tow <- Fish_allT %>% 
  filter(Survey %in% c("EDSM", "Townet", "FMWT")) %>%
  filter(Year>2012) %>%
  arrange(Survey, Date, StationCode, Tow, Depth)

### Complete Cases
# For each Date, Survey, StationCode combination, make sure each fish species is represented with either positive count or zero. 
Tow_completecase <- Tow %>%
  group_by(Date, Survey, StationCode, Tow, Depth, CommonName) %>%
  summarize(sum.count = sum(totalCount)) %>%
  ungroup() %>%
  complete(CommonName, nesting(Date, Survey, StationCode, Tow, Depth), fill = list(sum.count = 0)) %>%
  arrange(Date, Survey, StationCode,CommonName)

### Merge back together with rest of data

# Get distinct samples for looking at Water Quality 
Tow_samples <- Tow %>% dplyr::select(-c(CommonName, totalCount)) %>% distinct()

# Merge 
Tow_complete <- left_join(Tow_completecase, Tow_samples, by = c("Date", "Survey", "StationCode", "Tow", "Depth"))

### Calculate CPUE
# Remove samples with no volume 
Tow_CPUE <- Tow_complete %>%
  filter(!is.na(VolumeSampled))%>%
  mutate(CPUE = sum.count/VolumeSampled)

### Rearrange columns 
Tow_f <- Tow_CPUE[, c("Date", "Year", "fYear", "Month", "Day", "Survey", "StationCode", "Latitude", "Longitude","Region","MethodCode", "WYType", "FlowPulseType", "NetFlowDays","PulsePhase", "Secchi",  "Conductivity", "WaterTemp",  "Tow", "Depth", "VolumeSampled",  "CommonName", "sum.count", "CPUE")]
```


```{r}
### Mean CPUE 
## Calculate means for each species by year-location-PulsePhase
CPUE_means_Tow_phase <- Tow_f %>%
  group_by(Year, Survey, StationCode, Tow, PulsePhase, CommonName) %>% 
  summarize(mean.CPUE = mean(CPUE)) 
```

## Filter for species sets

```{r Tow filters, warning = FALSE, message = FALSE}
# Lists
list_native <- c("Sacramento Pikeminnow", "Splittail", "Hitch", "Hardhead", "Sacramento Sucker", "Sacramento Blackfish")
list_smeltish <- c("Wakasagi", "Mississippi Silverside", "Delta Smelt")

Tow_natives0 <- Tow_f %>%filter(CommonName %in% list_native)
Tow_smeltish0 <- Tow_f %>% filter(CommonName %in% list_smeltish)
```

## Run Models {.tabset}

### Smeltish

#### Sum smeltish into one CPUE/Tow
```{r}
Tow_smeltish <- Tow_smeltish0 %>%
  group_by(Survey, fYear, StationCode, PulsePhase, Region, Date,  VolumeSampled, Tow) %>%
  summarize(sumCPUE = sum(CPUE),
            sumCount = sum(sum.count))


Tow_smeltish$PulsePhase <-  factor(Tow_smeltish$PulsePhase, levels(Tow_smeltish$PulsePhase)[c(2,3,1)])
levels(Tow_smeltish$PulsePhase) <- c("Before", "During", "After")
```

#### Look at data

* Mostly silversides
* More fish from EDSM
* More upstream

```{r}
# Proportional abundance
ggplot(Tow_smeltish0, aes(x = fYear, y = CPUE, fill = CommonName)) + geom_col()+ theme_bw()

# Histogram
ggplot(Tow_smeltish, aes(sumCPUE)) + geom_histogram(binwidth = .001) + theme_bw()

# Survey
ggplot(Tow_smeltish, aes(x = Date, y = sumCPUE)) + geom_point() + facet_wrap(Survey~.) + theme_bw()

# Action Phase
SmeltishAP <- ggplot(Tow_smeltish, aes(x = PulsePhase, y = sumCPUE)) + geom_jitter(aes(color = Region), size = 3) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

ggplotly(SmeltishAP)

# Region
ggplot(Tow_smeltish, aes(x = Region, y = sumCPUE)) + geom_jitter(aes(color = PulsePhase), size = 3) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

# Year 
ggplot(Tow_smeltish, aes(x = PulsePhase, y = sumCPUE)) + geom_jitter(aes(color = Region), size = 3) + facet_wrap(~fYear) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

Tow_summary_yr_smeltish1 <- Tow_smeltish %>% 
  group_by(fYear, PulsePhase) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary_yr_smeltish_1, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = PulsePhase)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")

Tow_summary_yr_smeltish2 <- Tow_smeltish %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary_yr_smeltish2, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")


```

Remove outlier
```{r}
Tow_smeltish1 <- Tow_smeltish %>%
  filter(sumCPUE<0.2)
```

#### Barplots for figures - PulsePhase/Region and yr/Region
Summarize - Mean + Standard Error
```{r}
Tow_summary_smeltish <- Tow_smeltish1 %>% 
  group_by(Region, PulsePhase) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

plot4 <- ggplot(Tow_summary_smeltish, aes(x = PulsePhase, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
    annotate("text", x = "After", y = .0055, label = "D", size = 8)  + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15))
plot4

Tow_summary_yr_smeltish2a <- Tow_smeltish1 %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

plotD <- ggplot(Tow_summary_yr_smeltish2a, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
    annotate("text", x = "2019", y = .009, label = "D", size = 8)  + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom")
plotD
```


## Modeling 
Remove station effects by averaging into Region. Use this data for model.
```{r}
Tow_smeltish2 <- Tow_smeltish1 %>%
  group_by(fYear, Region, PulsePhase) %>%
  summarize(sumCount2 = sum(sumCount),
            sumVol = sum(VolumeSampled),
            sumCPUE2 = sum(sumCPUE),
            meanCPUE = mean(sumCPUE))

Tow_smeltish2$PulsePhase <-  factor(Tow_smeltish2$PulsePhase, levels(Tow_smeltish2$PulsePhase)[c(2,3,1)])

```

Replot data - outliers gone
```{r}
ggplot(Tow_smeltish2, aes(fYear, meanCPUE, fill = PulsePhase)) + geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")
```


Run models
```{r}
(M.smeltish1 <- glm(meanCPUE~Region + fYear + PulsePhase, data  = Tow_smeltish2))
summary(M.smeltish1)

(M.smeltish2 <- glm(meanCPUE~Region + fYear * PulsePhase, data  = Tow_smeltish2))
summary(M.smeltish2)

(M.smeltish3 <- glm(meanCPUE~Region * fYear + PulsePhase, data  = Tow_smeltish2))
summary(M.smeltish3)

(M.smeltish4 <- glm(meanCPUE~Region * PulsePhase + fYear, data  = Tow_smeltish2))
summary(M.smeltish4)


AIC(M.smeltish1, M.smeltish2, M.smeltish3, M.smeltish4)
```

- singularity in the ANOVA when added interactions
- chose simple  model M.smeltish1 

Diagnostic Plots
```{r}
par(mfrow = c(2,2))
plot(M.smeltish1)
```

Anova function 
```{r}
library(car)
Anova(M.smeltish1)

1-(M.smeltish1$deviance/M.smeltish1$null.deviance)
```

Post-hoc
```{r}

emmip(M.smeltish3, fYear~PulsePhase|Region)
emmip(M.smeltish3, Region~PulsePhase|fYear)

m_means1a <- emmeans(M.smeltish1, pairwise~fYear, adjust="sidak")
m_means1a

m_means1b <- emmeans(M.smeltish1, pairwise~Region, adjust="sidak")
m_means1b

m_means1c <- emmeans(M.smeltish1, pairwise~PulsePhase, adjust="sidak")
m_means1c
```

**Conclusions**

* Region, fYear, Action Phase, Region:fYear all significant
* Action Phase Pre > During
* More smeltish upstream compared with downstream and Middle Sac River
- significant in 2019, 2017
* 2019>2013, 2014, 2015, 2016, 2017, 2018,| 2017>2015 UPSTREAM only
* EDSM started later. 

**Notes**

* Removed one high CPUE outlier from 2019 of EDSM catch. 








### Native fishes

#### Sum natives into one CPUE/tow
* Filtered to 2016-2019 since no natives prior! 
```{r}
Tow_natives <- Tow_natives0 %>%
  #filter(Year>2015) %>%
  group_by(Date, Survey, fYear, StationCode, FlowPulseType, PulsePhase, Region, VolumeSampled, Tow) %>%
  summarize(sumCPUE = sum(CPUE),
            sumCount = sum(sum.count))


Tow_natives$PulsePhase <-  factor(Tow_natives$PulsePhase, levels(Tow_natives$PulsePhase)[c(2,3,1)])
levels(Tow_natives$PulsePhase) <- c("Before", "During", "After")
```

Look at plots

* Very little catch
```{r}
# Proportion Species
ggplot(Tow_natives0, aes(x = fYear, y = CPUE, fill = CommonName)) + geom_col()+ theme_bw() + scale_fill_viridis(discrete = TRUE, option = "plasma")

# Histogram
ggplot(Tow_natives, aes(sumCPUE)) + geom_histogram(binwidth = .0001) + theme_bw()

# Survey
ggplot(Tow_natives, aes(x = Date, y = sumCPUE)) + geom_point() + facet_wrap(Survey~.) + theme_bw()

# Action Phase
ggplot(Tow_natives, aes(x = PulsePhase, y = sumCPUE)) + geom_jitter(aes(color = Region), size = 3) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

# Region
ggplot(Tow_natives, aes(x = Region, y = sumCPUE)) + geom_jitter(aes(color = PulsePhase), size = 3) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

# Year 
ggplot(Tow_natives, aes(x = PulsePhase, y = sumCPUE)) + geom_jitter(aes(color = Region), size = 3) + facet_wrap(~fYear) + scale_color_viridis(discrete = TRUE, option = "magma") + theme_bw()

Tow_summary_native_yr3 <- Tow_natives %>% 
  group_by(fYear, PulsePhase) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary_native_yr3, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = PulsePhase)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")

Tow_summary_native_yr4 <- Tow_natives %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary_native_yr4, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")
```

#### Barplot
Summarize - Mean + Standard Error
```{r}
Tow_summary_native2 <- Tow_natives %>% 
  group_by(Region, PulsePhase) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

plot5 <- ggplot(Tow_summary_native2, aes(x = PulsePhase, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + 
      annotate("text", x = "After", y = 3e-05, label = "E", size = 8)  + 
  theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15))
plot5




Tow_summary_native_yr4 <- Tow_natives %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

Tow_summary_native_yr4 <- Tow_natives %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

plotE <- ggplot(Tow_summary_native_yr4, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  annotate("text", x = "2019", y = 4e-05, label = "E", size = 8)  + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none")
plotE
```


## Models
Remove station effects by averaging into Region. Use this data for model.
```{r}
Tow_natives2 <- Tow_natives %>%
  group_by(fYear, Region, PulsePhase) %>%
  summarize(sumCount2 = sum(sumCount),
            sumVol = sum(VolumeSampled),
            sumCPUE2 = sum(sumCPUE),
            meanCPUE = mean(sumCPUE))

Tow_natives2$PulsePhase <-  factor(Tow_natives2$PulsePhase, levels(Tow_natives2$PulsePhase)[c(2,3,1)])

```

Run models
```{r}
(M.natives1 <- glm(meanCPUE~Region + fYear + PulsePhase, data  = Tow_natives2))
summary(M.natives1)

(M.natives2 <- glm(meanCPUE~Region + fYear * PulsePhase, data  = Tow_natives2))
summary(M.natives2)

(M.natives3 <- glm(meanCPUE~Region * fYear + PulsePhase, data  = Tow_natives2))
summary(M.natives3)

(M.natives4 <- glm(meanCPUE~Region * PulsePhase + fYear, data  = Tow_natives2))
summary(M.natives4)


AIC(M.natives1, M.natives2, M.natives3, M.natives4)
```

Diagnostic Plots
```{r}
par(mfrow = c(2,2))
plot(M.natives1)
```

**Should we remove Points 26,29,39 as outliers?**


Anova function 
```{r}
Anova(M.natives1, type = 3)
```

Post-hoc

```{r}
m_means2a <- emmeans(M.natives1, pairwise~fYear, adjust="sidak")
m_means2a

m_means2b <- emmeans(M.natives1, pairwise~Region, adjust="sidak")
m_means2b

m_means2c <- emmeans(M.natives1, pairwise~PulsePhase, adjust="sidak")
m_means2c
```


**Conclusions**

* No factors significant
* Not going to use this model - don't think there are enough samples! 29/3144








### Overall CPUE

#### Sum Overall CPUE
```{r}
Tow_filter <- filter(Tow_f, !(CommonName %in% c("Crangon spp.", "Siberian Prawn", "NoCatch", "Mississippi Grass Shrimp")))
Tow_all <- Tow_filter %>%
  group_by(Survey,fYear, Date, Region, StationCode, PulsePhase, MethodCode, VolumeSampled, Tow) %>%
  summarize(sumCPUE = sum(CPUE),
            sumCount = sum(sum.count))

Tow_all$PulsePhase <-  factor(Tow_all$PulsePhase, levels(Tow_all$PulsePhase)[c(2,3,1)])
levels(Tow_all$PulsePhase) <- c("Before", "During", "After")
```

#### Look at data
```{r}
pal36 <- palette36.colors(36)
pal36 <- as.vector(t(pal36))


# Proportion Species
TowSp <- ggplot(Tow_filter, aes(x = fYear, y = CPUE, fill = CommonName)) + 
  geom_col() + 
  scale_fill_manual(values = pal36)+
  theme_bw()  + theme(legend.position = "bottom") + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 11),
        legend.title = element_blank())
TowSp

```

Calculate CPUE to report
```{r}
TowtotalCPUE <- Tow_filter %>%
  group_by(fYear) %>%
  summarize(TCPUE = sum(CPUE))

totalCPUESp <- Tow_filter %>%
  group_by(fYear, CommonName) %>%
  summarize(TsCPUE = sum(CPUE)) %>%
  left_join(TowtotalCPUE) %>%
  mutate(TsCPUE/TCPUE * 100)
```


```{r}
# Histogram
ggplot(Tow_all, aes(sumCPUE)) + geom_histogram(binwidth = .001) + theme_bw()

# Survey
ggplot(Tow_all, aes(x = Date, y = sumCPUE)) + geom_point() + facet_wrap(Survey~.) + theme_bw()

# Action Phase
ggplot(Tow_all, aes(x = PulsePhase, y = sumCPUE)) + geom_jitter(aes(color = Region), size = 3) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

# Region
ggplot(Tow_all, aes(x = Region, y = sumCPUE)) + geom_jitter(aes(color = PulsePhase), size = 3) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

# Boxplot

Tow_all2 <- filter(Tow_all, sumCPUE<3)
ggplot(Tow_all2, aes(x = Region, y = sumCPUE)) + geom_boxplot(aes(fill = PulsePhase)) + scale_color_viridis(discrete = TRUE, option = "viridis") + theme_bw()

ggplot(Tow_all2, aes(x = PulsePhase, y = sumCPUE)) + geom_boxplot(aes(fill = Region)) + scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw()


# Year 
ggplot(Tow_all, aes(x = PulsePhase, y = sumCPUE)) + geom_jitter(aes(color = Region), size = 3) + facet_wrap(~fYear) + scale_color_viridis(discrete = TRUE, option = "magma") + theme_bw()

Tow_summary_all_yr3 <- Tow_all %>% 
  group_by(fYear, PulsePhase) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary__all_yr3, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = PulsePhase)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")

Tow_summary_all_yr4 <- Tow_all %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary_all_yr4, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")

Tow_summary_all_yr5 <- Tow_all %>% 
  group_by(fYear, PulsePhase, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            sumCPUE = sum(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

ggplot(Tow_summary_all_yr5, aes(x = fYear, y = sumCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  facet_wrap(~PulsePhase) +
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
        legend.position = "bottom")
```

Deal with outlier
```{r}
Tow_all <- filter(Tow_all, sumCPUE<2)
```

#### Barplot
Summarize - Mean + Standard Error
```{r}
Tow_summary_all3 <- Tow_all %>% 
  group_by(Region, PulsePhase) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

plot6 <- ggplot(Tow_summary_all3, aes(x = PulsePhase, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + 
      annotate("text", x = "After", y = .07, label = "F", size = 8)  + 
  theme_bw() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15))
plot6




Tow_summary_all_yr4b <- Tow_all %>% 
  group_by(fYear, Region) %>%
  summarize(meanCPUE = mean(sumCPUE),
            n = n(),
            seCPUE = sqrt(var(sumCPUE)/sqrt(n))) 

plotF <- ggplot(Tow_summary_all_yr4b, aes(x = fYear, y = meanCPUE, ymin = meanCPUE - seCPUE, ymax = meanCPUE + seCPUE, fill = Region)) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.4) + 
  scale_fill_viridis(discrete = TRUE, option = "viridis") + theme_bw() +
  annotate("text", x = "2019", y = .13, label = "F", size = 8)  + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 15),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none")

plotF
```

#### Model - normal glm
```{r}
m1all <- glm(sumCPUE~Region + fYear + PulsePhase, data = Tow_all); summary(m1all)

par(mfrow = c(2,2))
plot(m1all)

```

#### Model with log transformation
```{r}
Tow_all$logCPUE = log(Tow_all$sumCPUE + 1)
m2all <- glm(logCPUE~Region + fYear + PulsePhase, data = Tow_all); summary(m2all)

par(mfrow = c(2,2))
plot(m2all)
```

#### Model with poisson distribution
```{r}

# For CPUE, use offset to account for sampling effort
Tow_all$Samp <- log(Tow_all$VolumeSampled)

# Formula
f3 <- formula(sumCount ~ Region + fYear + PulsePhase + offset(Samp))

# Check overdispersion (var>mean)
var(Tow_all$sumCPUE)
mean(Tow_all$sumCPUE)

# Poisson distribution or quasipoisson (overdispersed)
m3all <- glm(f3, data = Tow_all, family = poisson); summary(m3all)
# See dispersion parameter

m3all$deviance/m3all$df.residual #overdispersed

```

#### Model with negative binomial distribution

* https://stats.idre.ucla.edu/r/dae/negative-binomial-regression/
```{r}
library(MASS)
m4all <- glm.nb(f3, data = Tow_all); summary(m4all)
```

# Compare the two
```{r}
BIC(m3all, m4all)
```
Negative Binomial preferred.

Mixed modeling
```{r}

# For CPUE, use offset to account for sampling effort
Tow_all$Samp <- log(Tow_all$VolumeSampled)

# Formula
f1 <- formula(sumCount ~ Region + fYear + PulsePhase + offset(Samp))
f2 <- formula(sumCount ~ Region + fYear + PulsePhase + offset(Samp) + (1|StationCode))
f3 <- formula(sumCount ~ Region + fYear*PulsePhase  + offset(Samp) + (1|StationCode))
f4 <- formula(sumCount ~ Region*fYear + PulsePhase  + offset(Samp) + (1|StationCode))
f5 <- formula(sumCount ~ Region *PulsePhase + fYear  + offset(Samp) + (1|StationCode))
f6 <- formula(sumCount~ Region + PulsePhase + fYear + Region:PulsePhase + PulsePhase:fYear + offset(Samp) + (1|StationCode))

library(glmmTMB)
M.glmer1b <- glmmTMB(f2, data = Tow_all, family = nbinom2)
M.glmer2b <- glmmTMB(f3, data = Tow_all, family = nbinom2)
M.glmer3b <- glmmTMB(f4, data = Tow_all, family = nbinom2)
M.glmer4b <- glmmTMB(f5, data = Tow_all, family = nbinom2)
M.glmer5b <- glmmTMB(f6, data = Tow_all, family = nbinom2);
summary(M.glmer5b)

AIC(M.glmer1b, M.glmer2b, M.glmer4b, M.glmer5b)
```

Diagnostics
https://cran.r-project.org/web/packages/glmmTMB/vignettes/model_evaluation.pdf
```{r}
library(DHARMa)
TowallRes <- simulateResiduals(M.glmer5b)
plot(TowallRes)

library(effects)
plot(allEffects(M.glmer5b))

```

Post-hoc
```{r}
towallcontrast <- emmeans(M.glmer5b, ~PulsePhase * fYear * Region)
result <- contrast(towallcontrast, "consec", simple = "each", combine = TRUE, adjust = "sidak")

print(result)

m_means3a <- emmeans(M.glmer1, pairwise~fYear, adjust="sidak")
m_means3a

m_means3b <- emmeans(M.glmer1, pairwise~Region, adjust="sidak")
m_means3b

m_means3c <- emmeans(M.glmer1, pairwise~PulsePhase, adjust="sidak")
m_means3c
```

```{r}
library(sjPlot)

table <- tab_model(M.glmer5b)
```

**Conclusions**

* More fish Downstream > Upstream > Middle Sac (not significant)
* More fish Pre and During > Post 
* Years: highest 2015, 2017-2019, 2016, 2013, 2014
  * 2015>2016, 2018>2014, 2017>2014, 2015>2014. 2015>2013, 2017>2013, 2018>2013 

- Could consider making year plot and color coding by water year/ action type briefly





Simplified 
Remove station effects by averaging into Region. Use this data for model.
```{r}
Tow_all2 <- Tow_all %>%
  group_by(fYear, Region, PulsePhase) %>%
  summarize(sumCount2 = sum(sumCount),
            sumVol = sum(VolumeSampled),
            sumCPUE2 = sum(sumCPUE),
            meanCPUE = mean(sumCPUE))

Tow_all2$PulsePhase <-  factor(Tow_all2$PulsePhase, levels(Tow_all2$PulsePhase)[c(2,3,1)])

```

Run models
```{r}
(M.all1 <- glm(meanCPUE~Region + fYear + PulsePhase, data  = Tow_all2))
summary(M.all1)

(M.all2 <- glm(meanCPUE~Region + fYear * PulsePhase, data  = Tow_all2))
summary(M.all2)

(M.all3 <- glm(meanCPUE~Region * fYear + PulsePhase, data  = Tow_all2))
summary(M.all3)

(M.all4 <- glm(meanCPUE~Region * PulsePhase + fYear, data  = Tow_all2))
summary(M.all4)


AIC(M.all1, M.all2, M.all3, M.all4)
```
UPDATED MODEL CODE
Diagnostic Plots
```{r}
par(mfrow = c(2,2))
plot(M.all1)
```

**Should we remove Point 49 as outlier?**
Post-hoc

```{r}
m_means4a <- emmeans(M.all1, pairwise~fYear, adjust="sidak")
m_means4a

m_means4b <- emmeans(M.all1, pairwise~Region, adjust="sidak")
m_means4b

m_means4c <- emmeans(M.all1, pairwise~PulsePhase, adjust="sidak")
m_means4c
```
UPDATED MODEL CODE
**Conclusions**

* More fish Downstream vs Upstream and Middle Sac
* More fish Pre > Post, not other significant differences 
* More fish 2018 but not significant in multcomp
