---
title: "Analysis_Multivariate_Tow"
author: "Catarina Pien"
date: "5/19/2020"
output: 
  html_document:
    code_folding: hide
    editor_options: 
      chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation {.tabset}
### Load Packages

```{r load, echo = FALSE, results = FALSE, warning = FALSE, message = FALSE}

rm(list=ls(all=TRUE))
if(!require(tidyverse)){install.packages("tidyverse")
    library(tidyverse)}
if(!require(lubridate)){install.packages("lubridate")
    library(lubridate)} # Datetime
if(!require(GGally)){install.packages("GGally")
    library(GGally)} # Correlation Plots
if(!require(gridExtra)){install.packages("gridExtra")
    library(gridExtra)} # Multiple Plots
if(!require(plotly)){install.packages("plotly")
    library(plotly)} # Multiple Plots
source("HighstatLibV10.R")
```

### Import data
* Fish_ndfa: integrated fish dataset
* Merge flow data into rest of data
* Define variable structures
* Order action phases

```{r import and filter dates, warning = FALSE, message = FALSE, results = FALSE}
Fish_ndfa_0 <- read.csv("R_write/FISH_MAN_allIEPsurveys_20201030.csv")
Fish_ndfa <- select(Fish_ndfa_0, -c(Turbidity, DO, Latitude, Longitude))
FlowDesignation <- read.csv("Data/FlowDatesDesignations.csv")
Regions <- read.csv("R_write/Stations_Fish_NDFA_2020-10-30.csv")
Regions_min <- Regions %>%
  rename(Survey = Survey.x) %>%
  select(-c(Intersect, geometry))

# Look at data
head(Fish_ndfa)
str(Fish_ndfa)

# Add variables
Fish_ndfa$Date <- ymd(Fish_ndfa$Date)
Fish_ndfa$Month <- month(Fish_ndfa$Date)
Fish_ndfa$Day <- day(Fish_ndfa$Date)
FlowDesignation$PreFlowStart <- mdy(FlowDesignation$PreFlowStart)
FlowDesignation$PreFlowEnd <- mdy(FlowDesignation$PreFlowEnd)
FlowDesignation$PostFlowStart <- mdy(FlowDesignation$PostFlowStart)
FlowDesignation$PostFlowEnd <- mdy(FlowDesignation$PostFlowEnd)

# Merge data from FlowDesignation Table (Water Year Type, Flow days and type)
# Filter only Pre-During-Post Flow Action Data. 
# We have a during action, and then Pre = 30 days before/Post = 30 days after
Fish_all0 <- inner_join(Fish_ndfa,FlowDesignation, by = "Year")
Fish_all1 <- left_join(Fish_all0, Regions_min, by = c("StationCode", "Survey"))

# anti_join(Fish_all1, Fish_all0, by = colnames(Fish_all0))

Fish_all <- Fish_all1 %>%
   mutate(ActionPhase = ifelse(Date > PreFlowStart & Date<PreFlowEnd, "Pre", NA)) %>%
   mutate(ActionPhase = replace(ActionPhase, Date > PreFlowEnd & Date < PostFlowStart, "During")) %>% 
   mutate(ActionPhase = replace(ActionPhase, Date > PostFlowStart & Date < PostFlowEnd, "Post")) %>%
  filter(!is.na(ActionPhase)) %>%
   select(-c(PreFlowStart:PostFlowEnd)) %>%
   arrange(Date, Survey, StationCode, CommonName)


# Order Action Phases
Fish_all$ActionPhase <- as.factor(Fish_all$ActionPhase)
Fish_all$ActionPhase <-  factor(Fish_all$ActionPhase, levels(Fish_all$ActionPhase)[c(3,1,2)])

# Define variable structures
str(Fish_all)

Fish_all$Date <- ymd(Fish_all$Date)
Fish_all$fMonth <- ordered(Fish_all$Month)
Fish_all$fYear <- ordered(Fish_all$Year)
Fish_all$WYType <- as.factor(Fish_all$WYType)
Fish_all$FlowPulseType <- as.factor(Fish_all$FlowPulseType)
Fish_all$ActionPhase <- as.factor(Fish_all$ActionPhase)
Fish_all$X1 <- NULL
Fish_all$Survey <- as.factor(Fish_all$Survey)
Fish_all$StationCode <- as.factor(Fish_all$StationCode)
Fish_all$MethodCode <- as.factor(Fish_all$MethodCode)
Fish_all$Region <- as.factor(Fish_all$Region)

str(Fish_all)

```

## Plotting Functions
```{r Functions, warning = FALSE, message = FALSE}
## FUNCTIONS FOR PLOTTING -------------------------------------------------------------
VisPoint <-  function(data,y) {
    y <- enquo(y)
  data %>%
    ggplot() +
    geom_point(mapping = aes(Date,!! y), size = 2) +
    theme_bw() +
  scale_colour_manual(values = c("coral3", "lightseagreen"))+
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 11),
        legend.position = "bottom")
} 

# Boxplots by variable of interest
VisBox <-  function(data, x, y) {
    x <- enquo(x)
    y <- enquo(y)
  data %>%
    ggplot() +
    geom_boxplot(mapping = aes(!! x,!! y), fill = "lightseagreen", colour = "lightgray") +
    theme_bw() +
  scale_colour_manual(values = c("coral3", "lightseagreen"))+
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 11))
} 
#----------------------------------------------------------------------------------------------------
###########

```

## Exploration of overall dataset
* Station distribution
* Time span


```{r explore, warning = FALSE, message = FALSE}
library(leaflet)
library(plotly)
# Look at locations

# Define palette
pal <- colorFactor(c("slateblue", "darkseagreen", "orange", "orangered", "hotpink"), domain = c("DJFMP", "Yolo", "Townet", "EDSM", "FMWT"))

leaflet(Fish_all) %>%
  addTiles() %>%
  addCircleMarkers(
    color = ~pal(Survey),
    opacity = 0.5,
    lng = ~Longitude,
    lat = ~Latitude,
    label = ~StationCode) %>%
  addLegend(pal = pal,
            values = ~Survey,
            position = "bottomright")

                                     
pal2 <- colorFactor(c("purple3", "blue3", "forestgreen", "orange", "orangered", "hotpink"), domain = c("MiddleSacRiver", "ColusaDrainRCS", "UpperYolo", "LowerYolo", "CacheSloughComplex", "LowerSacRiver"))

leaflet(Fish_all)%>%
  addTiles() %>%
  addCircleMarkers(
    color = ~pal2(Region),
    opacity = 0.5,
    lng = ~Longitude,
    lat = ~Latitude,
    label = ~paste(StationCode, "Lat:", Latitude, "Long:", Longitude)) %>%
  addLegend(pal = pal2,
            values = ~Region,
            position = "bottomright")

# Time span 
time <- Fish_all%>%
  group_by(Survey, fYear) %>%
  summarize(sum.count = sum(totalCount))
ggplot(time, aes(x = Survey, y = fYear, fill = sum.count)) + geom_tile() + theme_minimal()

```

* Number of fish per survey
* Total counts per survey
```{r}
# Look at datasets
Fishsp <- Fish_all %>%
  group_by(CommonName) %>% summarize(count = n())

Fishsp %>%
  plot_ly(x = ~CommonName, 
          y = ~count,
          type = "bar")


fishP <- ggplot(Fish_all, aes(x = CommonName, y = totalCount, fill = Survey)) + geom_col() + theme_bw() + theme(
                    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplotly(fishP)

Fishsp %>%
  plot_ly(x = ~CommonName, 
          y = ~count,
          type = "bar")
```

# Tow Data 
## Data Prep{.tabset}
* Filter to just tow data
* Fill in zeros for unlisted species
* Calculate CPUE 
* Organize data the way you want it

```{r seine, warning = FALSE, message = FALSE}
# Filter Dataset
Tow <- Fish_all %>% filter(Survey %in% c("EDSM", "Townet", "FMWT")) %>%
  arrange(Survey, Date, StationCode, Tow, Depth)

### Complete Cases
# For each Date, Survey, StationCode combination, make sure each fish species is represented with 
# either positive count or zero. 
Tow_completecase <- Tow %>%
  group_by(Date, Survey, StationCode, Tow, Depth, CommonName) %>%
  summarize(sum.count = sum(totalCount)) %>%
  ungroup() %>%
  complete(CommonName, nesting(Date, Survey, StationCode, Tow, Depth), fill = list(sum.count = 0)) %>%
  arrange(Date,Survey, StationCode,CommonName)

### Merge back together with rest of data

# Get distinct samples for looking at Water Quality 
Tow_samples <- Tow %>% select(-c(CommonName, totalCount)) %>% distinct()

# Merge 
Tow_complete <- left_join(Tow_completecase, Tow_samples, by = c("Date", "Survey", "StationCode", "Tow", "Depth"))

### Calculate CPUE
# Remove samples with no volume 
Tow_CPUE <- Tow_complete %>%
  filter(!is.na(VolumeSampled))%>%
  mutate(CPUE = round(sum.count/VolumeSampled,3))

### Rearrange columns 
Tow_f <- Tow_CPUE[, c("Date", "Year", "Month", "Day", "Survey", "StationCode", 
                          "Latitude", "Longitude","Region","MethodCode",
                         "WYType", "FlowPulseType", "NetFlowDays","ActionPhase", 
                         "Secchi",  "Conductivity", "WaterTemp",  "Tow",
                         "Depth", "VolumeSampled",  "CommonName", "sum.count", "CPUE")]

### Mean CPUE 
## Calculate means for each species by year-location-actionphase
CPUE_means_Tow_phase <- Tow_f %>%
  group_by(Year, Survey, StationCode, Tow, ActionPhase, CommonName) %>% 
  summarize(mean.CPUE = mean(CPUE)) 

```

## Data Exploration
### Water Quality
* Simple point plots and histograms, correlation matrix
```{r envexplore, warning = FALSE, message = FALSE, fig.height=11, fig.width=10}
# Look at locations

# Define palette
pal2 <- colorFactor(c("orange", "orangered", "hotpink"), domain = c("Townet", "EDSM", "FMWT"))

leaflet(Tow_f) %>%
  addTiles() %>%
  addCircleMarkers(
    color = ~pal2(Survey),
    opacity = 0.5,
    radius=2,
    lng = ~Longitude,
    lat = ~Latitude,
    label = ~StationCode) %>%
  addLegend(pal = pal2,
            values = ~Survey,
            position = "bottomright")

############ OUTLIERS #################
# Boxplots
WTvis1 <- VisPoint(Tow_samples, WaterTemp)
WTvis2 <- VisBox(Tow_samples, Month, WaterTemp)
WTvis3 <- VisBox(Tow_samples, Year, WaterTemp)
Cvis1 <- VisPoint(Tow_samples, Conductivity)
Cvis2 <- VisBox(Tow_samples, Month, Conductivity)
Cvis3 <- VisBox(Tow_samples, Year, Conductivity)
Svis1 <- VisPoint(Tow_samples, Secchi)
Svis2 <- VisBox(Tow_samples, Month, Secchi)
Svis3 <- VisBox(Tow_samples, Year, Secchi)

# Plot together
grid.arrange(WTvis1, WTvis2, WTvis3, Cvis1, Cvis2, Cvis3, Svis1, Svis2, Svis3,  ncol = 3)

# Boxplots
plot_ly(data = Tow_samples, x = ~StationCode, y = ~WaterTemp, color = ~Region, type = 'box')
plot_ly(data = Tow_samples, x = ~StationCode, y = ~Conductivity, color = ~Region, type = 'box')
plot_ly(data = Tow_samples, x = ~StationCode, y = ~Secchi, color = ~Region, type = 'box')

############ CORRELATIONS ####################
# Correlation Matrix WQ
Corr.wqT <- Tow_samples %>% select(WaterTemp, Conductivity, Secchi)
ggpairs(Corr.wqT)

# Variance Inflation Factor (VIF)
corvif(Corr.wqT) # Want to get rid of the variable if VIF > 4


```

## Clean up WQ data (QC)
* Edit anything that needs to be removed, flagged, changed.
* Replace Missing Data

```{r clean, warning = FALSE, message = FALSE}
# Clean up wq data
# QC Check - does anything look weird?
Tow_samples %>% filter(WaterTemp>30 | WaterTemp<1) # High but will leave in
Tow_samples %>% filter(Conductivity > 5000) # High but will leave in

# Function to fill in missing values with mean 
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

# Filling in missing values using impute.mean function from above. 
# Variables are renamed because using mutate adds on new columns to the matrix.
# Use impute.mean to fill in NAs, rename updated variables using mutate
Tow_f <- Tow_f %>%
  group_by(WYType) %>%
  mutate(
    Cond = impute.mean(Conductivity),
    WTemp = impute.mean(WaterTemp),
    SecDepth = impute.mean(Secchi)) %>%
  ungroup()
```

# Tow Data Analysis {.tabset}
## Preparation 
1. Filter for species or sets of species
```{r Tow filters, warning = FALSE, message = FALSE}
# Which species are relatively abundant?
Tow_f %>%
  plot_ly(x = ~CommonName, 
          y = ~sum.count,
          type = "bar")

fig <- plot_ly(Tow_f, x = ~ActionPhase, y = ~CPUE, color = ~CommonName, type = 'bar', name = ~CommonName)
fig <- fig %>% layout(yaxis = list(title = 'CPUE'), barmode = 'stack')
fig

# Lists
list_native <- c("Sacramento Pikeminnow", "Splittail", "Hitch", "Hardhead", "Sacramento Sucker", "Sacramento Blackfish")
list_smeltish <- c("Wakasagi", "Inland Silverside")
list_smeltpred <- c("Largemouth Bass", "Smallmouth Bass", "Striped Bass", "Spotted Bass")
list_cyprinid <- c("Sacramento Pikeminnow", "Sacramento Splittail", "Hitch", "Fathead Minnow",
                           "Golden Shiner", "Hardhead", "Carp", "Goldfish")
list_nmds <- c("Sacramento Pikeminnow", "Splittail", "Hitch", "Hardhead", "Sacramento Sucker", "Sacramento Blackfish",
               "Wakasagi", "Inland Silverside", "Delta Smelt", 
               "Carp", "Goldfish", "Hardhead", "Golden Shiner", "Fathead Minnow", "Hitch",
               "Rainwater Killifish", "Western Mosquitofish", "Black Crappie", "White Crappie", "Bluegill", "Bigscale Logperch",
               "Largemouth Bass", "Smallmouth Bass", "Striped Bass", "Spotted Bass",
               "Threadfin Shad", "American Shad", "Tridentiger spp_", "Shimofuri Goby", "White Catfish")
list_nmds_small <- c("Sacramento Pikeminnow", "Splittail", "Hitch", "Hardhead", "Sacramento Sucker", "Sacramento Blackfish",
               "Wakasagi", "Inland Silverside", "Delta Smelt", "Longfin Smelt", 
               "Largemouth Bass", "Smallmouth Bass", "Striped Bass", "Spotted Bass",
               "Threadfin Shad", "American Shad", "Bluegill", "Black Crappie", "Bigscale Logperch")

Tow_tf <- Tow_f %>% filter(CommonName == "Threadfin Shad")
Tow_iss <- Tow_f %>% filter(CommonName == "Inland Silverside")
Tow_ashad <- Tow_f %>% filter(CommonName == "American Shad")
Tow_natives <- Tow_f %>%filter(CommonName %in% list_native)
Tow_nmds_larger <- Tow_f%>% filter(CommonName %in%list_nmds)
Tow_nmds_0 <- Tow_f %>% filter(CommonName %in% list_nmds_small)

```

2. Some summary stats 

### Summary Stats
```{r stats, warning = FALSE, message = FALSE}
towstatsall <- Tow_nmds_larger%>% 
  group_by(CommonName) %>%
  summarize(mean = mean(CPUE), 
            max = max(CPUE),
            min = min(CPUE), 
            sum = sum(CPUE),
            CV = sd(CPUE)/mean,
            zeros = sum(CPUE == 0),
            n = n(), 
            prop.absent = round(zeros/n,2))

```

3. Remove low abundance, high abundance species. In this case:
* Remove species that are not present
* Species < 5% of samples, left Delta Smelt in at 0.96 since we are interested in them

```{r clean species}
notabund <- c("Bigscale Logperch", "Black Crappie", "Bluegill", "Golden Shiner", "Hitch", "Largemouth Bass", "Sacramento Blackfish", "Rainwater Killifish", "Sacramento Sucker",
              "Splittail", "Striped Bass", "Spotted Bass", "White Catfish", "White Crappie")
Tow_nmds <- filter(Tow_nmds_0, !(CommonName %in% notabund))
```


## Exploration/QC
1. Look for outliers in species data
* Click species to remove, look at each individual species for any obvious outliers
* Potential Threadfin outliers? 

```{r sp outliers, warning = FALSE, message = FALSE}
# Overall CPUE
index= seq(1,length(Tow_nmds$CPUE))

# Explore each species - look for outliers here by adding and removing species
Tow_nmds%>%
  plot_ly() %>%
  add_trace(x = ~index, 
            y = ~CPUE, 
            color = ~CommonName,
            colors = "Set3",
            type = "scatter")

Tow_nmds %>% 
  plot_ly() %>%
  add_trace(x = ~CommonName,
            y = ~CPUE,
            color = ~Survey,
            type = "box")


```

Plots of species composition
* Lots of species 
* Species used in NMDS
* No Threadfin Shad
```{r Plots of species comp, warning = FALSE, message = FALSE, fig.width= 10, fig.height = 6.5}
library(RColorBrewer)
Tow_noTFS <- Tow_nmds %>% filter(CommonName != "Threadfin Shad")

# More plots
ggplot(Tow_nmds_larger, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity", position = "fill") + 
    scale_fill_brewer(palette = "Paired") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = "bottom")
  

ggplot(Tow_nmds_larger, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity", position = "fill") + facet_wrap(~Region) +
    scale_fill_brewer(palette = "Paired") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = "bottom")

# Proportional CPUE
ggplot(Tow_nmds, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity", position = "fill") +
  labs(y = "Proportional CPUE") + 
  scale_fill_brewer(palette = "Paired") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(size = 15))

# Proportional CPUE - Regions
ggplot(Tow_nmds, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity", position = "fill") + facet_wrap(~Region) +
    scale_fill_brewer(palette = "Paired") +
  labs(y = "Proportional CPUE") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        legend.position = "bottom")

# CPUE - Regions
ggplot(Tow_nmds, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity") + facet_wrap(~Region) +
    scale_fill_brewer(palette = "Paired") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        legend.position = "bottom")

# Proportional CPUE - Regions
ggplot(Tow_noTFS, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity", position = "fill") + facet_wrap(~Region) +
    scale_fill_brewer(palette = "Paired") +
  labs(y = "Proportional CPUE") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.title = element_text(size = 15),
        
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        legend.position = "bottom")

# Actual CPUE values
ggplot(Tow_noTFS, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity") + facet_wrap(~Region) +
    scale_fill_brewer(palette = "Paired") +
  theme_bw()+
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 15),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.position = "bottom",
        strip.text = element_text(size = 15))


ggplot(Tow_noTFS, aes(x = ActionPhase, y = CPUE, fill = CommonName)) + 
  geom_bar(stat = "identity", position = "fill") +
    scale_fill_brewer(palette = "Paired") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 15), 
        axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.position = "bottom")
```

2. Look at initial trends - environment vs. species

```{r trends, warning = FALSE, message = FALSE}
# Threadfin vs environmental
threadfin1 <- ggplot(Tow_tf, aes(x = WTemp, y = CPUE)) + geom_point() + theme_bw()
threadfin2 <- ggplot(Tow_tf, aes(x = Cond, y = CPUE)) + geom_point() + theme_bw()
threadfin3 <- ggplot(Tow_tf, aes(x = SecDepth, y = CPUE)) + geom_point() + theme_bw()
grid.arrange(threadfin1, threadfin2, threadfin3, top = "Threadfin CPUE by Environmental Variable")

# Inland Silverside vs environmental
iss1 <- ggplot(Tow_iss, aes(x = WTemp, y = CPUE)) + geom_point() + theme_bw()
iss2 <- ggplot(Tow_iss, aes(x = Cond, y = CPUE)) + geom_point() + theme_bw()
iss3 <- ggplot(Tow_iss, aes(x = SecDepth, y = CPUE)) + geom_point() + theme_bw()
grid.arrange(iss1, iss2, iss3, top = "Inland Silverside CPUE by Environmental Variable")

# American Shad vs environmental
ashad1 <- ggplot(Tow_ashad, aes(x = WTemp, y = CPUE)) + geom_point() + theme_bw()
ashad2 <- ggplot(Tow_ashad, aes(x = Cond, y = CPUE)) + geom_point() + theme_bw()
ashad3 <- ggplot(Tow_ashad, aes(x = SecDepth, y = CPUE)) + geom_point() + theme_bw()
grid.arrange(ashad1, ashad2, ashad3,  top = "American Shad CPUE by Environmental Variable")
```

## Transformations
1. Log of sqrt for highly skewed data, or ranging >2 order magnitude
2. Arcsine sqrt for proportional data
3. Use same transformation for same variable set (e.g. species)
4. Consider binary transformation when percent of zero values is high (>50%) or distinct values low (<10)

```{r transformtests, warning = FALSE, message = FALSE}
# Environmental Matrix
Tow_env <- Tow_nmds %>% dplyr::select(c(24:26))

# Check for normality of variables
cond <- Tow_env$Cond
temp <- Tow_env$WTemp
sd <- Tow_env$SecDepth

hist(cond)
hist(log(cond+1)) # Use this
hist(sqrt(cond+0.5))
qqnorm(log(cond+1))

hist(temp)
hist(log(temp+1)) # Use this
hist(sqrt(temp))
qqnorm(temp)
qqnorm(log(temp+1))

hist(sd) # Keep it?
hist(log(sd+1))
qqnorm(sd)
qqnorm(log(sd+1))
```
 
## PCA {.tabset}
### Run PCA
1.  Apply transformations to variables 
2.  If you have variables with different units/scales, choose the correlation matrix (scale = TRUE). Then you do not need to scale your data beforehand.
3. Run PCA
4. Check PCA (Scree plot, Randomization Tests)

* https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html
```{r pca, warning = FALSE, message = FALSE}
source('biostats.R')
library(vegan)
library(ggfortify)

### Transform to get close to normality assumptions
Tow_env_t <- Tow_env %>% mutate(Cond = log(cond + 1),
                                    WTemp = log(WTemp + 1),
                                    SecDepth = log(SecDepth + 1)) %>%
  as.data.frame()
row.names(Tow_env_t) <- row.names(Tow_nmds)

### Run PCA. Center and Scale the data (correlation matrix).
pca.env <- prcomp(Tow_env_t, center=T, scale.=T)
summary(pca.env)

### Check how well PCA worked
# Scree Plot with Broken Stick values. 
# If eigenvalue is greater than broken stick value it is "significant"
screeplot(pca.env, bstick = TRUE) # According to this, don't keep PC1 and PC2

# Monte Carlo Randomization
ordi.monte(Tow_env_t, ord = 'pca', dim = 3) # 2 axes is good

### Check which variables matter
# Loadings. Generally, if magnitude >/= 0.3 this is important.
pca.env$rotation
# Structure coefficients: linear correlations between original variables and PC scores
pca.structure(pca.env, Tow_env_t, dim = 5, cutoff = 0.4)

```

### Plot
5. Plot and interpret PCA

```{r interpretPCA, warning = FALSE, message = FALSE}

# Autoplot has best customization, looks nice
  # use the "data = " to bring in original dataset so we can 
  # color-code by factors
autoplot(pca.env, data = Tow_nmds, colour = 'ActionPhase', loadings = TRUE,
         loadings.label = TRUE,
         loadings.colour = "black",
         loadings.label.vjust = -.5,
         loadings.label.col = "black",
         loadings.label.size = 4,
         loadings.label.font = 4) +
  theme_bw()
autoplot(pca.env, data = Tow_nmds, colour = 'Region', loadings = TRUE,
         loadings.label = TRUE,
         loadings.colour = "black",
         loadings.label.vjust = -.5,
         loadings.label.col = "black",
         loadings.label.size = 4,
         loadings.label.font = 4) + 
  theme_bw()
```

## NMDS {.tabset}
1. Pick your variables
2. Switch to wide format (species listed across the top)
3. Remove any row where nothing was caught

#### Matrix
```{r matrix for nmds, warning = FALSE, message = FALSE}
Tow_f_sp <-Tow_nmds %>% dplyr::select(c("Survey", "StationCode", "Tow", "Depth", "Date", "FlowPulseType", "WYType", "Region",
                                    "ActionPhase", "Month", "CommonName", "CPUE", "Cond",
                                    "WTemp", "SecDepth"))
# Species Matrix
Tow_sp_w <- Tow_f_sp %>% pivot_wider(names_from = CommonName, values_from = CPUE, values_fill = list(CPUE=0)) %>% ungroup()

# Remove any row where there is no catch for the day.
Tow_sp_sum <- Tow_sp_w %>% mutate(Total = dplyr::select(., 14:19) %>%  rowSums(na.rm = TRUE)) %>%
  filter(Total !=0)
```

#### Transform and/or standardize for nMDS
* Some options for transformations: sqrt, log, absence/presence
```{r nmdstransform, warning = FALSE, message = FALSE}
# Sqrt transform data
sqrt.tow <- Tow_sp_sum %>% mutate_if(is.numeric, function(x) {
  sqrt(x)
})
                   
# Log transform data
log.tow <- Tow_sp_sum %>% mutate_if(is.numeric, function(x) {
  log(x + 1) })

# Absence/Presence data
bin.Tow <- Tow_sp_sum %>% mutate_if(is.numeric, function(x) {
  case_when(x>0 ~0,
            x ==0 ~1)})

# Proportional data

```
 
### Run NMDS
1. Run
* Currently a subset of data since otherwise even 2000 iterations won't converge :(
* Went with sqrt transformation - least change


```{r Tow nmds, warning = FALSE, message = FALSE, results = FALSE}
library(vegan)

test.tow <- sample_n(sqrt.tow, 120)

# NMDS Tow
# This can be slow
tow.nmds <- metaMDS(test.tow[,14:19], distance="bray", k=2, trymax=500, autotransform = FALSE)
tow.nmds

```

### Check
2. Check NMDS
* Stress < 0.15 is acceptable fit.
* Check stressplot.

```{r nmdsInterp, message = FALSE, warning = FALSE}
# Scree Plot - what should k=? 
# This can be slow
nmds.scree(test.tow[,14:19], distance='bray', k=5, trymax = 200, autotransform = FALSE) 

### Check NMDS solution 
# Large scatter is not good
stressplot(tow.nmds)
```

### Plot
3. Plot Results. 
* Resources:
- https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/
- https://rpubs.com/collnell/manova
- https://chrischizinski.github.io/rstats/vegan-ggplot2/
```{r nmdsplot, message = FALSE, warning = FALSE}
### NMDS scores
tow.scores <- as.data.frame(scores(tow.nmds))
species.scores <- as.data.frame(scores(tow.nmds, "species"))
species.scores$species <- rownames(species.scores)

# Need a category to merge scores with the rest of env data
row.names(tow.scores) <- row.names(test.tow)
tow.nmdsplot <- cbind(tow.scores, test.tow)

# Prettier Plots
# Option 1
ggplot(tow.nmdsplot, aes(NMDS1, NMDS2, color = ActionPhase)) + 
  geom_point(position = position_jitter(.1)) +
  geom_text(aes(label = WYType)) +
  stat_ellipse(type = 't', size = 1) +
  annotate("text", x = min(tow.nmdsplot$NMDS1), y = min(tow.nmdsplot$NMDS2), label = paste('Stress = ', round(tow.nmds$stress,3))) + # Add stress to plot
  theme_minimal() 

# Option 2
# With species scores and boxes

# Make the datasets for the boxes
grp.a <- tow.nmdsplot[tow.nmdsplot$ActionPhase == "Pre", ][chull(tow.nmdsplot[tow.nmdsplot$ActionPhase == 
    "Pre", c("NMDS1", "NMDS2")]), ]  # hull values for grp A
grp.b <- tow.nmdsplot[tow.nmdsplot$ActionPhase == "During", ][chull(tow.nmdsplot[tow.nmdsplot$ActionPhase == 
    "During", c("NMDS1", "NMDS2")]), ]  # hull values for grp B
grp.c<- tow.nmdsplot[tow.nmdsplot$ActionPhase == "Post", ][chull(tow.nmdsplot[tow.nmdsplot$ActionPhase == 
    "Post", c("NMDS1", "NMDS2")]), ]  # hull values for grp C
hull.data0 <- rbind(grp.a, grp.b)  #combine grp.a and grp.b
hull.data <- rbind(hull.data0,grp.c)

# Plot
ggplot() + 
  geom_point(data = tow.nmdsplot, aes(x = NMDS1, y = NMDS2, color = ActionPhase),
             position = position_jitter(.1)) +
  geom_text(data = tow.nmdsplot, aes(NMDS1, NMDS2, label = WYType, color = ActionPhase)) +
  geom_text(data = species.scores, aes(x = NMDS1, y = NMDS2, label= species), color = "black", size = 5) +
  theme_minimal() +
  annotate("text", x = min(tow.nmdsplot$NMDS1), y = min(tow.nmdsplot$NMDS2), label = paste('Stress = ', round(tow.nmds$stress,3))) + # Add stress to plot
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=ActionPhase,group=ActionPhase),alpha=0.30) 

```

## PERMANOVA {.tabset}

### Run models
1. Standardize species data (basically becomes proportional data) to reduce influence of large numbers. Could also do sqrt transformation.
2. Run PERMANOVA with each variable separately. Only keep significant variables.
3. Create model that is significant with greatest R2, move down the list until new variables are not signficant.
4. For any categorical variables, test interactions as well.

```{r permanova, warning = FALSE, message = FALSE}
# Row standardization
tow.st <- data.stand(Tow_sp_sum[,14:19], method = 'total', margin = 'row', plot = F)

# Individual permanova models
(perm.all <- adonis(formula = tow.st ~ Region + Month + Survey + WYType + ActionPhase + FlowPulseType + WTemp + Cond + SecDepth + Region:Month, data = Tow_sp_sum, method = "bray",
                   permutations = 99))

(perm.WY <- adonis(formula = tow.st~Region, data = Tow_sp_sum, method = "bray", permutations = 99))

(perm.Month <- adonis(formula = tow.st~Month, data = Tow_sp_sum, method = "bray", permutations = 99))

(perm.Survey <- adonis(formula = tow.st~Survey, data = Tow_sp_sum, method = "bray", permutations = 99))

(perm.WY <- adonis(formula = tow.st~WYType, data = Tow_sp_sum, method = "bray", permutations = 99))

(perm.Action <- adonis(formula = tow.st~ActionPhase, data = Tow_sp_sum, method = "bray", permutations = 99))

(perm.FlowPulse <- adonis(formula = tow.st~FlowPulseType, data = Tow_sp_sum, method = "bray", permutations = 99))

```

### Homogeneity of Variance
5. Test for Homogeneity of Variance
* Choose variables in your model
* You want the anova not to be significant (significant means variances are not homogeneous)
* Boxplots should be pretty equal
```{r betadispersion, warning = FALSE, message = FALSE}
# Make dissimilarity matrix
spe.d <- vegdist(tow.st, "bray")

# StationCode
(sp.bdp <- betadisper(spe.d, Tow_sp_sum$Region))
anova(sp.bdp)
permutest(sp.bdp,pairwise=TRUE)
plot(sp.bdp, ellipse = TRUE)
boxplot(sp.bdp)

# Month
(sp.bdp2 <- betadisper(spe.d, Tow_sp_sum$Month))
anova(sp.bdp2)
permutest(sp.bdp2,pairwise=TRUE)
plot(sp.bdp2, ellipse = TRUE)
boxplot(sp.bdp2)

# WYType
sp.bdp3 <- betadisper(spe.d, Tow_sp_sum$WYType)
sp.bdp3
anova(sp.bdp3)
permutest(sp.bdp3,pairwise=TRUE)
plot(sp.bdp3, ellipse = TRUE)
boxplot(sp.bdp3)

# ActionPhase
sp.bdp4 <- betadisper(spe.d, Tow_sp_sum$ActionPhase)
sp.bdp4
anova(sp.bdp4)
permutest(sp.bdp4,pairwise=TRUE)
plot(sp.bdp4, ellipse = TRUE)
boxplot(sp.bdp4)

# FlowPulseType
sp.bdp5 <- betadisper(spe.d, Tow_sp_sum$FlowPulseType)
sp.bdp5
anova(sp.bdp5)
permutest(sp.bdp5,pairwise=TRUE)
plot(sp.bdp5, ellipse = TRUE)
boxplot(sp.bdp5)

```

## CCA {.tabset}
* Format: 1 matrix of predictors, 1 matrix of species data
1. Transform and standardize continuous variables
2. Transform species data - sqrt
3. Check if unimodal distribution. If yes, run CCA.

### Run
```{r cca, warning = FALSE, message = FALSE}

# Prepare continuous data
tow.env.scaled <- as.data.frame(scale(Tow_sp_sum[,11:13]),center = "TRUE", scale = "TRUE")
tow.env.f <- cbind(tow.env.scaled, Tow_sp_sum[,c("Region", "WYType", "ActionPhase","FlowPulseType", "Month")])

# Prepare species data
tow.sp <- sqrt.tow[,14:19]

# Check unimodal distribution
# Axis length > 4 = unimodal (CCA)
# 2-4 probably unimodal
# <2 = linear model
decorana(tow.sp, ira=0)

# Run CCA
spe.cca <- cca(tow.sp ~.,tow.env.f)
summary(spe.cca)
```

### Interpretation
4. Interpretation of CCA axis, terms
* Can use a step  model to create final CCA model
```{r cca.interp, message = FALSE, warning = FALSE}

# Test the significance of the CCA
anova(spe.cca)
anova(spe.cca, by = "axis")
anova(spe.cca, by = "terms")
anova(spe.cca, by = 'margin')

# Forward stepping
cca.step <- ordistep(cca(tow.sp ~ 1, data = tow.env.f), scope = formula(spe.cca), 
                        direction = "forward", pstep = 1000)
```

### Plot
5. Plot CCA
```{r ccaplot, message = FALSE, warning = FALSE, fig.width=10, fig.height = 7}
# Plot the CCA

cca.biplot = function(cca){
  
  #find plot dimensions (changed these so all the points would fit)
  xmin = min(summary((cca))$species[, 1]) * 1.2
  xmax = max(summary((cca))$species[, 1]) * 1.7
  ymin = min(summary((cca))$species[, 2]) * 1.1
  ymax = max(summary((cca))$species[, 2]) * 1.5
  
  par(bty='l')
  
  # plot(cca,disp='species',scaling=1) #scale to show species the best
  plot(summary((cca))$species[, 1], summary((cca))$species[, 2],
       type = 'n', xlim = c(xmin, xmax), ylim = c(ymin, ymax), 
       ylab = 'CA2 (3.3% Variation Explained)', xlab = 'CA1 (10.5% Variation Explained)')
  axis(side = 1, lwd = 2)
  axis(side = 2, lwd = 2)
  box(lwd = 2)
  
  #draw origin lines
  segments(-2, 0, 2, 0, lwd = 1, lty = 3)
  segments(0, -4, 0, 2, lwd = 1, lty = 3)
  
  #Add species names
  text(summary((cca))$species[, 1],summary((cca))$species[, 2], labels = rownames(summary((cca))$species), cex = 0.9)
  
  #define continuous variables possibly used
  cont.vars=c('Cond','WTemp','DOx','SecDepth','Turb')
  
  bi.names = row.names(summary(cca)$biplot) #names of environmental variables
  # centroid.names=row.names(summary(cca)$centroid) 
  centroid.names = data.frame(Rows = rownames(summary((cca))$centroid))
  
  #Pain in the a$$ way to do this, but I created abbreviated names for each level of
  #my categorical variable so that they would look nicer when plotted
  #Else I would have names like HabitatC on the biplot
  # 
  centroid.names$New[centroid.names$Rows == 'ActionPhasePre'] = 'PreAction'
  centroid.names$New[centroid.names$Rows == 'ActionPhaseDuring'] = 'DuringAction'
  centroid.names$New[centroid.names$Rows == 'ActionPhasePost'] = 'PostAction'

  centroid.names$New[centroid.names$Rows == 'WYTypeC'] = 'CriticalWY'
  centroid.names$New[centroid.names$Rows == 'WYTypeD'] = 'DryWY'
  centroid.names$New[centroid.names$Rows == 'WYTypeW'] = 'WetWY'

  centroid.names$New[centroid.names$Rows == 'FlowPulseTypeMA-Ag'] = 'AgPulse'
  centroid.names$New[centroid.names$Rows == 'FlowPulseTypeMA-SR'] = 'SRPulse'
  centroid.names$New[centroid.names$Rows == 'FlowPulseTypeNF'] = 'NoFA'
  
  centroid.names$New[centroid.names$Rows == 'Month4'] = 'Apr'
  centroid.names$New[centroid.names$Rows == 'Month5'] = 'May'
  centroid.names$New[centroid.names$Rows == 'Month6'] = 'June'
  centroid.names$New[centroid.names$Rows == 'Month7'] = 'July'
  centroid.names$New[centroid.names$Rows == 'Month8'] = 'Aug'
  centroid.names$New[centroid.names$Rows == 'Month9'] = 'Sep'
  centroid.names$New[centroid.names$Rows == 'Month10'] = 'Oct'
  centroid.names$New[centroid.names$Rows == 'Month11'] = 'Nov'
  centroid.names$New[centroid.names$Rows == 'Month12'] = 'Dec'
  centroid.names$New[centroid.names$Rows == 'Month1'] = 'Jan'
  centroid.names$New[centroid.names$Rows == 'Month2'] = 'Feb'
  centroid.names$New[centroid.names$Rows == 'Month3'] = 'Mar'
  
    centroid.names$New[centroid.names$Rows == 'RegionCacheSloughComplex'] = 'CacheSloughComplex'
  centroid.names$New[centroid.names$Rows == 'RegionCentralYolo'] = 'CentralYolo'
  centroid.names$New[centroid.names$Rows == 'RegionColusaDrainRCS'] = 'ColusaDrainRCS'
  centroid.names$New[centroid.names$Rows == 'RegionLowerSacRiver'] = 'LowerSacRiver'
  centroid.names$New[centroid.names$Rows == 'RegionLowerYolo'] = 'LowerYolo'
  centroid.names$New[centroid.names$Rows == 'RegionMiddleSacRiver'] = 'MiddleSacRiver'
  
  #This could possibly be slimmed down, but it plots arrows if the variable is continuous (part of the list above)
  # and plots a point for the categorical variables instead
  
  for(i in 1:length(summary(cca)$biplot[, 1])){
    
    #Test that the row name is one of the continuous variables before plotting arrows
    if(bi.names[i] %in% cont.vars){
      arrows(0, 0, summary(cca)$biplot[i, 1], summary(cca)$biplot[i, 2],
             lwd = 1, angle = 25, length = 0.10, col = "#F4aa42")
      text(summary(cca)$biplot[i, 1] * 1.1, 
           summary(cca)$biplot[i, 2] * 1.1, labels = bi.names[i], col = '#F4aa42', cex = 0.8, font = 7)
    }
    
  }
  
  #now plot the centered mean value of each nominal variable
  #Use if statement to test for case where I don't have any categorical variables
  
  if(is.numeric(summary(cca)$centroids) == TRUE){
    
    points(summary(cca)$centroids[, 1] * 0.9, summary(cca)$centroids[, 2] * 0.9, pch = 15, col = 'black')
    
    text(summary(cca)$centroids[, 1] * 0.9, (summary(cca)$centroids[, 2] * 0.9 + 0.15), labels = centroid.names$New,
         col = 'darkcyan', cex = 0.9, adj = .5)
  }
  
}

cca.biplot(spe.cca)



```

