---
title: "FISH_MAN_Analysis_Models"
author: "Catarina Pien"
date: "12/15/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
Code for fish models
Fish ~ Region + Action Phase + Year

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r load, echo = FALSE, results = FALSE, warning = FALSE, message = FALSE}

rm(list=ls(all=TRUE))

library(tidyverse)
library(lubridate) # Datetime
library(pscl) # Run zips and zinbs
library(visreg) # Visualize models
```

## Import data
* Fish_ndfa: integrated fish dataset
* FlowDesignation: Assigns action phases by date
* Regions: Assigns regions to fish stations
* Merge all together
* Define variable structures
* Order action phases

```{r import and filter dates, warning = FALSE, message = FALSE, results = FALSE}
Fish_ndfa_0 <- read.csv("fish data/FISH_MAN_allIEPsurveys_20201030.csv")
Fish_ndfa <- select(Fish_ndfa_0, -c(Turbidity, DO, Latitude, Longitude))
FlowDesignation <- read.csv("fish data/FlowDatesDesignations_45days.csv")
Regions <- read.csv("fish data/Stations_Fish_NDFA_2021-01-13.csv")
Regions_min <- Regions %>%
  rename(Survey = Survey.x) %>%
  select(-c(Intersect, geometry, RegionMult))

# Look at data
head(Fish_ndfa)
str(Fish_ndfa)

# Add variables
Fish_ndfa$Date <- ymd(Fish_ndfa$Date)
Fish_ndfa$Month <- month(Fish_ndfa$Date)
Fish_ndfa$Day <- day(Fish_ndfa$Date)
FlowDesignation$PreFlowStart <- mdy(FlowDesignation$PreFlowStart)
FlowDesignation$PreFlowEnd <- mdy(FlowDesignation$PreFlowEnd)
FlowDesignation$PostFlowStart <- mdy(FlowDesignation$PostFlowStart)
FlowDesignation$PostFlowEnd <- mdy(FlowDesignation$PostFlowEnd)

# Merge data from FlowDesignation Table (Water Year Type, Flow days and type)
# Filter only Pre-During-Post Flow Action Data. 
# We have a during action, and then Pre = 30 days before/Post = 30 days after
Fish_all0 <- inner_join(Fish_ndfa,FlowDesignation, by = "Year")
Fish_all1 <- left_join(Fish_all0, Regions_min, by = c("StationCode", "Survey"))

# anti_join(Fish_all1, Fish_all0, by = colnames(Fish_all0))

Fish_all <- Fish_all1 %>%
   mutate(ActionPhase = ifelse(Date > PreFlowStart & Date<PreFlowEnd, "Pre", NA)) %>%
   mutate(ActionPhase = replace(ActionPhase, Date > PreFlowEnd & Date < PostFlowStart, "During")) %>% 
   mutate(ActionPhase = replace(ActionPhase, Date > PostFlowStart & Date < PostFlowEnd, "Post")) %>%
  filter(!is.na(ActionPhase)) %>%
   select(-c(PreFlowStart:PostFlowEnd)) %>%
   arrange(Date, Survey, StationCode, CommonName)


# Order Action Phases
Fish_all$ActionPhase <- as.factor(Fish_all$ActionPhase)
Fish_all$ActionPhase <-  factor(Fish_all$ActionPhase, levels(Fish_all$ActionPhase)[c(3,1,2)])

# Define variable structures
str(Fish_all)

Fish_all$Date <- ymd(Fish_all$Date)
Fish_all$fMonth <- ordered(Fish_all$Month)
Fish_all$fYear <- ordered(Fish_all$Year)
Fish_all$WYType <- as.factor(Fish_all$WYType)
Fish_all$FlowPulseType <- as.factor(Fish_all$FlowPulseType)
Fish_all$ActionPhase <- as.factor(Fish_all$ActionPhase)
Fish_all$X1 <- NULL
Fish_all$Survey <- as.factor(Fish_all$Survey)
Fish_all$StationCode <- as.factor(Fish_all$StationCode)
Fish_all$MethodCode <- as.factor(Fish_all$MethodCode)
Fish_all$Region <- as.factor(Fish_all$Region)

str(Fish_all)

```


## Data Prep
* Filter to just tow data and year starting 2013
* Fill in zeros for unlisted species using complete cases
* Calculate CPUE 
* Organize data

```{r seine, warning = FALSE, message = FALSE}
# Filter Dataset
Tow <- Fish_all %>% 
  filter(Survey %in% c("EDSM", "Townet", "FMWT")) %>%
  filter(Year>2012) %>%
  arrange(Survey, Date, StationCode, Tow, Depth)

### Complete Cases
# For each Date, Survey, StationCode combination, make sure each fish species is represented with either positive count or zero. 
Tow_completecase <- Tow %>%
  group_by(Date, Survey, StationCode, Tow, Depth, CommonName) %>%
  summarize(sum.count = sum(totalCount)) %>%
  ungroup() %>%
  complete(CommonName, nesting(Date, Survey, StationCode, Tow, Depth), fill = list(sum.count = 0)) %>%
  arrange(Date, Survey, StationCode,CommonName)

### Merge back together with rest of data

# Get distinct samples for looking at Water Quality 
Tow_samples <- Tow %>% select(-c(CommonName, totalCount)) %>% distinct()

# Merge 
Tow_complete <- left_join(Tow_completecase, Tow_samples, by = c("Date", "Survey", "StationCode", "Tow", "Depth"))

### Calculate CPUE
# Remove samples with no volume 
Tow_CPUE <- Tow_complete %>%
  filter(!is.na(VolumeSampled))%>%
  mutate(CPUE = sum.count/VolumeSampled)

### Rearrange columns 
Tow_f <- Tow_CPUE[, c("Date", "Year", "fYear", "Month", "Day", "Survey", "StationCode", "Latitude", "Longitude","Region","MethodCode", "WYType", "FlowPulseType", "NetFlowDays","ActionPhase", "Secchi",  "Conductivity", "WaterTemp",  "Tow", "Depth", "VolumeSampled",  "CommonName", "sum.count", "CPUE")]
```


```{r}
### Mean CPUE 
## Calculate means for each species by year-location-actionphase
CPUE_means_Tow_phase <- Tow_f %>%
  group_by(Year, Survey, StationCode, Tow, ActionPhase, CommonName) %>% 
  summarize(mean.CPUE = mean(CPUE)) 
```

## Calculate how often fish is present in samples
```{r}
Tow_f %>%
  group_by(Survey, StationCode, CommonName) %>%
  summarize()
```


## Filter for species sets

```{r Tow filters, warning = FALSE, message = FALSE}
# Lists
list_native <- c("Sacramento Pikeminnow", "Splittail", "Hitch", "Hardhead", "Sacramento Sucker", "Sacramento Blackfish")
list_smeltish <- c("Wakasagi", "Inland Silverside")
list_nmds <- c("Sacramento Pikeminnow", "Splittail", "Hitch", "Hardhead", "Sacramento Sucker", "Sacramento Blackfish","Wakasagi", "Inland Silverside", "Delta Smelt", "Carp", "Goldfish", "Hardhead", "Golden Shiner", "Fathead Minnow", "Hitch","Rainwater Killifish", "Western Mosquitofish", "Black Crappie", "White Crappie", "Bluegill", "Bigscale Logperch","Largemouth Bass", "Smallmouth Bass", "Striped Bass", "Spotted Bass","Threadfin Shad", "American Shad", "Tridentiger spp_", "Shimofuri Goby", "White Catfish")

Tow_natives <- Tow_f %>%filter(CommonName %in% list_native)
Tow_nmds_larger <- Tow_f%>% filter(CommonName %in%list_nmds)
Tow_smeltish <- Tow_f %>% filter(CommonName %in% list_smeltish)
```

## Run ANOVA {.tabset}

### Smeltish

Using transformations - doesn't work well
```{r}
(M.smeltish1 <- glm(log(CPUE+1)~FlowPulseType + ActionPhase + Region, data  = Tow_smeltish))
summary(M.smeltish1)
par(mfrow = c(2,2))
plot(M.smeltish1)

logCPUE <-  log(Tow_smeltish$CPUE + 1)
hist(logCPUE)

```

Zero-inflated model
```{r}

# For CPUE, use offset to account for sampling effort
Tow_smeltish$Samp <- log(Tow_smeltish$VolumeSampled)

# Formula
f1 <- formula(sum.count ~ Region + fYear + ActionPhase + offset(Samp))

# Negbin
zinb1 <- zeroinfl(f1, na.action = "na.fail", dist="negbin", link = "logit", data = Tow_smeltish) ; summary(zinb1)

# Poisson
zip1 <- zeroinfl(f1, dist = "poisson", link = "logit", data = Tow_smeltish); summary(zip1)
```

Compare zero-inflated poisson with zero-inflated negbin
```{r}

```

Diagnostic Plots

* Still need to figure out how to do this

```{r}
# Diagnostic Plots -------------------------------------
# Model Validation 
# Calculate residuals
EP1 <- resid(zinb1, type = "pearson")

# Diagnostic Plots
par(mfrow = c(2,2))
plot(x=zinb1$fitted.values, y = EP1, main = "Pearson residuals")
# qqrplot(EP1) # Don't know why not working
plot(x=Tow_smeltish$ActionPhase,y = EP1, main = "Action Phase")
plot(x=Tow_smeltish$Region, y = EP1, main = "Region")

par(mfrow = c(2,2))
visreg(zinb1)
```

**Conclusions**

* More smeltish in phase "pre" followed by "during" follow by "post"
* More smeltish in Lower Yolo, less in Lower Sac River and Middle Sac River
* No effect of region on the extra zeroes
* More zeroes in post and pre vs. during

### Native fishes
Zero-inflated model
```{r}

# For CPUE, use offset to account for sampling effort
Tow_natives$Samp <- log(Tow_natives$VolumeSampled)

# Formula
f2 <- formula(sum.count ~ Region + fYear + ActionPhase + offset(Samp))

# Negbin
zinb2 <- zeroinfl(f2, na.action = "na.fail", dist="negbin", link = "logit", data = Tow_natives) ; summary(zinb2)

# Poisson
zip2 <- zeroinfl(f2, dist = "poisson", link = "logit", data = Tow_natives); summary(zip2)
```

Compare zero-inflated poisson with zero-inflated negbin
```{r}

```

Diagnostic Plots
```{r}
# Diagnostic Plots -------------------------------------
# Model Validation 
# Calculate residuals
EP2 <- resid(zinb2, type = "pearson")

# Diagnostic Plots
par(mfrow = c(2,2))
plot(x=zinb2$fitted.values, y = EP2, main = "Pearson residuals")
# qqrplot(EP2) # Don't know why not working
plot(x=Tow_natives$ActionPhase,y = EP2, main = "Action Phase")
plot(x=Tow_natives$Region, y = EP2, main = "Region")

par(mfrow = c(2,2))
visreg(zinb2)
```

### Overall CPUE
* Think about if you want to sum overall CPUE or do individual species? - can do regular ANOVA if you combine all species for overall CPUE
* If you do modeling as above, consider using the Tow_nmds_larger grouping or calculate species that are in at least 5% of samples so there aren't *quite* so many zeroes. 
* Below code is modeling for including **all** fish

Sum Overall CPUE
```{r}
Tow_allfish <- Tow_f %>%
  group_by(Survey,fYear, Date, Region, StationCode, ActionPhase, MethodCode, Tow, VolumeSampled) %>%
  summarize(CPUE_total = sum(CPUE),
            Count_total = sum(sum.count))
```

Zero-inflated model
```{r}

# For CPUE, use offset to account for sampling effort
Tow_f$Samp <- log(Tow_f$VolumeSampled)

# Formula
f3 <- formula(sum.count ~ Region + fYear + ActionPhase + offset(Samp))

# Negbin
zinb3 <- zeroinfl(f3, na.action = "na.fail", dist="negbin", link = "logit", data = Tow_f) ; summary(zinb3)

# Poisson
zip3 <- zeroinfl(f3, dist = "poisson", link = "logit", data = Tow_f); summary(zip3)
```

Compare zero-inflated poisson with zero-inflated negbin
```{r}

```

Diagnostic Plots
```{r}
# Diagnostic Plots -------------------------------------
# Model Validation 
# Calculate residuals
EP3 <- resid(zinb3, type = "pearson")

# Diagnostic Plots
par(mfrow = c(2,2))
plot(x=zinb3$fitted.values, y = EP3, main = "Pearson residuals")
# qqrplot(EP3) # Don't know why not working
plot(x=Tow_f$ActionPhase,y = EP3, main = "Action Phase")
plot(x=Tow_f$Region, y = EP3, main = "Region")

par(mfrow = c(2,2))
visreg(zinb3)
```


