---
title: 'NDFA contaminants Data: exploratory data analysis of zooplankton concentrations'
author: "Dave Bosworth"
date: "3/5/2021"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document contains plots and data summary tables used for exploratory data analysis (EDA) for the contaminants concentrations in zooplankton used in North Delta Flow Action synthesis project.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(DT)
library(here)
```

```{r load functions, message = FALSE, warning = FALSE}
# Source global NDFA functions
source(here("global_ndfa_funcs.R"))

# Source contaminants EDA functions
source(here("Water_Quality/Contaminants/Analyses/contam_EDA_funcs.R"))
```

# Import and Prepare Data

```{r import data, message = FALSE}
# Import contaminants data
load(here("Water_Quality/Data_Processed/contam_proc_data.RData"))
```

To prepare the contaminants data for the exploratory data analysis (EDA) plots and summary tables, we will need do the following initial steps:

* Add flow action periods and remove any data outside of the period of interest (45 days before and after the flow pulse)
* Add pesticide type and class designations
* Create a `SamplingEvent` variable. Some of the stations were sampled 1-2 days apart. This variable will group these within one sampling event.

```{r prepare data 1, message = FALSE}
contam_zoop_clean_v1 <- contam_zoop %>% 
  # Add Date and Year variables
  mutate(
    Date = date(DateTime),
    Year = year(DateTime)
  ) %>% 
  # Add flow pulse periods
  ndfa_action_periods() %>% 
  # Add pesticide types and classes
  left_join(contam_type_class) %>% 
  # Clean up data frame
  select(
    StationCode,
    Date,
    Year,
    FlowActionPeriod,
    Analyte,
    PesticideType,
    PesticideClass,
    Result,
    Units
  )

# Create a tibble of dates and sampling events to join to main data frame
dates_se_cw <-
  tibble(
    Date = sort(unique(contam_zoop_clean_v1$Date)),
    SamplingEvent = c(
      rep("Aug 29-30, 2017", 2),
      rep("Sep 12-13, 2017", 2),
      rep("Sep 19-20, 2017", 2),
      rep("Oct 3-4, 2017", 2),
      rep("Oct 17-19, 2017", 2),
      rep("Nov 1-2, 2017", 2),
      rep("Jul 25-26, 2018", 2),
      rep("Aug 8-9, 2018", 2),
      rep("Aug 22-23, 2018", 2),
      "Aug 30, 2018",
      rep("Sep 5-6, 2018", 2),
      rep("Sep 19-20, 2018", 2),
      rep("Oct 10-11, 2018", 2),
      rep("Oct 25-26, 2018", 2),
      rep("Nov 8-9, 2018", 2),
      rep("Jul 15-16, 2019", 2),
      rep("Aug 5-7, 2019", 2),
      rep("Aug 19-21, 2019", 2),
      rep("Sep 4-5, 2019", 2),
      rep("Sep 16-18, 2019", 2),
      rep("Sep 30-Oct 2, 2019", 2),
      rep("Oct 14-16, 2019", 2),
      "Oct 28, 2019"
    )
  )

# Add SamplingEvent variable to contam_zoop_clean_v1
contam_zoop_clean_v2 <- contam_zoop_clean_v1 %>% 
  left_join(dates_se_cw) %>% 
  select(-Date) %>% 
  relocate(SamplingEvent, .after = StationCode)
```

There is one `SamplingEvent` and `StationCode` combination when samples were not collected. We need to record this combination so that it can be identified as "No Samples Collected" in the plots.

```{r find missing data combinations, message = FALSE}
# Find SamplingEvent and StationCode combinations when samples were not collected
contam_zoop_miss_v1 <- contam_zoop_clean_v2 %>%
  group_by(Year) %>%
  expand(StationCode, SamplingEvent) %>%
  ungroup() %>%
  anti_join(contam_zoop_clean_v2) %>% 
  # Add text for label in plots and grouping variable for colors of text labels
  mutate(
    txt_label = "No Samples\nCollected",
    label_grp = "Not sampled"
  )
```

There are also a couple of `SamplingEvent` and `StationCode` combinations with all contaminants having values below their Method Detection Limits (MDL's). We need to find these combinations so that they can be identified as "All values below their MDL's" in the plots by `PesticideType`.

```{r find samples with all below MDL}
# Find SamplingEvent and StationCode combinations with all contaminants having values below their MDL's
contam_zoop_nd <- contam_zoop_clean_v2 %>% 
  group_by(StationCode, SamplingEvent, Year) %>%
  filter(all(Result == "< MDL")) %>% 
  ungroup() %>% 
  distinct(Year, StationCode, SamplingEvent) %>% 
  # Add text for label in plots and grouping variable for colors of text labels
  mutate(
    txt_label = "All values below\ntheir MDL's",
    label_grp = "All below MDL"
  )
```

Finally, there are a significant number of `SamplingEvent`, `StationCode`, and `PesticideType` combinations with all contaminants having values below their Method Detection Limits (MDL's). We need to find these combinations so that they can be identified as "All values below their MDL's" in the plots by `PesticideClass`. This will only be done for the Fungicide, Herbicide, Insecticide, and their degradate types since these are the only types that have some values above their MDL's. For simplicity, we will combine the pesticide types with their corresponding degradate categories.

```{r find samples with all below MDL incl type}
# Find SamplingEvent, StationCode, and PesticideType combinations with all contaminants having values below their MDL's
contam_zoop_nd_type <- contam_zoop_clean_v2 %>%
  # Group the Herbicide and Insecticide pesticide types with their degradate categories
  mutate(
    PesticideType = case_when(
      str_detect(PesticideType, "^Herbicide") ~ "Herbicide",
      str_detect(PesticideType, "^Insecticide") ~ "Insecticide",
      TRUE ~ PesticideType
    )
  ) %>% 
  filter(PesticideType %in% c("Fungicide", "Herbicide", "Insecticide")) %>% 
  group_by(StationCode, SamplingEvent, Year, PesticideType) %>% 
  filter(all(Result == "< MDL")) %>% 
  ungroup() %>% 
  distinct(Year, StationCode, SamplingEvent, PesticideType) %>% 
  # Add text for label in plots and grouping variable for colors of text labels
  mutate(
    txt_label = "All values below\ntheir MDL's",
    label_grp = "All below MDL"
  )
```

The missing data due to lack of sampling or all values being below their MDL's create problems in the EDA plots if you still want to include all of the sampling events on the x-axis. To help with this, we will pull out the sampling event information and modify it to define the factor order for each year.

```{r define sampling events for each year}
# Define factor order for SamplingEvent
se_order <- dates_se_cw %>% 
  distinct(SamplingEvent) %>% 
  pull(SamplingEvent)

# Pull out Sampling Event information to be added to the plots later
contam_zoop_se <- contam_zoop_clean_v2 %>% 
  distinct(Year, SamplingEvent, FlowActionPeriod) %>% 
  mutate(SamplingEvent = factor(SamplingEvent, levels = se_order)) %>% 
  nest(df_se = -Year) %>% 
  mutate(df_se = map(df_se, ~mutate(.x, SamplingEvent = fct_drop(SamplingEvent))))
```

Now the `contam_zoop_clean_v2` data frame is ready for a few final steps to prepare it for summarizing for plots and summary tables:

* Assume that < MDL values are zero and remove them
* Convert `StationCode` to a factor for plotting order. `SamplingEvent` will be converted to a factor later in the script.

```{r prepare data 2}
# Finish preparing the contaminants data
contam_zoop_clean_f <- contam_zoop_clean_v2 %>%
  # Remove < MDL values
  filter(!str_detect(Result, "^<")) %>%
  # Convert Result to numeric variable
  mutate(Result = as.numeric(Result)) %>% 
# Apply factor order for StationCode
  conv_fact_sta_code()
```

Next, we will sum the contaminant concentrations by `PesticideType`.

```{r sum conc pesticide type, message = FALSE}
# Sum by PesticideType
contam_sum_type <- contam_zoop_clean_f %>% 
  group_by(
    StationCode, 
    SamplingEvent, 
    Year, 
    FlowActionPeriod, 
    PesticideType,
    Units
  ) %>% 
  summarize(TotalConc = sum(Result)) %>% 
  ungroup()
```

We will also sum the contaminant concentrations by `PesticideClass` keeping `PesticideType` as a higher level category in the data frame. This will allow for us to create plots of total contaminant concentration by pesticide class for each pesticide type. Only the Fungicide, Herbicide, Insecticide, and their degradate types will be plotted by class since these are the only types that have some values above their MDL's. For simplicity, we will combine the pesticide types with their corresponding degradate categories.

```{r sum conc pesticide class, message = FALSE}
# Sum by PesticideClass
contam_sum_class <- contam_zoop_clean_f %>% 
  group_by(
    StationCode, 
    SamplingEvent, 
    Year, 
    FlowActionPeriod, 
    PesticideType, 
    PesticideClass,
    Units
  ) %>% 
  summarize(TotalConc = sum(Result)) %>% 
  ungroup() %>%
  # Group the Herbicide and Insecticide pesticide types with their degradate categories
  mutate(
    PesticideType = case_when(
      str_detect(PesticideType, "^Herbicide") ~ "Herbicide",
      str_detect(PesticideType, "^Insecticide") ~ "Insecticide",
      TRUE ~ PesticideType
    )
  )
```

# Create Plots

The data frames that identify the `SamplingEvent` and `StationCode` combinations with missing data because of either lack of sampling or all values were below their MDL's need to be combined and modified to be used properly in the plotting function.

```{r combine and modify missing df}
# For the plots by Pesticide Type:
contam_zoop_miss_type_f <- 
  bind_rows(contam_zoop_miss_v1, contam_zoop_nd) %>% 
  # Apply factor order for StationCode
  conv_fact_sta_code() %>% 
  # Nest the data frame by year in order to join it contam_sum_type
  nest(data_miss = -Year)

# For the plots by Pesticide Class:
contam_zoop_miss_class_f <- contam_zoop_miss_v1 %>% 
  # Add all Pesticide Type combinations to contam_zoop_miss_v1
  expand_grid(PesticideType = c("Fungicide", "Herbicide", "Insecticide")) %>% 
  bind_rows(contam_zoop_nd_type) %>% 
  # Apply factor order for StationCode
  conv_fact_sta_code() %>% 
  # Nest the data frame by year and PesticideType in order to join it contam_sum_class
  nest(data_miss = -c(Year, PesticideType))
```

We are now ready to create the plots both by `PesticideType` and `PesticideClass`. The factor levels for `SamplingEvent` as defined in `contam_ss_se` will be applied to each plot so that the sampling events are displayed in the proper order.

```{r create plots, warning = FALSE, message = FALSE}
# Plots of Pesticide Types:
contam_plot_type <- contam_sum_type %>% 
  nest(data = -Year) %>%
  arrange(Year) %>% 
  left_join(contam_zoop_miss_type_f) %>% 
  left_join(contam_zoop_se) %>% 
  # Apply factor order for SamplingEvent to data and data_miss
  mutate(
    data = map2(data, df_se, .f = conv_fact_se),
    data_miss = map2(data_miss, df_se, .f = conv_fact_se)
  ) %>% 
  # Create plots
  mutate(
    plot_stack = pmap(
      list(data, data_miss, df_se),
      .f = create_barplot, 
      fill_var = PesticideType, 
      type = "stack"
    ),
    plot_fill = pmap(
      list(data, data_miss, df_se), 
      .f = create_barplot, 
      fill_var = PesticideType, 
      type = "fill"
    )
  )
  
# Plots of Pesticide Classes:
contam_plot_class <- contam_sum_class %>% 
  nest(data = -c(Year, PesticideType)) %>%
  arrange(Year, PesticideType) %>% 
  left_join(contam_zoop_miss_class_f) %>% 
  left_join(contam_zoop_se) %>% 
  # Apply factor order for SamplingEvent to data and data_miss
  mutate(
    data = map2(data, df_se, .f = conv_fact_se),
    data_miss = map2(data_miss, df_se, .f = conv_fact_se)
  ) %>% 
  # Create plots
  mutate(
    plot_stack = pmap(
      list(data, data_miss, df_se),
      .f = create_barplot, 
      fill_var = PesticideClass, 
      type = "stack"
    ),
    plot_fill = pmap(
      list(data, data_miss, df_se), 
      .f = create_barplot, 
      fill_var = PesticideClass, 
      type = "fill"
    )
  ) %>% 
  # structure data frame for more efficient printing of plots
  nest(yr_group = -Year)
```

# Plots by Pesticide Type {.tabset .tabset-pills}

For all plots, the colors of the x-axis labels indicate the flow action period during which the samples were collected:

* Blue: Before
* Green: During
* Black: After

```{r print plots pesticide type, echo = FALSE, results = "asis"}
for (i in 1:nrow(contam_plot_type)) {
  # Create subheadings for each Year
  cat("## ", as.character(contam_plot_type$Year[i]), " {.tabset .tabset-pills}\n\n")
  # Create subheading for stacked barplot and print it
  cat("### Stacked Barplot\n\n")
  print(contam_plot_type$plot_stack[[i]])
  cat("\n\n")
  # Create subheading for filled barplot and print it
  cat("### Filled Barplot\n\n")
  print(contam_plot_type$plot_fill[[i]])
  cat("\n\n")
}
```

# Plots by Pesticide Class {.tabset .tabset-pills}

For all plots, the colors of the x-axis labels indicate the flow action period during which the samples were collected:

* Blue: Before
* Green: During
* Black: After

```{r print plots pesticide class, echo = FALSE, results = "asis"}
for (i in 1:nrow(contam_plot_class)) {
  # Create subheadings for each Year
  cat("## ", as.character(contam_plot_class$Year[i]), " {.tabset .tabset-pills}\n\n")
  for (j in 1:nrow(contam_plot_class$yr_group[[i]])) {
    # Create subheadings for each Pesticide Type
    cat(
      "### ", 
      as.character(contam_plot_class$yr_group[[i]]$PesticideType[j]), 
      " {.tabset .tabset-pills}\n\n"
    ) 
   # Create subheading for stacked barplot and print it
    cat("#### Stacked Barplot\n\n")
    print(contam_plot_class$yr_group[[i]]$plot_stack[[j]])
    cat("\n\n")
    # Create subheading for filled barplot and print it
    cat("#### Filled Barplot\n\n")
    print(contam_plot_class$yr_group[[i]]$plot_fill[[j]])
    cat("\n\n")  
  }
}
```

# Data Summary Tables

## Total Conc by Pesticide Type

```{r data summ tbl pesticide type}
contam_sum_type %>% 
  mutate(
    FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")),
    PesticideType = factor(PesticideType)
  ) %>% 
  select(-Units) %>% 
  arrange(SamplingEvent, StationCode, PesticideType) %>% 
  datatable(filter = "top")
```

## Total Conc by Pesticide Class

```{r data summ tbl pesticide class}
contam_sum_class %>% 
  mutate(
    FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")),
    PesticideType = factor(PesticideType),
    PesticideClass = factor(PesticideClass)
  ) %>% 
  select(-Units) %>% 
  arrange(SamplingEvent, StationCode, PesticideType, PesticideClass) %>% 
  datatable(filter = "top")
```

## All Detected Data

```{r data summ tbl all detected data}
contam_zoop_clean_f %>% 
  mutate(
    FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")),
    PesticideType = factor(PesticideType),
    PesticideClass = factor(PesticideClass),
    Analyte = factor(Analyte)
  ) %>% 
  select(-Units) %>% 
  arrange(SamplingEvent, StationCode, PesticideType, PesticideClass, Analyte) %>% 
  datatable(filter = "top")
```

