---
title: "NDFA Contaminants Data: EPA Benchmark Exceedences"
author: "Dave Bosworth"
date: "`r Sys.Date()`"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Produce a simple plots showing the number of EPA benchmark exceedances for the contaminant concentrations in water and suspended sediment data sets collected for the NDFA project. Includes data collected from 2015-2019 and benchmarks for acute and chronic toxicity for both fish and invertebrates.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(scales)
library(here)

# Source functions
source(here("global_ndfa_funcs.R"))
```

# Import and Prepare Data

```{r import data, message = FALSE}
# Import data
load(here("Water_Quality/Data_Processed/contam_proc_data.RData"))
df_epa_exceed <- read_csv(here("Water_Quality/Contaminants/Analyses/epa_benchmark_exceedances_flowperiod.csv"))
```

```{r prepare data, message = FALSE}
# Combine and prepare data for plot
# Count the number of samples collected by year and flow pulse period
contam_n_samples <- 
  bind_rows(contam_water, contam_ss) %>% 
  mutate(
    Date = date(DateTime),
    Year = year(DateTime)
  ) %>% 
  ndfa_action_periods() %>% 
  distinct(StationCode, Year, Date, FlowActionPeriod) %>% 
  count(Year, FlowActionPeriod, name = "N_Samples")

# Calculate the percent of samples that exceed EPA benchmarks grouped by year
  # and flow pulse period
df_epa_exceed_perc <- df_epa_exceed %>% 
  left_join(contam_n_samples) %>% 
  mutate(Perc_Exceedance = N_Exceedance/N_Samples) %>% 
  # Apply plotting order for flow pulse period
  mutate(FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")))
```

# Create Plots

Number of samples that exceed at least one EPA benchmark:

```{r plot N}
df_epa_exceed_perc %>% 
  ggplot(aes(x = FlowActionPeriod, y = N_Exceedance)) +
  geom_col() +
  facet_grid(cols = vars(Benchmark), rows = vars(Year)) +
  labs(
    x = "Flow Action Period",
    y = "Number of samples with at least one exceedance"
  )
```

Percentage of samples that exceed at least one EPA benchmark:

```{r plot perc}
df_epa_exceed_perc %>% 
  ggplot(aes(x = FlowActionPeriod, y = Perc_Exceedance)) +
  geom_col() +
  facet_grid(cols = vars(Benchmark), rows = vars(Year)) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    x = "Flow Action Period",
    y = "Percent of samples with at least one exceedance"
  )
```

