---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Final results from selected model"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Look at results of the preferred model during model selection: GAM using smooths for weekly average flow by Station and Year with a cyclic penalized cubic regression spline smooth term for week number to account for seasonality. Create summary figures and tables for the manuscript. The model selection process and diagnostics can be found in the [manuscript_synthesis/notebooks/rtm_chl_models_flow_weekly_avg.Rmd](https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html) document.

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(mgcv)
library(broom)
library(ggeffects)
library(emmeans)
library(multcomp)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

# Import Data

```{r import data}
# Define file path for processed data
fp_data <- here("manuscript_synthesis/data/processed")

# Import weekly average water quality data
df_wq <- readRDS(file.path(fp_data, "wq_week_avg_2013-2019.rds"))

# Import weekly average flow data
df_flow <- readRDS(file.path(fp_data, "flow_week_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare chla data, message = FALSE}
# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

# We will use LIS flow data as a proxy for STTD
df_flow_c <- df_flow %>% mutate(StationCode = if_else(StationCode == "LIS", "STTD", StationCode))

# Prepare chlorophyll and flow data for analysis
df_chla_c <- df_wq %>% 
  select(StationCode, Year, Week, Chla) %>% 
  drop_na(Chla) %>% 
  # Filter to only include representative stations for 4 habitat types (RD22,
    # STTD, LIB, RVB) and only include years 2015-2019
  filter(
    StationCode %in% sta_order,
    Year %in% 2015:2019
  ) %>% 
  # Join flow data to chlorophyll data
  left_join(df_flow_c, by = join_by(StationCode, Year, Week)) %>% 
  # Remove all NA flow values
  drop_na(Flow) %>% 
  mutate(
    # Scale and log transform chlorophyll values
    Chla_log = log(Chla * 1000),
    # Apply factor order to StationCode
    StationCode = factor(StationCode, levels = sta_order),
    # Add a column for Year as a factor for the model
    Year_fct = factor(Year)
  ) %>% 
  arrange(StationCode, Year, Week) %>% 
  # Add 2 lag variables for chlorophyll to be used in the model to help with
    # serial autocorrelation
  group_by(StationCode, Year) %>% 
  mutate(
    lag1 = lag(Chla_log),
    lag2 = lag(Chla_log, 2)
  ) %>% 
  ungroup() %>% 
  drop_na(lag1, lag2)
```

# Create GAM model

```{r gam sflow lag2, warning = FALSE}
m_gam_sflow_lag2 <- gam(
  Chla_log ~ s(Flow, by = StationCode) + s(Flow, by = Year_fct) + Year_fct * StationCode + s(Week, bs = "cc", k = 5) + lag1 + lag2, 
  data = df_chla_c,
  method = "REML", 
  knots = list(week = c(0, 52))
)
```

# Model Results

## ANOVA table

```{r anova gam sflow lag2, warning = FALSE}
# Create ANOVA table for GAM
an_gam_sflow_lag2 <- anova(m_gam_sflow_lag2)
an_gam_sflow_lag2

# Extract elements from ANOVA table for export
# Parametric terms
df_an_gam_ptable <- an_gam_sflow_lag2$pTerms.table %>% as_tibble(rownames = "Term")

# Smooth terms
df_an_gam_stable <- an_gam_sflow_lag2$s.table %>% 
  as_tibble(rownames = "Term") %>% 
  rename(df = edf) %>% 
  select(-Ref.df)

# Combine and format for export
df_an_gam_sflow_lag2 <- 
  bind_rows(df_an_gam_ptable, df_an_gam_stable) %>% 
  transmute(
    Term,
    across(
      c("df", "F"),
      ~ if_else(
        abs(.x) < 0.1,
        formatC(signif(.x, 4), digits = 3, format = "e"),
        formatC(signif(.x, 4), digits = 4, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    )
  ) %>% 
  mutate(across(c("df", "F", "p_value"), ~ paste0(.x, "##")))
```

## Effect of Flow by Station

```{r gam sflow lag2 flow station effects plot}
# Calculate min and max flows for each station to narrow down x-axis in the plot
df_chla_flow_sta_summ <- df_chla_c %>% 
  summarize(
    Flow_min = min(Flow),
    Flow_max = max(Flow),
    .by = c(StationCode)
  ) %>% 
  mutate(
    Flow_buffer = (Flow_max - Flow_min) * 0.05,
    Flow_min = Flow_min - Flow_buffer,
    Flow_max = Flow_max + Flow_buffer
  )

# Calculate effects of flow on chlorophyll for each station holding the
  # non-focal variables constant - marginal effects/adjusted predictions
df_gam_flow_sta_eff <- 
  as.data.frame(
    ggemmeans(m_gam_sflow_lag2, terms = c("Flow", "StationCode")),
    terms_to_colnames = TRUE
  ) %>% 
  as_tibble() %>% 
  # Narrow down range of flow values for each station
  left_join(df_chla_flow_sta_summ, by = join_by(StationCode)) %>% 
  filter(Flow >= Flow_min & Flow <= Flow_max) %>% 
  transmute(
    StationCode,
    Flow,
    # Back calculate model fits and confidence levels
    across(c("predicted", "conf.low", "conf.high"), ~ exp(.x) / 1000)
  )

# Create effects plot
plt_gam_flow_sta_eff <- df_gam_flow_sta_eff %>% 
  ggplot(aes(x = Flow, y = predicted)) +
  geom_point(
    data = df_chla_c,
    aes(y = Chla, color = Year_fct),
    alpha = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
  facet_wrap(vars(StationCode), scales = "free") +
  theme_bw() +
  labs(
    x = "Flow (cfs)",
    y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
  ) +
  scale_x_continuous(breaks = breaks_extended(6)) +
  scale_color_viridis_d(name = "Year", option = "C")

plt_gam_flow_sta_eff
```

## Effect of Flow by Year

```{r gam sflow lag2 flow year effects plot, warning = FALSE}
# Calculate min and max flows for each station to narrow down x-axis in the plot
df_chla_flow_yr_summ <- df_chla_c %>% 
  summarize(
    Flow_min = min(Flow),
    Flow_max = max(Flow),
    .by = c(Year_fct)
  ) %>% 
  mutate(
    Flow_buffer = (Flow_max - Flow_min) * 0.05,
    Flow_min = Flow_min - Flow_buffer,
    Flow_max = Flow_max + Flow_buffer
  )

# Calculate effects of flow on chlorophyll for each year holding the
  # non-focal variables constant - marginal effects/adjusted predictions
df_gam_flow_yr_eff <- 
  as.data.frame(
    ggemmeans(m_gam_sflow_lag2, terms = c("Flow", "Year_fct")),
    terms_to_colnames = TRUE
  ) %>% 
  as_tibble() %>% 
  # Narrow down range of flow values for each station
  left_join(df_chla_flow_yr_summ, by = join_by(Year_fct)) %>% 
  filter(Flow >= Flow_min & Flow <= Flow_max) %>% 
  transmute(
    Year_fct,
    Flow,
    # Back calculate model fits and confidence levels
    across(c("predicted", "conf.low", "conf.high"), ~ exp(.x) / 1000)
  )

# Create effects plot
plt_gam_flow_yr_eff <- df_gam_flow_yr_eff %>% 
  ggplot(aes(x = Flow, y = predicted)) +
  geom_point(
    data = df_chla_c,
    aes(y = Chla, color = StationCode),
    alpha = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
  facet_wrap(vars(Year_fct), scales = "free") +
  theme_bw() +
  labs(
    x = "Flow (cfs)",
    y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
  ) +
  scale_x_continuous(breaks = breaks_extended(6)) +
  scale_color_viridis_d(name = "Station", option = "C") +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.position = c(0.8, 0.3)
  )

plt_gam_flow_yr_eff
```

## Station by Year Contrasts

```{r gam sflow lag2 contrasts station by year, warning = FALSE}
# Estimated marginal means for station by year
em_gam_sta_yr <- emmeans(m_gam_sflow_lag2, ~ StationCode | Year_fct)

# Tukey post-hoc contrasts
pairs(em_gam_sta_yr)

# Create table of contrasts and convert it to a tibble for plot
df_gam_sta_yr <- em_gam_sta_yr %>% 
  cld(sort = FALSE, Letters = letters) %>% 
  as_tibble() %>% 
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    # back transform log-transformed results
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
  ) %>% 
  # Add min and max values of observed data to the Tukey post-hoc results and
    # calculate vertical positioning of letters
  left_join(
    df_chla_c %>% 
      summarize(
        max_val = max(Chla),
        min_val = min(Chla),
        .by = Year_fct
    ), 
    by = join_by(Year_fct)
  ) %>% 
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>% 
  group_by(Year_fct) %>% 
  mutate(max_val = max(max_val)) %>% 
  ungroup() %>% 
  mutate(y_pos = max_val + (max_val - min_val) / 10) %>% 
  select(
    StationCode,
    Year_fct,
    emmean,
    lower.CL,
    upper.CL,
    group,
    y_pos
  )

# Create boxplot showing Tukey post-hoc results
plt_gam_sta_yr <- df_gam_sta_yr %>% 
  ggplot(
    aes(
      x = StationCode,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL
    )
  ) +
  geom_boxplot(
    data = df_chla_c,
    aes(x = StationCode, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos, label = group), size = 3.5) +
  facet_wrap(vars(Year_fct), scales = "free_y") +
  xlab("Station") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw()

plt_gam_sta_yr
```

## Year by Station Contrasts

```{r gam sflow lag2 contrasts year by station, warning = FALSE}
# Estimated marginal means for year by station
em_gam_yr_sta <- emmeans(m_gam_sflow_lag2, ~ Year_fct | StationCode)

# Tukey post-hoc contrasts
pairs(em_gam_yr_sta)

# Create table of contrasts and convert it to a tibble for plot
df_gam_yr_sta <- em_gam_yr_sta %>% 
  cld(sort = FALSE, Letters = letters) %>% 
  as_tibble() %>% 
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    # back transform log-transformed results
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
  ) %>% 
  # Add min and max values of observed data to the Tukey post-hoc results and
    # calculate vertical positioning of letters
  left_join(
    df_chla_c %>% 
      summarize(
        max_val = max(Chla),
        min_val = min(Chla),
        .by = StationCode
    ), 
    by = join_by(StationCode)
  ) %>% 
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>% 
  group_by(StationCode) %>% 
  mutate(max_val = max(max_val)) %>% 
  ungroup() %>% 
  mutate(y_pos = max_val + (max_val - min_val) / 10) %>% 
  select(
    StationCode,
    Year_fct,
    emmean,
    lower.CL,
    upper.CL,
    group,
    y_pos
  )

# Create boxplot showing Tukey post-hoc results
plt_gam_yr_sta <- df_gam_yr_sta %>% 
  ggplot(
    aes(
      x = Year_fct,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL
    )
  ) +
  geom_boxplot(
    data = df_chla_c,
    aes(x = Year_fct, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos, label = group), size = 3.5) +
  facet_wrap(vars(StationCode), scales = "free_y") +
  xlab("Year") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw()

plt_gam_yr_sta
```

# Export Results

```{r export results}
# Define file path for figures for the manuscript
fp_figures <- here("manuscript_synthesis/results/figures")

# Export effects plot of flow by station
ggsave(
  file.path(fp_figures, "chl_flow_effects_plot_by_station.jpg"),
  plot = plt_gam_flow_sta_eff,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export effects plot of flow by year
ggsave(
  file.path(fp_figures, "chl_flow_effects_plot_by_year.jpg"),
  plot = plt_gam_flow_yr_eff,
  width = 7,
  height = 5,
  units = "in"
)

# Export boxplot of station contrasts by year
ggsave(
  file.path(fp_figures, "chl_station_contrasts_by_year.jpg"),
  plot = plt_gam_sta_yr,
  width = 7,
  height = 5,
  units = "in"
)

# Export boxplot of station contrasts by year
ggsave(
  file.path(fp_figures, "chl_year_contrasts_by_station.jpg"),
  plot = plt_gam_yr_sta,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export ANOVA table for GAM
df_an_gam_sflow_lag2 %>% write_csv(here("manuscript_synthesis/results/tables/chl_GAM_anova.csv"))
```

