---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Final results from selected models"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Look at results of the preferred models during model selection: separate GAMs for each station (RD22, STTD, LIB, and RVB) using an interaction between weekly average flow and Year with a cyclic penalized cubic regression spline smooth term for week number to account for seasonality. Create summary figures and tables for the manuscript. The model selection process and diagnostics can be found in the [manuscript_synthesis/notebooks/rtm_chl_models_flow_weekly_avg.Rmd](https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html) document.

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(mgcv)
library(ggeffects)
library(emmeans)
library(multcomp)
library(gratia)
library(patchwork)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

```{r functions and global code}
# Calculate effects of flow on chlorophyll for each year holding the
  # non-focal variables constant - marginal effects/adjusted predictions
calc_flow_yr_marg_eff <- function(df_data, m_gam) {
  # Calculate min and max flows for each year to narrow down x-axis in the
    # effects plots
  df_flow_yr_summ <- df_data %>% 
    summarize(
      Flow_min = min(Flow),
      Flow_max = max(Flow),
      .by = c(Year_fct)
    ) %>% 
    mutate(
      Flow_buffer = (Flow_max - Flow_min) * 0.05,
      Flow_min = Flow_min - Flow_buffer,
      Flow_max = Flow_max + Flow_buffer
    ) %>% 
    select(-Flow_buffer)
  
  # Calculate marginal effects
  as.data.frame(
    predict_response(m_gam, terms = c("Flow", "Year_fct"), margin = "marginalmeans"),
    terms_to_colnames = TRUE
  ) %>% 
  as_tibble() %>% 
  # Narrow down range of flow values for each station
  left_join(df_flow_yr_summ, by = join_by(Year_fct)) %>% 
  filter(Flow >= Flow_min & Flow <= Flow_max) %>% 
  transmute(
    Year_fct,
    Flow,
    # Back calculate model fits and confidence levels
    across(c(predicted, conf.low, conf.high), \(x) exp(x) / 1000)
  )
}

# Create effects plot of flow on chlorophyll by year
plot_eff_flow_by_yr <- function(df_eff, df_data) {
  df_eff %>% 
    ggplot(aes(x = Flow, y = predicted)) +
    geom_point(data = df_data, aes(y = Chla), alpha = 0.6) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
    facet_wrap(vars(Year_fct), scales = "free_x", nrow = 1) +
    theme_bw() +
    labs(
      title = df_data$StationCode[1],
      x = "Flow (cfs)",
      y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
    ) +
    scale_x_continuous(labels = label_comma(), breaks = breaks_extended()) +
    theme(
      axis.text = element_text(size = 8),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
    )
}

# Create boxplots of chlorophyll by year showing Tukey post-hoc results
boxplot_cont_yr <- function(em_obj, df_data, xlab = TRUE) {
  # Create table of contrasts and convert it to a tibble for plot
  df_yr_cont <- em_obj %>% 
    cld(sort = FALSE, Letters = letters) %>% 
    as_tibble() %>% 
    mutate(
      group = str_remove_all(.group, fixed(" ")),
      # back transform log-transformed results
      across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
    ) %>% 
    # Add min and max values of observed data to the Tukey post-hoc results and
      # calculate vertical positioning of letters
    left_join(
      df_data %>% 
        summarize(
          max_val = max(Chla),
          min_val = min(Chla),
          .by = Year_fct
      ), 
      by = join_by(Year_fct)
    ) %>% 
    mutate(
      max_val = if_else(upper.CL > max_val, upper.CL, max_val),
      y_pos = max_val + (max_val - min_val) / 10,
      # Make all post-hoc contrast letters at same height equal to max of all
      y_pos = max(y_pos)
    ) %>% 
    select(
      Year_fct,
      emmean,
      lower.CL,
      upper.CL,
      group,
      y_pos
    )
  
  # Create boxplot showing Tukey post-hoc results
  df_yr_cont %>% 
    ggplot(
      aes(
        x = Year_fct,
        y = emmean,
        ymin = lower.CL,
        ymax = upper.CL
      )
    ) +
    geom_boxplot(
      data = df_data,
      aes(x = Year_fct, y = Chla),
      inherit.aes = FALSE
    ) +
    geom_crossbar(
      color = "grey82", 
      fill = "grey", 
      alpha = 0.7, 
      linewidth = 0.1
    ) +
    geom_point(color = "red") +
    geom_text(aes(y = y_pos, label = group), size = 3) +
    labs(
      title = df_data$StationCode[1],
      x = "Year",
      y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
    ) +
    theme_bw() +
    theme(axis.text = element_text(size = 8))
}

# Create smooth of Week plot
plot_smooth_week <- function(m_gam, station) {
  ls_plt <- draw(m_gam, residuals = TRUE, rug = FALSE, wrap = FALSE)
  
  # Extract plot from list and format
  chuck(ls_plt, 1) + 
    theme_bw() + 
    labs(title = station, caption = NULL)
}
```

# Import Data

```{r import data}
# Define file path for processed data
fp_data <- here("manuscript_synthesis/data/processed")

# Import weekly average water quality data
df_wq <- readRDS(file.path(fp_data, "wq_week_avg_2013-2019.rds"))

# Import weekly average flow data
df_flow <- readRDS(file.path(fp_data, "flow_week_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare chla data}
# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

# We will use LIS flow data as a proxy for STTD
df_flow_c <- df_flow %>% mutate(StationCode = if_else(StationCode == "LIS", "STTD", StationCode))

# Prepare chlorophyll and flow data for analysis
ls_chla <- df_wq %>% 
  select(StationCode, Year, Week, Chla) %>% 
  drop_na(Chla) %>% 
  # Filter to only include representative stations for 4 habitat types (RD22,
    # STTD, LIB, RVB) and only include years 2015-2019
  filter(
    StationCode %in% sta_order,
    Year %in% 2015:2019
  ) %>% 
  # Join flow data to chlorophyll data
  left_join(df_flow_c, by = join_by(StationCode, Year, Week)) %>% 
  # Remove all NA flow values
  drop_na(Flow) %>% 
  mutate(
    # Scale and log transform chlorophyll values
    Chla_log = log(Chla * 1000),
    # Add a column for Year as a factor for the models
    Year_fct = factor(Year)
  ) %>% 
  arrange(StationCode, Year, Week) %>% 
  split(~ StationCode)

# Add 2 lag variables for chlorophyll to be used in the LIB model to help with
  # serial autocorrelation
df_chla_lib <- ls_chla$LIB %>% 
  group_by(Year) %>% 
  mutate(
    lag1 = lag(Chla_log),
    lag2 = lag(Chla_log, 2)
  ) %>% 
  ungroup() %>% 
  drop_na(lag1, lag2)

# Add 1 lag variables for chlorophyll to be used in the RVB model to help with
  # serial autocorrelation
df_chla_rvb <- ls_chla$RVB %>% 
  group_by(Year) %>% 
  mutate(lag1 = lag(Chla_log)) %>% 
  ungroup() %>% 
  drop_na(lag1)

# Combine all data sets into a nested data frame to be used for analysis
ndf_chla <- 
  tibble(
    StationCode = sta_order,
    df_data = list(ls_chla$RD22, ls_chla$STTD, df_chla_lib, df_chla_rvb)
  ) %>% 
  # Apply factor order to StationCode
  mutate(StationCode = factor(StationCode, levels = sta_order)) %>% 
  arrange(StationCode)
```

# Create GAM models

```{r create gam models, warning = FALSE}
# Define base model formula
f_base <- "Chla_log ~ Year_fct * Flow + s(Week, bs = 'cc', k = 5)"

# Create gam models for each station
ndf_gam <- ndf_chla %>% 
  mutate(
    str_formula = case_when(
      StationCode == "LIB" ~ paste(f_base, "+ lag1 + lag2"),
      StationCode == "RVB" ~ paste(f_base, "+ lag1"),
      TRUE ~ f_base
    ),
    model = map2(
      str_formula,
      df_data,
      \(x, y) gam(
        formula = as.formula(x),
        data = y,
        method = "REML", 
        knots = list(week = c(0, 52))
      )
    )
  )
```

# Model Results

## ANOVA Tables

```{r gam anova tables, warning = FALSE}
# Create ANOVA tables for each GAM model and extract elements for export
ndf_gam_anova <- ndf_gam %>% 
  mutate(
    an_gam = map(model, anova),
    # Parametric terms
    df_parametric = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "pTerms.table"), rownames = "Term")
    ),
    # Smooth terms
    df_smooth = map(
      an_gam, 
      \(x) as_tibble(chuck(x, "s.table"), rownames = "Term") %>% 
        rename(df = edf) %>% 
        select(-Ref.df)
    ),
    df_anova_tbl = map2(df_parametric, df_smooth, bind_rows)
  )

# Combine ANOVA tables and format for export
df_gam_anova_tbl_c <- ndf_gam_anova %>% 
  select(StationCode, df_anova_tbl) %>% 
  unnest(df_anova_tbl) %>% 
  transmute(
    Station = StationCode,
    Term,
    across(
      c("df", "F"),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    ),
    p_value = if_else(
      `p-value` < 0.001, 
      "< 0.001", 
      formatC(`p-value`, digits = 3, format = "f")
    )
  )

df_gam_anova_tbl_c
```

## Effect of Flow by Year

### Effect Plots

```{r gam flow year effect plots, warning = FALSE, fig.height = 9}
# Create effects plot of flow on chlorophyll by year for each station
ndf_gam_eff_flow_plt <- ndf_gam %>% 
  mutate(
    df_marg_eff = map2(df_data, model, calc_flow_yr_marg_eff),
    plt_marg_eff = map2(df_marg_eff, df_data, plot_eff_flow_by_yr)
  )

# Combine effects plots together using patchwork
plt_gam_eff_flow <- 
  wrap_plots(ndf_gam_eff_flow_plt$plt_marg_eff, ncol = 1) + 
  plot_layout(axes = "collect")

plt_gam_eff_flow
```

### Slopes and p-values

```{r gam flow year effect stats}
# Calculate slopes and p-values of the modeled effects of flow on chlorophyll by
  # year for each station
ndf_gam_eff_flow_stat <- ndf_gam %>% 
  mutate(
    em_gam_flow = map(model, \(x) emtrends(x, ~ Year_fct, var = "Flow")),
    df_stat = map(em_gam_flow, \(x) as_tibble(test(x)))
  )

# Combine slopes and p-values into one data frame for export
df_gam_eff_flow_stat <- ndf_gam_eff_flow_stat %>% 
  select(StationCode, df_stat) %>% 
  unnest(df_stat) %>% 
  # Format for export
  rename(
    Station = StationCode,
    Year = Year_fct,
    Flow_slope = Flow.trend,
    t_ratio = t.ratio,
    p_value = p.value
  ) %>% 
  mutate(
    p_value = if_else(
      p_value < 0.001, 
      "< 0.001", 
      formatC(p_value, digits = 3, format = "f")
    ),
    across(
      where(is.numeric),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    )
  )

df_gam_eff_flow_stat
```

## Year Contrasts

### Boxplots

```{r gam year contrast boxplots, warning = FALSE}
# Calculate estimated marginal means of year for each station
ndf_gam_cont_yr <- ndf_gam %>% 
  mutate(em_gam_yr = map(model, \(x) emmeans(x, ~ Year_fct)))

# Create boxplots showing Tukey post-hoc results of year contrasts for each station
ndf_gam_cont_yr_plt <- ndf_gam_cont_yr %>% 
  mutate(plt_cont_yr = map2(em_gam_yr, df_data, boxplot_cont_yr))

# Combine boxplots together using patchwork
plt_gam_cont_yr <- 
  wrap_plots(ndf_gam_cont_yr_plt$plt_cont_yr, ncol = 2) + 
  plot_layout(axes = "collect")

plt_gam_cont_yr
```

### Statistics

```{r gam year contrast stats}
# Generate pairwise Tukey post-hoc year contrasts for each station
ndf_gam_cont_yr_stat <- ndf_gam_cont_yr %>% 
  mutate(df_stat = map(em_gam_yr, \(x) as_tibble(pairs(x))))

# Combine pairwise Tukey post-hoc year contrasts into one data frame for export
df_gam_cont_yr_stat <- ndf_gam_cont_yr_stat %>% 
  select(StationCode, df_stat) %>% 
  unnest(df_stat) %>% 
  # Format for export
  rename(
    Station = StationCode,
    Contrast = contrast,
    Estimate = estimate,
    t_ratio = t.ratio,
    p_value = p.value
  ) %>% 
  mutate(
    p_value = if_else(
      p_value < 0.001, 
      "< 0.001", 
      formatC(p_value, digits = 3, format = "f")
    ),
    across(
      where(is.numeric),
      \(x) if_else(
        abs(x) < 0.01,
        formatC(signif(x, 3), digits = 2, format = "e"),
        formatC(signif(x, 3), digits = 3, format = "fg", flag = "#")
      )
    )
  )

print(df_gam_cont_yr_stat, n = Inf)
```

## s(Week) Plots

```{r gam sWeek plots, warning = FALSE}
# Create smooth of Week plots for each station
ndf_gam_s_week_plt <- ndf_gam %>% 
  mutate(plt_s_week = map2(model, StationCode, plot_smooth_week))

# Combine smooth of Week plots together using patchwork
plt_gam_s_week <- 
  wrap_plots(ndf_gam_s_week_plt$plt_s_week, ncol = 2) + 
  plot_layout(axis_titles = "collect") +
  plot_annotation(caption = "Basis: Cyclic CRS")

plt_gam_s_week
```

# Export Results

```{r export results}
# Define file path for figures and tables for the manuscript
fp_figures <- here("manuscript_synthesis/results/figures")
fp_tables <- here("manuscript_synthesis/results/tables")

# Export effects plot of flow by year for each station
ggsave(
  file.path(fp_figures, "chl_flow_eff_plot_by_year_each_sta.jpg"),
  plot = plt_gam_eff_flow,
  width = 7,
  height = 9,
  units = "in"
)

ggsave(
  file.path(fp_figures, "chl_flow_eff_plot_by_year_each_sta.pdf"),
  plot = plt_gam_eff_flow,
  width = 7,
  height = 9,
  units = "in"
)

# Export boxplots of year contrasts for each station
ggsave(
  file.path(fp_figures, "chl_year_contrasts_each_sta.jpg"),
  plot = plt_gam_cont_yr,
  width = 6.5,
  height = 5,
  units = "in"
)

ggsave(
  file.path(fp_figures, "chl_year_contrasts_each_sta.pdf"),
  plot = plt_gam_cont_yr,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export effects plots of smooth of Week
ggsave(
  file.path(fp_figures, "chl_smooth_week.jpg"),
  plot = plt_gam_s_week,
  width = 6.5,
  height = 5,
  units = "in"
)

ggsave(
  file.path(fp_figures, "chl_smooth_week.pdf"),
  plot = plt_gam_s_week,
  width = 6.5,
  height = 5,
  units = "in"
)

# Export combined ANOVA tables for GAMs
df_gam_anova_tbl_c %>% 
  mutate(across(c("df", "F", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "chl_GAM_anova.csv"))

# Export slopes and p-values of the modeled effects of flow by year for each station
df_gam_eff_flow_stat %>% 
  mutate(across(c("Flow_slope", "SE", "df", "t_ratio", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "chl_GAM_eff_flow_stat.csv"))

# Export pairwise Tukey post-hoc year contrasts
df_gam_cont_yr_stat %>% 
  mutate(across(c("Estimate", "SE", "df", "t_ratio", "p_value"), ~ paste0(.x, "##"))) %>% 
  write_csv(file.path(fp_tables, "chl_GAM_year_contrasts_stat.csv"))
```

