---
title: "Raw Data Plots - RYI"
author: "Amanda Maguire and Dave Bosworth"
date: "11/23/2020"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Purpose

Create plots of the raw continuous water quality data for the RYI station. The data in these plots has been cleaned and formatted, but not inspected for suspect values. The purpose of these plots is to allow for us to visually inspect the data for suspect or unreliable values during the study period (2011-2019). The final QA/QC steps for this data can be found in the rtm_QA_and_combine_data_all.Rmd file.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
```

```{r load functions, message = FALSE, warning = FALSE}
# Define directory for local repository
fp_local_repo <- "C:/Repositories/ND-FASTR/"

# Source global WQ functions
source(paste0(fp_local_repo, "global_ndfa_funcs.R"))
source(paste0(fp_local_repo, "Water_Quality/global_wq_funcs.R"))

# Source continuous WQ raw data plot functions - look at this R script file for greater detail 
  # regarding the functions used in this Rmd file
source(paste0(fp_local_repo, "Water_Quality/Continuous_WQ/Plots_Raw_Data/rtm_raw_plots_funcs.R"))
```

```{r define file paths}
# Define relative file path for processed continuous WQ data file for RYI
fp_rel_rtm_data <- 
  "WQ_Subteam/Processed_Data/Continuous/Filtered_Dates/RTM_OUTPUT_RYI_formatted_filt.csv"

# Define absolute file path
fp_abs_rtm_data <- ndfa_abs_sp_path(fp_rel_rtm_data)
```

```{r set system tz as PST}
# Set System Timezone as "Etc/GMT+8" (PST) to make it consistent with the data 
Sys.setenv(TZ = "Etc/GMT+8")
```

# Import and Prepare Data

```{r import data}
# Import data
df_orig <- import_rtm_data(fp_abs_rtm_data, 10)
```

```{r prepare data}
# Clean original data and nest it by year and parameter for plots of individual years 
  # for each parameter
df_clean_nest_indiv_yr <- df_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # pivot data longer for easier plotting
  restructure_df_long() %>% 
  # add a year variable
  mutate(yr = year(DateTime)) %>% 
  # remove data with NA values
  filter(!is.na(value)) %>% 
  # group nest by year and parameter
  nest(data = -c(yr, parameter)) %>%
  # add extra datetime values for gaps of missing data greater than a half a day
  mutate(
    data_full_dt = if_else(
      parameter != "FlowTF",
      map(data, add_extra_dt_halfday, time_int = "15min"),
      map(data, add_extra_dt_halfday, time_int = "1hr")
    )
  ) %>% 
  select(-data)

# Create another dataframe nested by parameter for the plots of all years of data 
  # for each parameter
df_clean_nest_all_yrs <- df_clean_nest_indiv_yr %>% 
  unnest(data_full_dt) %>% 
  # Add extra datetime values so that the lines connecting years are eliminated in the plots 
  add_extra_dt_yr() %>% 
  select(-yr) %>% 
  nest(data_all_yrs = -parameter)
```

# Create Plots and Summary Tables

```{r create plots and summary tables, warning = FALSE}
# Create interactive plotly plots of individual years for each parameter
df_plots_indiv_yr <- df_clean_nest_indiv_yr %>% 
  mutate(plot = map2(data_full_dt, parameter, create_ts_plotly))

# Create static ggplots of all years of data for each parameter
df_plots_all_yrs <- df_clean_nest_all_yrs %>% 
  mutate(plot = map2(data_all_yrs, parameter, create_ts_ggplot))

# Create summary tables with counts of Qual Codes by year for each parameter
df_qual_code_summ <- df_clean_nest_indiv_yr %>% 
  unnest(data_full_dt) %>% 
   # remove data with NA values
  filter(!is.na(value)) %>%
  count(parameter, yr, qual_code) %>% 
  arrange(qual_code) %>% 
  pivot_wider(names_from = qual_code, values_from = n) %>% 
  rename(Year = yr) %>% 
  arrange(Year) %>% 
  nest(data = -parameter) %>% 
  arrange(parameter) %>% 
  mutate(
    data_kable = map(
      data,
      ~kable(
        .x, 
        format = "html", 
        align = "l",
        caption = "Count of Qual Codes"
      ) %>% 
        kable_styling(
          bootstrap_options = "bordered", 
          position = "left",
          full_width = FALSE
        )
    )
  )
```

# Raw Data Plots {.tabset .tabset-pills}

## Chla {.tabset .tabset-pills}

```{r filter indiv chla plots, include = FALSE}
filt_param <- "Chla"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot chla all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2015

```{r plot chla 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot chla 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot chla 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot chla 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


## DO {.tabset .tabset-pills}

```{r filter indiv do plots, include = FALSE}
filt_param <- "DO"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot do all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2015

```{r plot do 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot do 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot do 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot do 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


## fDOM {.tabset .tabset-pills}

```{r filter indiv fdom plots, include = FALSE}
filt_param <- "fDOM"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot fdom all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2015

```{r plot fdom 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot fdom 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot fdom 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot fdom 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


```{r print flow plots, include = FALSE}
# Create a list of plot templates to use for the flow plots
produce_plots_flow <- list(
  knit_expand(file = "rtm_raw_plots_template_2011.Rmd", parameter = "Flow"),
  knit_expand(file = "rtm_raw_plots_template_2011.Rmd", parameter = "FlowTF")
)
```

`r knit(text = unlist(produce_plots_flow))`


## NitrateNitrite {.tabset .tabset-pills}

```{r filter indiv nitrate_nitrite plots, include = FALSE}
filt_param <- "NitrateNitrite"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot nitrate_nitrite all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2013

```{r plot nitrate_nitrite 2013, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2013) %>% 
  print_plot()
```

### 2014

```{r plot nitrate_nitrite 2014, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2014) %>% 
  print_plot()
```

### 2015

```{r plot nitrate_nitrite 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot nitrate_nitrite 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot nitrate_nitrite 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot nitrate_nitrite 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


## pH {.tabset .tabset-pills}

```{r filter indiv ph plots, include = FALSE}
filt_param <- "pH"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot ph all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2015

```{r plot ph 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot ph 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot ph 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot ph 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


## SpCnd {.tabset .tabset-pills}

```{r filter indiv spcnd plots, include = FALSE}
filt_param <- "SpCnd"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot spcnd all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2015

```{r plot spcnd 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot spcnd 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot spcnd 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot spcnd 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


## Turbidity {.tabset .tabset-pills}

```{r filter indiv turbidity plots, include = FALSE}
filt_param <- "Turbidity"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot turbidity all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2011

```{r plot turbidity 2011, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2011) %>% 
  print_plot()
```

### 2012

```{r plot turbidity 2012, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2012) %>% 
  print_plot()
```

### 2013

```{r plot turbidity 2013, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2013) %>% 
  print_plot()
```

### 2014

```{r plot turbidity 2014, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2014) %>% 
  print_plot()
```

### 2015

```{r plot turbidity 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot turbidity 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot turbidity 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot turbidity 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


## WaterTemp {.tabset .tabset-pills}

```{r filter indiv water temp plots, include = FALSE}
filt_param <- "WaterTemp"

df_plots_indiv_yr_filt <- filter(df_plots_indiv_yr, parameter == filt_param)
```

### All Years

```{r plot water temp all yrs, echo = FALSE}
df_plots_all_yrs %>% 
  filter(parameter == filt_param) %>% 
  print_plot()
```

### 2011

```{r plot water temp 2011, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2011) %>% 
  print_plot()
```

### 2012

```{r plot water temp 2012, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2012) %>% 
  print_plot()
```

### 2013

```{r plot water temp 2013, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2013) %>% 
  print_plot()
```

### 2015

```{r plot water temp 2015, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2015) %>% 
  print_plot()
```

### 2016

```{r plot water temp 2016, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2016) %>% 
  print_plot()
```

### 2018

```{r plot water temp 2018, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2018) %>% 
  print_plot()
```

### 2019

```{r plot water temp 2019, echo = FALSE}
df_plots_indiv_yr_filt %>% 
  filter(yr == 2019) %>% 
  print_plot()
```


# Summary of Qual Codes {.tabset .tabset-pills}

```{r print summ kables, echo = FALSE, results = "asis"}
for (i in 1:nrow(df_qual_code_summ)) {
  # Create heading for each parameter
  cat("## ", as.character(df_qual_code_summ$parameter[i]), "\n\n")
  # Print kable
  print(df_qual_code_summ$data_kable[[i]])
  cat("\n\n")
}
```

