---
title: "Raw Data Plots - I80"
author: "Dave Bosworth & Amanda Maguire"
date: "11/6/2020"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Create plots of the raw continuous water quality data for the I-80 station. The data in these plots has been cleaned and formatted, but not inspected for suspect values. The purpose of these plots is to allow for us to visually inspect the data for suspect or unreliable values during the study period (2011-2019). The final QA/QC steps for this data can be found in the xxx.Rmd file.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(knitr)
```

```{r load functions, message = FALSE, warning = FALSE}
# Source global WQ functions
source("../../../global_ndfa_funcs.R")

# Source continuous WQ raw data plot functions - look at this R script file for greater detail 
  # regarding the functions used in this Rmd file
source("rtm_raw_plots_funcs.R")
```

```{r define file paths}
# Define relative file path for processed continuous WQ data file for I-80
fp_rel_rtm_data <- 
  "WQ_Subteam/Processed_Data/Continuous/Filtered_Dates/RTM_OUTPUT_I80_formatted_filt.csv"

# Define absolute file path
fp_abs_rtm_data <- ndfa_abs_sp_path(fp_rel_rtm_data)
```

```{r set system tz as PST}
# Set System Timezone as "Etc/GMT+8" (PST) to make it consistent with the data 
Sys.setenv(TZ = "Etc/GMT+8")
```

# Import and Prepare Data

```{r import data}
# Import data
df_orig <- import_rtm_data(fp_abs_rtm_data, 6)
```

```{r prepare data, message = FALSE}
# Clean original data
df_clean <- df_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # pivot data longer for easier plotting
  restructure_df_long() %>% 
  # add a year variable
  mutate(yr = year(DateTime)) %>% 
  # remove data before the first year when data was collected for each parameter so that 
    # the x-axis is centered correctly on the plots of all years of data
  trim_to_first_yr() %>% 
  # Add extra datetime values so that the lines connecting years are eliminated in the plots 
    # of all years of data
  add_extra_dt_values()
```

# Raw Data Plots {.tabset .tabset-pills}

```{r print plots, include = FALSE}
# Find the starting year for the period of record for each parameter
df_start_yr <- df_clean %>% 
  group_by(parameter) %>% 
  summarize(start_yr = min(yr)) %>% 
  arrange(parameter)

# Create a list of plot templates to use for each parameter
produce_plots <- 
  map2(
    df_start_yr$start_yr,
    df_start_yr$parameter,
    ~knit_expand(file = paste0("rtm_raw_plots_template_", .x, ".Rmd"), parameter = .y)
  )
```

`r knit(text = unlist(produce_plots))`

