---
title: "Raw Data Plots - RMB"
author: "Dave Bosworth & Amanda Maguire"
date: "9/03/2020"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Create plots of the raw continuous water quality data for the RMB station. The data in these plots has been cleaned and formatted, but not inspected for suspect values. The purpose of these plots is to allow for us to visually inspect the data for suspect or unreliable values during the study period (2011-2019). The final QA/QC steps for this data can be found in the xxx.Rmd file.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

```{r load functions, message = FALSE, warning = FALSE}
# Source global WQ functions
source("Water_Quality/global_wq_funcs.R")

# Source continuous WQ raw data plot functions
source("Water_Quality/Continuous_WQ/Plots_Raw_Data/rtm_raw_plots_funcs.R")
```

```{r define file paths}
# Define main NDFA file path for WQ subteam (assumes synced with SharePoint)
fp_fastr <- "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/"

# Define relative file path for processed continuous WQ data file for RMB
fp_rel_rtm_data <- paste0(fp_fastr, "Processed_Data/Continuous/RTM_OUTPUT_RMB_formatted.csv")

# Define absolute file path
fp_abs_rtm_data <- get_abs_path(fp_rel_rtm_data)
```

```{r set system tz as PST}
# Set System Timezone as "Etc/GMT+8" (PST) to make it consistent with the data 
Sys.setenv(TZ = "Etc/GMT+8")
```

# Import and Prepare Data

```{r import data}
# Import data
data_orig <- import_rtm_data(fp_abs_rtm_data, 6)
```

```{r prepare data, message = FALSE}
# Clean original data
data_clean <- data_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # pivot data longer for easier plotting
  select(-c(StationCode, ends_with("_Qual"))) %>% 
  pivot_longer(
    cols = -DateTime,
    names_to = "parameter",
    values_to = "value"
  ) %>% 
  # add yr and y-axis label variables
  mutate(yr = year(DateTime)) %>% 
  add_yaxis_var(parameter)
```

# Raw Data Plots {.tabset .tabset-pills}

## Chla

```{r plot chla data, echo = FALSE, warning = FALSE}
data_clean %>% 
  filter(parameter == "Chla") %>% 
  create_ts_plot(.$yaxis_lab[1])
```

## DO

```{r plot do data, echo = FALSE, warning = FALSE}
data_clean %>% 
  filter(parameter == "DO") %>% 
  create_ts_plot(.$yaxis_lab[1])
```

## pH

```{r plot pH data, echo = FALSE, warning = FALSE}
data_clean %>% 
  filter(parameter == "pH") %>% 
  create_ts_plot(.$yaxis_lab[1])
```

## SpCnd

```{r plot SpCnd data, echo = FALSE, warning = FALSE}
data_clean %>% 
  filter(parameter == "SpCnd") %>% 
  create_ts_plot(.$yaxis_lab[1])
```

## Turbidity

```{r plot turb data, echo = FALSE, warning = FALSE}
data_clean %>% 
  filter(parameter == "Turbidity") %>% 
  create_ts_plot(.$yaxis_lab[1])
```

## WaterTemp

```{r plot wt data, echo = FALSE, warning = FALSE}
data_clean %>% 
  filter(parameter == "WaterTemp") %>% 
  create_ts_plot(.$yaxis_lab[1])
```
