---
title: "Raw Data Plots - LIS Year-round"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Create plots of the raw continuous water quality data collected year-round at the LIS station. All data in these plots has been cleaned and formatted, but not all of it has been inspected for suspect values. Data has already been QA'ed and validated for the time period during the flow actions and 45 days before and after. Data outside of these time periods is still considered raw and not validated. The purpose of these plots is to allow for us to visually inspect the raw data for suspect or unreliable values.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(knitr)
```

```{r load functions, message = FALSE, warning = FALSE}
# Define directory for local repository
fp_local_repo <- "C:/Repositories/ND-FASTR"

# Source global WQ functions
source(file.path(fp_local_repo, "global_ndfa_funcs.R"))
source(file.path(fp_local_repo, "Water_Quality/global_wq_funcs.R"))

# Source continuous WQ raw data plot functions - look at this R script file for greater detail 
  # regarding the functions used in this Rmd file
source(
  file.path(
    fp_local_repo, 
    "Water_Quality/Continuous_WQ/Plots_Raw_Data/rtm_raw_plots_year-round_funcs.R"
  )
)
```

```{r define file paths}
# Define relative file path for processed continuous WQ data file for LIS - year-round data
fp_rel_rtm_proc_all <- 
  "WQ_Subteam/Processed_Data/Continuous/All_Dates/RTM_OUTPUT_LIS_formatted_all.csv"

# Define relative file path for file containing all QA'ed and cleaned continuous WQ data
fp_rel_rtm_data_qa <- "WQ_Subteam/Processed_Data/Continuous/RTM_INPUT_all_2021-04-20.csv"

# Define absolute file paths
fp_abs_rtm_proc_all <- ndfa_abs_sp_path(fp_rel_rtm_proc_all)
fp_abs_rtm_data_qa <- ndfa_abs_sp_path(fp_rel_rtm_data_qa)
```

```{r set system tz as PST}
# Set System Timezone as "Etc/GMT+8" (PST) to make it consistent with the data 
Sys.setenv(TZ = "Etc/GMT+8")
```

# Import Data

```{r import data}
# Import processed continuous WQ data for LIS - year-round data
rtm_wq_proc_orig <- import_rtm_data(fp_abs_rtm_proc_all, 7)

# Import QA'ed and cleaned continuous WQ data
rtm_wq_qa_orig <- import_rtm_data(fp_abs_rtm_data_qa, 10)
```

# Prepare Data

```{r prepare raw data yr round}
# Prepare data for plotting
rtm_wq_proc_clean <- rtm_wq_proc_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # only keep a subset of the parameters (Chla, DO, pH, SpCnd, Turbidity, WaterTemp)
  select(DateTime, Chla, DO, pH, SpCnd, Turbidity, WaterTemp) %>%  
  # pivot data longer for easier plotting
  pivot_longer(
    cols = where(is.numeric),
    names_to = "Parameter",
    values_to = "Value"
  ) %>% 
  # remove data with NA values
  filter(!is.na(Value)) %>% 
  # complete all date-time stamps for each Parameter
  group_by(Parameter) %>% 
  complete(DateTime = seq.POSIXt(min(DateTime), max(DateTime), by = "15 min")) %>% 
  ungroup() %>% 
  # Add variable to identify which data set it came from
  mutate(Status = "Raw")
```

```{r prepare qa data}
# Prepare QA'ed and cleaned continuous WQ data to be joined to the raw year-round data
rtm_wq_qa_clean <- rtm_wq_qa_orig %>% 
  # parse date-time variable and define tz as PST
  mutate(DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8")) %>% 
  # only include LIS
  filter(StationCode == "LIS") %>% 
  # only keep a subset of the parameters (Chla, DO, pH, SpCnd, Turbidity, WaterTemp)
  select(DateTime, Chla, DO, pH, SpCnd, Turbidity, WaterTemp) %>%
  # pivot data longer for easier plotting
  pivot_longer(
    cols = where(is.numeric),
    names_to = "Parameter",
    values_to = "Value"
  ) %>% 
  # remove data with NA values
  filter(!is.na(Value)) %>% 
  # add a temporary year variable
  mutate(yr_temp = year(DateTime)) %>% 
  # complete all date-time stamps for each Parameter and Year combination
  group_by(Parameter, yr_temp) %>% 
  complete(DateTime = seq.POSIXt(min(DateTime), max(DateTime), by = "15 min")) %>% 
  ungroup() %>% 
  # add variable to identify which data set it came from
  mutate(Status = "QA") %>% 
  # remove temporary year variable
  select(-yr_temp)
```

```{r join data together}
# Remove data from the raw year-round data frame that already exists in the QA'ed 
  # and cleaned data frame
rtm_wq_proc_extra <- anti_join(rtm_wq_proc_clean, rtm_wq_qa_clean, by = c("Parameter", "DateTime"))

# Add in the QA'ed and cleaned data
rtm_wq_all <- bind_rows(rtm_wq_proc_extra, rtm_wq_qa_clean)
```

```{r prepare all data}
# Finish preparing all joined data for plotting
rtm_wq_all_clean <- rtm_wq_all %>% 
  # add a year variable
  mutate(Year = year(DateTime)) %>% 
  # remove data with NA values
  filter(!is.na(Value)) %>% 
  # sort data by Parameter and DateTime
  arrange(Parameter, DateTime) %>% 
  # group nest by Parameter and Year
  nest(df_data = -c(Parameter, Year)) %>%
  # add extra datetime values for gaps of missing data greater than a half a day and at the 
    # transitions between raw and QA'ed data - this is to eliminate the lines connecting the
    # data in the plots
  mutate(
    df_data = map(
      df_data, 
      ~add_extra_dt_halfday(.x) %>% 
        add_extra_dt_btw_status() %>% 
        # define a group variable to eliminate lines when using a color grouping variable
        mutate(grp = cumsum(is.na(Value)))
    )
  )
```

# Create Plots

```{r create plots, warning = FALSE}
# Create interactive plotly plots of individual years for each parameter
rtm_wq_plots <- rtm_wq_all_clean %>% mutate(plt = map2(df_data, Parameter, create_ts_plotly)) 
```

# Plots {.tabset .tabset-pills}

```{r prepare plot templates, include = FALSE}
# Find the starting year for the period of record for each parameter
df_start_yr <- rtm_wq_plots %>% 
  group_by(Parameter) %>% 
  summarize(start_yr = min(Year)) %>% 
  arrange(Parameter)

# Define file path for the location of the plot templates
fp_plot_templates <- "C:/Repositories/ND-FASTR/Water_Quality/Continuous_WQ/Plots_Raw_Data/"

# Create a list of plot templates to use for each parameter
produce_plots <- 
  map2(
    df_start_yr$start_yr,
    df_start_yr$Parameter,
    ~knit_expand(
      file = paste0(fp_plot_templates, "rtm_raw_plots_year-round_template_", .x, ".Rmd"), 
      parameter = .y
    )
  )
```

In the plots below, data in the **<span style="color: gray;">gray</span>** color has already been QA'ed and validated and data in **<span style="color: brown;">brown</span>** is still considered raw.

`r knit(text = unlist(produce_plots))`

