---
title: "Code_cont_WQ_LIS_plots"
author: "Amanda Maguire"
date: "July 29, 2020"
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document provides an example of how to use Rmarkdown to bring together text, code, and its output in way to communicate code and results to others. This document also introduces some useful code to use to create simple plots of continuous water quality data.

Code is placed in code chunks like below:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
```

# Import Data

To start plotting, first we need to import some data of course:
```{r import data}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/Office of Water Quality and Estuarine Ecology - North Delta Flow Action/WQ_Subteam/Processed_Data/Continuous"
  )
)

# Import Lisbon data
lis_orig <- 
  read_csv(
    file = paste0(sharepoint_path, "/RTM_OUTPUT_LIS_formatted.csv"),
    col_types = paste0("T", str_c(rep("d", 14), collapse = ""), "c")
  )
```

The data looks as we expect it to:
```{r glimpse data}
glimpse(lis_orig)
```

# Create Plots

## One parameter

Simple line plot of one parameter collected at the Lisbon station. We'll use the `ggplot` function from the ggplot2 package.

WaterTemp plot
```{r plot one param, warning = FALSE}
p <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = WaterTemp)) +
  geom_line()

print(p)
```

Zoom x axis of WaterTemp data
```{r plot one param change x axis lim, warning = FALSE}
p + scale_x_datetime(limits = c(as_datetime("2013-01-01 00:00:00"), NA))
```

Turbidity plot
```{r plot Turbidity param, warning = FALSE}
pt <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = Turbidity)) +
  geom_line()

print(pt)
```

Zoom x axis of Turbidity data
```{r plot Turbidity param change x axis lim, warning = FALSE}
pt + scale_x_datetime(limits = c(as_datetime("2013-01-01 00:00:00"), NA))
```

SpCnd plot
```{r plot SpCnd param, warning = FALSE}
ps <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = SpCnd)) +
  geom_line()

print(pt)
```

Zoom x axis of SpCnd data
```{r plot SpCnd param change x axis lim, warning = FALSE}
ps + scale_x_datetime(limits = c(as_datetime("2013-01-01 00:00:00"), NA))
```

DO plot
```{r plot DO param, warning = FALSE}
po <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = DO)) +
  geom_line()

print(po)
```

Zoom x axis of DO data
```{r plot DO param change x axis lim, warning = FALSE}
po + scale_x_datetime(limits = c(as_datetime("2013-01-01 00:00:00"), NA))
```

Chla plot
```{r plot Chla param, warning = FALSE}
pc <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = Chla)) +
  geom_line()

print(pc)
```

Zoom x axis of Chla data
```{r plot Chla param change x axis lim, warning = FALSE}
pc + scale_x_datetime(limits = c(as_datetime("2013-01-01 00:00:00"), NA))
```

Flow plot
```{r plot Flow param, warning = FALSE}
pflo <- lis_orig %>% 
  ggplot(aes(x = DateTime, y = Flow)) +
  geom_line()

print(pflo)
```


## Multiple parameters

This is nice, but we can also plot multiple parameters at once on a single plot with `facet_grid`. But, first we have to remove the Quality Code variables and restructure the data to a long format using `pivot_longer`.
```{r pivot data longer}
lis_long <- lis_orig %>% 
  select(!ends_with("_Qual")) %>% 
  pivot_longer(
    cols = Flow:Chla,
    names_to = "parameter",
    values_to = "value"
  )
```

Now the data is structured so that the parameter names are in the `parameter` variable and the data values are in the `value` variable. 
```{r display data in long format}
head(lis_long)
```

Let's create a plot with facets. Use the `scales` argument set to `"free_y"` in `facet_grid` to adjust y-axis limits to the data of each parameter.
```{r plot with facets, warning = FALSE}
p2 <- lis_long %>%
  ggplot(aes(x = DateTime, y = value)) +
  geom_line() +
  facet_grid(
    rows = vars(parameter),
    scales = "free_y"
  )

p2
```

This is nice, but it would be even nicer to make the plot larger so we can see it better. This is done with the `fig.width` and `fig.height` options in the top of the code chunk.
```{r plot with facets larger, warning = FALSE, fig.width = 8.5, fig.height = 11}
p2
```

If you would like to zoom in on one particular time period, we can use the `scale_x_datetime` function as we did above.
```{r plot with facets change x axis lim, warning = FALSE, fig.height = 8}
p2 + scale_x_datetime(
  limits = c(
    as_datetime("2017-01-01 00:00:00"), 
    as_datetime("2018-01-01 00:00:00")
  )
)
```

