---
title: "Raw Data Plots - Chlorophyll"
subtitle: "July 1 - Nov 15"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Create plots of the raw continuous chlorophyll data. Data for all stations have already been QA/QC'ed for the restricted dates for the period of interest for the NDFS synthesis study; however, we need to expand the data set to July 1 - Nov 15 for each year. This additional data needs to be visually inspected for suspect or unreliable values. Additionally, stations LIB, LIS, RVB, and STTD have already been QA/QC'ed for the expanded time period.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(fs)
library(plotly)
library(htmltools)
library(scales)
library(knitr)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter())
```

```{r load functions, message = FALSE, warning = FALSE}
# Source functions
source(here("global_ndfa_funcs.R"))

# Define root directory for the WQ_Subteam on the NDFA SharePoint
fp_wq_root <- ndfa_abs_sp_path("2011-2019 Synthesis Study-FASTR/WQ_Subteam")
```

```{r create functions}
# Import continuous chlorophyll data from formatted csv files
import_rtm_chla <- function(fp) {
  read_csv(
    file = fp, 
    col_types = cols_only(
      StationCode = "c",
      DateTime = "c",
      Chla = "d"
    )
  )
}

# Create simple interactive timeseries plotly plot of continuous chlorophyll Data
create_ts_plotly <- function(df, plt_title) {
  # create plot
  p <- df %>% 
    ggplot(aes(x = DateTime, y = Chla, color = Status)) +
    geom_point(size = 1) +
    scale_x_datetime(
      name = "Date",
      breaks = breaks_pretty(15),
      labels = label_date_short()
    ) +
    ylab("Chlorophyll") +
    scale_color_manual(values = c("QA" = "gray40", "Raw" = "brown")) +
    ggtitle({{ plt_title }})
  
  ggplotly(p)
}
```

# Import Data

```{r import data}
# Import QA'ed and cleaned continuous chlorophyll data for the NDFA period of interest
df_chla_qa <- import_rtm_chla(file.path(fp_wq_root, "Processed_Data/Continuous/RTM_INPUT_all_2021-04-20.csv"))

# Create a vector of all relevant file paths for the processed continuous chlorophyll data
fp_rtm_wq_proc <- dir_ls(
  path = file.path(fp_wq_root, "Processed_Data/Continuous/All_Dates"),
  regexp = "RTM_OUTPUT_(I80|RCS|RD22|RYI)_formatted_all\\.csv$"
)

# Import processed continuous chlorophyll data into a dataframe
df_chla_proc <- 
  map(fp_rtm_wq_proc, import_rtm_chla) %>% 
  list_rbind()
```

# Prepare Data

We need to prepare both the processed data and the QA'ed data so that they can be combined only adding the processed data outside of the NDFS period of interest.

```{r prepare processed data, message = FALSE}
# Prepare processed data
df_chla_proc_c <- df_chla_proc %>% 
  # parse date-time variable and define tz as PST; add date and year variables
  mutate(
    DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8"),
    Date = date(DateTime),
    Year = year(DateTime)
  ) %>% 
  # add flow action periods to the data frame but keep all data
  ndfa_action_periods(na_action_remove = FALSE) %>%
  # only keep data outside of the NDFS period of interest
  filter(is.na(FlowActionPeriod)) %>%
  # remove NA Chla values
  drop_na(Chla) %>% 
  # filter data to July 1 - Nov 15 for each year
  filter(
    month(Date) %in% 7:11,
    !(month(Date) == 11 & day(Date) > 15)
  ) %>% 
  # Add variable to identify which data set it came from
  mutate(Status = "Raw") %>% 
  # remove unnecessary variables
  select(-c(Date, FlowActionPeriod))
```

```{r prepare qa data}
# Prepare QA'ed and cleaned data to be combined to the processed data
df_chla_qa_c <- df_chla_qa %>% 
  # parse date-time variable and define tz as PST; add year variable
  mutate(
    DateTime = ymd_hms(DateTime, tz = "Etc/GMT+8"),
    Year = year(DateTime)
  ) %>% 
  # remove NA Chla values
  drop_na(Chla) %>% 
  # add variable to identify which data set it came from
  mutate(Status = "QA")
```

The data is now ready to be combined.

```{r combine data}
# Combine the raw the QA'ed and cleaned data
df_chla_all <- bind_rows(df_chla_proc_c, df_chla_qa_c) %>% arrange(StationCode, DateTime)
```

The entire timeseries of some Station-Year combinations have been QC'ed already, so we'll only include Station-Year combinations that contain some raw data.

```{r filter station yr combinations}
df_chla_all_c <- df_chla_all %>% 
  group_by(StationCode, Year) %>% 
  mutate(Keep = if_else(any(Status == "Raw"), TRUE, FALSE)) %>% 
  ungroup() %>% 
  filter(Keep) %>% 
  select(-Keep)
```

# Create Plots

```{r create plots}
# Create interactive plotly plots of individual years for each Station
ndf_plots <- df_chla_all_c %>% 
  nest(df_data = c(DateTime, Chla, Status)) %>% 
  mutate(plt = map2(df_data, Year, create_ts_plotly)) %>% 
  select(-df_data)
```

# Plots {.tabset .tabset-pills}

```{r prepare plot templates, include = FALSE}
# Create a list of plot templates to use for each station
produce_plots <- 
  map(
    unique(ndf_plots$StationCode),
    ~ knit_expand(
      file = here("manuscript_synthesis/notebooks/plot_rtm_data_for_QA_template.Rmd"), 
      station = .x
    )
  )
```

In the plots below, data in the **<span style="color: gray;">gray</span>** color has already been QA'ed and validated and data in **<span style="color: brown;">brown</span>** is still considered raw.

`r knit(text = unlist(produce_plots))`

