---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Model selection"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Purpose

Explore and analyze the continuous chlorophyll data to be included in the NDFS synthesis manuscript. We will attempt to fit multiple models to the data set and use a model selection process to determine the best one. At a minimum the model will contain the three categorical variables (Year, Station, and Flow Pulse Period) and optionally a term to account for seasonality (either daily average water temperature or a GAM smooth for DOY). These models will only include representative stations for 3 habitat types - upstream (RD22), lower Yolo Bypass (STTD), and downstream (LIB).

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(knitr)
library(mgcv)
library(car)
library(gratia)
library(rlang)
library(patchwork)
library(here)
library(conflicted)

# Source functions
source(here("manuscript_synthesis/src/global_functions.R"))

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

Create function to plot linear model diagnostics:

```{r global funcs}
# Create model diagnostic plots to check assumptions
plot_lm_diag <- function(df_data, param_var, unit_label, model, trans = c("none", "log", "sqrt"), ...) {
  trans <- match.arg(trans)
  
  df_data <- df_data %>%
    mutate(
      Residuals = resid(model),
      Fitted = predict(model)
    )
  
  param_name <- as_name(ensym(param_var))
  
  if (trans == "log") {
    unit_label <- paste0("log[", unit_label, "]")
  } else if (trans == "sqrt") {
    unit_label <- paste0("sqrt[", unit_label, "]")
  }
  
  plt_hist <- ggplot(df_data, aes(x = Residuals)) +
    geom_histogram(...) +
    labs(
      title = "Residual Histogram", 
      x = "Residuals"
    ) +
    theme_bw()
  
  plt_qq <- ggplot(df_data, aes(sample = Residuals)) + 
    labs(
      title = "Residual Probability Plot", 
      x = "Normal Quantiles", 
      y = "Residuals"
    ) + 
    stat_qq() + 
    stat_qq_line(linewidth = 1, color = 'red') + 
    theme_bw()
  
  plt_res_fit <- ggplot(df_data, aes(x = Fitted, y = Residuals)) +
    geom_point() +
    labs(
      title = "Residuals vs. Fitted Values", 
      x = (paste0("Predicted ", param_name, " (", unit_label, ")")),
      y = paste0("Residuals (", unit_label, ")")
    ) +
    theme_bw()
  
  plt_obs_fit <- ggplot(df_data, aes(x = Fitted, y = {{ param_var}})) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0, color = "red") +
    labs(
      title = "Observed vs. Fitted Values", 
      x = paste0("Predicted ", param_name, " (", unit_label, ")"),
      y = paste0("Observed ", param_name, " (", unit_label, ")")
    ) +
    theme_bw()
  
  (plt_hist + plt_qq) / (plt_res_fit + plt_obs_fit)
}
```


# Import Data

```{r import data}
# Define file path for processed data
fp_data <- here("manuscript_synthesis/data/processed")

# Import daily average chlorophyll and water temperature data
df_chla_wt <- readRDS(file.path(fp_data, "chla_wt_daily_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare chla data, message = FALSE}
# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB")

# Prepare chlorophyll and water temperature data for exploration and analysis
df_chla_wt_c <- df_chla_wt %>% 
  # Filter to only include representative stations for 3 habitat types - RD22, STTD, LIB
  filter(StationCode %in% sta_order) %>% 
  mutate(
    # Scale and log transform chlorophyll values
    Chla_log = log(Chla * 1000),
    # Add Year and DOY variables
    Year = year(Date),
    DOY = yday(Date)
  ) %>% 
  # Add Flow Action Periods
  ndfa_action_periods() %>% 
  mutate(
    # Apply factor orders to FlowActionPeriod and StationCode
    FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")),
    StationCode = factor(StationCode, levels = sta_order),
    # Add a column for Year as a factor for the model
    Year_fct = factor(Year)
  ) %>% 
  arrange(StationCode, Date)
  
# Create another dataframe that has up to 3 lag variables for chlorophyll to be
  # used in the linear models
df_chla_wt_lag <- df_chla_wt_c %>% 
  # Fill in missing days for each StationCode-Year combination
  group_by(StationCode, Year) %>% 
  complete(Date = seq.Date(min(Date), max(Date), by = "1 day")) %>% 
  # Create lag variables of scaled log transformed chlorophyll values
  mutate(
    lag1 = lag(Chla_log),
    lag2 = lag(Chla_log, 2),
    lag3 = lag(Chla_log, 3)
  ) %>% 
  ungroup()
```

# Explore sample counts by Station

```{r chla sample counts station}
df_chla_wt_c %>% 
  summarize(
    min_date = min(Date),
    max_date = max(Date),
    num_samples = n(),
    .by = c(StationCode, Year, FlowActionPeriod)
  ) %>% 
  complete(StationCode, Year, FlowActionPeriod) %>% 
  arrange(StationCode, Year, FlowActionPeriod) %>% 
  kable()
```

Except for a few Station-Year-FlowPeriod combinations with low sample counts, it appears we have adequate data.

# Plots

```{r chla boxplot station, fig.width = 8, fig.height = 6}
df_chla_wt_c %>%
  ggplot(aes(x = FlowActionPeriod, y = Chla, fill = FlowActionPeriod)) +
  geom_boxplot() +
  facet_grid(cols = vars(Year), rows = vars(StationCode)) +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides = "l") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position = "top"
  )
```

# Models

## Model 1: GAM 3-way interactions with s(DOY)

### Initial Model

We'll try running a GAM using a three-way interaction between Year, Flow Action Period, and Station, and a smooth term for day of year to account for seasonality (restricting the k-value to 5 to reduce overfitting). First we'll run the GAM without accounting for serial autocorrelation.

```{r gam cat3 no autocorr, warning = FALSE}
m_gam_cat3 <- gam(
  Chla_log ~ Year_fct * FlowActionPeriod * StationCode + s(DOY, k = 5), 
  data = df_chla_wt_c,
  method = "REML"
)
```

Lets look at the model summary and diagnostics:

```{r gam cat3 no autocorr diag, warning = FALSE}
summary(m_gam_cat3)
appraise(m_gam_cat3)
k.check(m_gam_cat3)
draw(m_gam_cat3, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_cat3, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_cat3))
Box.test(residuals(m_gam_cat3), lag = 20, type = 'Ljung-Box')
```

### Model with AR() correlation structure

Now, we'll try to deal with the residual autocorrelation. We'll run AR(1), AR(2), and AR(3) models and compare them using AIC.

```{r gamm cat3 with AR comparison, warning = FALSE}
# Define model formula as an object
f_gam_cat3 <- as.formula("Chla_log ~ Year_fct * FlowActionPeriod * StationCode + s(DOY, k = 5)")

# Fit original model with k = 5 and three successive AR(p) models
m_gamm_cat3 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c,
  method = "REML"
)

m_gamm_cat3_ar1 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 1), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat3_ar2 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 2), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat3_ar3 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 3), # grouped by Year_fct and StationCode
  method = "REML"
)

# Compare models
anova(
  m_gamm_cat3$lme, 
  m_gamm_cat3_ar1$lme, 
  m_gamm_cat3_ar2$lme, 
  m_gamm_cat3_ar3$lme
)
```

It looks like the AR(1) model has the best fit according to the AIC values and backed up by the BIC values. Let's take a closer look at that one.

### AR(1) Model

```{r gamm cat3 ar1 diag, warning = FALSE}
# Pull out the residuals and the GAM model
resid_cat3_ar1 <- residuals(m_gamm_cat3_ar1$lme, type = "normalized")
m_gamm_cat3_ar1_gam <- m_gamm_cat3_ar1$gam

summary(m_gamm_cat3_ar1_gam)
appraise(m_gamm_cat3_ar1_gam)
k.check(m_gamm_cat3_ar1_gam)
draw(m_gamm_cat3_ar1_gam, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gamm_cat3_ar1_gam, pages = 1, all.terms = TRUE)
acf(resid_cat3_ar1)
Box.test(resid_cat3_ar1, lag = 20, type = 'Ljung-Box')
```

The AR(1) model has much less residual autocorrelation, and the diagnostics plots look pretty good. We'll take a look at the ANOVA table to see if we need all the interaction terms included in the model.

```{r gamm cat3 ar1 anova, warning = FALSE}
# the anova.gam function is similar to a type 3 ANOVA
anova(m_gamm_cat3_ar1_gam)
```

The 3-way interaction term is significant in this model, so we'll use this one for the model selection process. Note that the s(DOY) term is marginally significant (p = 0.0511) in this model.

```{r gamm cat3 clean up}
rm(m_gam_cat3, m_gamm_cat3, m_gamm_cat3_ar2, m_gamm_cat3_ar3)
```


## Model 2: LM 3-way interactions with water temperature

### Initial Model

We'll try running a linear model using a three-way interaction between Year, Flow Action Period, and Station, and a covariate of water temperature to account for seasonality and its effect of primary production. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat3 wt no autocorr}
m_lm_cat3_wt <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat3 wt no autocorr diag}
summary(m_lm_cat3_wt)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat3_wt, 
    trans = "log"
  )

shapiro.test(m_lm_cat3_wt$residuals)
acf(residuals(m_lm_cat3_wt))
Box.test(residuals(m_lm_cat3_wt), lag = 20, type = 'Ljung-Box')
```

Hmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat3 wt lag1}
m_lm_cat3_wt_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp + lag1, data = .)

acf(residuals(m_lm_cat3_wt_lag1))
Box.test(residuals(m_lm_cat3_wt_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat3 wt lag2}
m_lm_cat3_wt_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp + lag1 + lag2, data = .)

acf(residuals(m_lm_cat3_wt_lag2))
Box.test(residuals(m_lm_cat3_wt_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat3 wt lag3}
m_lm_cat3_wt_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat3_wt_lag3))
Box.test(residuals(m_lm_cat3_wt_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat3 wt lag comparison, warning = FALSE}
AIC(m_lm_cat3_wt_lag1, m_lm_cat3_wt_lag2, m_lm_cat3_wt_lag3)
BIC(m_lm_cat3_wt_lag1, m_lm_cat3_wt_lag2, m_lm_cat3_wt_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat3 wt lag1 summary and diag}
summary(m_lm_cat3_wt_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat3_wt_lag1, 
    trans = "log"
  )

shapiro.test(m_lm_cat3_wt_lag1$residuals)
```

Hmm, the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat3 wt lag1}
Anova(m_lm_cat3_wt_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right and the water temperature term isn't significant.

```{r lm cat3 wt clean up}
rm(m_lm_cat3_wt, m_lm_cat3_wt_lag2, m_lm_cat3_wt_lag3)
```


## Model 3: LM 3-way interactions without seasonal term

### Initial Model

We'll try running a linear model using a three-way interaction between Year, Flow Action Period, and Station without including a term to account for seasonality. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat3 no autocorr}
m_lm_cat3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat3 no autocorr diag}
summary(m_lm_cat3)

df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat3, 
    trans = "log"
  )

shapiro.test(m_lm_cat3$residuals)
acf(residuals(m_lm_cat3))
Box.test(residuals(m_lm_cat3), lag = 20, type = 'Ljung-Box')
```

Hmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat3 lag1}
m_lm_cat3_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + lag1, data = .)

acf(residuals(m_lm_cat3_lag1))
Box.test(residuals(m_lm_cat3_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat3 lag2}
m_lm_cat3_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + lag1 + lag2, data = .)

acf(residuals(m_lm_cat3_lag2))
Box.test(residuals(m_lm_cat3_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat3 lag3}
m_lm_cat3_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat3_lag3))
Box.test(residuals(m_lm_cat3_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat3 lag comparison, warning = FALSE}
AIC(m_lm_cat3_lag1, m_lm_cat3_lag2, m_lm_cat3_lag3)
BIC(m_lm_cat3_lag1, m_lm_cat3_lag2, m_lm_cat3_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat3 lag1 summary and diag}
summary(m_lm_cat3_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat3_lag1, 
    trans = "log"
  )

shapiro.test(m_lm_cat3_lag1$residuals)
```

Hmm, again the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat3 lag1}
Anova(m_lm_cat3_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right.

```{r lm cat3 clean up}
rm(m_lm_cat3, m_lm_cat3_lag2, m_lm_cat3_lag3)
```


## Model 4: GAM 2-way interactions with s(DOY)

### Initial Model

We'll try running a GAM using all two-way interactions between Year, Flow Action Period, and Station, and a smooth term for day of year to account for seasonality (restricting the k-value to 5 to reduce overfitting). First we'll run the GAM without accounting for serial autocorrelation.

```{r gam cat2 no autocorr, warning = FALSE}
m_gam_cat2 <- gam(
  Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + s(DOY, k = 5), 
  data = df_chla_wt_c,
  method = "REML"
)
```

Lets look at the model summary and diagnostics:

```{r gam cat2 no autocorr diag, warning = FALSE}
summary(m_gam_cat2)
appraise(m_gam_cat2)
k.check(m_gam_cat2)
draw(m_gam_cat2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_cat2, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_cat2))
Box.test(residuals(m_gam_cat2), lag = 20, type = 'Ljung-Box')
```

### Model with AR() correlation structure

Now, we'll try to deal with the residual autocorrelation. We'll run AR(1), AR(2), and AR(3) models and compare them using AIC.

```{r gamm cat2 with AR comparison, warning = FALSE}
# Define model formula as an object
f_gam_cat2 <- as.formula("Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + s(DOY, k = 5)")

# Fit original model with k = 5 and three successive AR(p) models
m_gamm_cat2 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c,
  method = "REML"
)

m_gamm_cat2_ar1 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 1), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat2_ar2 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 2), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat2_ar3 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 3), # grouped by Year_fct and StationCode
  method = "REML"
)

# Compare models
anova(
  m_gamm_cat2$lme, 
  m_gamm_cat2_ar1$lme, 
  m_gamm_cat2_ar2$lme, 
  m_gamm_cat2_ar3$lme
)
```

It looks like the AR(1) model has the best fit according to the AIC values and backed up by the BIC values. Let's take a closer look at that one.

### AR(1) Model

```{r gamm cat2 ar1 diag, warning = FALSE}
# Pull out the residuals and the GAM model
resid_cat2_ar1 <- residuals(m_gamm_cat2_ar1$lme, type = "normalized")
m_gamm_cat2_ar1_gam <- m_gamm_cat2_ar1$gam

summary(m_gamm_cat2_ar1_gam)
appraise(m_gamm_cat2_ar1_gam)
k.check(m_gamm_cat2_ar1_gam)
draw(m_gamm_cat2_ar1_gam, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gamm_cat2_ar1_gam, pages = 1, all.terms = TRUE)
acf(resid_cat2_ar1)
Box.test(resid_cat2_ar1, lag = 20, type = 'Ljung-Box')
```

The AR(1) model has much less residual autocorrelation, and the diagnostics plots look pretty good. We'll take a look at the ANOVA table to see if we need all the interaction terms included in the model.

```{r gamm cat2 ar1 anova, warning = FALSE}
# the anova.gam function is similar to a type 3 ANOVA
anova(m_gamm_cat2_ar1_gam)
```

All the interaction terms are significant in this model, so we'll use this one for the model selection process. Note that the s(DOY) term is significant in this model, but is effectively a straight line.

```{r gamm cat2 clean up}
rm(m_gam_cat2, m_gamm_cat2, m_gamm_cat2_ar2, m_gamm_cat2_ar3)
```


## Model 5: LM 2-way interactions with water temperature

### Initial Model

We'll try running a linear model using all two-way interactions between Year, Flow Action Period, and Station, and a covariate of water temperature to account for seasonality and its effect of primary production. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat2 wt no autocorr}
m_lm_cat2_wt <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat2 wt no autocorr diag}
summary(m_lm_cat2_wt)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat2_wt, 
    trans = "log"
  )

shapiro.test(m_lm_cat2_wt$residuals)
acf(residuals(m_lm_cat2_wt))
Box.test(residuals(m_lm_cat2_wt), lag = 20, type = 'Ljung-Box')
```

Hmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat2 wt lag1}
m_lm_cat2_wt_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp + lag1, data = .)

acf(residuals(m_lm_cat2_wt_lag1))
Box.test(residuals(m_lm_cat2_wt_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat2 wt lag2}
m_lm_cat2_wt_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp + lag1 + lag2, data = .)

acf(residuals(m_lm_cat2_wt_lag2))
Box.test(residuals(m_lm_cat2_wt_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat2 wt lag3}
m_lm_cat2_wt_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat2_wt_lag3))
Box.test(residuals(m_lm_cat2_wt_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat2 wt lag comparison, warning = FALSE}
AIC(m_lm_cat2_wt_lag1, m_lm_cat2_wt_lag2, m_lm_cat2_wt_lag3)
BIC(m_lm_cat2_wt_lag1, m_lm_cat2_wt_lag2, m_lm_cat2_wt_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat2 wt lag1 summary and diag}
summary(m_lm_cat2_wt_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat2_wt_lag1, 
    trans = "log"
  )

shapiro.test(m_lm_cat2_wt_lag1$residuals)
```

Hmm, the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat2 wt lag1}
Anova(m_lm_cat2_wt_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right and the water temperature term isn't significant.

```{r lm cat2 wt clean up}
rm(m_lm_cat2_wt, m_lm_cat2_wt_lag2, m_lm_cat2_wt_lag3)
```


## Model 6: LM 2-way interactions without seasonal term

### Initial Model

We'll try running a linear model using all two-way interactions between Year, Flow Action Period, and Station without including a term to account for seasonality. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat2 no autocorr}
m_lm_cat2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat2 no autocorr diag}
summary(m_lm_cat2)

df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat2, 
    trans = "log"
  )

shapiro.test(m_lm_cat2$residuals)
acf(residuals(m_lm_cat2))
Box.test(residuals(m_lm_cat2), lag = 20, type = 'Ljung-Box')
```

Hmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat2 lag1}
m_lm_cat2_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + lag1, data = .)

acf(residuals(m_lm_cat2_lag1))
Box.test(residuals(m_lm_cat2_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat2 lag2}
m_lm_cat2_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + lag1 + lag2, data = .)

acf(residuals(m_lm_cat2_lag2))
Box.test(residuals(m_lm_cat2_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat2 lag3}
m_lm_cat2_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat2_lag3))
Box.test(residuals(m_lm_cat2_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat2 lag comparison, warning = FALSE}
AIC(m_lm_cat2_lag1, m_lm_cat2_lag2, m_lm_cat2_lag3)
BIC(m_lm_cat2_lag1, m_lm_cat2_lag2, m_lm_cat2_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat2 lag1 summary and diag}
summary(m_lm_cat2_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  plot_lm_diag(
    Chla_log, 
    "ug/L", 
    m_lm_cat2_lag1, 
    trans = "log"
  )

shapiro.test(m_lm_cat2_lag1$residuals)
```

Hmm, again the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat2 lag1}
Anova(m_lm_cat2_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right.

```{r lm cat2 clean up}
rm(m_lm_cat2, m_lm_cat2_lag2, m_lm_cat2_lag3)
```


# Model selection - GAM and LM models

As a summary, here are the 6 models we are comparing:

* Model 1 - `m_gamm_cat3_ar1` - GAM 3-way interactions with s(DOY) <br>
Formula: `r format(m_gamm_cat3_ar1_gam$formula) %>% str_c(collapse = "") %>% str_squish()` <br><br>
* Model 2 - `m_lm_cat3_wt_lag1` - LM 3-way interactions with water temperature <br>
Formula: `r format(m_lm_cat3_wt_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br><br>
* Model 3 - `m_lm_cat3_lag1` - LM 3-way interactions without seasonal term <br>
Formula: `r format(m_lm_cat3_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br><br>
* Model 4 - `m_gamm_cat2_ar1` - GAM 2-way interactions with s(DOY) <br>
Formula: `r format(m_gamm_cat2_ar1_gam$formula) %>% str_c(collapse = "") %>% str_squish()` <br><br>
* Model 5 - `m_lm_cat2_wt_lag1` - LM 2-way interactions with water temperature <br>
Formula: `r format(m_lm_cat2_wt_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br><br>
* Model 6 - `m_lm_cat2_lag1` - LM 2-way interactions without seasonal term <br>
Formula: `r format(m_lm_cat2_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br>

```{r model sel gam, warning = FALSE}
# AIC values
df_m_aic <- 
  AIC(
    m_gamm_cat3_ar1$lme,
    m_lm_cat3_wt_lag1,
    m_lm_cat3_lag1,
    m_gamm_cat2_ar1$lme,
    m_lm_cat2_wt_lag1,
    m_lm_cat2_lag1
  ) %>% 
  as_tibble(rownames = "Model")

# BIC values
df_m_bic <- 
  BIC(
    m_gamm_cat3_ar1$lme,
    m_lm_cat3_wt_lag1,
    m_lm_cat3_lag1,
    m_gamm_cat2_ar1$lme,
    m_lm_cat2_wt_lag1,
    m_lm_cat2_lag1
  ) %>% 
  as_tibble(rownames = "Model")

# Combine AIC and BIC
df_m_aic_bic <- left_join(df_m_aic, df_m_bic, by = join_by(Model, df))

# Sort by AIC
df_m_aic_bic %>% arrange(AIC)

# Sort by BIC
df_m_aic_bic %>% arrange(BIC)
```

According to AIC, model 3 (LM 3-way interactions without seasonal term) was the best model, while BIC preferred model 6 (LM 2-way interactions without seasonal term). We'll select model 3 (LM 3-way interactions without seasonal term) because the 3-way interaction between categorical variables was significant in this model indicating that there are possible significant differences between Flow Pulse Periods within each Station-Year grouping. We'll export model 3 and its data to be used for further exploration in other notebooks. 

```{r export models and data}
# Define file path for model and data used in the model
fp_model <- here("manuscript_synthesis/model/interim")

# Export model and data used in the model
m_lm_cat3_lag1 %>% saveRDS(file.path(fp_model, "chla_cat3_lm_model.rds"))

df_chla_wt_lag %>% 
  select(StationCode, Year, Date, Chla, Chla_log, FlowActionPeriod, lag1) %>% 
  drop_na(Chla_log, lag1) %>% 
  saveRDS(file.path(fp_model, "chla_model_data.rds"))
```

