---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Explore relationship with daily average percent flow pulse water"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Purpose

Explore the relationship between continuous chlorophyll data and the modelled percent flow pulse water to see if should be included in the NDFS synthesis manuscript.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(knitr)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

# Import Data

```{r import data}
# Define file path for processed data
fp_data <- here("manuscript_synthesis/data/processed")

# Import daily average water quality data
df_wq <- readRDS(file.path(fp_data, "wq_daily_avg_2013-2019.rds"))
  
# Import daily average percent flow pulse water data
df_pfp <- readRDS(file.path(fp_data, "perc_flow_pulse_daily_avg_2016-2019.rds"))
```

# Prepare Data

```{r prepare chla data, message = FALSE}
# Create a vector for the factor order of StationCode
sta_order <- c("LIS", "STTD", "LIB", "RYI", "RVB")

# Prepare chlorophyll and percent flow pulse water data for exploration and analysis
df_chla_c <- df_wq %>% 
  select(StationCode, Date, Chla) %>% 
  drop_na(Chla) %>% 
  # Join flow data to chlorophyll data
  left_join(df_pfp, by = join_by(StationCode, Date)) %>% 
  # Remove all NA flow values
  drop_na(PercFlowPulseAvg) %>% 
  mutate(
    # Add Year variable
    Year = year(Date),
    # Apply factor order to StationCode
    StationCode = factor(StationCode, levels = sta_order),
    # Convert percent flow pulse water values to proportions so they scale
      # correctly with scales::label_percent()
    PercFlowPulseAvg = PercFlowPulseAvg / 100
  ) %>% 
  arrange(StationCode, Date)
```

# Explore sample counts by Station

```{r chla sample counts station}
df_chla_c %>% 
  summarize(
    min_date = min(Date),
    max_date = max(Date),
    num_samples = n(),
    .by = c(StationCode, Year)
  ) %>% 
  arrange(StationCode, Year) %>% 
  kable()
```

Except for RYI missing data for 2017 and RVB having limited data for 2017, it looks like we pretty good data coverage.

# Plots

Let's explore the data with some plots. First, lets plot the data in scatterplots of chlorophyll and percent flow pulse water facetted by Station and grouping all years together.

```{r chla scatterplot all yrs, message = FALSE}
df_chla_c %>% 
  ggplot(aes(x = PercFlowPulseAvg, y = Chla)) +
  geom_point() +
  geom_smooth(formula = "y ~ x") +
  facet_wrap(vars(StationCode), scales = "free") +
  theme_bw() +
  ylab("Chlorophyll (ug/L)") +
  scale_x_continuous(name = "Percent Flow Pulse Water", labels = label_percent())
```

There's some strange patterns in the data. Let's break these scatterplots apart by year to see if they can be explained by annual differences.

```{r chla scatterplot facet yrs, message = FALSE, warning = FALSE, fig.height = 11, fig.width = 8}
df_chla_c %>% 
  ggplot(aes(x = PercFlowPulseAvg, y = Chla)) +
  geom_point() +
  geom_smooth(formula = "y ~ x") +
  facet_wrap(
    vars(StationCode, Year),
    ncol = 4,
    scales = "free",
    labeller = labeller(.multi_line = FALSE)
  ) +
  theme_bw() +
  ylab("Chlorophyll (ug/L)") +
  scale_x_continuous(name = "Percent Flow Pulse Water", labels = label_percent()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Hmmm, a lot of these relationships look complicated. I'm not sure we'll try using the daily average percent flow pulse water as a continuous predictor to model chlorophyll concentrations. 

