---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Models using weekly average flow as continuous predictor"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Purpose

Explore and analyze the continuous chlorophyll data to be included in the NDFS synthesis manuscript. We will attempt to fit multiple models to predict weekly average chlorophyll fluorescence values.  All models will only include representative stations for 4 habitat types - upstream (RD22), lower Yolo Bypass (STTD), Cache Slough complex (LIB), and downstream (RVB). At a minimum, the models will contain the two categorical variables - Year and Station - as predictor variables. In some of the models, we will add weekly average flow as a continuous predictor which replaces the categorical predictor - flow action period - in the original analysis. Additionally, we'll add a GAM smooth for Week number term to account for seasonality in some of the models. After fitting multiple models, we'll use a model selection process to determine the best one.

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(knitr)
library(mgcv)
library(car)
library(gratia)
library(ggeffects)
library(emmeans)
library(multcomp)
library(here)
library(conflicted)

# Source functions
source(here("manuscript_synthesis/src/global_functions.R"))

# Declare package conflict preferences
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

# Import Data

```{r import data}
# Define file path for processed data
fp_data <- here("manuscript_synthesis/data/processed")

# Import weekly average water quality data
df_wq <- readRDS(file.path(fp_data, "wq_week_avg_2013-2019.rds"))

# Import weekly average flow data
df_flow <- readRDS(file.path(fp_data, "flow_week_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare chla data, message = FALSE}
# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

# We will use LIS flow data for all stations except for RD22
df_flow_c <- bind_rows(
  expand_grid(
    StationCode = c("STTD", "LIB", "RVB"),
    df_flow |> filter(StationCode == "LIS") |> select(-StationCode)
  ),
  df_flow |> filter(StationCode == "RD22")
)


# Prepare chlorophyll and flow data for exploration and analysis
df_chla_c1 <- df_wq %>%
  select(StationCode, Year, Week, Chla) %>%
  drop_na(Chla) %>%
  # Filter to only include representative stations for 4 habitat types - RD22, STTD, LIB, RVB
  filter(StationCode %in% sta_order) %>%
  # Join flow data to chlorophyll data
  left_join(df_flow_c, by = join_by(StationCode, Year, Week)) %>%
  # Remove all NA flow values
  drop_na(Flow) %>%
  mutate(
    # Scale and log transform chlorophyll values
    Chla_log = log(Chla * 1000),
    # Apply factor order to StationCode
    StationCode = factor(StationCode, levels = sta_order),
    # Add a column for Year as a factor for the model
    Year_fct = factor(Year)
  ) %>%
  arrange(StationCode, Year, Week)
```

# Explore sample counts by Station

```{r chla sample counts station}
df_chla_c1 %>%
  summarize(
    min_week = min(Week),
    max_week = max(Week),
    num_samples = n(),
    .by = c(StationCode, Year)
  ) %>%
  arrange(StationCode, Year) %>%
  kable()
```

Looking at the sample counts and date ranges, we'll only include years 2015-2019 for the analysis.

```{r chla remove under sampled}
df_chla_c2 <- df_chla_c1 %>%
  filter(Year %in% 2015:2019) %>%
  mutate(Year_fct = fct_drop(Year_fct))
```

We'll create another dataframe that has up to 2 lag variables for chlorophyll to be used in the models to help with serial autocorrelation.

```{r chla lag}
df_chla_c2_lag <- df_chla_c2 %>%
  # Fill in missing weeks for each StationCode-Year combination
  group_by(StationCode, Year) %>%
  # Create lag variables of scaled log transformed chlorophyll values
  mutate(
    lag1 = lag(Chla_log),
    lag2 = lag(Chla_log, 2)
  ) %>%
  ungroup()
```

# Plots

Let's explore the data with some plots. First, lets plot the data in scatter plots of chlorophyll and flow faceted by Station and grouping all years together.

```{r chla scatterplot all yrs, message = FALSE, fig.height = 6}
df_chla_c2 %>%
  ggplot(aes(x = Flow, y = Chla)) +
  geom_point() +
  geom_smooth(formula = "y ~ x") +
  facet_wrap(vars(StationCode), scales = "free") +
  theme_bw()
```

At first glance, I'm not sure how well flow is going to be able to predict chlorophyll concentrations. At the furthest upstream station - RD22 - chlorophyll appears to be highest at the lowest flows, but the variation is at its maximum at the lowest flows. There may be some dilution effect going on here at the higher flows. At STTD, there does seem to be a modest increase in chlorophyll concentrations at the mid-range flows. There appears to be no effect of flow on chlorophyll at RVB, but the range of chlorophyll concentrations is narrow at this station (between 0 and 5).

Let's break these scatterplots apart by year to see how these patterns vary annually.

```{r chla scatterplot facet yrs, message = FALSE, fig.height = 8, fig.width = 8.5}
df_chla_c2 %>%
  ggplot(aes(x = Flow, y = Chla)) +
  geom_point() +
  geom_smooth(formula = "y ~ x") +
  facet_wrap(
    vars(StationCode, Year),
    ncol = 5,
    scales = "free",
    labeller = labeller(.multi_line = FALSE)
  ) +
  theme_bw()
```

The patterns appear to vary annually at each station, which may justify using a 3-way interaction.

# Model 1: GAM Model with Flow and 3-way interactions

First, we will attempt to fit a generalized additive model (GAM) to the data set to help account for seasonality in the data. We'll try running a GAM using a three-way interaction between Year, Weekly Average Flow, and Station, and a cyclic penalized cubic regression spline smooth term for week number to account for seasonality (restricting the k-value to 5 to reduce overfitting). Initially, we'll run the GAM without accounting for serial autocorrelation.

## Initial Model

```{r gam flow3 no autocorr, warning = FALSE}
m_gam_flow3 <- gam(
  Chla_log ~ Year_fct * Flow * StationCode + s(Week,  bs = "cc", k = 5),
  data = df_chla_c2,
  method = "REML",
  knots = list(week = c(0, 52))
)
```

Lets look at the model summary and diagnostics:

```{r gam flow3 no autocorr diag, warning = FALSE}
summary(m_gam_flow3)
appraise(m_gam_flow3)
shapiro.test(residuals(m_gam_flow3))
k.check(m_gam_flow3)
draw(m_gam_flow3, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_flow3, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_flow3))
Box.test(residuals(m_gam_flow3), lag = 20, type = 'Ljung-Box')
```

Besides the Shapiro-Wilk normality test showing that the residuals aren't normal, the diagnostic plots look pretty good. However, the residuals are autocorrelated.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation. We'll run a series of models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r gam flow3 lag1, warning = FALSE}
m_gam_flow3_lag1 <- gam(
  Chla_log ~ Year_fct * Flow * StationCode + s(Week, bs = "cc", k = 5) + lag1,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_flow3_lag1))
Box.test(residuals(m_gam_flow3_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r gam flow3 lag2, warning = FALSE}
m_gam_flow3_lag2 <- gam(
  Chla_log ~ Year_fct * Flow * StationCode + s(Week, bs = "cc", k = 5) + lag1 + lag2,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_flow3_lag2))
Box.test(residuals(m_gam_flow3_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 1 lag term already seems to address the serial autocorrelation. Let's use AIC to see how they compare.

### Compare Models

```{r gam flow3 compare}
AIC(m_gam_flow3, m_gam_flow3_lag1, m_gam_flow3_lag2)
```

It looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r gam flow3 lag2 summary and diag, warning = FALSE}
summary(m_gam_flow3_lag2)
appraise(m_gam_flow3_lag2)
shapiro.test(residuals(m_gam_flow3_lag2))
k.check(m_gam_flow3_lag2)
draw(m_gam_flow3_lag2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_flow3_lag2, pages = 1, all.terms = TRUE)
anova(m_gam_flow3_lag2)
```

The model diagnostics look pretty good. Note that the 2-way interaction between Year and Flow isn't significant. We'll use `m_gam_flow3_lag2` in the model selection process.

```{r gam flow3 clean up}
rm(m_gam_flow3, m_gam_flow3_lag1)
```

# Model 2: GAM Model with Flow and 2-way interactions

Now we'll try running a GAM using all two-way interactions between Year, Flow, and Station.

## Initial Model

```{r gam flow2 no autocorr, warning = FALSE}
m_gam_flow2 <- gam(
  Chla_log ~ (Year_fct + Flow + StationCode)^2 + s(Week, bs = "cc", k = 5),
  data = df_chla_c2,
  method = "REML",
  knots = list(week = c(0, 52))
)
```

Lets look at the model summary and diagnostics:

```{r gam flow2 no autocorr diag, warning = FALSE}
summary(m_gam_flow2)
appraise(m_gam_flow2)
shapiro.test(residuals(m_gam_flow2))
k.check(m_gam_flow2)
draw(m_gam_flow2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_flow2, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_flow2))
Box.test(residuals(m_gam_flow2), lag = 20, type = 'Ljung-Box')
```

Besides the Shapiro-Wilk normality test showing that the residuals aren't normal, the diagnostic plots look really good. However, the residuals are autocorrelated.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation. We'll run a series of models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r gam flow2 lag1, warning = FALSE}
m_gam_flow2_lag1 <- gam(
  Chla_log ~ (Year_fct + Flow + StationCode)^2 + s(Week, bs = "cc", k = 5) + lag1,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_flow2_lag1))
Box.test(residuals(m_gam_flow2_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r gam flow2 lag2, warning = FALSE}
m_gam_flow2_lag2 <- gam(
  Chla_log ~ (Year_fct + Flow + StationCode)^2 + s(Week, bs = "cc", k = 5) + lag1 + lag2,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_flow2_lag2))
Box.test(residuals(m_gam_flow2_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 1 lag term already seems to address the serial autocorrelation, but the lag2 model is even better. Let's use AIC to see how they compare.

### Compare Models

```{r gam flow2 compare}
AIC(m_gam_flow2, m_gam_flow2_lag1, m_gam_flow2_lag2)
```

Again, it looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r gam flow2 lag2 summary and diag, warning = FALSE}
summary(m_gam_flow2_lag2)
appraise(m_gam_flow2_lag2)
shapiro.test(residuals(m_gam_flow2_lag2))
k.check(m_gam_flow2_lag2)
draw(m_gam_flow2_lag2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_flow2_lag2, pages = 1, all.terms = TRUE)
anova(m_gam_flow2_lag2)
```

The model diagnostics look pretty good. Note that the 2-way interaction between Year and Flow isn't significant. We'll use `m_gam_flow2_lag2` in the model selection process.

```{r gam flow2 clean up}
rm(m_gam_flow2, m_gam_flow2_lag1)
```

# Model 3: GAM Model with 2-way interaction between Station and Year but without Flow

Next we'll try running a GAM using a two-way interaction between Year and Station but not including flow as a predictor.

## Initial Model

```{r gam cat2 no autocorr, warning = FALSE}
m_gam_cat2 <- gam(
  Chla_log ~ Year_fct * StationCode + s(Week, bs = "cc", k = 5),
  data = df_chla_c2,
  method = "REML",
  knots = list(week = c(0, 52))
)
```

Lets look at the model summary and diagnostics:

```{r gam cat2 no autocorr diag, warning = FALSE}
summary(m_gam_cat2)
appraise(m_gam_cat2)
shapiro.test(residuals(m_gam_cat2))
k.check(m_gam_cat2)
draw(m_gam_cat2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_cat2, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_cat2))
Box.test(residuals(m_gam_cat2), lag = 20, type = 'Ljung-Box')
```

Besides the Shapiro-Wilk normality test showing that the residuals aren't normal, the diagnostic plots look really good. However, the residuals are autocorrelated.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation. We'll run a series of models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r gam cat2 lag1, warning = FALSE}
m_gam_cat2_lag1 <- gam(
  Chla_log ~ Year_fct * StationCode + s(Week, bs = "cc", k = 5) + lag1,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_cat2_lag1))
Box.test(residuals(m_gam_cat2_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r gam cat2 lag2, warning = FALSE}
m_gam_cat2_lag2 <- gam(
  Chla_log ~ Year_fct * StationCode + s(Week, bs = "cc", k = 5) + lag1 + lag2,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_cat2_lag2))
Box.test(residuals(m_gam_cat2_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 1 lag term already seems to address the serial autocorrelation, but the lag2 model is even better. Let's use AIC to see how they compare.

### Compare Models

```{r gam cat2 compare}
AIC(m_gam_cat2, m_gam_cat2_lag1, m_gam_cat2_lag2)
```

Again, it looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r gam cat2 lag2 summary and diag, warning = FALSE}
summary(m_gam_cat2_lag2)
appraise(m_gam_cat2_lag2)
shapiro.test(residuals(m_gam_cat2_lag2))
k.check(m_gam_cat2_lag2)
draw(m_gam_cat2_lag2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_cat2_lag2, pages = 1, all.terms = TRUE)
anova(m_gam_cat2_lag2)
```

The model diagnostics look pretty good but not quite as good as with the initial model. We'll use `m_gam_cat2_lag2` in the model selection process.

```{r gam cat2 clean up}
rm(m_gam_cat2, m_gam_cat2_lag1)
```

# Model 4: Linear Model with Flow and 3-way interactions

Let's try the weekly average model as a linear model with a three-way interaction between Year, Weekly Average Flow, and Station but without the smooth term for week number. Initially, we'll run the model without accounting for serial autocorrelation.

## Initial Model

```{r lm flow3 no autocorr}
m_lm_flow3 <- lm(Chla_log ~ Year_fct * Flow * StationCode, data = df_chla_c2)
```

Lets look at the model summary and diagnostics:

```{r lm flow3 no autocorr diag}
summary(m_lm_flow3)

df_chla_c2 %>% plot_lm_diag(Chla_log, m_lm_flow3)

shapiro.test(residuals(m_lm_flow3))
acf(residuals(m_lm_flow3))
Box.test(residuals(m_lm_flow3), lag = 20, type = 'Ljung-Box')
```

The residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, model definitely has residual autocorrelation as indicated by the ACF plot and the Box-Ljung test.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r lm flow3 lag1}
m_lm_flow3_lag1 <- df_chla_c2_lag %>%
  drop_na(Chla_log, lag1) %>%
  lm(Chla_log ~ Year_fct * Flow * StationCode + lag1, data = .)

acf(residuals(m_lm_flow3_lag1))
Box.test(residuals(m_lm_flow3_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r lm flow3 lag2}
m_lm_flow3_lag2 <- df_chla_c2_lag %>%
  drop_na(Chla_log, lag1, lag2) %>%
  lm(Chla_log ~ Year_fct * Flow * StationCode + lag1 + lag2, data = .)

acf(residuals(m_lm_flow3_lag2))
Box.test(residuals(m_lm_flow3_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 1 lag term already has better ACF and Box-Ljung test results than the initial model. Let's use AIC to see how they compare.

### Compare Models

```{r lm flow3 compare, warning = FALSE}
AIC(m_lm_flow3, m_lm_flow3_lag1, m_lm_flow3_lag2)
```

Again, it looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r lm flow3 lag2 summary and diag}
summary(m_lm_flow3_lag2)

df_chla_c2_lag %>%
  drop_na(Chla_log, lag1, lag2) %>%
  plot_lm_diag(Chla_log, m_lm_flow3_lag2)

shapiro.test(residuals(m_lm_flow3_lag2))

Anova(m_lm_flow3_lag2, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

The model diagnostics look okay, but not as good as with the GAM models. We'll use `m_lm_flow3_lag2` in the model selection process.

```{r lm flow3 clean up}
rm(m_lm_flow3, m_lm_flow3_lag1)
```

# Model 5: Linear Model with Flow and 2-way interactions

Let's try a linear model using all two-way interactions between Year, Weekly Average Flow, and Station. Initially, we'll run the model without accounting for serial autocorrelation.

## Initial Model

```{r lm flow2 no autocorr}
m_lm_flow2 <- lm(Chla_log ~ (Year_fct + Flow + StationCode)^2, data = df_chla_c2)
```

Lets look at the model summary and diagnostics:

```{r lm flow2 no autocorr diag}
summary(m_lm_flow2)

df_chla_c2 %>% plot_lm_diag(Chla_log, m_lm_flow2)

shapiro.test(residuals(m_lm_flow2))
acf(residuals(m_lm_flow2))
Box.test(residuals(m_lm_flow2), lag = 20, type = 'Ljung-Box')
```

The residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, model definitely has residual autocorrelation as indicated by the ACF plot and the Box-Ljung test.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r lm flow2 lag1}
m_lm_flow2_lag1 <- df_chla_c2_lag %>%
  drop_na(Chla_log, lag1) %>%
  lm(Chla_log ~ (Year_fct + Flow + StationCode)^2 + lag1, data = .)

acf(residuals(m_lm_flow2_lag1))
Box.test(residuals(m_lm_flow2_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r lm flow2 lag2}
m_lm_flow2_lag2 <- df_chla_c2_lag %>%
  drop_na(Chla_log, lag1, lag2) %>%
  lm(Chla_log ~ (Year_fct + Flow + StationCode)^2 + lag1 + lag2, data = .)

acf(residuals(m_lm_flow2_lag2))
Box.test(residuals(m_lm_flow2_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 2 lag terms seems to be okay in terms of serial autocorrelation. Let's use AIC to see how they compare.

### Compare Models

```{r lm flow2 compare, warning = FALSE}
AIC(m_lm_flow2, m_lm_flow2_lag1, m_lm_flow2_lag2)
```

Again, it looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r lm flow2 lag2 summary and diag}
summary(m_lm_flow2_lag2)

df_chla_c2_lag %>%
  drop_na(Chla_log, lag1, lag2) %>%
  plot_lm_diag(Chla_log, m_lm_flow2_lag2)

shapiro.test(residuals(m_lm_flow2_lag2))

Anova(m_lm_flow2_lag2, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

The model diagnostics look somewhat worse than those for the 3-way interaction model. Note that the 2-way interaction between Year and Flow isn't significant. We'll use `m_lm_flow2_lag2` in the model selection process.

```{r lm flow2 clean up}
rm(m_lm_flow2, m_lm_flow2_lag1)
```

# Model 6: Linear Model with 2-way interaction between Station and Year but without Flow

We'll try running a linear model using a two-way interaction between Year and Station but not including flow as a predictor. Initially, we'll run the model without accounting for serial autocorrelation.

## Initial Model

```{r lm cat2 no autocorr}
m_lm_cat2 <- lm(Chla_log ~ Year_fct * StationCode, data = df_chla_c2)
```

Lets look at the model summary and diagnostics:

```{r lm cat2 no autocorr diag}
summary(m_lm_cat2)

df_chla_c2 %>% plot_lm_diag(Chla_log, m_lm_cat2)

shapiro.test(residuals(m_lm_cat2))
acf(residuals(m_lm_cat2))
Box.test(residuals(m_lm_cat2), lag = 20, type = 'Ljung-Box')
```

Besides the Shapiro-Wilk normality test showing that the residuals aren't normal, the diagnostic plots look pretty good. However, the residuals are autocorrelated.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r lm cat2 lag1}
m_lm_cat2_lag1 <- df_chla_c2_lag %>%
  drop_na(Chla_log, lag1) %>%
  lm(Chla_log ~ Year_fct * StationCode + lag1, data = .)

acf(residuals(m_lm_cat2_lag1))
Box.test(residuals(m_lm_cat2_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r lm cat2 lag2}
m_lm_cat2_lag2 <- df_chla_c2_lag %>%
  drop_na(Chla_log, lag1, lag2) %>%
  lm(Chla_log ~ Year_fct * StationCode + lag1 + lag2, data = .)

acf(residuals(m_lm_cat2_lag2))
Box.test(residuals(m_lm_cat2_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 2 lag terms seems to be okay in terms of serial autocorrelation. Let's use AIC to see how they compare.

### Compare Models

```{r lm cat2 compare, warning = FALSE}
AIC(m_lm_cat2, m_lm_cat2_lag1, m_lm_cat2_lag2)
```

Again, it looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r lm cat2 lag2 summary and diag}
summary(m_lm_cat2_lag2)

df_chla_c2_lag %>%
  drop_na(Chla_log, lag1, lag2) %>%
  plot_lm_diag(Chla_log, m_lm_cat2_lag2)

shapiro.test(residuals(m_lm_cat2_lag2))

Anova(m_lm_cat2_lag2, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

The model diagnostics don't look that great. However, we'll use `m_lm_cat2_lag2` in the model selection process.

```{r lm cat2 clean up}
rm(m_lm_cat2, m_lm_cat2_lag1)
```

# Model 7: GAM model using smooths for Flow

Finally, we'll try running a GAM model using smooths for weekly average flow by Station and Year and a smooth term for week number to account for seasonality. Initially, we'll run the model without accounting for serial autocorrelation.

## Initial Model

```{r gam sflow no autocorr, warning = FALSE}
m_gam_sflow <- gam(
  Chla_log ~ s(Flow, by = StationCode) + s(Flow, by = Year_fct) + Year_fct * StationCode + s(Week, bs = "cc", k = 5),
  data = df_chla_c2,
  method = "REML",
  knots = list(week = c(0, 52))
)
```

Lets look at the model summary and diagnostics:

```{r gam sflow no autocorr diag, warning = FALSE}
summary(m_gam_sflow)
appraise(m_gam_sflow)
shapiro.test(residuals(m_gam_sflow))
k.check(m_gam_sflow)
concurvity(m_gam_sflow, full = FALSE)$worst
draw(m_gam_sflow, select = 10, residuals = TRUE, rug = FALSE)
plot(m_gam_sflow, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_sflow))
Box.test(residuals(m_gam_sflow), lag = 20, type = 'Ljung-Box')
```

The diagnostic plots look really good. However, the residuals are autocorrelated.

## Model with lag terms

Now, we'll try to deal with the residual autocorrelation. We'll run a series of models adding 1 and 2 lag terms and compare how well they correct for autocorrelation.

### Lag 1

```{r gam sflow lag1, warning = FALSE}
m_gam_sflow_lag1 <- gam(
  Chla_log ~ s(Flow, by = StationCode) + s(Flow, by = Year_fct) + Year_fct * StationCode + s(Week, bs = "cc", k = 5) + lag1,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_sflow_lag1))
Box.test(residuals(m_gam_sflow_lag1), lag = 20, type = 'Ljung-Box')
```

### Lag 2

```{r gam sflow lag2, warning = FALSE}
m_gam_sflow_lag2 <- gam(
  Chla_log ~ s(Flow, by = StationCode) + s(Flow, by = Year_fct) + Year_fct * StationCode + s(Week, bs = "cc", k = 5) + lag1 + lag2,
  data = df_chla_c2_lag,
  method = "REML",
  knots = list(week = c(0, 52))
)

acf(residuals(m_gam_sflow_lag2))
Box.test(residuals(m_gam_sflow_lag2), lag = 20, type = 'Ljung-Box')
```

The model with 2 lag terms seems to be okay in terms of serial autocorrelation. Let's use AIC to see how they compare.

### Compare Models

```{r gam sflow compare}
AIC(m_gam_sflow, m_gam_sflow_lag1, m_gam_sflow_lag2)
```

It looks like the lag2 model has the best fit according to the AIC values. Let's take a closer look at that one.

### Lag 2 model summary

```{r gam sflow lag2 summary and diag, warning = FALSE}
summary(m_gam_sflow_lag2)
appraise(m_gam_sflow_lag2)
shapiro.test(residuals(m_gam_sflow_lag2))
k.check(m_gam_sflow_lag2)
concurvity(m_gam_sflow_lag2, full = FALSE)$worst
draw(m_gam_sflow_lag2, select = 1:4, residuals = TRUE, rug = FALSE)
draw(m_gam_sflow_lag2, select = 5:9, residuals = TRUE, rug = FALSE)
draw(m_gam_sflow_lag2, select = 10, residuals = TRUE, rug = FALSE)
anova(m_gam_sflow_lag2)
```

The model diagnostics look a little worse than with the initial model but they still look pretty good. We'll use `m_gam_sflow_lag2` in the model selection process.

```{r gam sflow clean up}
rm(m_gam_sflow, m_gam_sflow_lag1)
```

# Model selection

Now we'll compare the seven candidate models with AIC to select the one with the best fit. As a summary, here are the 7 models we are comparing:

* Model 1 - `m_gam_flow3_lag2` - GAM 3-way interactions with s(Week) <br>
Formula: `r deparse1(m_gam_flow3_lag2$formula)` <br><br>
* Model 2 - `m_gam_flow2_lag2` - GAM 2-way interactions with s(Week) <br>
Formula: `r deparse1(m_gam_flow2_lag2$formula)` <br><br>
* Model 3 - `m_gam_cat2_lag2` - GAM 2-way interaction between Station and Year with s(Week) but without Flow <br>
Formula: `r deparse1(m_gam_cat2_lag2$formula)` <br><br>
* Model 4 - `m_lm_flow3_lag2` - LM 3-way interactions without seasonal term <br>
Formula: `r deparse1(m_lm_flow3_lag2$call) %>% str_sub(start = 14, end = -12)` <br><br>
* Model 5 - `m_lm_flow2_lag2` - LM 2-way interactions without seasonal term <br>
Formula: `r deparse1(m_lm_flow2_lag2$call) %>% str_sub(start = 14, end = -12)` <br><br>
* Model 6 - `m_lm_cat2_lag2` - LM 2-way interaction between Station and Year but without Flow and seasonal term <br>
Formula: `r deparse1(m_lm_cat2_lag2$call) %>% str_sub(start = 14, end = -12)` <br><br>
* Model 7 - `m_gam_sflow_lag2` - GAM using smooths for Flow with s(Week) <br>
Formula: `r deparse1(m_gam_sflow_lag2$formula)` <br>

```{r model selection}
# AIC values
df_m_aic <-
  AIC(
    m_gam_flow3_lag2,
    m_gam_flow2_lag2,
    m_gam_cat2_lag2,
    m_lm_flow3_lag2,
    m_lm_flow2_lag2,
    m_lm_cat2_lag2,
    m_gam_sflow_lag2
  ) %>%
  as_tibble(rownames = "Model") %>%
  mutate(Model_Number = 1:7, .before = Model)

# BIC values
df_m_bic <-
  BIC(
    m_gam_flow3_lag2,
    m_gam_flow2_lag2,
    m_gam_cat2_lag2,
    m_lm_flow3_lag2,
    m_lm_flow2_lag2,
    m_lm_cat2_lag2,
    m_gam_sflow_lag2
  ) %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
df_m_aic_bic <-
  left_join(df_m_aic, df_m_bic, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
df_m_aic_bic %>% arrange(AIC)
```

Export the AIC/BIC table.

```{r export aic table}
df_m_aic_bic %>%
  arrange(AIC) %>%
  mutate(across(where(is.numeric), ~ paste0(formatC(.x, digits = 1, format = "f"), "##"))) %>%
  write_csv(here("manuscript_synthesis/results/tables/chl_aic_weekly_models.csv"))
```

According to AIC, Model 1 (GAM 3-way interactions with s(Week)) was the model with the best fit. BIC preferred Model 2 (GAM 2-way interactions with s(Week)). BIC prefers simpler models, so I'm inclined to go with Model 2, especially since we really don't need to look at the 3-way interaction. Before we proceed with either Models 1 or 2, let's revisit the model diagnostics for each and take a closer look at how the back-transformed fitted values from the models match the observed values.

# Model 1 diagnostics

```{r gam flow3 lag2 diag revist, warning = FALSE}
appraise(m_gam_flow3_lag2)
shapiro.test(residuals(m_gam_flow3_lag2))
k.check(m_gam_flow3_lag2)
draw(m_gam_flow3_lag2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_flow3_lag2, pages = 1, all.terms = TRUE)
anova(m_gam_flow3_lag2)
```

# Model 1 observed vs fitted values

```{r gam flow3 lag2 obs vs fit, warning = FALSE}
df_chla_c2_lag2 <- df_chla_c2_lag %>% drop_na(lag1, lag2)

df_m_gam_flow3_lag2_fit <- df_chla_c2_lag2 %>%
  fitted_values(m_gam_flow3_lag2, data = .) %>%
  mutate(fitted_bt = exp(.fitted) / 1000)

plt_m_gam_flow3_lag2_fit <- df_m_gam_flow3_lag2_fit %>%
  ggplot(aes(x = fitted_bt, y = Chla)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_bw() +
  labs(
    x = "Back-transformed Fitted Values",
    y = "Observed Values"
  )

plt_m_gam_flow3_lag2_fit
```

Let's group by station.

```{r gam flow3 lag2 obs vs fit facet sta}
plt_m_gam_flow3_lag2_fit + facet_wrap(vars(StationCode), scales = "free")
```

Now, group by year.

```{r gam flow3 lag2 obs vs fit facet yr}
plt_m_gam_flow3_lag2_fit + facet_wrap(vars(Year_fct), scales = "free")
```

# Model 2 diagnostics

```{r gam flow2 lag2 diag revist, warning = FALSE}
appraise(m_gam_flow2_lag2)
shapiro.test(residuals(m_gam_flow2_lag2))
k.check(m_gam_flow2_lag2)
draw(m_gam_flow2_lag2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_flow2_lag2, pages = 1, all.terms = TRUE)
anova(m_gam_flow2_lag2)
```

# Model 2 observed vs fitted values

```{r gam flow2 lag2 obs vs fit, warning = FALSE}
df_m_gam_flow2_lag2_fit <- df_chla_c2_lag2 %>%
  fitted_values(m_gam_flow2_lag2, data = .) %>%
  mutate(fitted_bt = exp(.fitted) / 1000)

plt_m_gam_flow2_lag2_fit <- df_m_gam_flow2_lag2_fit %>%
  ggplot(aes(x = fitted_bt, y = Chla)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme_bw() +
  labs(
    x = "Back-transformed Fitted Values",
    y = "Observed Values"
  )

plt_m_gam_flow2_lag2_fit
```

Let's group by station.

```{r gam flow2 lag2 obs vs fit facet sta}
plt_m_gam_flow2_lag2_fit + facet_wrap(vars(StationCode), scales = "free")
```

Now, group by year.

```{r gam flow2 lag2 obs vs fit facet yr}
plt_m_gam_flow2_lag2_fit + facet_wrap(vars(Year_fct), scales = "free")
```

Everything looks pretty decent with both models. Not perfect, but pretty good given the number of data points. Note that variability does increase as the chlorophyll values increase. Before proceeding with either Model 1 or 2, let's look more closely at their results.

# Model 1 Results

## Effect of Flow by Station

```{r gam flow3 lag2 flow station effects plot}
# Calculate min and max flows for each station to narrow down x-axis in the plot
df_chla_flow_sta_summ <- df_chla_c2_lag2 %>%
  summarize(
    Flow_min = min(Flow),
    Flow_max = max(Flow),
    .by = c(StationCode)
  ) %>%
  mutate(
    Flow_buffer = (Flow_max - Flow_min) * 0.05,
    Flow_min = Flow_min - Flow_buffer,
    Flow_max = Flow_max + Flow_buffer
  )

# Calculate effects of flow on chlorophyll for each station holding the
  # non-focal variables constant - marginal effects/adjusted predictions
df_gam_flow3_sta_eff <-
  as.data.frame(
    predict_response(
      m_gam_flow3_lag2,
      terms = c("Flow", "StationCode"),
      margin = "marginalmeans"
    ),
    terms_to_colnames = TRUE
  ) %>%
  as_tibble() %>%
  # Narrow down range of flow values for each station
  left_join(df_chla_flow_sta_summ, by = join_by(StationCode)) %>%
  filter(Flow >= Flow_min & Flow <= Flow_max) %>%
  transmute(
    StationCode,
    Flow,
    # Back calculate model fits and confidence levels
    across(c(predicted, conf.low, conf.high), ~ exp(.x) / 1000)
  )

# Create effects plot
plt_gam_flow3_sta_eff <- df_gam_flow3_sta_eff %>%
  ggplot(aes(x = Flow, y = predicted)) +
  geom_point(
    data = df_chla_c2_lag2,
    aes(y = Chla, color = Year_fct),
    alpha = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
  facet_wrap(vars(StationCode), scales = "free") +
  theme_bw() +
  labs(
    x = "Flow (cfs)",
    y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
  ) +
  scale_x_continuous(breaks = breaks_extended(6)) +
  scale_color_viridis_d(name = "Year", option = "C")

plt_gam_flow3_sta_eff
```

There is a lot of uncertainty in the model results at the highest flows. This seems problematic.

## Effect of Flow by Year

```{r gam flow3 lag2 flow year effects plot, warning = FALSE, fig.width = 8}
# Calculate min and max flows for each station to narrow down x-axis in the plot
df_chla_flow_yr_summ <- df_chla_c2_lag2 %>%
  summarize(
    Flow_min = min(Flow),
    Flow_max = max(Flow),
    .by = c(Year_fct)
  ) %>%
  mutate(
    Flow_buffer = (Flow_max - Flow_min) * 0.05,
    Flow_min = Flow_min - Flow_buffer,
    Flow_max = Flow_max + Flow_buffer
  )

# Calculate effects of flow on chlorophyll for each year holding the
  # non-focal variables constant - marginal effects/adjusted predictions
df_gam_flow3_yr_eff <-
  as.data.frame(
    predict_response(
      m_gam_flow3_lag2,
      terms = c("Flow", "Year_fct"),
      margin = "marginalmeans"
    ),
    terms_to_colnames = TRUE
  ) %>%
  as_tibble() %>%
  # Narrow down range of flow values for each station
  left_join(df_chla_flow_yr_summ, by = join_by(Year_fct)) %>%
  filter(Flow >= Flow_min & Flow <= Flow_max) %>%
  transmute(
    Year_fct,
    Flow,
    # Back calculate model fits and confidence levels
    across(c(predicted, conf.low, conf.high), ~ exp(.x) / 1000)
  )

# Create effects plot
plt_gam_flow3_yr_eff <- df_gam_flow3_yr_eff %>%
  ggplot(aes(x = Flow, y = predicted)) +
  geom_point(
    data = df_chla_c2_lag2,
    aes(y = Chla, color = StationCode),
    alpha = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
  facet_wrap(vars(Year_fct), scales = "free") +
  theme_bw() +
  labs(
    x = "Flow (cfs)",
    y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
  ) +
  scale_x_continuous(breaks = breaks_extended(6)) +
  scale_color_viridis_d(name = "Station", option = "C") +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.3)
  )

plt_gam_flow3_yr_eff
```

These results look reasonable.

## Station by Year Contrasts

```{r gam flow3 lag2 contrasts station by year, warning = FALSE}
# Estimated marginal means for station by year
em_gam3_sta_yr <- emmeans(m_gam_flow3_lag2, ~ StationCode | Year_fct)

# Tukey post-hoc contrasts
pairs(em_gam3_sta_yr)

# Create table of contrasts and convert it to a tibble for plot
df_gam3_sta_yr <- em_gam3_sta_yr %>%
  cld(sort = FALSE, Letters = letters) %>%
  as_tibble() %>%
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    # back transform log-transformed results
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
  ) %>%
  # Add min and max values of observed data to the Tukey post-hoc results and
    # calculate vertical positioning of letters
  left_join(
    df_chla_c2_lag2 %>%
      summarize(
        max_val = max(Chla),
        min_val = min(Chla),
        .by = Year_fct
    ),
    by = join_by(Year_fct)
  ) %>%
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>%
  group_by(Year_fct) %>%
  mutate(max_val = max(max_val)) %>%
  ungroup() %>%
  mutate(y_pos = max_val + (max_val - min_val) / 10) %>%
  select(
    StationCode,
    Year_fct,
    emmean,
    lower.CL,
    upper.CL,
    group,
    y_pos
  )

# Create boxplot showing Tukey post-hoc results
plt_gam3_sta_yr <- df_gam3_sta_yr %>%
  ggplot(
    aes(
      x = StationCode,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL
    )
  ) +
  geom_boxplot(
    data = df_chla_c2_lag2,
    aes(x = StationCode, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos, label = group), size = 3.5) +
  facet_wrap(vars(Year_fct), scales = "free_y") +
  xlab("Station") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw()

plt_gam3_sta_yr
```

Many of the contrasts are significant. The model under predicts some values, and 2017 has a lot of uncertainty.

## Year by Station Contrasts

```{r gam flow3 lag2 contrasts year by station, warning = FALSE}
# Estimated marginal means for year by station
em_gam3_yr_sta <- emmeans(m_gam_flow3_lag2, ~ Year_fct | StationCode)

# Tukey post-hoc contrasts
pairs(em_gam3_yr_sta)

# Create table of contrasts and convert it to a tibble for plot
df_gam3_yr_sta <- em_gam3_yr_sta %>%
  cld(sort = FALSE, Letters = letters) %>%
  as_tibble() %>%
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    # back transform log-transformed results
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
  ) %>%
  # Add min and max values of observed data to the Tukey post-hoc results and
    # calculate vertical positioning of letters
  left_join(
    df_chla_c2_lag2 %>%
      summarize(
        max_val = max(Chla),
        min_val = min(Chla),
        .by = StationCode
    ),
    by = join_by(StationCode)
  ) %>%
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>%
  group_by(StationCode) %>%
  mutate(max_val = max(max_val)) %>%
  ungroup() %>%
  mutate(y_pos = max_val + (max_val - min_val) / 10) %>%
  select(
    StationCode,
    Year_fct,
    emmean,
    lower.CL,
    upper.CL,
    group,
    y_pos
  )

# Create boxplot showing Tukey post-hoc results
plt_gam3_yr_sta <- df_gam3_yr_sta %>%
  ggplot(
    aes(
      x = Year_fct,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL
    )
  ) +
  geom_boxplot(
    data = df_chla_c2_lag2,
    aes(x = Year_fct, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos, label = group), size = 3.5) +
  facet_wrap(vars(StationCode), scales = "free_y") +
  xlab("Year") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw()

plt_gam3_yr_sta
```

Again, the model under predicts some values, and 2017 has a lot of uncertainty.

# Model 2 Results

## Effect of Flow by Station

```{r gam flow2 lag2 flow station effects plot}
# Calculate effects of flow on chlorophyll for each station holding the
  # non-focal variables constant - marginal effects/adjusted predictions
df_gam_flow2_sta_eff <-
  as.data.frame(
    predict_response(
      m_gam_flow2_lag2,
      terms = c("Flow", "StationCode"),
      margin = "marginalmeans"
    ),
    terms_to_colnames = TRUE
  ) %>%
  as_tibble() %>%
  # Narrow down range of flow values for each station
  left_join(df_chla_flow_sta_summ, by = join_by(StationCode)) %>%
  filter(Flow >= Flow_min & Flow <= Flow_max) %>%
  transmute(
    StationCode,
    Flow,
    # Back calculate model fits and confidence levels
    across(c(predicted, conf.low, conf.high), ~ exp(.x) / 1000)
  )

# Create effects plot
plt_gam_flow2_sta_eff <- df_gam_flow2_sta_eff %>%
  ggplot(aes(x = Flow, y = predicted)) +
  geom_point(
    data = df_chla_c2_lag2,
    aes(y = Chla, color = Year_fct),
    alpha = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
  facet_wrap(vars(StationCode), scales = "free") +
  theme_bw() +
  labs(
    x = "Flow (cfs)",
    y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
  ) +
  scale_x_continuous(breaks = breaks_extended(6)) +
  scale_color_viridis_d(name = "Year", option = "C")

plt_gam_flow2_sta_eff
```

There is much less uncertainty in the model results at the highest flows when compared to model 1. These results seem more reasonable.

## Effect of Flow by Year

```{r gam flow2 lag2 flow year effects plot, warning = FALSE, fig.width = 8}
# Calculate effects of flow on chlorophyll for each year holding the
  # non-focal variables constant - marginal effects/adjusted predictions
df_gam_flow2_yr_eff <-
  as.data.frame(
    predict_response(
      m_gam_flow2_lag2,
      terms = c("Flow", "Year_fct"),
      margin = "marginalmeans"
    ),
    terms_to_colnames = TRUE
  ) %>%
  as_tibble() %>%
  # Narrow down range of flow values for each station
  left_join(df_chla_flow_yr_summ, by = join_by(Year_fct)) %>%
  filter(Flow >= Flow_min & Flow <= Flow_max) %>%
  transmute(
    Year_fct,
    Flow,
    # Back calculate model fits and confidence levels
    across(c(predicted, conf.low, conf.high), ~ exp(.x) / 1000)
  )

# Create effects plot
plt_gam_flow2_yr_eff <- df_gam_flow2_yr_eff %>%
  ggplot(aes(x = Flow, y = predicted)) +
  geom_point(
    data = df_chla_c2_lag2,
    aes(y = Chla, color = StationCode),
    alpha = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25) +
  facet_wrap(vars(Year_fct), scales = "free") +
  theme_bw() +
  labs(
    x = "Flow (cfs)",
    y = expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))
  ) +
  scale_x_continuous(breaks = breaks_extended(6)) +
  scale_color_viridis_d(name = "Station", option = "C") +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.3)
  )

plt_gam_flow2_yr_eff
```

These results look reasonable, but not interesting.

## Station by Year Contrasts

```{r gam flow2 lag2 contrasts station by year, warning = FALSE}
# Estimated marginal means for station by year
em_gam2_sta_yr <- emmeans(m_gam_flow2_lag2, ~ StationCode | Year_fct)

# Tukey post-hoc contrasts
pairs(em_gam2_sta_yr)

# Create table of contrasts and convert it to a tibble for plot
df_gam2_sta_yr <- em_gam2_sta_yr %>%
  cld(sort = FALSE, Letters = letters) %>%
  as_tibble() %>%
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    # back transform log-transformed results
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
  ) %>%
  # Add min and max values of observed data to the Tukey post-hoc results and
    # calculate vertical positioning of letters
  left_join(
    df_chla_c2_lag2 %>%
      summarize(
        max_val = max(Chla),
        min_val = min(Chla),
        .by = Year_fct
    ),
    by = join_by(Year_fct)
  ) %>%
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>%
  group_by(Year_fct) %>%
  mutate(max_val = max(max_val)) %>%
  ungroup() %>%
  mutate(y_pos = max_val + (max_val - min_val) / 10) %>%
  select(
    StationCode,
    Year_fct,
    emmean,
    lower.CL,
    upper.CL,
    group,
    y_pos
  )

# Create boxplot showing Tukey post-hoc results
plt_gam2_sta_yr <- df_gam2_sta_yr %>%
  ggplot(
    aes(
      x = StationCode,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL
    )
  ) +
  geom_boxplot(
    data = df_chla_c2_lag2,
    aes(x = StationCode, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos, label = group), size = 3.5) +
  facet_wrap(vars(Year_fct), scales = "free_y") +
  xlab("Station") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw()

plt_gam2_sta_yr
```

Many of the contrasts are significant. The model under predicts RD22, but the uncertainty looks reasonable for all values.

## Year by Station Contrasts

```{r gam flow2 lag2 contrasts year by station, warning = FALSE}
# Estimated marginal means for year by station
em_gam2_yr_sta <- emmeans(m_gam_flow2_lag2, ~ Year_fct | StationCode)

# Tukey post-hoc contrasts
pairs(em_gam2_yr_sta)

# Create table of contrasts and convert it to a tibble for plot
df_gam2_yr_sta <- em_gam2_yr_sta %>%
  cld(sort = FALSE, Letters = letters) %>%
  as_tibble() %>%
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    # back transform log-transformed results
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000)
  ) %>%
  # Add min and max values of observed data to the Tukey post-hoc results and
    # calculate vertical positioning of letters
  left_join(
    df_chla_c2_lag2 %>%
      summarize(
        max_val = max(Chla),
        min_val = min(Chla),
        .by = StationCode
    ),
    by = join_by(StationCode)
  ) %>%
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>%
  group_by(StationCode) %>%
  mutate(max_val = max(max_val)) %>%
  ungroup() %>%
  mutate(y_pos = max_val + (max_val - min_val) / 10) %>%
  select(
    StationCode,
    Year_fct,
    emmean,
    lower.CL,
    upper.CL,
    group,
    y_pos
  )

# Create boxplot showing Tukey post-hoc results
plt_gam2_yr_sta <- df_gam2_yr_sta %>%
  ggplot(
    aes(
      x = Year_fct,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL
    )
  ) +
  geom_boxplot(
    data = df_chla_c2_lag2,
    aes(x = Year_fct, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos, label = group), size = 3.5) +
  facet_wrap(vars(StationCode), scales = "free_y") +
  xlab("Year") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw()

plt_gam2_yr_sta
```

Again, the model under predicts RD22, but the uncertainty looks reasonable for all values.

## Conclusions

After looking at the results more closely from Models 1 and 2, I think we should use Model 2 (GAM 2-way interactions with s(Week)) because it's the simpler model, supported by BIC, and the uncertainty around model predictions is less than with Model 1.
