---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Models using categorical predictors - Revised Analysis"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Purpose

Explore and analyze the continuous chlorophyll data to be included in the NDFS synthesis manuscript. We will attempt to fit multiple models to the data set and use a model selection process to determine the best one. At a minimum the model will contain the three categorical variables (Year, Station, and Flow Pulse Period) and optionally a term to account for seasonality (either daily average water temperature or a GAM smooth for DOY). These models will only include representative stations for 4 habitat types - upstream (RD22), lower Yolo Bypass (STTD), Cache Slough complex (LIB), and downstream (RVB).

# Global code and functions

```{r load packages and functions, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(knitr)
library(mgcv)
library(car)
library(gratia)
library(emmeans)
library(multcomp)
library(here)
library(conflicted)

# Source functions
source(here("manuscript_synthesis/src/global_functions.R"))

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::lag(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

Create functions for document:

```{r global funcs}
# Create summary figure for model results and observed data
plot_model_summary <- function(df_data, em_tuk_obj) {
  # Calculate min and max values of observed data for each Station - Year group to
    # determine vertical positioning of letters for figure
  df_data_summ <- df_data %>% 
    summarize(
      max_val = max(Chla),
      min_val = min(Chla),
      .by = c(StationCode, Year)
    )

  # Add significance grouping letters to the Tukey post-hoc results
  df_tuk <- em_tuk_obj %>% 
    cld(sort = FALSE, Letters = letters) %>% 
    as_tibble() %>% 
    mutate(
      group = str_remove_all(.group, fixed(" ")),
      # back transform log-transformed results
      across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000),
      Year = as.numeric(as.character(Year_fct))
    ) %>% 
    # Add min and max values of observed data to the Tukey post-hoc results and
      # calculate vertical positioning of letters
    left_join(df_data_summ, by = join_by(StationCode, Year)) %>% 
    mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>% 
    group_by(StationCode, Year) %>% 
    mutate(max_val = max(max_val)) %>% 
    ungroup() %>% 
    mutate(y_pos = max_val + (max_val - min_val) / 10) %>% 
    select(
      StationCode,
      Year,
      FlowActionPeriod,
      emmean,
      lower.CL,
      upper.CL,
      group,
      y_pos
    )
  
  # Create summary figure
  df_tuk %>%
    ggplot(
      aes(
        x = FlowActionPeriod,
        y = emmean,
        ymin = lower.CL,
        ymax = upper.CL
      )
    ) +
    geom_boxplot(
      data = df_data,
      aes(x = FlowActionPeriod, y = Chla),
      inherit.aes = FALSE
    ) +
    geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
    geom_point(color = "red") +
    geom_text(aes(y = y_pos, label = group), size = 3) +
    facet_wrap(
      vars(StationCode, Year), 
      scales = "free_y",
      nrow = 4,
      labeller = labeller(.multi_line = FALSE)
    ) +
    xlab("Flow Pulse Period") +
    ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
}
```


# Import Data

```{r import data}
# Define file path for processed data
fp_data <- here("manuscript_synthesis/data/processed")

# Import daily average water quality data
df_wq <- readRDS(file.path(fp_data, "wq_daily_avg_2013-2019.rds"))
```

# Prepare Data

```{r prepare chla data, message = FALSE}
# Create a vector for the factor order of StationCode
sta_order <- c("RD22", "STTD", "LIB", "RVB")

# Prepare chlorophyll and water temperature data for exploration and analysis
df_chla_wt_c <- df_wq %>% 
  select(StationCode, Date, Chla, WaterTemp) %>% 
  drop_na(Chla) %>% 
  # Filter to only include representative stations for 3 habitat types - RD22, STTD, LIB
  filter(StationCode %in% sta_order) %>% 
  mutate(
    # Scale and log transform chlorophyll values
    Chla_log = log(Chla * 1000),
    # Add Year and DOY variables
    Year = year(Date),
    DOY = yday(Date)
  ) %>% 
  # Add Flow Action Periods
  ndfa_action_periods() %>% 
  mutate(
    # Apply factor orders to FlowActionPeriod and StationCode
    FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")),
    StationCode = factor(StationCode, levels = sta_order),
    # Add a column for Year as a factor for the model
    Year_fct = factor(Year)
  ) %>% 
  arrange(StationCode, Date)
  
# Create another dataframe that has up to 3 lag variables for chlorophyll to be
  # used in the linear models
df_chla_wt_lag <- df_chla_wt_c %>% 
  # Fill in missing days for each StationCode-Year combination
  group_by(StationCode, Year) %>% 
  complete(Date = seq.Date(min(Date), max(Date), by = "1 day")) %>% 
  # Create lag variables of scaled log transformed chlorophyll values
  mutate(
    lag1 = lag(Chla_log),
    lag2 = lag(Chla_log, 2),
    lag3 = lag(Chla_log, 3)
  ) %>% 
  ungroup()
```

# Explore sample counts by Station

```{r chla sample counts station}
df_chla_wt_c %>% 
  summarize(
    min_date = min(Date),
    max_date = max(Date),
    num_samples = n(),
    .by = c(StationCode, Year, FlowActionPeriod)
  ) %>% 
  complete(StationCode, Year, FlowActionPeriod) %>% 
  arrange(StationCode, Year, FlowActionPeriod) %>% 
  kable()
```

Except for a few Station-Year-FlowPeriod combinations with low sample counts, it appears we have adequate data.

# Plots

```{r chla boxplot station, fig.width = 8, fig.height = 8}
df_chla_wt_c %>%
  ggplot(aes(x = FlowActionPeriod, y = Chla, fill = FlowActionPeriod)) +
  geom_boxplot() +
  facet_grid(cols = vars(Year), rows = vars(StationCode)) +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides = "l") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position = "top"
  )
```

# Models

## Model 1: GAM 3-way interactions with s(DOY)

### Initial Model

We'll try running a GAM using a three-way interaction between Year, Flow Action Period, and Station, and a smooth term for day of year to account for seasonality (restricting the k-value to 5 to reduce overfitting). First we'll run the GAM without accounting for serial autocorrelation.

```{r gam cat3 no autocorr, warning = FALSE}
m_gam_cat3 <- gam(
  Chla_log ~ Year_fct * FlowActionPeriod * StationCode + s(DOY, k = 5), 
  data = df_chla_wt_c,
  method = "REML"
)
```

Lets look at the model summary and diagnostics:

```{r gam cat3 no autocorr diag, warning = FALSE}
summary(m_gam_cat3)
appraise(m_gam_cat3)
k.check(m_gam_cat3)
draw(m_gam_cat3, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_cat3, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_cat3))
Box.test(residuals(m_gam_cat3), lag = 20, type = 'Ljung-Box')
```

### Model with AR() correlation structure

Now, we'll try to deal with the residual autocorrelation. We'll run AR(1), AR(2), and AR(3) models and compare them using AIC.

```{r gamm cat3 with AR comparison, warning = FALSE}
# Define model formula as an object
f_gam_cat3 <- as.formula("Chla_log ~ Year_fct * FlowActionPeriod * StationCode + s(DOY, k = 5)")

# Fit original model with k = 5 and three successive AR(p) models
m_gamm_cat3 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c,
  method = "REML"
)

m_gamm_cat3_ar1 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 1), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat3_ar2 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 2), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat3_ar3 <- gamm(
  f_gam_cat3, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 3), # grouped by Year_fct and StationCode
  method = "REML"
)

# Compare models
anova(
  m_gamm_cat3$lme, 
  m_gamm_cat3_ar1$lme, 
  m_gamm_cat3_ar2$lme, 
  m_gamm_cat3_ar3$lme
)
```

It looks like the AR(1) model has the best fit according to the AIC values and backed up by the BIC values. Let's take a closer look at that one.

### AR(1) Model

```{r gamm cat3 ar1 diag, warning = FALSE}
# Pull out the residuals and the GAM model
resid_cat3_ar1 <- residuals(m_gamm_cat3_ar1$lme, type = "normalized")
m_gamm_cat3_ar1_gam <- m_gamm_cat3_ar1$gam

summary(m_gamm_cat3_ar1_gam)
appraise(m_gamm_cat3_ar1_gam)
k.check(m_gamm_cat3_ar1_gam)
draw(m_gamm_cat3_ar1_gam, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gamm_cat3_ar1_gam, pages = 1, all.terms = TRUE)
acf(resid_cat3_ar1)
Box.test(resid_cat3_ar1, lag = 20, type = 'Ljung-Box')
```

The AR(1) model has much less residual autocorrelation, and the diagnostics plots look pretty good. We'll take a look at the ANOVA table to see if we need all the interaction terms included in the model.

```{r gamm cat3 ar1 anova, warning = FALSE}
# the anova.gam function is similar to a type 3 ANOVA
anova(m_gamm_cat3_ar1_gam)
```

The 3-way interaction term is significant in this model, so we'll use this one for the model selection process.

```{r gamm cat3 clean up}
rm(m_gam_cat3, m_gamm_cat3, m_gamm_cat3_ar2, m_gamm_cat3_ar3)
```


## Model 2: LM 3-way interactions with water temperature

### Initial Model

We'll try running a linear model using a three-way interaction between Year, Flow Action Period, and Station, and a covariate of water temperature to account for seasonality and its effect of primary production. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat3 wt no autocorr}
m_lm_cat3_wt <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat3 wt no autocorr diag}
summary(m_lm_cat3_wt)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  plot_lm_diag(Chla_log, m_lm_cat3_wt)

shapiro.test(m_lm_cat3_wt$residuals)
acf(residuals(m_lm_cat3_wt))
Box.test(residuals(m_lm_cat3_wt), lag = 20, type = 'Ljung-Box')
```

Hmmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat3 wt lag1}
m_lm_cat3_wt_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp + lag1, data = .)

acf(residuals(m_lm_cat3_wt_lag1))
Box.test(residuals(m_lm_cat3_wt_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat3 wt lag2}
m_lm_cat3_wt_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp + lag1 + lag2, data = .)

acf(residuals(m_lm_cat3_wt_lag2))
Box.test(residuals(m_lm_cat3_wt_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat3 wt lag3}
m_lm_cat3_wt_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + WaterTemp + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat3_wt_lag3))
Box.test(residuals(m_lm_cat3_wt_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat3 wt lag comparison, warning = FALSE}
AIC(m_lm_cat3_wt_lag1, m_lm_cat3_wt_lag2, m_lm_cat3_wt_lag3)
BIC(m_lm_cat3_wt_lag1, m_lm_cat3_wt_lag2, m_lm_cat3_wt_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat3 wt lag1 summary and diag}
summary(m_lm_cat3_wt_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  plot_lm_diag(Chla_log, m_lm_cat3_wt_lag1)

shapiro.test(m_lm_cat3_wt_lag1$residuals)
```

Hmmm, the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat3 wt lag1}
Anova(m_lm_cat3_wt_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right and the water temperature term isn't significant.

```{r lm cat3 wt clean up}
rm(m_lm_cat3_wt, m_lm_cat3_wt_lag2, m_lm_cat3_wt_lag3)
```


## Model 3: LM 3-way interactions without seasonal term

### Initial Model

We'll try running a linear model using a three-way interaction between Year, Flow Action Period, and Station without including a term to account for seasonality. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat3 no autocorr}
m_lm_cat3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat3 no autocorr diag}
summary(m_lm_cat3)

df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  plot_lm_diag(Chla_log, m_lm_cat3)

shapiro.test(m_lm_cat3$residuals)
acf(residuals(m_lm_cat3))
Box.test(residuals(m_lm_cat3), lag = 20, type = 'Ljung-Box')
```

Hmmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat3 lag1}
m_lm_cat3_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + lag1, data = .)

acf(residuals(m_lm_cat3_lag1))
Box.test(residuals(m_lm_cat3_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat3 lag2}
m_lm_cat3_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + lag1 + lag2, data = .)

acf(residuals(m_lm_cat3_lag2))
Box.test(residuals(m_lm_cat3_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat3 lag3}
m_lm_cat3_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ Year_fct * FlowActionPeriod * StationCode + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat3_lag3))
Box.test(residuals(m_lm_cat3_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat3 lag comparison, warning = FALSE}
AIC(m_lm_cat3_lag1, m_lm_cat3_lag2, m_lm_cat3_lag3)
BIC(m_lm_cat3_lag1, m_lm_cat3_lag2, m_lm_cat3_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat3 lag1 summary and diag}
summary(m_lm_cat3_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  plot_lm_diag(Chla_log, m_lm_cat3_lag1)

shapiro.test(m_lm_cat3_lag1$residuals)
```

Hmmm, again the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat3 lag1}
Anova(m_lm_cat3_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right.

```{r lm cat3 clean up}
rm(m_lm_cat3, m_lm_cat3_lag2, m_lm_cat3_lag3)
```


## Model 4: GAM 2-way interactions with s(DOY)

### Initial Model

We'll try running a GAM using all two-way interactions between Year, Flow Action Period, and Station, and a smooth term for day of year to account for seasonality (restricting the k-value to 5 to reduce overfitting). First we'll run the GAM without accounting for serial autocorrelation.

```{r gam cat2 no autocorr, warning = FALSE}
m_gam_cat2 <- gam(
  Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + s(DOY, k = 5), 
  data = df_chla_wt_c,
  method = "REML"
)
```

Lets look at the model summary and diagnostics:

```{r gam cat2 no autocorr diag, warning = FALSE}
summary(m_gam_cat2)
appraise(m_gam_cat2)
k.check(m_gam_cat2)
draw(m_gam_cat2, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gam_cat2, pages = 1, all.terms = TRUE)
acf(residuals(m_gam_cat2))
Box.test(residuals(m_gam_cat2), lag = 20, type = 'Ljung-Box')
```

### Model with AR() correlation structure

Now, we'll try to deal with the residual autocorrelation. We'll run AR(1), AR(2), and AR(3) models and compare them using AIC.

```{r gamm cat2 with AR comparison, warning = FALSE}
# Define model formula as an object
f_gam_cat2 <- as.formula("Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + s(DOY, k = 5)")

# Fit original model with k = 5 and three successive AR(p) models
m_gamm_cat2 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c,
  method = "REML"
)

m_gamm_cat2_ar1 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 1), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat2_ar2 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 2), # grouped by Year_fct and StationCode
  method = "REML"
)

m_gamm_cat2_ar3 <- gamm(
  f_gam_cat2, 
  data = df_chla_wt_c, 
  correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 3), # grouped by Year_fct and StationCode
  method = "REML"
)

# Compare models
anova(
  m_gamm_cat2$lme, 
  m_gamm_cat2_ar1$lme, 
  m_gamm_cat2_ar2$lme, 
  m_gamm_cat2_ar3$lme
)
```

It looks like the AR(1) model has the best fit according to the AIC values and backed up by the BIC values. Let's take a closer look at that one.

### AR(1) Model

```{r gamm cat2 ar1 diag, warning = FALSE}
# Pull out the residuals and the GAM model
resid_cat2_ar1 <- residuals(m_gamm_cat2_ar1$lme, type = "normalized")
m_gamm_cat2_ar1_gam <- m_gamm_cat2_ar1$gam

summary(m_gamm_cat2_ar1_gam)
appraise(m_gamm_cat2_ar1_gam)
k.check(m_gamm_cat2_ar1_gam)
draw(m_gamm_cat2_ar1_gam, select = 1, residuals = TRUE, rug = FALSE)
plot(m_gamm_cat2_ar1_gam, pages = 1, all.terms = TRUE)
acf(resid_cat2_ar1)
Box.test(resid_cat2_ar1, lag = 20, type = 'Ljung-Box')
```

The AR(1) model has much less residual autocorrelation, and the diagnostics plots look pretty good. We'll take a look at the ANOVA table to see if we need all the interaction terms included in the model.

```{r gamm cat2 ar1 anova, warning = FALSE}
# the anova.gam function is similar to a type 3 ANOVA
anova(m_gamm_cat2_ar1_gam)
```

All the interaction terms are significant in this model, so we'll use this one for the model selection process.

```{r gamm cat2 clean up}
rm(m_gam_cat2, m_gamm_cat2, m_gamm_cat2_ar2, m_gamm_cat2_ar3)
```


## Model 5: LM 2-way interactions with water temperature

### Initial Model

We'll try running a linear model using all two-way interactions between Year, Flow Action Period, and Station, and a covariate of water temperature to account for seasonality and its effect of primary production. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat2 wt no autocorr}
m_lm_cat2_wt <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat2 wt no autocorr diag}
summary(m_lm_cat2_wt)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp) %>% 
  plot_lm_diag(Chla_log, m_lm_cat2_wt)

shapiro.test(m_lm_cat2_wt$residuals)
acf(residuals(m_lm_cat2_wt))
Box.test(residuals(m_lm_cat2_wt), lag = 20, type = 'Ljung-Box')
```

Hmmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat2 wt lag1}
m_lm_cat2_wt_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp + lag1, data = .)

acf(residuals(m_lm_cat2_wt_lag1))
Box.test(residuals(m_lm_cat2_wt_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat2 wt lag2}
m_lm_cat2_wt_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp + lag1 + lag2, data = .)

acf(residuals(m_lm_cat2_wt_lag2))
Box.test(residuals(m_lm_cat2_wt_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat2 wt lag3}
m_lm_cat2_wt_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + WaterTemp + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat2_wt_lag3))
Box.test(residuals(m_lm_cat2_wt_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat2 wt lag comparison, warning = FALSE}
AIC(m_lm_cat2_wt_lag1, m_lm_cat2_wt_lag2, m_lm_cat2_wt_lag3)
BIC(m_lm_cat2_wt_lag1, m_lm_cat2_wt_lag2, m_lm_cat2_wt_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat2 wt lag1 summary and diag}
summary(m_lm_cat2_wt_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, WaterTemp, lag1) %>% 
  plot_lm_diag(Chla_log, m_lm_cat2_wt_lag1)

shapiro.test(m_lm_cat2_wt_lag1$residuals)
```

Hmmm, the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat2 wt lag1}
Anova(m_lm_cat2_wt_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right and the water temperature term isn't significant.

```{r lm cat2 wt clean up}
rm(m_lm_cat2_wt, m_lm_cat2_wt_lag2, m_lm_cat2_wt_lag3)
```


## Model 6: LM 2-way interactions without seasonal term

### Initial Model

We'll try running a linear model using all two-way interactions between Year, Flow Action Period, and Station without including a term to account for seasonality. First we'll run the linear model without accounting for serial autocorrelation.

```{r lm cat2 no autocorr}
m_lm_cat2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2, data = .)
```

Lets look at the model summary and diagnostics:

```{r lm cat2 no autocorr diag}
summary(m_lm_cat2)

df_chla_wt_lag %>% 
  drop_na(Chla_log) %>% 
  plot_lm_diag(Chla_log, m_lm_cat2)

shapiro.test(m_lm_cat2$residuals)
acf(residuals(m_lm_cat2))
Box.test(residuals(m_lm_cat2), lag = 20, type = 'Ljung-Box')
```

Hmmm, the residuals deviate from a normal distribution according to visual inspection and the Shapiro-Wilk normality test. Also, the residuals are autocorrelated.

### Model with lag terms

Now, we'll try to deal with the residual autocorrelation and the non-normal residuals. We'll run a series of linear models adding 1, 2, and 3 lag terms and compare how well they correct for autocorrelation.

#### Lag 1

```{r lm cat2 lag1}
m_lm_cat2_lag1 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + lag1, data = .)

acf(residuals(m_lm_cat2_lag1))
Box.test(residuals(m_lm_cat2_lag1), lag = 20, type = 'Ljung-Box')
```

#### Lag 2

```{r lm cat2 lag2}
m_lm_cat2_lag2 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + lag1 + lag2, data = .)

acf(residuals(m_lm_cat2_lag2))
Box.test(residuals(m_lm_cat2_lag2), lag = 20, type = 'Ljung-Box')
```

#### Lag 3

```{r lm cat2 lag3}
m_lm_cat2_lag3 <- df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1, lag2, lag3) %>% 
  lm(Chla_log ~ (Year_fct + FlowActionPeriod + StationCode)^2 + lag1 + lag2 + lag3, data = .)

acf(residuals(m_lm_cat2_lag3))
Box.test(residuals(m_lm_cat2_lag3), lag = 20, type = 'Ljung-Box')
```

The model with 3 lag terms looks to have slightly better ACF and Box-Ljung test results than the model with 1 lag term. Let's compare the 3 models using AIC and BIC to see which model those prefer.

#### Compare models

```{r lm cat2 lag comparison, warning = FALSE}
AIC(m_lm_cat2_lag1, m_lm_cat2_lag2, m_lm_cat2_lag3)
BIC(m_lm_cat2_lag1, m_lm_cat2_lag2, m_lm_cat2_lag3)
```

Both AIC and BIC prefer the model with 1 lag term. Let's take a closer look at that one.

### Lag 1 model summary

```{r lm cat2 lag1 summary and diag}
summary(m_lm_cat2_lag1)

df_chla_wt_lag %>% 
  drop_na(Chla_log, lag1) %>% 
  plot_lm_diag(Chla_log, m_lm_cat2_lag1)

shapiro.test(m_lm_cat2_lag1$residuals)
```

Hmmm, again the residuals deviate from a normal distribution particularly at the both tails of the distribution. I'm not so sure about using this model. Let's take a look at its ANOVA table using type 3 sum of squares.

```{r anova cat2 lag1}
Anova(m_lm_cat2_lag1, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

All the interaction terms are significant, so we can use this model in our model selection process; however, I'm somewhat leery of using this model since the model residuals don't look right.

```{r lm cat2 clean up}
rm(m_lm_cat2, m_lm_cat2_lag2, m_lm_cat2_lag3)
```


# Model selection - GAM and LM models

As a summary, here are the 6 models we are comparing:

* Model 1 - `m_gamm_cat3_ar1` - GAM 3-way interactions with s(DOY) <br>
Formula: `r format(m_gamm_cat3_ar1_gam$formula) %>% str_c(collapse = "") %>% str_squish()` <br><br>
* Model 2 - `m_lm_cat3_wt_lag1` - LM 3-way interactions with water temperature <br>
Formula: `r format(m_lm_cat3_wt_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br><br>
* Model 3 - `m_lm_cat3_lag1` - LM 3-way interactions without seasonal term <br>
Formula: `r format(m_lm_cat3_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br><br>
* Model 4 - `m_gamm_cat2_ar1` - GAM 2-way interactions with s(DOY) <br>
Formula: `r format(m_gamm_cat2_ar1_gam$formula) %>% str_c(collapse = "") %>% str_squish()` <br><br>
* Model 5 - `m_lm_cat2_wt_lag1` - LM 2-way interactions with water temperature <br>
Formula: `r format(m_lm_cat2_wt_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br><br>
* Model 6 - `m_lm_cat2_lag1` - LM 2-way interactions without seasonal term <br>
Formula: `r format(m_lm_cat2_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)` <br>

```{r model sel gam, warning = FALSE}
# AIC values
df_m_aic <- 
  AIC(
    m_gamm_cat3_ar1$lme,
    m_lm_cat3_wt_lag1,
    m_lm_cat3_lag1,
    m_gamm_cat2_ar1$lme,
    m_lm_cat2_wt_lag1,
    m_lm_cat2_lag1
  ) %>% 
  as_tibble(rownames = "Model")

# BIC values
df_m_bic <- 
  BIC(
    m_gamm_cat3_ar1$lme,
    m_lm_cat3_wt_lag1,
    m_lm_cat3_lag1,
    m_gamm_cat2_ar1$lme,
    m_lm_cat2_wt_lag1,
    m_lm_cat2_lag1
  ) %>% 
  as_tibble(rownames = "Model")

# Combine AIC and BIC
df_m_aic_bic <- left_join(df_m_aic, df_m_bic, by = join_by(Model, df))

# Sort by AIC
df_m_aic_bic %>% arrange(AIC)

# Sort by BIC
df_m_aic_bic %>% arrange(BIC)
```

According to AIC, model 3 (LM 3-way interactions without seasonal term) was the best model, while BIC preferred model 6 (LM 2-way interactions without seasonal term). We'll select model 3 (LM 3-way interactions without seasonal term) because the 3-way interaction between categorical variables was significant in this model indicating that there are possible significant differences between Flow Pulse Periods within each Station-Year grouping. 

# Model Results

## Model 3

Lets take a closer look at model 3: LM 3-way interactions without seasonal term <br>
Formula: `r format(m_lm_cat3_lag1$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)`

### Pairwise Contrasts

Tukey pairwise contrasts of Flow Pulse Period for each Station - Year combination:

```{r lm cat3 lag1 tukey flow action period, warning = FALSE}
em_lm_cat3_lag1 <- emmeans(m_lm_cat3_lag1, ~ FlowActionPeriod | StationCode * Year_fct)
pairs(em_lm_cat3_lag1)
```

### Summary Figure

```{r lm cat3 lag 1 summary figure, fig.width = 8.5, fig.height = 7.5}
df_chla_wt_lag %>% 
  drop_na(Chla, lag1) %>% 
  plot_model_summary(em_lm_cat3_lag1)
```

Observed daily average chlorophyll fluorescence values (boxplots) and model results (model means as red points ±95% confidence intervals as gray boxes) for the Flow Pulse Period comparisons by Station and Year. Different letters above boxplots identify statistically significant (p < 0.05) differences from a Tukey post-hoc test.

The model results don't match the observed values well at all. Unfortunately, this linear model is not a good candidate to use for our analysis.

## Model 1

Since the linear model using 3-way interactions doesn't fit the observed data that well, let's take a closer look at the GAM version of this model, model 1: GAM 3-way interactions with s(DOY) <br>
Formula: `r format(m_gamm_cat3_ar1_gam$formula) %>% str_c(collapse = "") %>% str_squish()`

### Pairwise Contrasts

Tukey pairwise contrasts of Flow Pulse Period for each Station - Year combination:

```{r gam cat3 ar1 tukey flow action period, warning = FALSE}
# Add the model call back to the gam object so it works with emmeans
m_gamm_cat3_ar1_gam$call <- quote(
  gamm(
    f_gam_cat3, 
    data = df_chla_wt_c, 
    correlation = corARMA(form = ~ 1|Year_fct/StationCode, p = 1),
    method = "REML"
  )
)

em_gamm_cat3_ar1 <- emmeans(m_gamm_cat3_ar1_gam, ~ FlowActionPeriod | StationCode * Year_fct)
pairs(em_gamm_cat3_ar1)
```

### Summary Figure

```{r gam cat3 ar1 summary figure, fig.width = 8.5, fig.height = 7.5}
plot_model_summary(df_chla_wt_c, em_gamm_cat3_ar1)
```

Observed daily average chlorophyll fluorescence values (boxplots) and model results (model means as red points ±95% confidence intervals as gray boxes) for the Flow Pulse Period comparisons by Station and Year. Different letters above boxplots identify statistically significant (p < 0.05) differences from a Tukey post-hoc test.

The model means seem to match the observed values better than Model 3, but there is a large amount of uncertainty around the means resulting in the model not seeing any significant differences among the pairwise comparisons. Unfortunately, this linear model does not seem like a good candidate to use for our analysis as well.

