---
title: "NDFS Synthesis Manuscript: Chlorophyll analysis"
subtitle: "Results from Final Model"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Look at results of the preferred model during model selection: Linear model with 3-way interactions between categorical variables - Year, Station, and Flow Pulse Period - without seasonal term.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(car)
library(emmeans)
library(multcomp)
library(here)
library(conflicted)

# Declare package conflict preferences 
conflicts_prefer(dplyr::filter(), dplyr::select())
```

Display current versions of R and packages used for this analysis:

```{r print session info}
devtools::session_info()
```

# Import Model and Data

```{r import data}
# Define file path for model and data used in the model
fp_model <- here("manuscript_synthesis/model")

# Import model and data used in the model
m_lm_cat3 <- read_rds(file.path(fp_model, "chla_cat3_lm_model.rds"))
df_chla <- read_rds(file.path(fp_model, "chla_model_data.rds"))
```

# Model Results

Model formula: `r format(m_lm_cat3$call) %>% str_c(collapse = "") %>% str_squish() %>% str_sub(start = 14, end = -12)`

## ANOVA table

```{r anova lm cat3, warning = FALSE}
Anova(m_lm_cat3, type = 3, contrasts = list(topic = contr.sum, sys = contr.sum))
```

## Pairwise Contrasts

Tukey pairwise contrasts of Flow Pulse Period for each Station - Year combination:

```{r tukey flow action period, warning = FALSE}
em_cat3 <- emmeans(m_lm_cat3, ~ FlowActionPeriod | StationCode * Year_fct)
pairs(em_cat3)
```

## Summary Figure

```{r summary figure, fig.width = 8.5, fig.height = 7.5}
# Add significance grouping letters from the Tukey post-hoc results and back
  # transform log-transformed results
df_cat3_tuk <- em_cat3 %>% 
  cld(sort = FALSE, Letters = letters) %>% 
  as_tibble() %>% 
  mutate(
    group = str_remove_all(.group, fixed(" ")),
    across(c(emmean, lower.CL, upper.CL), ~ exp(.x) / 1000),
    Year = as.numeric(as.character(Year_fct))
  ) %>% 
  select(
    StationCode,
    Year,
    FlowActionPeriod,
    emmean,
    lower.CL,
    upper.CL,
    group
  )

# Determine vertical positioning of letters for figure
# Calculate min and max values of observed data for each Station - Year group
df_chla_summ <- df_chla %>% 
  summarize(
    max_val = max(Chla),
    min_val = min(Chla),
    .by = c(StationCode, Year)
  )

# Add min and max values of observed data to the Tukey post-hoc results and
  # calculate vertical positioning of letters
df_cat3_tuk_c <- df_cat3_tuk %>% 
  left_join(df_chla_summ, by = join_by(StationCode, Year)) %>% 
  mutate(max_val = if_else(upper.CL > max_val, upper.CL, max_val)) %>% 
  group_by(StationCode, Year) %>% 
  mutate(max_val = max(max_val)) %>% 
  ungroup() %>% 
  mutate(y_pos = max_val + (max_val - min_val) / 10)

# Create summary figure
df_cat3_tuk_c %>%
  ggplot(
    aes(
      x = FlowActionPeriod,
      y = emmean,
      ymin = lower.CL,
      ymax = upper.CL,
      label = group
    )
  ) +
  geom_boxplot(
    data = df_chla,
    aes(x = FlowActionPeriod, y = Chla),
    inherit.aes = FALSE
  ) +
  geom_crossbar(color = "grey82", fill = "grey", alpha = 0.7, linewidth = 0.1) +
  geom_point(color = "red") +
  geom_text(aes(y = y_pos), size = 3) +
  facet_wrap(
    vars(StationCode, Year), 
    scales = "free_y",
    nrow = 4,
    labeller = labeller(.multi_line = FALSE)
  ) +
  xlab("Flow Pulse Period") +
  ylab(expression(Chlorophyll~Fluoresence~(mu*g~L^{-1}))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Observed daily average chlorophyll fluorescence values (boxplots) and model results (model means as red points ±95% confidence intervals as gray boxes) for the Flow Pulse Period comparisons by Station and Year. Different letters above boxplots identify statistically significant (p < 0.05) differences from a Tukey post-hoc test.

