---
title: "Summary Boxplots for the NDFA Synthesis report"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Create boxplots summarizing daily average upstream and downstream continuous values separated by low- and high-flow pulse years. These are intended to be used as summary figures in the NDFA Synthesis report, but may or may not be included in the end. These include data for all years in the study period (2011-2019).

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(scales)
```

```{r load functions, message = FALSE, warning = FALSE}
# Source functions
source("global_ndfa_funcs.R")
source("Water_Quality/global_wq_funcs.R")
```

# Import and Prepare Data

```{r import data}
# Define relative file path for file containing all QA'ed and cleaned continuous WQ data
fp_rel_rtm_data <- "WQ_Subteam/Processed_Data/Continuous/RTM_INPUT_all_2021-04-20.csv"

# Define absolute file path
fp_abs_rtm_data <- ndfa_abs_sp_path(fp_rel_rtm_data)

# Import continuous WQ data
df_rtm_orig <- import_rtm_data(fp_abs_rtm_data, 10)
```

```{r prepare data, message = FALSE}
# Clean original continuous WQ data and calculate daily averages
df_rtm_clean <- df_rtm_orig %>% 
  # Only keep a subset of the parameters (Chla, DO, pH, SpCnd, Turbidity, WaterTemp)
  select(StationCode, DateTime, Chla, DO, pH, SpCnd, Turbidity, WaterTemp) %>%
  mutate(
    DateTime = ymd_hms(DateTime),
    Date = date(DateTime),
    Year = year(DateTime)
  ) %>%
  # Exclude SDI, SGG, and SRH from the plots
  filter(!StationCode %in% c("SDI", "SGG", "SRH")) %>%
  # Pivot parameters in the long format
  pivot_longer(cols = where(is.numeric) & !Year, names_to = "Parameter", values_to = "Value") %>%
  # Remove all NA values
  filter(!is.na(Value)) %>%
  # Calculate daily averages
  group_by(Parameter, Year, StationCode, Date) %>%
  summarize(Daily_avg = mean(Value)) %>%
  ungroup() %>% 
  # Add variables for Region, Flow Pulse Period, and Pulse Type
  ndfa_action_periods() %>% 
  mutate(
    Region = if_else(
      StationCode %in% c("RMB", "RCS", "RD22", "I80", "LIS", "TOE", "STTD"),
      "Upstream",
      "Downstream"
    ),
    PulseType = if_else(Year %in% c(2013, 2014, 2017), "Low-flow", "High-flow")
  ) %>% 
  # Convert PulseType, FlowActionPeriod, and Region to factors
  mutate(
    PulseType = factor(PulseType, levels = c("Low-flow", "High-flow")),
    FlowActionPeriod = factor(FlowActionPeriod, levels = c("Before", "During", "After")),
    Region = factor(Region, levels = c("Upstream", "Downstream"))
  )
```

# Create Functions

```{r int plot function}
# Internal function to define y-axis labels in plots
int_define_yaxis_lab <- function(param) {
  yaxis_lab <- case_when(
    param == "Chla" ~ "Chlorophyll (µg/L)",
    param == "DO" ~ "Dissolved Oxygen (mg/L)",
    param == "fDOM" ~ "fDOM (µg/L as QSE)",
    param == "NitrateNitrite" ~ "Nitrate + Nitrite (mg/L as N)",
    param == "pH" ~ "pH (pH units)",
    param == "SpCnd" ~ "Specific Conductance (µS/cm)",
    param == "Turbidity" ~ "Turbidity (FNU)",
    param == "WaterTemp" ~ "Water Temperature (degrees C)"
  )
  
  return(yaxis_lab)
}
```

```{r boxplot function sep fa}
# Function to create a summary boxplot separated by flow pulse period on x-axis
create_boxplot_fa <- function(df, param) {
  # define y-axis label
  y_lab <- int_define_yaxis_lab(param)
  
  # create plot
  p <- df %>% 
    ggplot(aes(x = FlowActionPeriod, y = Daily_avg)) +
    geom_boxplot() +
    facet_grid(
      rows = vars(Region), 
      cols = vars(PulseType),
      scales = "free_y"
    ) +
    stat_summary( 
      fun = mean, 
      color = "red", 
      geom = "point", 
      shape = 8, 
      size = 2 
    ) +
    theme_light() +
    theme(
      strip.text = element_text(color = "black"),
      panel.grid.minor = element_blank()
    ) +
    xlab("Flow Pulse Period") +
    scale_y_continuous(name = y_lab, labels = label_comma())
  
  return(p)
}
```

```{r boxplot function sep region}
# Function to create a summary boxplot separated by region on x-axis
create_boxplot_region <- function(df, param) {
  # define y-axis label
  y_lab <- int_define_yaxis_lab(param)
  
  # create plot
  p <- df %>% 
    ggplot(aes(x = Region, y = Daily_avg)) +
    geom_boxplot() +
    facet_grid(cols = vars(PulseType)) +
    stat_summary( 
      fun = mean, 
      color = "red", 
      geom = "point", 
      shape = 8, 
      size = 2 
    ) +
    theme_light() +
    theme(
      strip.text = element_text(color = "black"),
      panel.grid.minor = element_blank()
    ) +
    xlab("Region") +
    scale_y_continuous(name = y_lab, labels = label_comma())
  
  return(p)
}
```

# Create Plots

```{r create boxplots sep fa}
df_boxplot_fa <- df_rtm_clean %>% 
  nest(df_data = -Parameter) %>% 
  mutate(plt = map2(df_data, Parameter, create_boxplot_fa))
```

```{r create boxplots sep region}
df_boxplot_region <- df_rtm_clean %>% 
  nest(df_data = -Parameter) %>% 
  mutate(plt = map2(df_data, Parameter, create_boxplot_region))
```

# View Plots

These are boxplots of medians and quartiles of daily average continuous values. The red asterisks represent the mean of each group.

## Separated by Flow Pulse Period {.tabset .tabset-pills}

```{r print plots sep fa, echo = FALSE, results = "asis", fig.width = 6.5}
for (i in 1:nrow(df_boxplot_fa)) {
  # Create subheadings for each Parameter
  cat("### ", df_boxplot_fa$Parameter[i], "\n\n")
  # Print plot
  print(df_boxplot_fa$plt[[i]])
  cat("\n\n")
}
```

## Not separated by Flow Pulse Period {.tabset .tabset-pills}

```{r print plots sep region, echo = FALSE, results = "asis", fig.width = 6.5, fig.height = 3}
for (i in 1:nrow(df_boxplot_region)) {
  # Create subheadings for each Parameter
  cat("### ", df_boxplot_region$Parameter[i], "\n\n")
  # Print plot
  print(df_boxplot_region$plt[[i]])
  cat("\n\n")
}
```

# Export Plots

```{r export plots}
# Define file path to export plots to
fp_abs_boxplot <- ndfa_abs_sp_path("WQ_Subteam/Plots/Continuous/Report/Boxplot_summary")

# Export boxplots separated by flow pulse period
walk2(
  df_boxplot_fa$plt,
  df_boxplot_fa$Parameter,
  ~ggsave(
    plot = .x,
    filename = paste0(fp_abs_boxplot, "/", .y, "_boxplot_by_flow_pulse.jpg"),
    width = 6.5, 
    height = 5, 
    units = "in", 
    dpi = 300
  )
)

# Export boxplots separated by region
walk2(
  df_boxplot_region$plt,
  df_boxplot_region$Parameter,
  ~ggsave(
    plot = .x,
    filename = paste0(fp_abs_boxplot, "/", .y, "_boxplot_by_region.jpg"),
    width = 6.5, 
    height = 3, 
    units = "in", 
    dpi = 300
  )
)
```

